---
title: "Monthly Satellite Soil Combined Three Cohort File"
author: "Gillian Goobie"
date: "12/22/2021"
output: 
  html_document: 
    toc: yes
    toc_float: yes
    number_sections: yes
    code_folding: hide
  bookdown::html_document2:
    toc: yes
    toc_float: yes
    number_sections: yes
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE, warning=FALSE, echo=F}
library(tidyverse)
library(readxl)
library(here)
library(writexl)
library(knitr)
library(lintr)
library(tidync)
library(maps)
library(ncmeta)
library(devtools)
library(RNetCDF)
library(ncdf4)
library(survival)
library(survminer)
library(lubridate)
library(cmprsk)
library(riskRegression)
library(prodlim)
library(RColorBrewer)
library(prodlim)
library(lme4)
library(performance)
library(psycho)
library(report)
library(rms)
library(splines)
library(Greg)
library(rstpm2)
library(pspline)
library(smoothHR)
```

# Importing Datasets
Here I am importing the file which contains cleaned Simmons Data with Soil Levels Matched
```{r}
outfile1 <- here("Simmons_fILD_Soil_BaselineData_2021_12_17.xlsx")
Simm <- read_excel(outfile1)
```

Here I am importing the file which contains cleaned PFF Data with Soil Levels Matched
```{r}
outfile2 <- here("PFF_fILD_Soil_BaselineData_ConsentDateRef_2021_12_17.xlsx")
PFF <- read_excel(outfile2)
```

Here I am importing the file which contains cleaned CARE-PF Data with Soil Levels Matched
```{r}
outfile3 <- here("CAREPF_fILD_Soil_BaselineData_2021_12_17.xlsx")
CARE <- read_excel(outfile3)
```

# Create Cohort Variable
```{r}
Simm <- Simm %>% mutate(cohort="Simmons")
PFF <- PFF %>% mutate(cohort="PFF")
CARE <- CARE %>% mutate(cohort="CARE-PF")
```

# Simplify included variables in each dataframe
```{r}
Simm <- Simm %>% rename("lat"="lat.x", "lon"="lon.x")
Simm <- Simm %>% dplyr::select(ID, sex, race, dich_Race, ethnicity, smokeHx, dich_smoking, age_dx, metro, cohort, lat, lon, ADI_nat, natADI_quart, dx, dx_group, dx_IPF, fvc_pct, dlco_pct, deadORtx, status, time_DeathTxCensor, dx_date, death_date, tx_date, pft_date,  Soil_dx, Soildx_quart, Soildx_dec, Soildx_dich, Soil_1yrPreCensor, Soil1yrCensor_quart, Soil1yrCensor_dec, Soil1yrCensor_dich, Soil_5yrPreCensor, Soil5yrCensor_quart, Soil5yrCensor_dec, Soil5yrCensor_dich, Soil_1yrPreDx, Soil1yr_quart, Soil1yr_dec, Soil1yr_dich, Soil_5yrPreDx, Soil5yr_quart, Soil5yr_dec, Soil5yr_dich)
str(Simm)
```

```{r}
PFF$ID <- str_pad(PFF$ID, width=9, side="right", pad="0")
PFF$ID <- as.numeric(PFF$ID)
PFF$deadORtx <- as.character(PFF$deadORtx)
PFF$status <- as.numeric(PFF$status)
PFF$dich_smoking <- PFF$smokeHx
PFF <- PFF %>% rename("lat"="lat.x", "lon"="lon.x")
PFF <- PFF %>% dplyr::select(ID, sex, race, dich_Race, ethnicity, smokeHx, dich_smoking, age_dx, metro, cohort, lat, lon, dx, dx_group, dx_IPF, fvc_pct, dlco_pct, deadORtx, status, time_DeathTxCensor, dx_date, death_date, tx_date, fvc_date, dlco_date, Soil_dx, Soildx_quart, Soildx_dec, Soildx_dich, Soil_1yrPreCensor, Soil1yrCensor_quart, Soil1yrCensor_dec, Soil1yrCensor_dich, Soil_5yrPreCensor, Soil5yrCensor_quart, Soil5yrCensor_dec, Soil5yrCensor_dich, Soil_1yrPreDx, Soil1yr_dec, Soil1yr_dich, Soil_5yrPreDx, Soil5yr_quart, Soil5yr_dec, Soil5yr_dich)
str(PFF)
```

```{r}
CARE <- CARE %>% rename("lat"="lat.x")
CARE <- CARE %>% dplyr::select(ID, sex, race, dich_Race, ethnicity, smokeHx, dich_smoking, age_dx, metro, cohort, lat, lon, dx, dx_group, dx_IPF, fvc_pct, dlco_pct, deadORtx, status, time_DeathTxCensor, dx_date, death_date, tx_date, pft_date, Soil_dx, Soildx_quart, Soildx_dec, Soildx_dich, Soil_1yrPreCensor, Soil1yrCensor_quart, Soil1yrCensor_dec, Soil1yrCensor_dich, Soil_5yrPreCensor, Soil5yrCensor_quart, Soil5yrCensor_dec, Soil5yrCensor_dich, Soil_1yrPreDx, Soil1yr_quart, Soil1yr_dec, Soil1yr_dich, Soil_5yrPreDx, Soil5yr_quart, Soil5yr_dec, Soil5yr_dich)
str(CARE)
```

# Combine Cohorts into one file
```{r}
Soil <- full_join(Simm, CARE)
Soil <- full_join(Soil, PFF)
```


# Visual EDA
Overlapping histograms of Soil by Cohort
```{r}
(Soil %>% ggplot(aes(x=Soil_dx, group=cohort, fill=cohort))+
    geom_density(adjust=1.5, alpha=0.4)+
   scale_fill_manual(values=c("#FF0000", "#00B0F0", "#FFFF00"))+
   labs(x="Soil During Disease", y="Density of Patients in Each Cohort", title="Soil Exposure During Disease by Cohort")+
   theme(plot.title = element_text(hjust = 0.5))+
    xlim(0,3))
```
PFF has some higher outliers.

Overlapping histograms of Soil by Cohort
```{r}
(Soil %>% ggplot(aes(x=Soil_5yrPreCensor, group=cohort, fill=cohort))+
    geom_density(adjust=1.5, alpha=0.4)+
   scale_fill_manual(values=c("#FF0000", "#00B0F0", "#FFFF00"))+
   labs(x="Soil in 5yrs Pre-Censoring", y="Density of Patients in Each Cohort", title="Soil Exposure in 5yrs Pre-Censoring by Cohort")+
   theme(plot.title = element_text(hjust = 0.5))+
    xlim(0,3))
```
PFF has some higher outliers.

Overlapping histograms of Soil by Cohort
```{r}
(Soil %>% ggplot(aes(x=Soil_5yrPreDx, group=cohort, fill=cohort))+
    geom_density(adjust=1.5, alpha=0.4)+
   scale_fill_manual(values=c("#FF0000", "#00B0F0", "#FFFF00"))+
   labs(x="Soil in 5yrs Pre-Diagnosis", y="Density of Patients in Each Cohort", title="Soil Exposure in 5yrs Pre-Diagnosis by Cohort")+
   theme(plot.title = element_text(hjust = 0.5))+
    xlim(0,3))
```
PFF has some higher outliers.