---
title: "Monthly Satellite PM Analysis with CARE-PF Data - Revision 1 Final Analysis"
author: "Gillian Goobie"
date: "07/27/2022"
output: 
  html_document: 
    toc: yes
    toc_float: yes
    number_sections: yes
    code_folding: hide
  bookdown::html_document2:
    toc: yes
    toc_float: yes
    number_sections: yes
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE, warning=FALSE, echo=F}
library(tidyverse)
library(readxl)
library(here)
library(writexl)
library(knitr)
library(lintr)
library(tidync)
library(maps)
library(ncmeta)
library(devtools)
library(RNetCDF)
library(ncdf4)
library(survival)
library(survminer)
library(lubridate)
library(cmprsk)
library(riskRegression)
library(prodlim)
library(RColorBrewer)
library(prodlim)
library(lme4)
library(performance)
library(psycho)
library(report)
library(rms)
library(splines)
library(Greg)
library(rstpm2)
library(pspline)
library(smoothHR)
library(raster)
```

# Importing Datasets
Here I am importing the file which contains monthly PM level estimates by satellite at nearest lon/lat to CAREPF patient residential addresses. These are linked to the patient ID.
```{r}
outfile1 <- here("MonthlyPM25Data/CAREPF_fILD_2000_2018_PM25_2021_09_16.xlsx")
PM <- read_excel(outfile1)
```

Here I am importing the file which I used for my CIMD work that contains the baseline clinical and demographic data for 3389 patients who have CIMD and full records otherwise
```{r}
outfile2 <- here("MonthlyPM25Data/CAREPF_fILDPts_BaselineData_2021_10_22.xlsx")
CARE <- read_excel(outfile2)
```

# Merging Datasets
I used a inner_join here so that the complete "PM" dataframe only includes patients with fILD that have all baseline demographics and CIMD available as well as PM data.
```{r}
PM <- inner_join(CARE, PM, by="ID")
```

# Extracting year of diagnosis and year of death/transplant/censoring
Since the PM Satellite data is only averages by month, I want to create a column for each patient where just the month/year of diagnosis and the month/year of death or lung transplant is extracted. Then I can use this to determine the average PM exposure in the 5yrs prior to diagnosis and in the years from diagnosis until death/tx.
```{r}
#Start with the year of diagnosis
PM <- PM %>% 
  mutate(dx_yrmo = format(as.Date(PM$dx_date, format="%Y-%m-%d"),"%Y-%m"))
PM <- PM %>% 
  mutate(dx_yr = format(as.Date(PM$dx_date, format="%Y-%m-%d"),"%Y"))
PM$dx_yr <- as.numeric(PM$dx_yr)

#Then the year of death or lung transplant
PM <- PM %>% 
  mutate(deathORtx_date = if_else(!is.na(tx_date), tx_date, death_date))
PM <- PM %>% 
  mutate(deathORtx_yrmo = format(as.Date(PM$deathORtx_date, format="%Y-%m-%d"),"%Y-%m"))

#Then the year the records were last updated (i.e. year of censoring)
PM <- PM %>% 
  mutate(DeathTxCensor_date = if_else(!is.na(deathORtx_date), deathORtx_date, last_updated))
PM <- PM %>% 
  mutate(censor_yrmo = format(as.Date(PM$DeathTxCensor_date, format="%Y-%m-%d"),"%Y-%m"))
```


# Censored vs Lost to Followup
Determining how many patients were lost to followup vs censored with data extraction
```{r}
PMx <- PM %>% filter(status==0)
PMx <- PMx %>% 
  mutate_at(c("last_updated","reg_update"), as.Date)
str(PMx$last_updated)
str(PMx$reg_update)
```

```{r}
#Patients were censored for data collection on March 23, 2021
PMx <- PMx %>% mutate(censor_date=as.Date("2021-03-23"))
str(PMx$censor_date)
```


If no last_updated within 1yr, considered lost to f/u
```{r}
PMx <- PMx %>% mutate(diff_censor_lastvisit=(last_updated-censor_date))
str(PMx$diff_censor_lastvisit)
PMx <- PMx %>% mutate(losttofu=ifelse((diff_censor_lastvisit>365 | is.na(diff_censor_lastvisit)), 1, 0))
PMx$losttofu <- as.factor(PMx$losttofu)
str(PMx$losttofu)
table(PMx$losttofu)
table(PM$status)
```
So 335 patients were seen within a year of the censor date and 174 were seen >1yr past the censor date, indicating lost to followup.


# Pivoting to Long Format
First we need to convert the PM dataframe into the long rather than the wide format, which will allow us to use it more easily in R's tidyverse as this is "tidy" formatting.
```{r}
PM <- PM %>% 
  pivot_longer(cols=c(66:292), names_to="PM_date", names_prefix="PM25_")
```

## Convert date PM_date to same format as above
```{r}
PMx <- PM 
PMx$PM_date <- gsub("jan", "01-01-20", PMx$PM_date)
PMx$PM_date <- gsub("feb", "01-02-20", PMx$PM_date)
PMx$PM_date <- gsub("mar", "01-03-20", PMx$PM_date)
PMx$PM_date <- gsub("apr", "01-04-20", PMx$PM_date)
PMx$PM_date <- gsub("may", "01-05-20", PMx$PM_date)
PMx$PM_date <- gsub("jun", "01-06-20", PMx$PM_date)
PMx$PM_date <- gsub("jul", "01-07-20", PMx$PM_date)
PMx$PM_date <- gsub("aug", "01-08-20", PMx$PM_date)
PMx$PM_date <- gsub("sep", "01-09-20", PMx$PM_date)
PMx$PM_date <- gsub("oct", "01-10-20", PMx$PM_date)
PMx$PM_date <- gsub("nov", "01-11-20", PMx$PM_date)
PMx$PM_date <- gsub("dec", "01-12-20", PMx$PM_date)

PMx$PM_date <- format(as.Date(PMx$PM_date, format="%d-%m-%Y"),"%Y-%m-%d")
PMx$PM_date <- as.Date(PMx$PM_date)
PM <- PMx
```


# Convert Date Columns to Date Format
Next I need to convert all date columns to proper format
```{r}
PM <- PM %>% 
  mutate_at(c("dob", "dx_date", "initial_visit_date", "formal_dx_date", "reg_update", "death_date", "tx_date", "pft_date", "last_updated", "deathORtx_date", "DeathTxCensor_date"), as.Date)
str(PM)
```

# Creating Exposure Variables
Now I'm creating new variables where I am matching up PM averages per year to years of major events for patients (year of diagnosis, year of death/lung transplant/censoring, etc)

## 5yrs Pre-Death/Transplant/Censoring
Here I am calculating the average PM value in the 5yrs prior to death/lung transplant/censoring.
```{r}
PM <- PM %>% 
  group_by(ID) %>% 
  mutate(PM_5yrPreCensor = mean(value[ymd(PM_date)>=(ymd(DeathTxCensor_date) - years(5)) & 
                                        ymd(PM_date)<=ymd(DeathTxCensor_date)]))
           
           
#mean(value[ymd(PM_date) == ymd(DeathTxCensor_date) | ymd(PM_date) == (ymd(DeathTxCensor_date) - years(1))]))
```

### 5yrs Pre-Censoring Warm Only (April-Sept Months Averaged)
Filter out warm months only
```{r}
PM_warm <- PM %>% mutate(month=str_sub(PM_date, -5, -4)) 
str(PM_warm$month)
warm_months <- c("04","05","06","07","08","09")
PM_warm <- PM_warm %>% filter(month %in% warm_months)
```

Average of 5yr Pre-Censoring PM2.5 in Warm Months
```{r}
PM_warm <- PM_warm %>% 
  group_by(ID) %>% 
  mutate(PM_5yrPreCensor = mean(value[ymd(PM_date)>=(ymd(DeathTxCensor_date) - years(5)) & 
                                        ymd(PM_date)<=ymd(DeathTxCensor_date)]))
```

### 5yrs Pre-Censoring Winter Only (Oct-March Months Averaged)
Filter out cold months only
```{r}
PM_cold <- PM %>% mutate(month=str_sub(PM_date, -5, -4)) 
str(PM_cold$month)
cold_months <- c("01","02","03","10","11","12")
PM_cold <- PM_cold %>% filter(month %in% cold_months)
```

Average of 5yr Pre-Censoring PM2.5 in cold Months
```{r}
PM_cold <- PM_cold %>% 
  group_by(ID) %>% 
  mutate(PM_5yrPreCensor = mean(value[ymd(PM_date)>=(ymd(DeathTxCensor_date) - years(5)) & 
                                        ymd(PM_date)<=ymd(DeathTxCensor_date)]))
```


## 5 years Prior to Diagnosis
Here I am calculating the average PM value in the year of diagnosis and the 5 years leading up to diagnosis (so if all years available, it will be the average of 6 years).
```{r}
PM <- PM %>% 
  group_by(ID) %>% 
  mutate(PM_5yrPreDx = mean(value[ymd(PM_date) <= ymd(dx_date) & 
                                    ymd(PM_date) >= (ymd(dx_date)-years(5))]))
```

### 5yrs Pre-Diagnosis Warm Months Only
```{r}
PM_warm <- PM_warm %>% 
  group_by(ID) %>% 
  mutate(PM_5yrPreDx = mean(value[ymd(PM_date) <= ymd(dx_date) & 
                                    ymd(PM_date) >= (ymd(dx_date)-years(5))]))
```

### 5yrs Pre-Diagnosis Cold Months Only
```{r}
PM_cold <- PM_cold %>% 
  group_by(ID) %>% 
  mutate(PM_5yrPreDx = mean(value[ymd(PM_date) <= ymd(dx_date) & 
                                    ymd(PM_date) >= (ymd(dx_date)-years(5))]))
```

## Remove Dataframes not in use
```{r}
rm(list=c("PMx", "CARE"))
```

## Getting Rid of Duplicated IDs
Now that we have calculated the patient-specific PM exposures, I can get rid of all rows other than the first row for each patient.
```{r}
PM <-  PM %>% 
  distinct_at(vars(ID), .keep_all=T)
```
This takes us down to our  observations that we had before.

```{r}
PM_warm <-  PM_warm %>% 
  distinct_at(vars(ID), .keep_all=T)
```

```{r}
PM_cold <-  PM_cold %>% 
  distinct_at(vars(ID), .keep_all=T)
```


# Creating Low vs High Exposure Groups

## PM 5yrs Pre-Censor High vs Low
```{r}
PM$PM5yrCensor_dich <- cut(PM$PM_5yrPreCensor,
                      breaks=c(0, 8, 50),
                      labels=c("Low", "High"))
summary(PM$PM5yrCensor_dich)
class(PM$PM5yrCensor_dich)
```

```{r}
PM_warm$PM5yrCensor_dich <- cut(PM_warm$PM_5yrPreCensor,
                      breaks=c(0, 8, 50),
                      labels=c("Low", "High"))
summary(PM_warm$PM5yrCensor_dich)
class(PM_warm$PM5yrCensor_dich)
```

```{r}
PM_cold$PM5yrCensor_dich <- cut(PM_cold$PM_5yrPreCensor,
                      breaks=c(0, 8, 50),
                      labels=c("Low", "High"))
summary(PM_cold$PM5yrCensor_dich)
class(PM_cold$PM5yrCensor_dich)
```

## PM 5yrs Pre-Dx High vs Low
```{r}
PM$PM5yr_dich <- cut(PM$PM_5yrPreDx,
                      breaks=c(0, 8, 50),
                      labels=c("Low", "High"))
summary(PM$PM5yr_dich)
class(PM$PM5yr_dich)
```


```{r}
PM_warm$PM5yr_dich <- cut(PM_warm$PM_5yrPreDx,
                      breaks=c(0, 8, 50),
                      labels=c("Low", "High"))
summary(PM_warm$PM5yr_dich)
class(PM_warm$PM5yr_dich)
```

```{r}
PM_cold$PM5yr_dich <- cut(PM_cold$PM_5yrPreDx,
                      breaks=c(0, 8, 50),
                      labels=c("Low", "High"))
summary(PM_cold$PM5yr_dich)
class(PM_cold$PM5yr_dich)
```



# Creating Other New Variables
## IPF vs Other Diagnosis
```{r}
PM <- PM %>% 
  mutate(dx_IPF=ifelse(dx=="IPF", "IPF", "not_IPF"))    
```

# Releveling Factors
Here I am releveling factors so that they are in an intuitive order for my later analyses.
```{r}
PM$site <- as.factor(PM$site)
PM$sex <- fct_relevel(PM$sex, c("M","F"))
PM$race <- fct_relevel(PM$race, c("W","B","A","N","U"))
PM$ethnicity <- fct_relevel(PM$ethnicity, c("N","H","D","U"))
PM$dich_Race <- fct_relevel(PM$dich_Race, c("White","Non-White"))
PM$smokeHx <- fct_relevel(PM$smokeHx, c("Never","Former","Always"))

#For dx and dx_group, I just want IPF to be first and then the rest of the categories are alphabetical
PM$dx <- fct_relevel(PM$dx, c("IPF"))
PM$dx_group <- fct_relevel(PM$dx_group, c("IPF"))
PM$dx_IPF <- fct_relevel(PM$dx_IPF, c("IPF"))
str(PM)
```

```{r}
PM_warm$site <- as.factor(PM_warm$site)
PM_warm$sex <- fct_relevel(PM_warm$sex, c("M","F"))
PM_warm$race <- fct_relevel(PM_warm$race, c("W","B","A","N","U"))
PM_warm$ethnicity <- fct_relevel(PM_warm$ethnicity, c("N","H","D","U"))
PM_warm$dich_Race <- fct_relevel(PM_warm$dich_Race, c("White","Non-White"))
PM_warm$smokeHx <- fct_relevel(PM_warm$smokeHx, c("Never","Former","Always"))

#For dx and dx_group, I just want IPF to be first and then the rest of the categories are alphabetical
PM_warm$dx <- fct_relevel(PM_warm$dx, c("IPF"))
PM_warm$dx_group <- fct_relevel(PM_warm$dx_group, c("IPF"))
```

```{r}
PM_cold$site <- as.factor(PM_cold$site)
PM_cold$sex <- fct_relevel(PM_cold$sex, c("M","F"))
PM_cold$race <- fct_relevel(PM_cold$race, c("W","B","A","N","U"))
PM_cold$ethnicity <- fct_relevel(PM_cold$ethnicity, c("N","H","D","U"))
PM_cold$dich_Race <- fct_relevel(PM_cold$dich_Race, c("White","Non-White"))
PM_cold$smokeHx <- fct_relevel(PM_cold$smokeHx, c("Never","Former","Always"))

#For dx and dx_group, I just want IPF to be first and then the rest of the categories are alphabetical
PM_cold$dx <- fct_relevel(PM_cold$dx, c("IPF"))
PM_cold$dx_group <- fct_relevel(PM_cold$dx_group, c("IPF"))
```

# Creating Datasets with only Metropolitan Patients
```{r}
PM_metro <- PM %>% filter(metro=="metropolitan")
```
This results in a dataset of  patients

# Creating Datasets with only IPF Patients
```{r}
PM_IPF <- PM %>% filter(dx=="IPF")
```
This results in a dataset of  pts

# Creating Datasets with only CTD Patients
```{r}
PM_CTD <- PM %>% filter(dx_group=="CTD-ILD")
```
This results in a dataset of  pts

# Creating individual site subgroups
```{r}
PM_toronto <- PM %>% filter(site==101)
PM_ab <- PM %>% filter(site==102)
PM_sph <- PM %>% filter(site==103)
PM_vgh <- PM %>%  filter(site==104)
PM_bc <- PM %>% filter(site==103 | site==104)
PM_laval <- PM %>% filter(site==105)
PM_hamilton <- PM %>% filter(site==106)
PM_on <- PM %>% filter(site==101 | site==106)
PM_mcgill <- PM %>% filter(site==107)
PM_qc <- PM %>% filter(site==105 | site==107)
PM_sk <- PM %>% filter(site==108)
```


# Export Excel File with PM Data Matched to Patients and Full Cohort Data
```{r}
write_xlsx(PM, path="Final_CAREPF_fILD_PM_BaselineData_2022_02_12.xlsx")
```

```{r}
write_xlsx(PM_warm, path="Final_CAREPF_fILD_PMwarm_BaselineData_2022_02_12.xlsx")
```

```{r}
write_xlsx(PM_cold, path="Final_CAREPF_fILD_PMcold_BaselineData_2022_02_12.xlsx")
```

# Exploratory Data Analysis
This is a function that allows me to make tables which summarize the count and percentages of each level of factor variables
```{r}
n_prop_tbl <- function(x) {
  tbl <- table(x)
  res <- cbind(tbl, round(prop.table(tbl)*100,2))
  colnames(res) <-  c('Count', 'Percentage')
  res
}
```


## PM Breakdown
```{r}
summary(PM$PM_5yrPreCensor)
summary(PM$PM_5yrPreDx)
summary(PM_warm$PM_5yrPreCensor)
summary(PM_warm$PM_5yrPreDx)
summary(PM_cold$PM_5yrPreCensor)
summary(PM_cold$PM_5yrPreDx)
```


## Sex Breakdown
```{r}
n_prop_tbl(PM$sex)
```

## Race and Ethnicity Breakdown 
```{r}
n_prop_tbl(PM$race)

n_prop_tbl(PM$ethnicity)
```

## Smoking History Breakdown
```{r}
n_prop_tbl(PM$smokeHx)
summary(PM$pkyrs)
```


## Diagnostic Group Breakdown
```{r}
n_prop_tbl(PM$dx_group)
```


## Province and Site Breakdown
```{r}
n_prop_tbl(PM$province)
n_prop_tbl(PM$site)
```

### Metropolitan Breakdown
```{r}
n_prop_tbl(PM$metro)
```
69% are metropolitan

## Age at Diagnosis Breakdown
```{r}
shapiro.test(PM$age_dx)
#Shapiro tests tells us that age_dx is not normally distributed, so should report median, IQR
summary(PM$age_dx)
```

## CIMD Breakdown
```{r}
summary(PM$avg_s)
```


## Vital Status Breakdown
```{r}
n_prop_tbl(PM$status)
```
Majority of the patients in CARE-PF are still alive.

## Baseline Lung Function Breakdown
```{r}
shapiro.test(PM$fvc_pct)
summary(PM$fvc_pct)
sd(PM$fvc_pct, na.rm=T)
```

```{r}
shapiro.test(PM$dlco_pct)
summary(PM$dlco_pct)
sd(PM$dlco_pct, na.rm=T)
```

## Time to Censoring Breakdown
```{r}
summary(PM$time_DeathTxCensor)
sd(PM$time_DeathTxCensor, na.rm=T)
```


## Breakdown of Characteristics by Low vs High
### PM_5yrPreCensor Low vs High
```{r}
table(PM$sex, PM$PM5yrCensor_dich)
table(PM$race, PM$PM5yrCensor_dich)
table(PM$dich_Race, PM$PM5yrCensor_dich)
table(PM$dx_group, PM$PM5yrCensor_dich)
table(PM$smokeHx, PM$PM5yrCensor_dich)
table(PM$metro, PM$PM5yrCensor_dich)
table(PM$province, PM$PM5yrCensor_dich)
table(PM$status, PM$PM5yrCensor_dich)
group_by(PM, PM5yrCensor_dich) %>%
  summarise(
    count = n(),
    mean = mean(age_dx, na.rm = TRUE),
    sd = sd(age_dx, na.rm = TRUE),
    median = median(age_dx, na.rm = TRUE),
    quantile = quantile(age_dx, probs=c(0.25), na.rm = TRUE),
    quantile2 = quantile(age_dx, probs=c(0.75), na.rm = TRUE)
  )
group_by(PM, PM5yrCensor_dich) %>%
  summarise(
    count = n(),
    mean = mean(avg_s, na.rm = TRUE),
    sd = sd(avg_s, na.rm = TRUE),
    median = median(avg_s, na.rm = TRUE),
    quantile = quantile(avg_s, probs=c(0.25), na.rm = TRUE),
    quantile2 = quantile(avg_s, probs=c(0.75), na.rm = TRUE)
  )
group_by(PM, PM5yrCensor_dich) %>%
  summarise(
    count = n(),
    mean = mean(fvc_pct, na.rm = TRUE),
    sd = sd(fvc_pct, na.rm = TRUE),
    median = median(fvc_pct, na.rm = TRUE),
    quantile = quantile(fvc_pct, probs=c(0.25), na.rm = TRUE),
    quantile2 = quantile(fvc_pct, probs=c(0.75), na.rm = TRUE)
  )
group_by(PM, PM5yrCensor_dich) %>%
  summarise(
    count = n(),
    mean = mean(dlco_pct, na.rm = TRUE),
    sd = sd(dlco_pct, na.rm = TRUE),
    median = median(dlco_pct, na.rm = TRUE),
    quantile = quantile(dlco_pct, probs=c(0.25), na.rm = TRUE),
    quantile2 = quantile(dlco_pct, probs=c(0.75), na.rm = TRUE)
  )
group_by(PM, PM5yrCensor_dich) %>%
  summarise(
    count = n(),
    mean = mean(time_DeathTxCensor, na.rm = TRUE),
    sd = sd(time_DeathTxCensor, na.rm = TRUE),
    median = median(time_DeathTxCensor, na.rm = TRUE),
    quantile = quantile(time_DeathTxCensor, probs=c(0.25), na.rm = TRUE),
    quantile2 = quantile(time_DeathTxCensor, probs=c(0.75), na.rm = TRUE)
  )
```

### PM_5yrPreDx Low vs High
```{r}
table(PM$sex, PM$PM5yr_dich)
table(PM$race, PM$PM5yr_dich)
table(PM$dich_Race, PM$PM5yr_dich)
table(PM$dx_group, PM$PM5yr_dich)
table(PM$smokeHx, PM$PM5yr_dich)
table(PM$metro, PM$PM5yr_dich)
table(PM$province, PM$PM5yr_dich)
table(PM$status, PM$PM5yr_dich)
group_by(PM, PM5yr_dich) %>%
  summarise(
    count = n(),
    mean = mean(age_dx, na.rm = TRUE),
    sd = sd(age_dx, na.rm = TRUE),
    median = median(age_dx, na.rm = TRUE),
    quantile = quantile(age_dx, probs=c(0.25), na.rm = TRUE),
    quantile2 = quantile(age_dx, probs=c(0.75), na.rm = TRUE)
  )
group_by(PM, PM5yr_dich) %>%
  summarise(
    count = n(),
    mean = mean(avg_s, na.rm = TRUE),
    sd = sd(avg_s, na.rm = TRUE),
    median = median(avg_s, na.rm = TRUE),
    quantile = quantile(avg_s, probs=c(0.25), na.rm = TRUE),
    quantile2 = quantile(avg_s, probs=c(0.75), na.rm = TRUE)
  )
group_by(PM, PM5yr_dich) %>%
  summarise(
    count = n(),
    mean = mean(fvc_pct, na.rm = TRUE),
    sd = sd(fvc_pct, na.rm = TRUE),
    median = median(fvc_pct, na.rm = TRUE),
    quantile = quantile(fvc_pct, probs=c(0.25), na.rm = TRUE),
    quantile2 = quantile(fvc_pct, probs=c(0.75), na.rm = TRUE)
  )
group_by(PM, PM5yr_dich) %>%
  summarise(
    count = n(),
    mean = mean(dlco_pct, na.rm = TRUE),
    sd = sd(dlco_pct, na.rm = TRUE),
    median = median(dlco_pct, na.rm = TRUE),
    quantile = quantile(dlco_pct, probs=c(0.25), na.rm = TRUE),
    quantile2 = quantile(dlco_pct, probs=c(0.75), na.rm = TRUE)
  )
group_by(PM, PM5yr_dich) %>%
  summarise(
    count = n(),
    mean = mean(time_DeathTxCensor, na.rm = TRUE),
    sd = sd(time_DeathTxCensor, na.rm = TRUE),
    median = median(time_DeathTxCensor, na.rm = TRUE),
    quantile = quantile(time_DeathTxCensor, probs=c(0.25), na.rm = TRUE),
    quantile2 = quantile(time_DeathTxCensor, probs=c(0.75), na.rm = TRUE)
  )
```


## Visual EDA
### Histograms
Histogram of PM Exposure 5yr Pre-Death/Transplant/Censoring
```{r}
(PM %>% ggplot(aes(x=PM_5yrPreCensor))+
   geom_histogram(fill="blue", color="black")+
   labs(x="Average PM Level 5yr Prior to Death/Transplant/Censoring", y="Number of patients with fILD", title="Average PM 5yr Prior to Death/Transplant/Censoring")+
   theme(plot.title = element_text(hjust = 0.5))+
   xlim(0,20))
```


Histogram of PM Exposure 5yr Pre-Dx
```{r}
(PM %>% ggplot(aes(x=PM_5yrPreDx))+
   geom_histogram(fill="blue", color="black")+
   labs(x="Average PM Level 5yr Prior to Diagnosis", y="Number of patients with fILD", title="Average PM 5yr Prior to Diagnosis")+
   theme(plot.title = element_text(hjust = 0.5))+
   xlim(0,20))
```


### Violin plots of PM vs other demographic variables
Violin plot wrapping boxplot to visualize low vs high PM5yrCensor_dich vs CIMD
```{r}
(PM %>% ggplot(aes(x=PM5yrCensor_dich, y=avg_s, fill=PM5yrCensor_dich))+
   geom_boxplot(width=0.2, color="black", alpha=1.0)+
   geom_violin(width=1.0, alpha=0.5)+
   labs(x="Low vs High PM in 5yrs Pre-Censoring", y="National CIMD", title="National CIMD by PM Exposure")+
   theme(plot.title = element_text(hjust = 0.5))+
   ylim(-2.5,2.5))
```

Violin plot wrapping boxplot to visualize sex vs PM_5yrPreCensor
```{r}
(PM %>% ggplot(aes(x=sex, y=PM_5yrPreCensor, fill=sex))+
   geom_boxplot(width=0.2, color="black", alpha=1.0)+
   geom_violin(width=1.0, alpha=0.5)+
   labs(x="Sex", y="Average PM 5 Years Pre-Censor", title="PM Exposure Pre-Censor by Sex")+
   theme(plot.title = element_text(hjust = 0.5))+
   ylim(0,20))
```


Violin plot wrapping boxplot to visualize sex vs PM_5yrPreDx
```{r}
(PM %>% ggplot(aes(x=sex, y=PM_5yrPreDx, fill=sex))+
   geom_boxplot(width=0.2, color="black", alpha=1.0)+
   geom_violin(width=1.0, alpha=0.5)+
   labs(x="Sex", y="Average PM 5 Years Pre-Diagnosis", title="PM Exposure Pre-Diagnosis by Sex")+
   theme(plot.title = element_text(hjust = 0.5))+
   ylim(0,20))
```

Violin plot wrapping boxplot to visualize dich_Race vs PM_5yrPreCensor
```{r}
(PM %>% ggplot(aes(x=dich_Race, y=PM_5yrPreCensor, fill=dich_Race))+
   geom_boxplot(width=0.2, color="black", alpha=1.0)+
   geom_violin(width=1.0, alpha=0.5)+
   labs(x="Race", y="Average PM 5 Years Pre-Censoring", title="PM Exposure Pre-Censoring by Race")+
   theme(plot.title = element_text(hjust = 0.5))+
   ylim(0,20))
```


Violin plot wrapping boxplot to visualize race vs PM_5yrPreCensor
```{r}
(PM %>% ggplot(aes(x=race, y=PM_5yrPreCensor, fill=race))+
   geom_boxplot(width=0.2, color="black", alpha=1.0)+
   geom_violin(width=1.0, alpha=0.5)+
   labs(x="Race", y="Average PM 5 Years Pre-Censoring", title="PM Exposure Pre-Censoring by Race")+
   theme(plot.title = element_text(hjust = 0.5))+
   ylim(0,20))
```

Violin plot wrapping boxplot to visualize dich_Race vs PM_5yrPreDx
```{r}
(PM %>% ggplot(aes(x=dich_Race, y=PM_5yrPreDx, fill=dich_Race))+
   geom_boxplot(width=0.2, color="black", alpha=1.0)+
   geom_violin(width=1.0, alpha=0.5)+
   labs(x="Race", y="Average PM 5 Years Pre-Diagnosis", title="PM Exposure Pre-Diagnosis by Race")+
   theme(plot.title = element_text(hjust = 0.5))+
   ylim(0,20))
```

Violin plot wrapping boxplot to visualize race vs PM_5yrPreDx
```{r}
(PM %>% ggplot(aes(x=race, y=PM_5yrPreDx, fill=race))+
   geom_boxplot(width=0.2, color="black", alpha=1.0)+
   geom_violin(width=1.0, alpha=0.5)+
   labs(x="Race", y="Average PM 5 Years Pre-Diagnosis", title="PM Exposure Pre-Diagnosis by Race")+
   theme(plot.title = element_text(hjust = 0.5))+
   ylim(0,20))
```

Violin plot wrapping boxplot to visualize metro vs PM_5yrPreCensor
```{r}
(PM %>% ggplot(aes(x=metro, y=PM_5yrPreCensor, fill=metro))+
   geom_boxplot(width=0.2, color="black", alpha=1.0)+
   geom_violin(width=1.0, alpha=0.5)+
   labs(x="Residential Location", y="Average PM in 5yrs Pre-Censoring", title="PM Exposure in 5yrs Pre-Censoring by Residential Location")+
   theme(plot.title = element_text(hjust = 0.5))+
   ylim(0,20))
```


Violin plot wrapping boxplot to visualize metro vs PM_5yrPreDx
```{r}
(PM %>% ggplot(aes(x=metro, y=PM_5yrPreDx, fill=metro))+
   geom_boxplot(width=0.2, color="black", alpha=1.0)+
   geom_violin(width=1.0, alpha=0.5)+
   labs(x="Residential Location", y="Average PM in 5yrs Pre-Diagnosis", title="PM Exposure in 5yrs Pre-Diagnosis by Residential Location")+
   theme(plot.title = element_text(hjust = 0.5))+
   ylim(0,20))
```

### Scatterplots of PM over time
Scatterplot showing average PM 5yr pre-death/transplant/censoring
```{r}
(PM %>% ggplot(aes(x=dx_date, y=PM_5yrPreCensor))+
   geom_point()+
   labs(x="Date of Diagnosis", y="PM Exposure 5yr Prior Death/Transplant/Censoring", title="PM 5yr Prior to Death/Transplant/Censoring vs Date of Diagnosis")+
   geom_smooth(method="lm", se = FALSE) +
   theme(plot.title = element_text(hjust = 0.5))+
   ylim(0,20))
```


```{r}
(PM %>% ggplot(aes(x=dx_date, y=PM_5yrPreCensor, color=site))+
   geom_point(alpha=0.5)+
   labs(x="Date of Diagnosis", y="PM Exposure 5yr Prior Death/Transplant/Censoring", title="PM 5yr Prior to Death/Transplant/Censoring vs Date of Diagnosis")+
   geom_smooth(method="lm", se = FALSE) +
   theme(plot.title = element_text(hjust = 0.5))+
   ylim(0,20))
```

Scatterplot showing average PM during 5yrs pre-diagnosis level by year of diagnosis
```{r}
(PM %>% ggplot(aes(x=dx_date, y=PM_5yrPreDx))+
   geom_point()+
   labs(x="Date of Diagnosis", y="PM Exposure During 5yrs Pre-Diagnosis", title="Pre-Diagnosis PM vs Date of Diagnosis")+
   geom_smooth(method="lm", se = FALSE) +
   theme(plot.title = element_text(hjust = 0.5))+
   ylim(0,20)+
   scale_x_date(limits = as.Date(c("2000-01-01","2020-01-01"))))
```


Scatterplot showing average PM during 5yrs pre-diagnosis level by year of diagnosis
```{r}
(PM %>% ggplot(aes(x=dx_date, y=PM_5yrPreDx, color=site))+
   geom_point(alpha=0.5)+
   labs(x="Date of Diagnosis", y="PM Exposure During 5yrs Pre-Diagnosis", title="Pre-Diagnosis PM vs Date of Diagnosis")+
   geom_smooth(method="lm", se = FALSE) +
   theme(plot.title = element_text(hjust = 0.5))+
   ylim(0,20)+
   scale_x_date(limits = as.Date(c("2000-01-01","2020-01-01"))))
```


# Cox PH Models for Association between PM and Death or Lung Transplant
## PM 5yrs Pre-Censoring Models
### Continuous PM 5yrs pre-censoring until Death/Transplant/Censoring
Start with the simplest model just looking at PM in the year of and 5yrs prior to death/transplant/censoring
```{r}
coxPH_model1 <- coxph(Surv(time_DeathTxCensor, deadORtx==1) ~ PM_5yrPreCensor, data=PM, id=ID)
summary(coxPH_model1)
```


Complete model
```{r}
coxPH_model2 <- coxph(Surv(time_DeathTxCensor, deadORtx==1) ~ PM_5yrPreCensor + sex + age_dx + smokeHx + dich_Race + avg_s + site, data=PM, id=ID)
summary(coxPH_model2)
```


Complete model + dx_yr
```{r}
coxPH_model3 <- coxph(Surv(time_DeathTxCensor, deadORtx==1) ~ PM_5yrPreCensor + sex + age_dx + smokeHx + dich_Race + avg_s + site + dx_yr, data=PM, id=ID)
summary(coxPH_model3)
```



### Smoothed HR Function for 5yr Pre-Censor Continuous
Base model
```{r}
coxPH_model1_tvc <- cox.tvc(coxPH_model1, "PM_5yrPreCensor")
print(coxPH_model1_tvc)
plot(coxPH_model1_tvc)
```


Complete model (all covariates except lung function)
```{r}
coxPH_model2_tvc <- cox.tvc(coxPH_model2, "PM_5yrPreCensor")
print(coxPH_model2_tvc)
plot(coxPH_model2_tvc)
```


```{r}
coxPH_model3_tvc <- cox.tvc(coxPH_model3, "PM_5yrPreCensor")
print(coxPH_model3_tvc)
plot(coxPH_model3_tvc)
```


### Variable HR for PM risk in 5yr Pre-Cenosr
Want a plot with x-axis of PM level and y-axis of HR

Unadjusted model
```{r}
#First need to make dataframe that only includes patients with time_DeathTxCensor
PMx <- PM %>% filter(!is.na(PM_5yrPreCensor) & !is.na(time_DeathTxCensor))

#Then make survival function
surv1 <- Surv(PMx$time_DeathTxCensor, PMx$deadORtx==1)
fit1 <- coxph(surv1 ~ pspline(PMx$PM_5yrPreCensor, df=3))
predicted <- predict(fit1, type="terms", se.fit=T, terms=1)

#Then plot
plot(PMx$PM_5yrPreCensor, exp(predicted$fit), type="n", xlim=c(2,16), ylim=c(0,6))
lines(sm.spline(PMx$PM_5yrPreCensor, exp(predicted$fit)), col = "red" , lty = 1 )
lines(sm.spline(PMx$PM_5yrPreCensor, exp(predicted$fit + 1.96 * predicted$se)), col = "orange" , lty = 2 )
lines(sm.spline(PMx$PM_5yrPreCensor, exp(predicted$fit - 1.96 * predicted$se)), col = "orange" , lty = 2 )
```

Complete model
```{r}
#First need to make dataframe that only includes patients with time_DeathTxCensor
PMx <- PM %>% filter(!is.na(PM_5yrPreCensor) & !is.na(time_DeathTxCensor) & !is.na(age_dx) & !is.na(sex) & !is.na(smokeHx) & !is.na(dich_Race) & !is.na(avg_s) & !is.na(site))

#Then survival function
surv1 <- Surv(PMx$time_DeathTxCensor, PMx$deadORtx==1)
fit1 <- coxph(surv1 ~ pspline(PMx$PM_5yrPreCensor, df=3) + PMx$age_dx + PMx$sex + PMx$smokeHx + PMx$dich_Race + PMx$avg_s + PMx$site)
predicted <- predict(fit1, type="terms", se.fit=T, terms=1)

#Then plot
plot(PMx$PM_5yrPreCensor, exp(predicted$fit), type="n", xlim=c(2,16), ylim=c(0,6))
lines(sm.spline(PMx$PM_5yrPreCensor, exp(predicted$fit)), col = "red" , lty = 1 )
lines(sm.spline(PMx$PM_5yrPreCensor, exp(predicted$fit + 1.96 * predicted$se)), col = "orange" , lty = 2 )
lines(sm.spline(PMx$PM_5yrPreCensor, exp(predicted$fit - 1.96 * predicted$se)), col = "orange" , lty = 2 )
```

Complete model + dx_yr
```{r}
#First need to make dataframe that only includes patients with time_DeathTxCensor
PMx <- PM %>% filter(!is.na(PM_5yrPreCensor) & !is.na(time_DeathTxCensor) & !is.na(age_dx) & !is.na(sex) & !is.na(smokeHx) & !is.na(dich_Race) & !is.na(avg_s) & !is.na(site) & !is.na(dx_yr))

#Then survival function
surv1 <- Surv(PMx$time_DeathTxCensor, PMx$deadORtx==1)
fit1 <- coxph(surv1 ~ pspline(PMx$PM_5yrPreCensor, df=4) + PMx$age_dx + PMx$sex + PMx$smokeHx + PMx$dich_Race + PMx$avg_s + PMx$site)
predicted <- predict(fit1, type="terms", se.fit=T, terms=1)

#Then plot
plot(PMx$PM_5yrPreCensor, exp(predicted$fit), type="n", xlim=c(2,16), ylim=c(0,6))
lines(sm.spline(PMx$PM_5yrPreCensor, exp(predicted$fit)), col = "red" , lty = 1 )
lines(sm.spline(PMx$PM_5yrPreCensor, exp(predicted$fit + 1.96 * predicted$se)), col = "orange" , lty = 2 )
lines(sm.spline(PMx$PM_5yrPreCensor, exp(predicted$fit - 1.96 * predicted$se)), col = "orange" , lty = 2 )
```


### PM Low vs High 5yrs Pre-Censoring
Base Model:
```{r}
coxPH_model7 <- coxph(Surv(time_DeathTxCensor, deadORtx==1) ~ PM5yrCensor_dich, data=PM, id=ID)
summary(coxPH_model7)
```

Complete model
```{r}
coxPH_model8 <- coxph(Surv(time_DeathTxCensor, deadORtx==1) ~ PM5yrCensor_dich + sex + age_dx + smokeHx + dich_Race + avg_s + site, data=PM, id=ID)
summary(coxPH_model8)
```

Complete model + dx_yr
```{r}
coxPH_model9 <- coxph(Surv(time_DeathTxCensor, deadORtx==1) ~ PM5yrCensor_dich + sex + age_dx + smokeHx + dich_Race + avg_s + site + dx_yr, data=PM, id=ID)
summary(coxPH_model9)
```


Now I will make a Kaplan-Meier survival curve, discretized by PM during disease deciles:
```{r, fig.height=5, fig.width=7}
surv1 <- survfit(Surv(time_DeathTxCensor, deadORtx==1) ~ PM5yrCensor_dich, data=PM)
(ggsurvplot(surv1, 
            palette=c("dodgerblue4","firebrick4"),
            censor=TRUE,
            pval=TRUE,
            xlim=c(0,10),
            legend="right",
           data=PM,
           legend.title="PM Exposure",
           legend.labs=c("Low", "High"),
           risk.table=T,
           title="Time to Death or Transplant by PM 5yrs Pre-Censoring",
           xlab="Time (years)",
           ylab="Survival Probability (Death or Lung Transplant)",
           ggtheme = theme_grey()
           ))
```


Now we want to check the proportionality assumption by plotting the scaled Schoenfield residuals against time for each covariate in our final model.
```{r}
residtest <- cox.zph(coxPH_model8)
residtest
ggcoxzph(residtest)
```


Now I will present a forest plot with the model
```{r}
ggforest(coxPH_model8)
```



# Association between PM and Baseline FVC
## Visual EDA 5yrs pre-Dx
Scatterplot of continuous PM 5yrs pre-diagnosis vs FVC, with sex represented by the color of points
```{r}
(PM %>% ggplot(aes(x=PM_5yrPreDx, y=fvc_pct, color=sex))+
   geom_point()+
   labs(x="PM 5yrs Pre-Diagnosis", y="Baseline FVC % Predicted", title="Baseline FVC % Predicted by PM 5yrs Pre-Diagnosis")+
   geom_smooth(method="lm", se = FALSE) +
   theme(plot.title = element_text(hjust = 0.5))+
   xlim(0,7))
```


Violin plot wrapping boxplot to visualize PM 5yrs pre-dx low vs high vs FVC
```{r}
PMb <- PM %>% filter(!is.na(PM5yr_dich))
(PMb %>% ggplot(aes(x=PM5yr_dich, y=fvc_pct, fill=PM5yr_dich))+
   geom_boxplot(width=0.2, color="black", alpha=1.0)+
   geom_violin(width=1.0, alpha=0.5)+
   labs(x="PM 5yrs Pre-Diagnosis Low vs High", y="Baseline FVC % Predicted", title="Baseline FVC % Predicted by PM 5yrs Pre-Diagnosis Low vs High")+
   theme_light()+
   theme(legend.position = "none", plot.title = element_text(hjust = 0.5))+
   scale_fill_brewer(type="seq", palette="YlOrRd"))
```



## PM 5yrs Pre-Diagnosis and FVC Models
Looking at continuous PM_5yrPreDx vs baseline FVC

Simplest model just looking at PM in 5yrs pre-diagnosis
```{r}
FVC_model1 <- lm(fvc_pct ~ PM_5yrPreDx, data=PM)
summary(FVC_model1)
confint(FVC_model1)
```

Complete model
```{r}
FVC_model2 <- lm(fvc_pct ~ PM_5yrPreDx + sex + age_dx + smokeHx + dich_Race + avg_s + site, data=PM)
summary(FVC_model2)
confint(FVC_model2)
```

Complete model + dx_yr
```{r}
FVC_model3 <- lm(fvc_pct ~ PM_5yrPreDx + sex + age_dx + smokeHx + dich_Race + avg_s + site + dx_yr, data=PM)
summary(FVC_model3)
confint(FVC_model3)
```


### PM 5yrs Pre-Diagnosis Low vs High and Baseline FVC
Looking at PM5yr_dich vs baseline FVC

Simplest model just looking at PM in 5yrs pre-diagnosis
```{r}
FVC_model1 <- lm(fvc_pct ~ PM5yr_dich, data=PM)
summary(FVC_model1)
confint(FVC_model1)
```

Complete model
```{r}
FVC_model2 <- lm(fvc_pct ~ PM5yr_dich + sex + age_dx + smokeHx + dich_Race + avg_s + site, data=PM)
summary(FVC_model2)
confint(FVC_model2)
```

Complete model + dx_yr
```{r}
FVC_model3 <- lm(fvc_pct ~ PM5yr_dich + sex + age_dx + smokeHx + dich_Race + avg_s + site + dx_yr, data=PM)
summary(FVC_model3)
confint(FVC_model3)
```

# Association between PM and Baseline DLCO
## Visual EDA 5yrs pre-Dx
Scatterplot of continuous PM 5yrs pre-diagnosis vs DLCO, with sex represented by the color of points
```{r}
(PM %>% ggplot(aes(x=PM_5yrPreDx, y=dlco_pct, color=sex))+
   geom_point()+
   labs(x="PM 5yrs Pre-Diagnosis", y="Baseline DLCO % Predicted", title="Baseline DLCO % Predicted by PM 5yrs Pre-Diagnosis")+
   geom_smooth(method="lm", se = FALSE) +
   theme(plot.title = element_text(hjust = 0.5))+
   xlim(0,12))
```


Violin plot wrapping boxplot to visualize PM 5yrs pre-dx low vs high vs dlco
```{r}
(PMb %>% ggplot(aes(x=PM5yr_dich, y=dlco_pct, fill=PM5yr_dich))+
   geom_boxplot(width=0.2, color="black", alpha=1.0)+
   geom_violin(width=1.0, alpha=0.5)+
   labs(x="PM 5yrs Pre-Diagnosis Low vs High", y="Baseline DLCO % Predicted", title="Baseline DLCO % Predicted by PM 5yrs Pre-Diagnosis Low vs High")+
   theme_light()+
   theme(legend.position = "none", plot.title = element_text(hjust = 0.5))+
   scale_fill_brewer(type="seq", palette="YlOrRd"))
```



## PM 5yrs Pre-Diagnosis and DLCO Models
### PM 5yrs Pre-Diagnosis and Baseline DLCO
Looking at continuous PM_5yrPreDx vs baseline DLCO

Simplest model just looking at PM in 5yrs pre-diagnosis
```{r}
DLCO_model1 <- lm(dlco_pct ~ PM_5yrPreDx, data=PM)
summary(DLCO_model1)
confint(DLCO_model1)
```

Complete model
```{r}
DLCO_model2 <- lm(dlco_pct ~ PM_5yrPreDx + sex + age_dx + smokeHx + dich_Race + avg_s + site, data=PM)
summary(DLCO_model2)
confint(DLCO_model2)
```

Complete model + dx_yr
```{r}
DLCO_model3 <- lm(dlco_pct ~ PM_5yrPreDx + sex + age_dx + smokeHx + dich_Race + avg_s + site + dx_yr, data=PM)
summary(DLCO_model3)
confint(DLCO_model3)
```


### PM 5yrs Pre-Diagnosis Low vs High and Baseline DLCO
Looking at PM5yr_dich vs baseline DLCO

Simplest model just looking at PM in 5yrs pre-diagnosis
```{r}
DLCO_model1 <- lm(dlco_pct ~ PM5yr_dich, data=PM)
summary(DLCO_model1)
confint(DLCO_model1)
```

Complete model
```{r}
DLCO_model2 <- lm(dlco_pct ~ PM5yr_dich + sex + age_dx + smokeHx + dich_Race + avg_s + site, data=PM)
summary(DLCO_model2)
confint(DLCO_model2)
```


Complete model + dx_yr
```{r}
DLCO_model3 <- lm(dlco_pct ~ PM5yr_dich + sex + age_dx + smokeHx + dich_Race + avg_s + site + dx_yr, data=PM)
summary(DLCO_model3)
confint(DLCO_model3)
```


# IPF-Only - Cox PH Models for Association between PM and Death or Lung Transplant
## IPF-Only - PM 5yrs Pre-Censoring Models
### IPF-Only - Continuous PM 5yrs pre-censoring until Death/Transplant/Censoring
Start with the simplest model just looking at PM in the year of and 5yrs prior to death/transplant/censoring
```{r}
coxPH_model1 <- coxph(Surv(time_DeathTxCensor, deadORtx==1) ~ PM_5yrPreCensor, data=PM_IPF, id=ID)
summary(coxPH_model1)
```


Complete model
```{r}
coxPH_model2 <- coxph(Surv(time_DeathTxCensor, deadORtx==1) ~ PM_5yrPreCensor + sex + age_dx + smokeHx + dich_Race + avg_s + site, data=PM_IPF, id=ID)
summary(coxPH_model2)
```


Complete model + dx_yr
```{r}
coxPH_model3 <- coxph(Surv(time_DeathTxCensor, deadORtx==1) ~ PM_5yrPreCensor + sex + age_dx + smokeHx + dich_Race + avg_s + site + dx_yr, data=PM_IPF, id=ID)
summary(coxPH_model3)
```



### IPF-Only - Smoothed HR Function for 5yr Pre-Censor Continuous
Base model
```{r}
coxPH_model1_tvc <- cox.tvc(coxPH_model1, "PM_5yrPreCensor")
print(coxPH_model1_tvc)
plot(coxPH_model1_tvc)
```


Complete model (all covariates except lung function)
```{r}
coxPH_model2_tvc <- cox.tvc(coxPH_model2, "PM_5yrPreCensor")
print(coxPH_model2_tvc)
plot(coxPH_model2_tvc)
```


```{r}
coxPH_model3_tvc <- cox.tvc(coxPH_model3, "PM_5yrPreCensor")
print(coxPH_model3_tvc)
plot(coxPH_model3_tvc)
```


### IPF-Only - Variable HR for PM risk in 5yr Pre-Cenosr
Want a plot with x-axis of PM level and y-axis of HR

Unadjusted model
```{r}
#First need to make dataframe that only includes patients with time_DeathTxCensor
PMx <- PM_IPF %>% filter(!is.na(PM_5yrPreCensor) & !is.na(time_DeathTxCensor))

#Then make survival function
surv1 <- Surv(PMx$time_DeathTxCensor, PMx$deadORtx==1)
fit1 <- coxph(surv1 ~ pspline(PMx$PM_5yrPreCensor, df=3))
predicted <- predict(fit1, type="terms", se.fit=T, terms=1)

#Then plot
plot(PMx$PM_5yrPreCensor, exp(predicted$fit), type="n")
lines(sm.spline(PMx$PM_5yrPreCensor, exp(predicted$fit)), col = "red" , lty = 1 )
lines(sm.spline(PMx$PM_5yrPreCensor, exp(predicted$fit + 1.96 * predicted$se)), col = "orange" , lty = 2 )
lines(sm.spline(PMx$PM_5yrPreCensor, exp(predicted$fit - 1.96 * predicted$se)), col = "orange" , lty = 2 )
```

Complete model
```{r}
#First need to make dataframe that only includes patients with time_DeathTxCensor
PMx <- PM_IPF %>% filter(!is.na(PM_5yrPreCensor) & !is.na(time_DeathTxCensor) & !is.na(age_dx) & !is.na(sex) & !is.na(smokeHx) & !is.na(dich_Race) & !is.na(avg_s) & !is.na(site))

#Then survival function
surv1 <- Surv(PMx$time_DeathTxCensor, PMx$deadORtx==1)
fit1 <- coxph(surv1 ~ pspline(PMx$PM_5yrPreCensor, df=3) + PMx$age_dx + PMx$sex + PMx$smokeHx + PMx$dich_Race + PMx$avg_s + PMx$site)
predicted <- predict(fit1, type="terms", se.fit=T, terms=1)

#Then plot
plot(PMx$PM_5yrPreCensor, exp(predicted$fit), type="n")
lines(sm.spline(PMx$PM_5yrPreCensor, exp(predicted$fit)), col = "red" , lty = 1 )
lines(sm.spline(PMx$PM_5yrPreCensor, exp(predicted$fit + 1.96 * predicted$se)), col = "orange" , lty = 2 )
lines(sm.spline(PMx$PM_5yrPreCensor, exp(predicted$fit - 1.96 * predicted$se)), col = "orange" , lty = 2 )
```

Complete Model + dx_yr
```{r}
#First need to make dataframe that only includes patients with time_DeathTxCensor
PMx <- PM_IPF %>% filter(!is.na(PM_5yrPreCensor) & !is.na(time_DeathTxCensor) & !is.na(age_dx) & !is.na(sex) & !is.na(smokeHx) & !is.na(dich_Race) & !is.na(avg_s) & !is.na(site) & !is.na(dx_yr))

#Then survival function
surv1 <- Surv(PMx$time_DeathTxCensor, PMx$deadORtx==1)
fit1 <- coxph(surv1 ~ pspline(PMx$PM_5yrPreCensor, df=4) + PMx$age_dx + PMx$sex + PMx$smokeHx + PMx$dich_Race + PMx$avg_s + PMx$site + PMx$dx_yr)
predicted <- predict(fit1, type="terms", se.fit=T, terms=1)

#Then plot
plot(PMx$PM_5yrPreCensor, exp(predicted$fit), type="n", xlim=c(0,20), ylim=c(0,4))
lines(sm.spline(PMx$PM_5yrPreCensor, exp(predicted$fit)), col = "red" , lty = 1 )
lines(sm.spline(PMx$PM_5yrPreCensor, exp(predicted$fit + 1.96 * predicted$se)), col = "orange" , lty = 2 )
lines(sm.spline(PMx$PM_5yrPreCensor, exp(predicted$fit - 1.96 * predicted$se)), col = "orange" , lty = 2 )
```


### IPF-Only - PM Low vs High 5yrs Pre-Censoring
Base Model:
```{r}
coxPH_model7 <- coxph(Surv(time_DeathTxCensor, deadORtx==1) ~ PM5yrCensor_dich, data=PM_IPF, id=ID)
summary(coxPH_model7)
```


Complete model
```{r}
coxPH_model8 <- coxph(Surv(time_DeathTxCensor, deadORtx==1) ~ PM5yrCensor_dich + sex + age_dx + smokeHx + dich_Race + avg_s + site, data=PM_IPF, id=ID)
summary(coxPH_model8)
```


Complete model + dx_yr
```{r}
coxPH_model9 <- coxph(Surv(time_DeathTxCensor, deadORtx==1) ~ PM5yrCensor_dich + sex + age_dx + smokeHx + dich_Race + avg_s + site + dx_yr, data=PM_IPF, id=ID)
summary(coxPH_model9)
```


Now I will make a Kaplan-Meier survival curve, discretized by PM during disease deciles:
```{r, fig.height=5, fig.width=7}
surv1 <- survfit(Surv(time_DeathTxCensor, deadORtx==1) ~ PM5yrCensor_dich, data=PM_IPF)
(ggsurvplot(surv1, 
            palette=c("dodgerblue4","firebrick4"),
            censor=TRUE,
            pval=TRUE,
            xlim=c(0,10),
            legend="right",
           data=PM_IPF,
           legend.title="PM Exposure",
           legend.labs=c("Low", "High"),
           risk.table=T,
           title="Time to Death or Transplant by PM 5yrs Pre-Censoring",
           xlab="Time (years)",
           ylab="Survival Probability (Death or Lung Transplant)",
           ggtheme = theme_grey()
           ))
```


Now we want to check the proportionality assumption by plotting the scaled Schoenfield residuals against time for each covariate in our final model.
```{r}
residtest <- cox.zph(coxPH_model9)
residtest
ggcoxzph(residtest)
```


Now I will present a forest plot with the model
```{r}
ggforest(coxPH_model9)
```



# IPF-Only - Association between PM and Baseline FVC
## IPF-Only - Visual EDA 5yrs pre-Dx
Scatterplot of continuous PM 5yrs pre-diagnosis vs FVC, with sex represented by the color of points
```{r}
(PM %>% ggplot(aes(x=PM_5yrPreDx, y=fvc_pct, color=sex))+
   geom_point()+
   labs(x="PM 5yrs Pre-Diagnosis", y="Baseline FVC % Predicted", title="Baseline FVC % Predicted by PM 5yrs Pre-Diagnosis")+
   geom_smooth(method="lm", se = FALSE) +
   theme(plot.title = element_text(hjust = 0.5))+
   xlim(0,7))
```


Violin plot wrapping boxplot to visualize PM 5yrs pre-dx low vs high vs FVC
```{r}
PMb <- PM %>% filter(!is.na(PM5yr_dich))
(PMb %>% ggplot(aes(x=PM5yr_dich, y=fvc_pct, fill=PM5yr_dich))+
   geom_boxplot(width=0.2, color="black", alpha=1.0)+
   geom_violin(width=1.0, alpha=0.5)+
   labs(x="PM 5yrs Pre-Diagnosis Low vs High", y="Baseline FVC % Predicted", title="Baseline FVC % Predicted by PM 5yrs Pre-Diagnosis Low vs High")+
   theme_light()+
   theme(legend.position = "none", plot.title = element_text(hjust = 0.5))+
   scale_fill_brewer(type="seq", palette="YlOrRd"))
```



## IPF-Only - PM 5yrs Pre-Diagnosis and FVC Models
Looking at continuous PM_5yrPreDx vs baseline FVC

Simplest model just looking at PM in 5yrs pre-diagnosis
```{r}
FVC_model1 <- lm(fvc_pct ~ PM_5yrPreDx, data=PM_IPF)
summary(FVC_model1)
confint(FVC_model1)
```

Complete model
```{r}
FVC_model2 <- lm(fvc_pct ~ PM_5yrPreDx + sex + age_dx + smokeHx + dich_Race + avg_s + site, data=PM_IPF)
summary(FVC_model2)
confint(FVC_model2)
```

Complete model + dx_yr
```{r}
FVC_model3 <- lm(fvc_pct ~ PM_5yrPreDx + sex + age_dx + smokeHx + dich_Race + avg_s + site + dx_yr, data=PM_IPF)
summary(FVC_model3)
confint(FVC_model3)
```


### IPF-Only - PM 5yrs Pre-Diagnosis Low vs High and Baseline FVC
Looking at PM5yr_dich vs baseline FVC

Simplest model just looking at PM in 5yrs pre-diagnosis
```{r}
FVC_model1 <- lm(fvc_pct ~ PM5yr_dich, data=PM_IPF)
summary(FVC_model1)
confint(FVC_model1)
```

Complete model
```{r}
FVC_model2 <- lm(fvc_pct ~ PM5yr_dich + sex + age_dx + smokeHx + dich_Race + avg_s + site, data=PM_IPF)
summary(FVC_model2)
confint(FVC_model2)
```

Complete model + dx_yr
```{r}
FVC_model3 <- lm(fvc_pct ~ PM5yr_dich + sex + age_dx + smokeHx + dich_Race + avg_s + site + dx_yr, data=PM_IPF)
summary(FVC_model3)
confint(FVC_model3)
```

# IPF-Only - Association between PM and Baseline DLCO
## IPF-Only - Visual EDA 5yrs pre-Dx
Scatterplot of continuous PM 5yrs pre-diagnosis vs DLCO, with sex represented by the color of points
```{r}
(PM %>% ggplot(aes(x=PM_5yrPreDx, y=dlco_pct, color=sex))+
   geom_point()+
   labs(x="PM 5yrs Pre-Diagnosis", y="Baseline DLCO % Predicted", title="Baseline DLCO % Predicted by PM 5yrs Pre-Diagnosis")+
   geom_smooth(method="lm", se = FALSE) +
   theme(plot.title = element_text(hjust = 0.5))+
   xlim(0,12))
```


Violin plot wrapping boxplot to visualize PM 5yrs pre-dx low vs high vs DLCO
```{r}
(PMb %>% ggplot(aes(x=PM5yr_dich, y=dlco_pct, fill=PM5yr_dich))+
   geom_boxplot(width=0.2, color="black", alpha=1.0)+
   geom_violin(width=1.0, alpha=0.5)+
   labs(x="PM 5yrs Pre-Diagnosis Low vs High", y="Baseline DLCO % Predicted", title="Baseline DLCO % Predicted by PM 5yrs Pre-Diagnosis Low vs High")+
   theme_light()+
   theme(legend.position = "none", plot.title = element_text(hjust = 0.5))+
   scale_fill_brewer(type="seq", palette="YlOrRd"))
```



## IPF-Only - PM 5yrs Pre-Diagnosis and DLCO Models
### IPF-Only - PM 5yrs Pre-Diagnosis and Baseline DLCO
Looking at continuous PM_5yrPreDx vs baseline DLCO

Simplest model just looking at PM in 5yrs pre-diagnosis
```{r}
DLCO_model1 <- lm(dlco_pct ~ PM_5yrPreDx, data=PM_IPF)
summary(DLCO_model1)
confint(DLCO_model1)
```

Complete model
```{r}
DLCO_model2 <- lm(dlco_pct ~ PM_5yrPreDx + sex + age_dx + smokeHx + dich_Race + avg_s + site, data=PM_IPF)
summary(DLCO_model2)
confint(DLCO_model2)
```

Complete model + dx_yr
```{r}
DLCO_model3 <- lm(dlco_pct ~ PM_5yrPreDx + sex + age_dx + smokeHx + dich_Race + avg_s + site + dx_yr, data=PM_IPF)
summary(DLCO_model3)
confint(DLCO_model3)
```


### IPF-Only - PM 5yrs Pre-Diagnosis Low vs High and Baseline DLCO
Looking at PM5yr_dich vs baseline DLCO

Simplest model just looking at PM in 5yrs pre-diagnosis
```{r}
DLCO_model1 <- lm(dlco_pct ~ PM5yr_dich, data=PM_IPF)
summary(DLCO_model1)
confint(DLCO_model1)
```

Complete model
```{r}
DLCO_model2 <- lm(dlco_pct ~ PM5yr_dich + sex + age_dx + smokeHx + dich_Race + avg_s + site, data=PM_IPF)
summary(DLCO_model2)
confint(DLCO_model2)
```


Complete model + dx_yr
```{r}
DLCO_model3 <- lm(dlco_pct ~ PM5yr_dich + sex + age_dx + smokeHx + dich_Race + avg_s + site + dx_yr, data=PM_IPF)
summary(DLCO_model3)
confint(DLCO_model3)
```



# Pre-2015 vs Post-2015 Models
```{r}
pre2015 <- PM %>% filter(dx_yr<=2015)

post2015 <- PM %>% filter(dx_yr>2015)
```

# Pre/Post-2015 Only - Cox PH Models for Association between PM and Death or Lung Transplant
## Pre/Post-2015 Only - PM 5yrs Pre-Censoring Models
### Pre/Post-2015 Only - Continuous PM 5yrs pre-censoring until Death/Transplant/Censoring
Start with the simplest model just looking at PM in the year of and 5yrs prior to death/transplant/censoring
```{r}
coxPH_model1a <- coxph(Surv(time_DeathTxCensor, deadORtx==1) ~ PM_5yrPreCensor, data=pre2015, id=ID)
summary(coxPH_model1a)
```

```{r}
coxPH_model1b <- coxph(Surv(time_DeathTxCensor, deadORtx==1) ~ PM_5yrPreCensor, data=post2015, id=ID)
summary(coxPH_model1b)
```

Complete model
```{r}
coxPH_model2a <- coxph(Surv(time_DeathTxCensor, deadORtx==1) ~ PM_5yrPreCensor + sex + age_dx + smokeHx + dich_Race + avg_s + site, data=pre2015, id=ID)
summary(coxPH_model2a)
```

Complete model
```{r}
coxPH_model2b <- coxph(Surv(time_DeathTxCensor, deadORtx==1) ~ PM_5yrPreCensor + sex + age_dx + smokeHx + dich_Race + avg_s + site, data=post2015, id=ID)
summary(coxPH_model2b)
```

### Pre/Post-2015 Only - Variable HR for PM risk in 5yr Pre-Cenosr
Want a plot with x-axis of PM level and y-axis of HR

Unadjusted model - Pre-2015
```{r}
#First need to make dataframe that only includes patients with time_DeathTxCensor
pre2015x <- pre2015 %>% filter(!is.na(PM_5yrPreCensor) & !is.na(time_DeathTxCensor))

#Then make survival function
surv1 <- Surv(pre2015x$time_DeathTxCensor, pre2015x$deadORtx==1)
fit1 <- coxph(surv1 ~ pspline(pre2015x$PM_5yrPreCensor, df=3))
predicted <- predict(fit1, type="terms", se.fit=T, terms=1)

#Then plot
plot(pre2015x$PM_5yrPreCensor, exp(predicted$fit), type="n")
lines(sm.spline(pre2015x$PM_5yrPreCensor, exp(predicted$fit)), col = "red" , lty = 1 )
lines(sm.spline(pre2015x$PM_5yrPreCensor, exp(predicted$fit + 1.96 * predicted$se)), col = "orange" , lty = 2 )
lines(sm.spline(pre2015x$PM_5yrPreCensor, exp(predicted$fit - 1.96 * predicted$se)), col = "orange" , lty = 2 )
```

Unadjusted model - Post-2015
```{r}
#First need to make dataframe that only includes patients with time_DeathTxCensor
post2015x <- post2015 %>% filter(!is.na(PM_5yrPreCensor) & !is.na(time_DeathTxCensor))

#Then make survival function
surv1 <- Surv(post2015x$time_DeathTxCensor, post2015x$deadORtx==1)
fit1 <- coxph(surv1 ~ pspline(post2015x$PM_5yrPreCensor, df=3))
predicted <- predict(fit1, type="terms", se.fit=T, terms=1)

#Then plot
plot(post2015x$PM_5yrPreCensor, exp(predicted$fit), type="n")
lines(sm.spline(post2015x$PM_5yrPreCensor, exp(predicted$fit)), col = "red" , lty = 1 )
lines(sm.spline(post2015x$PM_5yrPreCensor, exp(predicted$fit + 1.96 * predicted$se)), col = "orange" , lty = 2 )
lines(sm.spline(post2015x$PM_5yrPreCensor, exp(predicted$fit - 1.96 * predicted$se)), col = "orange" , lty = 2 )
```

Complete model Pre-2015
```{r}
#First need to make dataframe that only includes patients with time_DeathTxCensor
pre2015x <- pre2015 %>% filter(!is.na(PM_5yrPreCensor) & !is.na(time_DeathTxCensor) & !is.na(age_dx) & !is.na(sex) & !is.na(smokeHx) & !is.na(dich_Race) & !is.na(avg_s) & !is.na(site))

#Then survival function
surv1 <- Surv(pre2015x$time_DeathTxCensor, pre2015x$deadORtx==1)
fit1 <- coxph(surv1 ~ pspline(pre2015x$PM_5yrPreCensor, df=3) + pre2015x$age_dx + pre2015x$sex + pre2015x$smokeHx + pre2015x$dich_Race + pre2015x$avg_s + pre2015x$site)
predicted <- predict(fit1, type="terms", se.fit=T, terms=1)

#Then plot
plot(pre2015x$PM_5yrPreCensor, exp(predicted$fit), type="n")
lines(sm.spline(pre2015x$PM_5yrPreCensor, exp(predicted$fit)), col = "red" , lty = 1 )
lines(sm.spline(pre2015x$PM_5yrPreCensor, exp(predicted$fit + 1.96 * predicted$se)), col = "orange" , lty = 2 )
lines(sm.spline(pre2015x$PM_5yrPreCensor, exp(predicted$fit - 1.96 * predicted$se)), col = "orange" , lty = 2 )
```

Complete model Post-2015
```{r}
#First need to make dataframe that only includes patients with time_DeathTxCensor
post2015x <- post2015 %>% filter(!is.na(PM_5yrPreCensor) & !is.na(time_DeathTxCensor) & !is.na(age_dx) & !is.na(sex) & !is.na(smokeHx) & !is.na(dich_Race) & !is.na(avg_s) & !is.na(site))

#Then survival function
surv1 <- Surv(post2015x$time_DeathTxCensor, post2015x$deadORtx==1)
fit1 <- coxph(surv1 ~ pspline(post2015x$PM_5yrPreCensor, df=3) + post2015x$age_dx + post2015x$sex + post2015x$smokeHx + post2015x$dich_Race + post2015x$avg_s + post2015x$site)
predicted <- predict(fit1, type="terms", se.fit=T, terms=1)

#Then plot
plot(post2015x$PM_5yrPreCensor, exp(predicted$fit), type="n")
lines(sm.spline(post2015x$PM_5yrPreCensor, exp(predicted$fit)), col = "red" , lty = 1 )
lines(sm.spline(post2015x$PM_5yrPreCensor, exp(predicted$fit + 1.96 * predicted$se)), col = "orange" , lty = 2 )
lines(sm.spline(post2015x$PM_5yrPreCensor, exp(predicted$fit - 1.96 * predicted$se)), col = "orange" , lty = 2 )
```

### Pre/Post-2015 Only - PM Low vs High 5yrs Pre-Censoring
Base Model:
```{r}
coxPH_model7a <- coxph(Surv(time_DeathTxCensor, deadORtx==1) ~ PM5yrCensor_dich, data=pre2015, id=ID)
summary(coxPH_model7a)
```

Base Model:
```{r}
coxPH_model7b <- coxph(Surv(time_DeathTxCensor, deadORtx==1) ~ PM5yrCensor_dich, data=post2015, id=ID)
summary(coxPH_model7b)
```

Complete model
```{r}
coxPH_model8a <- coxph(Surv(time_DeathTxCensor, deadORtx==1) ~ PM5yrCensor_dich + sex + age_dx + smokeHx + dich_Race + avg_s + site, data=pre2015, id=ID)
summary(coxPH_model8a)
```

Complete model
```{r}
coxPH_model8b <- coxph(Surv(time_DeathTxCensor, deadORtx==1) ~ PM5yrCensor_dich + sex + age_dx + smokeHx + dich_Race + avg_s + site, data=post2015, id=ID)
summary(coxPH_model8b)
```


Now I will make a Kaplan-Meier survival curve, discretized by PM during disease deciles:
```{r, fig.height=5, fig.width=7}
surv1 <- survfit(Surv(time_DeathTxCensor, deadORtx==1) ~ PM5yrCensor_dich, data=pre2015)
(ggsurvplot(surv1, 
            palette=c("dodgerblue4","firebrick4"),
            censor=TRUE,
            pval=TRUE,
            xlim=c(0,10),
            legend="right",
           data=PM_IPF,
           legend.title="PM Exposure",
           legend.labs=c("Low", "High"),
           risk.table=T,
           title="Time to Death or Transplant by PM 5yrs Pre-Censoring",
           xlab="Time (years)",
           ylab="Survival Probability (Death or Lung Transplant)",
           ggtheme = theme_grey()
           ))
```

Now I will make a Kaplan-Meier survival curve, discretized by PM during disease deciles:
```{r, fig.height=5, fig.width=7}
surv1 <- survfit(Surv(time_DeathTxCensor, deadORtx==1) ~ PM5yrCensor_dich, data=post2015)
(ggsurvplot(surv1, 
            palette=c("dodgerblue4","firebrick4"),
            censor=TRUE,
            pval=TRUE,
            xlim=c(0,10),
            legend="right",
           data=PM_IPF,
           legend.title="PM Exposure",
           legend.labs=c("Low", "High"),
           risk.table=T,
           title="Time to Death or Transplant by PM 5yrs Pre-Censoring",
           xlab="Time (years)",
           ylab="Survival Probability (Death or Lung Transplant)",
           ggtheme = theme_grey()
           ))
```

Now we want to check the proportionality assumption by plotting the scaled Schoenfield residuals against time for each covariate in our final model.
```{r}
residtest <- cox.zph(coxPH_model8a)
residtest
ggcoxzph(residtest)
```

```{r}
residtest <- cox.zph(coxPH_model8b)
residtest
ggcoxzph(residtest)
```


Now I will present a forest plot with the model
```{r}
ggforest(coxPH_model8a)
```

```{r}
ggforest(coxPH_model8b)
```


# Pre/Post-2015 Only - Association between PM and Baseline FVC
## Pre/Post-2015 Only - Visual EDA 5yrs pre-Dx
Scatterplot of continuous PM 5yrs pre-diagnosis vs FVC, with sex represented by the color of points - Pre-2015
```{r}
(pre2015 %>% ggplot(aes(x=PM_5yrPreDx, y=fvc_pct, color=sex))+
   geom_point()+
   labs(x="PM 5yrs Pre-Diagnosis", y="Baseline FVC % Predicted", title="Baseline FVC % Predicted by PM 5yrs Pre-Diagnosis")+
   geom_smooth(method="lm", se = FALSE) +
   theme(plot.title = element_text(hjust = 0.5))+
   xlim(5,20))
```

Scatterplot of continuous PM 5yrs pre-diagnosis vs FVC, with sex represented by the color of points - Post-2015
```{r}
(post2015 %>% ggplot(aes(x=PM_5yrPreDx, y=fvc_pct, color=sex))+
   geom_point()+
   labs(x="PM 5yrs Pre-Diagnosis", y="Baseline FVC % Predicted", title="Baseline FVC % Predicted by PM 5yrs Pre-Diagnosis")+
   geom_smooth(method="lm", se = FALSE) +
   theme(plot.title = element_text(hjust = 0.5))+
   xlim(5,20))
```

Violin plot wrapping boxplot to visualize PM 5yrs pre-dx low vs high vs FVC - Pre-2015
```{r}
pre2015b <- pre2015 %>% filter(!is.na(PM5yr_dich))
(pre2015b %>% ggplot(aes(x=PM5yr_dich, y=fvc_pct, fill=PM5yr_dich))+
   geom_boxplot(width=0.2, color="black", alpha=1.0)+
   geom_violin(width=1.0, alpha=0.5)+
   labs(x="PM 5yrs Pre-Diagnosis Low vs High", y="Baseline FVC % Predicted", title="Baseline FVC % Predicted by PM 5yrs Pre-Diagnosis Low vs High")+
   theme_light()+
   theme(legend.position = "none", plot.title = element_text(hjust = 0.5))+
   scale_fill_brewer(type="seq", palette="YlOrRd"))
```

Violin plot wrapping boxplot to visualize PM 5yrs pre-dx low vs high vs FVC - Post-2015
```{r}
post2015b <- post2015 %>% filter(!is.na(PM5yr_dich))
(post2015b %>% ggplot(aes(x=PM5yr_dich, y=fvc_pct, fill=PM5yr_dich))+
   geom_boxplot(width=0.2, color="black", alpha=1.0)+
   geom_violin(width=1.0, alpha=0.5)+
   labs(x="PM 5yrs Pre-Diagnosis Low vs High", y="Baseline FVC % Predicted", title="Baseline FVC % Predicted by PM 5yrs Pre-Diagnosis Low vs High")+
   theme_light()+
   theme(legend.position = "none", plot.title = element_text(hjust = 0.5))+
   scale_fill_brewer(type="seq", palette="YlOrRd"))
```



## Pre/Post-2015 Only - PM 5yrs Pre-Diagnosis and FVC Models
Looking at continuous PM_5yrPreDx vs baseline FVC

Simplest model just looking at PM in 5yrs pre-diagnosis
```{r}
FVC_model1a <- lm(fvc_pct ~ PM_5yrPreDx, data=pre2015)
summary(FVC_model1a)
confint(FVC_model1a)
```

Simplest model just looking at PM in 5yrs pre-diagnosis
```{r}
FVC_model1b <- lm(fvc_pct ~ PM_5yrPreDx, data=post2015)
summary(FVC_model1b)
confint(FVC_model1b)
```

Complete model
```{r}
FVC_model2a <- lm(fvc_pct ~ PM_5yrPreDx + sex + age_dx + smokeHx + dich_Race + avg_s + site, data=pre2015)
summary(FVC_model2a)
confint(FVC_model2a)
```

Complete model
```{r}
FVC_model2b <- lm(fvc_pct ~ PM_5yrPreDx + sex + age_dx + smokeHx + dich_Race + avg_s + site, data=post2015)
summary(FVC_model2b)
confint(FVC_model2b)
```


### Pre/Post-2015 Only - PM 5yrs Pre-Diagnosis Low vs High and Baseline FVC
Looking at PM5yr_dich vs baseline FVC

Simplest model just looking at PM in 5yrs pre-diagnosis
```{r}
FVC_model1a <- lm(fvc_pct ~ PM5yr_dich, data=pre2015)
summary(FVC_model1a)
confint(FVC_model1a)
```

Simplest model just looking at PM in 5yrs pre-diagnosis
```{r}
FVC_model1b <- lm(fvc_pct ~ PM5yr_dich, data=post2015)
summary(FVC_model1b)
confint(FVC_model1b)
```

Complete model
```{r}
FVC_model2a <- lm(fvc_pct ~ PM5yr_dich + sex + age_dx + smokeHx + dich_Race + avg_s + site, data=pre2015)
summary(FVC_model2a)
confint(FVC_model2a)
```

Complete model
```{r}
FVC_model2b <- lm(fvc_pct ~ PM5yr_dich + sex + age_dx + smokeHx + dich_Race + avg_s + site, data=post2015)
summary(FVC_model2b)
confint(FVC_model2b)
```

# Pre/Post-2015 Only - Association between PM and Baseline DLCO
## Pre/Post-2015 Only - Visual EDA 5yrs pre-Dx
Scatterplot of continuous PM 5yrs pre-diagnosis vs DLCO, with sex represented by the color of points - Pre-2015
```{r}
(pre2015 %>% ggplot(aes(x=PM_5yrPreDx, y=dlco_pct, color=sex))+
   geom_point()+
   labs(x="PM 5yrs Pre-Diagnosis", y="Baseline DLCO % Predicted", title="Baseline DLCO % Predicted by PM 5yrs Pre-Diagnosis")+
   geom_smooth(method="lm", se = FALSE) +
   theme(plot.title = element_text(hjust = 0.5))+
   xlim(5,20))
```

Scatterplot of continuous PM 5yrs pre-diagnosis vs DLCO, with sex represented by the color of points - Post-2015
```{r}
(post2015 %>% ggplot(aes(x=PM_5yrPreDx, y=dlco_pct, color=sex))+
   geom_point()+
   labs(x="PM 5yrs Pre-Diagnosis", y="Baseline DLCO % Predicted", title="Baseline DLCO % Predicted by PM 5yrs Pre-Diagnosis")+
   geom_smooth(method="lm", se = FALSE) +
   theme(plot.title = element_text(hjust = 0.5))+
   xlim(5,20))
```

Violin plot wrapping boxplot to visualize PM 5yrs pre-dx low vs high vs DLCO - Pre-2015
```{r}
(pre2015b %>% ggplot(aes(x=PM5yr_dich, y=dlco_pct, fill=PM5yr_dich))+
   geom_boxplot(width=0.2, color="black", alpha=1.0)+
   geom_violin(width=1.0, alpha=0.5)+
   labs(x="PM 5yrs Pre-Diagnosis Low vs High", y="Baseline DLCO % Predicted", title="Baseline DLCO % Predicted by PM 5yrs Pre-Diagnosis Low vs High")+
   theme_light()+
   theme(legend.position = "none", plot.title = element_text(hjust = 0.5))+
   scale_fill_brewer(type="seq", palette="YlOrRd"))
```

Violin plot wrapping boxplot to visualize PM 5yrs pre-dx low vs high vs DLCO - Post-2015
```{r}
(post2015b %>% ggplot(aes(x=PM5yr_dich, y=dlco_pct, fill=PM5yr_dich))+
   geom_boxplot(width=0.2, color="black", alpha=1.0)+
   geom_violin(width=1.0, alpha=0.5)+
   labs(x="PM 5yrs Pre-Diagnosis Low vs High", y="Baseline DLCO % Predicted", title="Baseline DLCO % Predicted by PM 5yrs Pre-Diagnosis Low vs High")+
   theme_light()+
   theme(legend.position = "none", plot.title = element_text(hjust = 0.5))+
   scale_fill_brewer(type="seq", palette="YlOrRd"))
```

## Pre/Post-2015 Only - PM 5yrs Pre-Diagnosis and DLCO Models
### Pre/Post-2015 Only - PM 5yrs Pre-Diagnosis and Baseline DLCO
Looking at continuous PM_5yrPreDx vs baseline DLCO

Simplest model just looking at PM in 5yrs pre-diagnosis
```{r}
DLCO_model1a <- lm(dlco_pct ~ PM_5yrPreDx, data=pre2015)
summary(DLCO_model1a)
confint(DLCO_model1a)
```

Simplest model just looking at PM in 5yrs pre-diagnosis
```{r}
DLCO_model1b <- lm(dlco_pct ~ PM_5yrPreDx, data=post2015)
summary(DLCO_model1b)
confint(DLCO_model1b)
```

Complete model
```{r}
DLCO_model2a <- lm(dlco_pct ~ PM_5yrPreDx + sex + age_dx + smokeHx + dich_Race + avg_s + site, data=pre2015)
summary(DLCO_model2a)
confint(DLCO_model2a)
```

Complete model
```{r}
DLCO_model2b <- lm(dlco_pct ~ PM_5yrPreDx + sex + age_dx + smokeHx + dich_Race + avg_s + site, data=post2015)
summary(DLCO_model2b)
confint(DLCO_model2b)
```

### Pre/Post-2015 Only - PM 5yrs Pre-Diagnosis Low vs High and Baseline DLCO
Looking at PM5yr_dich vs baseline DLCO

Simplest model just looking at PM in 5yrs pre-diagnosis
```{r}
DLCO_model1a <- lm(dlco_pct ~ PM5yr_dich, data=pre2015)
summary(DLCO_model1a)
confint(DLCO_model1a)
```

Simplest model just looking at PM in 5yrs pre-diagnosis
```{r}
DLCO_model1b <- lm(dlco_pct ~ PM5yr_dich, data=post2015)
summary(DLCO_model1b)
confint(DLCO_model1b)
```

Complete model
```{r}
DLCO_model2a <- lm(dlco_pct ~ PM5yr_dich + sex + age_dx + smokeHx + dich_Race + avg_s + site, data=pre2015)
summary(DLCO_model2a)
confint(DLCO_model2a)
```

Complete model
```{r}
DLCO_model2b <- lm(dlco_pct ~ PM5yr_dich + sex + age_dx + smokeHx + dich_Race + avg_s + site, data=post2015)
summary(DLCO_model2b)
confint(DLCO_model2b)
```



# Warm vs Cold Months - Cox PH Models for Association between PM and Death or Lung Transplant
## Warm vs Cold Months - PM 5yrs Pre-Censoring Models
### Warm vs Cold Months - Continuous PM 5yrs pre-censoring until Death/Transplant/Censoring
Start with the simplest model just looking at PM in the year of and 5yrs prior to death/transplant/censoring
```{r}
coxPH_model1a <- coxph(Surv(time_DeathTxCensor, deadORtx==1) ~ PM_5yrPreCensor, data=PM_warm, id=ID)
summary(coxPH_model1a)
```

```{r}
coxPH_model1b <- coxph(Surv(time_DeathTxCensor, deadORtx==1) ~ PM_5yrPreCensor, data=PM_cold, id=ID)
summary(coxPH_model1b)
```

Complete model
```{r}
coxPH_model2a <- coxph(Surv(time_DeathTxCensor, deadORtx==1) ~ PM_5yrPreCensor + sex + age_dx + smokeHx + dich_Race + avg_s + site, data=PM_warm, id=ID)
summary(coxPH_model2a)
```

Complete model
```{r}
coxPH_model2b <- coxph(Surv(time_DeathTxCensor, deadORtx==1) ~ PM_5yrPreCensor + sex + age_dx + smokeHx + dich_Race + avg_s + site, data=PM_cold, id=ID)
summary(coxPH_model2b)
```

### Warm vs Cold Months - Variable HR for PM risk in 5yr Pre-Cenosr
Want a plot with x-axis of PM level and y-axis of HR

Unadjusted model - Pre-2015
```{r}
#First need to make dataframe that only includes patients with time_DeathTxCensor
PM_warmx <- PM_warm %>% filter(!is.na(PM_5yrPreCensor) & !is.na(time_DeathTxCensor))

#Then make survival function
surv1 <- Surv(PM_warmx$time_DeathTxCensor, PM_warmx$deadORtx==1)
fit1 <- coxph(surv1 ~ pspline(PM_warmx$PM_5yrPreCensor, df=3))
predicted <- predict(fit1, type="terms", se.fit=T, terms=1)

#Then plot
plot(PM_warmx$PM_5yrPreCensor, exp(predicted$fit), type="n")
lines(sm.spline(PM_warmx$PM_5yrPreCensor, exp(predicted$fit)), col = "red" , lty = 1 )
lines(sm.spline(PM_warmx$PM_5yrPreCensor, exp(predicted$fit + 1.96 * predicted$se)), col = "orange" , lty = 2 )
lines(sm.spline(PM_warmx$PM_5yrPreCensor, exp(predicted$fit - 1.96 * predicted$se)), col = "orange" , lty = 2 )
```

Unadjusted model - Post-2015
```{r}
#First need to make dataframe that only includes patients with time_DeathTxCensor
PM_coldx <- PM_cold %>% filter(!is.na(PM_5yrPreCensor) & !is.na(time_DeathTxCensor))

#Then make survival function
surv1 <- Surv(PM_coldx$time_DeathTxCensor, PM_coldx$deadORtx==1)
fit1 <- coxph(surv1 ~ pspline(PM_coldx$PM_5yrPreCensor, df=3))
predicted <- predict(fit1, type="terms", se.fit=T, terms=1)

#Then plot
plot(PM_coldx$PM_5yrPreCensor, exp(predicted$fit), type="n")
lines(sm.spline(PM_coldx$PM_5yrPreCensor, exp(predicted$fit)), col = "red" , lty = 1 )
lines(sm.spline(PM_coldx$PM_5yrPreCensor, exp(predicted$fit + 1.96 * predicted$se)), col = "orange" , lty = 2 )
lines(sm.spline(PM_coldx$PM_5yrPreCensor, exp(predicted$fit - 1.96 * predicted$se)), col = "orange" , lty = 2 )
```

Complete model Pre-2015
```{r}
#First need to make dataframe that only includes patients with time_DeathTxCensor
PM_warmx <- PM_warm %>% filter(!is.na(PM_5yrPreCensor) & !is.na(time_DeathTxCensor) & !is.na(age_dx) & !is.na(sex) & !is.na(smokeHx) & !is.na(dich_Race) & !is.na(avg_s) & !is.na(site))

#Then survival function
surv1 <- Surv(PM_warmx$time_DeathTxCensor, PM_warmx$deadORtx==1)
fit1 <- coxph(surv1 ~ pspline(PM_warmx$PM_5yrPreCensor, df=3) + PM_warmx$age_dx + PM_warmx$sex + PM_warmx$smokeHx + PM_warmx$dich_Race + PM_warmx$avg_s + PM_warmx$site)
predicted <- predict(fit1, type="terms", se.fit=T, terms=1)

#Then plot
plot(PM_warmx$PM_5yrPreCensor, exp(predicted$fit), type="n")
lines(sm.spline(PM_warmx$PM_5yrPreCensor, exp(predicted$fit)), col = "red" , lty = 1 )
lines(sm.spline(PM_warmx$PM_5yrPreCensor, exp(predicted$fit + 1.96 * predicted$se)), col = "orange" , lty = 2 )
lines(sm.spline(PM_warmx$PM_5yrPreCensor, exp(predicted$fit - 1.96 * predicted$se)), col = "orange" , lty = 2 )
```

Complete model Post-2015
```{r}
#First need to make dataframe that only includes patients with time_DeathTxCensor
PM_coldx <- PM_cold %>% filter(!is.na(PM_5yrPreCensor) & !is.na(time_DeathTxCensor) & !is.na(age_dx) & !is.na(sex) & !is.na(smokeHx) & !is.na(dich_Race) & !is.na(avg_s) & !is.na(site))

#Then survival function
surv1 <- Surv(PM_coldx$time_DeathTxCensor, PM_coldx$deadORtx==1)
fit1 <- coxph(surv1 ~ pspline(PM_coldx$PM_5yrPreCensor, df=3) + PM_coldx$age_dx + PM_coldx$sex + PM_coldx$smokeHx + PM_coldx$dich_Race + PM_coldx$avg_s + PM_coldx$site)
predicted <- predict(fit1, type="terms", se.fit=T, terms=1)

#Then plot
plot(PM_coldx$PM_5yrPreCensor, exp(predicted$fit), type="n")
lines(sm.spline(PM_coldx$PM_5yrPreCensor, exp(predicted$fit)), col = "red" , lty = 1 )
lines(sm.spline(PM_coldx$PM_5yrPreCensor, exp(predicted$fit + 1.96 * predicted$se)), col = "orange" , lty = 2 )
lines(sm.spline(PM_coldx$PM_5yrPreCensor, exp(predicted$fit - 1.96 * predicted$se)), col = "orange" , lty = 2 )
```

### Warm vs Cold Months - PM Low vs High 5yrs Pre-Censoring
Base Model:
```{r}
coxPH_model7a <- coxph(Surv(time_DeathTxCensor, deadORtx==1) ~ PM5yrCensor_dich, data=PM_warm, id=ID)
summary(coxPH_model7a)
```

Base Model:
```{r}
coxPH_model7b <- coxph(Surv(time_DeathTxCensor, deadORtx==1) ~ PM5yrCensor_dich, data=PM_cold, id=ID)
summary(coxPH_model7b)
```

Complete model
```{r}
coxPH_model8a <- coxph(Surv(time_DeathTxCensor, deadORtx==1) ~ PM5yrCensor_dich + sex + age_dx + smokeHx + dich_Race + avg_s + site, data=PM_warm, id=ID)
summary(coxPH_model8a)
```

Complete model
```{r}
coxPH_model8b <- coxph(Surv(time_DeathTxCensor, deadORtx==1) ~ PM5yrCensor_dich + sex + age_dx + smokeHx + dich_Race + avg_s + site, data=PM_cold, id=ID)
summary(coxPH_model8b)
```


Now I will make a Kaplan-Meier survival curve, discretized by PM during disease deciles:
```{r, fig.height=5, fig.width=7}
surv1 <- survfit(Surv(time_DeathTxCensor, deadORtx==1) ~ PM5yrCensor_dich, data=PM_warm)
(ggsurvplot(surv1, 
            palette=c("dodgerblue4","firebrick4"),
            censor=TRUE,
            pval=TRUE,
            xlim=c(0,10),
            legend="right",
           data=PM_IPF,
           legend.title="PM Exposure",
           legend.labs=c("Low", "High"),
           risk.table=T,
           title="Time to Death or Transplant by PM 5yrs Pre-Censoring",
           xlab="Time (years)",
           ylab="Survival Probability (Death or Lung Transplant)",
           ggtheme = theme_grey()
           ))
```

Now I will make a Kaplan-Meier survival curve, discretized by PM during disease deciles:
```{r, fig.height=5, fig.width=7}
surv1 <- survfit(Surv(time_DeathTxCensor, deadORtx==1) ~ PM5yrCensor_dich, data=PM_cold)
(ggsurvplot(surv1, 
            palette=c("dodgerblue4","firebrick4"),
            censor=TRUE,
            pval=TRUE,
            xlim=c(0,10),
            legend="right",
           data=PM_IPF,
           legend.title="PM Exposure",
           legend.labs=c("Low", "High"),
           risk.table=T,
           title="Time to Death or Transplant by PM 5yrs Pre-Censoring",
           xlab="Time (years)",
           ylab="Survival Probability (Death or Lung Transplant)",
           ggtheme = theme_grey()
           ))
```

Now we want to check the proportionality assumption by plotting the scaled Schoenfield residuals against time for each covariate in our final model.
```{r}
residtest <- cox.zph(coxPH_model8a)
residtest
ggcoxzph(residtest)
```

```{r}
residtest <- cox.zph(coxPH_model8b)
residtest
ggcoxzph(residtest)
```


Now I will present a forest plot with the model
```{r}
ggforest(coxPH_model8a)
```

```{r}
ggforest(coxPH_model8b)
```


# Warm vs Cold Months - Association between PM and Baseline FVC
## Warm vs Cold Months - Visual EDA 5yrs pre-Dx
Scatterplot of continuous PM 5yrs pre-diagnosis vs FVC, with sex represented by the color of points - Pre-2015
```{r}
(PM_warm %>% ggplot(aes(x=PM_5yrPreDx, y=fvc_pct, color=sex))+
   geom_point()+
   labs(x="PM 5yrs Pre-Diagnosis", y="Baseline FVC % Predicted", title="Baseline FVC % Predicted by PM 5yrs Pre-Diagnosis")+
   geom_smooth(method="lm", se = FALSE) +
   theme(plot.title = element_text(hjust = 0.5))+
   xlim(5,20))
```

Scatterplot of continuous PM 5yrs pre-diagnosis vs FVC, with sex represented by the color of points - Post-2015
```{r}
(PM_cold %>% ggplot(aes(x=PM_5yrPreDx, y=fvc_pct, color=sex))+
   geom_point()+
   labs(x="PM 5yrs Pre-Diagnosis", y="Baseline FVC % Predicted", title="Baseline FVC % Predicted by PM 5yrs Pre-Diagnosis")+
   geom_smooth(method="lm", se = FALSE) +
   theme(plot.title = element_text(hjust = 0.5))+
   xlim(5,20))
```

Violin plot wrapping boxplot to visualize PM 5yrs pre-dx low vs high vs FVC - Pre-2015
```{r}
PM_warmb <- PM_warm %>% filter(!is.na(PM5yr_dich))
(PM_warmb %>% ggplot(aes(x=PM5yr_dich, y=fvc_pct, fill=PM5yr_dich))+
   geom_boxplot(width=0.2, color="black", alpha=1.0)+
   geom_violin(width=1.0, alpha=0.5)+
   labs(x="PM 5yrs Pre-Diagnosis Low vs High", y="Baseline FVC % Predicted", title="Baseline FVC % Predicted by PM 5yrs Pre-Diagnosis Low vs High")+
   theme_light()+
   theme(legend.position = "none", plot.title = element_text(hjust = 0.5))+
   scale_fill_brewer(type="seq", palette="YlOrRd"))
```

Violin plot wrapping boxplot to visualize PM 5yrs pre-dx low vs high vs FVC - Post-2015
```{r}
PM_coldb <- PM_cold %>% filter(!is.na(PM5yr_dich))
(PM_coldb %>% ggplot(aes(x=PM5yr_dich, y=fvc_pct, fill=PM5yr_dich))+
   geom_boxplot(width=0.2, color="black", alpha=1.0)+
   geom_violin(width=1.0, alpha=0.5)+
   labs(x="PM 5yrs Pre-Diagnosis Low vs High", y="Baseline FVC % Predicted", title="Baseline FVC % Predicted by PM 5yrs Pre-Diagnosis Low vs High")+
   theme_light()+
   theme(legend.position = "none", plot.title = element_text(hjust = 0.5))+
   scale_fill_brewer(type="seq", palette="YlOrRd"))
```



## Warm vs Cold Months - PM 5yrs Pre-Diagnosis and FVC Models
Looking at continuous PM_5yrPreDx vs baseline FVC

Simplest model just looking at PM in 5yrs pre-diagnosis
```{r}
FVC_model1a <- lm(fvc_pct ~ PM_5yrPreDx, data=PM_warm)
summary(FVC_model1a)
confint(FVC_model1a)
```

Simplest model just looking at PM in 5yrs pre-diagnosis
```{r}
FVC_model1b <- lm(fvc_pct ~ PM_5yrPreDx, data=PM_cold)
summary(FVC_model1b)
confint(FVC_model1b)
```

Complete model
```{r}
FVC_model2a <- lm(fvc_pct ~ PM_5yrPreDx + sex + age_dx + smokeHx + dich_Race + avg_s + site, data=PM_warm)
summary(FVC_model2a)
confint(FVC_model2a)
```

Complete model
```{r}
FVC_model2b <- lm(fvc_pct ~ PM_5yrPreDx + sex + age_dx + smokeHx + dich_Race + avg_s + site, data=PM_cold)
summary(FVC_model2b)
confint(FVC_model2b)
```


### Warm vs Cold Months - PM 5yrs Pre-Diagnosis Low vs High and Baseline FVC
Looking at PM5yr_dich vs baseline FVC

Simplest model just looking at PM in 5yrs pre-diagnosis
```{r}
FVC_model1a <- lm(fvc_pct ~ PM5yr_dich, data=PM_warm)
summary(FVC_model1a)
confint(FVC_model1a)
```

Simplest model just looking at PM in 5yrs pre-diagnosis
```{r}
FVC_model1b <- lm(fvc_pct ~ PM5yr_dich, data=PM_cold)
summary(FVC_model1b)
confint(FVC_model1b)
```

Complete model
```{r}
FVC_model2a <- lm(fvc_pct ~ PM5yr_dich + sex + age_dx + smokeHx + dich_Race + avg_s + site, data=PM_warm)
summary(FVC_model2a)
confint(FVC_model2a)
```

Complete model
```{r}
FVC_model2b <- lm(fvc_pct ~ PM5yr_dich + sex + age_dx + smokeHx + dich_Race + avg_s + site, data=PM_cold)
summary(FVC_model2b)
confint(FVC_model2b)
```

# Warm vs Cold Months - Association between PM and Baseline DLCO
## Warm vs Cold Months - Visual EDA 5yrs pre-Dx
Scatterplot of continuous PM 5yrs pre-diagnosis vs DLCO, with sex represented by the color of points - Pre-2015
```{r}
(PM_warm %>% ggplot(aes(x=PM_5yrPreDx, y=dlco_pct, color=sex))+
   geom_point()+
   labs(x="PM 5yrs Pre-Diagnosis", y="Baseline DLCO % Predicted", title="Baseline DLCO % Predicted by PM 5yrs Pre-Diagnosis")+
   geom_smooth(method="lm", se = FALSE) +
   theme(plot.title = element_text(hjust = 0.5))+
   xlim(5,20))
```

Scatterplot of continuous PM 5yrs pre-diagnosis vs DLCO, with sex represented by the color of points - Post-2015
```{r}
(PM_cold %>% ggplot(aes(x=PM_5yrPreDx, y=dlco_pct, color=sex))+
   geom_point()+
   labs(x="PM 5yrs Pre-Diagnosis", y="Baseline DLCO % Predicted", title="Baseline DLCO % Predicted by PM 5yrs Pre-Diagnosis")+
   geom_smooth(method="lm", se = FALSE) +
   theme(plot.title = element_text(hjust = 0.5))+
   xlim(5,20))
```

Violin plot wrapping boxplot to visualize PM 5yrs pre-dx low vs high vs DLCO - Pre-2015
```{r}
(PM_warmb %>% ggplot(aes(x=PM5yr_dich, y=dlco_pct, fill=PM5yr_dich))+
   geom_boxplot(width=0.2, color="black", alpha=1.0)+
   geom_violin(width=1.0, alpha=0.5)+
   labs(x="PM 5yrs Pre-Diagnosis Low vs High", y="Baseline DLCO % Predicted", title="Baseline DLCO % Predicted by PM 5yrs Pre-Diagnosis Low vs High")+
   theme_light()+
   theme(legend.position = "none", plot.title = element_text(hjust = 0.5))+
   scale_fill_brewer(type="seq", palette="YlOrRd"))
```

Violin plot wrapping boxplot to visualize PM 5yrs pre-dx low vs high vs DLCO - Post-2015
```{r}
(PM_coldb %>% ggplot(aes(x=PM5yr_dich, y=dlco_pct, fill=PM5yr_dich))+
   geom_boxplot(width=0.2, color="black", alpha=1.0)+
   geom_violin(width=1.0, alpha=0.5)+
   labs(x="PM 5yrs Pre-Diagnosis Low vs High", y="Baseline DLCO % Predicted", title="Baseline DLCO % Predicted by PM 5yrs Pre-Diagnosis Low vs High")+
   theme_light()+
   theme(legend.position = "none", plot.title = element_text(hjust = 0.5))+
   scale_fill_brewer(type="seq", palette="YlOrRd"))
```

## Warm vs Cold Months - PM 5yrs Pre-Diagnosis and DLCO Models
### Warm vs Cold Months - PM 5yrs Pre-Diagnosis and Baseline DLCO
Looking at continuous PM_5yrPreDx vs baseline DLCO

Simplest model just looking at PM in 5yrs pre-diagnosis
```{r}
DLCO_model1a <- lm(dlco_pct ~ PM_5yrPreDx, data=PM_warm)
summary(DLCO_model1a)
confint(DLCO_model1a)
```

Simplest model just looking at PM in 5yrs pre-diagnosis
```{r}
DLCO_model1b <- lm(dlco_pct ~ PM_5yrPreDx, data=PM_cold)
summary(DLCO_model1b)
confint(DLCO_model1b)
```

Complete model
```{r}
DLCO_model2a <- lm(dlco_pct ~ PM_5yrPreDx + sex + age_dx + smokeHx + dich_Race + avg_s + site, data=PM_warm)
summary(DLCO_model2a)
confint(DLCO_model2a)
```

Complete model
```{r}
DLCO_model2b <- lm(dlco_pct ~ PM_5yrPreDx + sex + age_dx + smokeHx + dich_Race + avg_s + site, data=PM_cold)
summary(DLCO_model2b)
confint(DLCO_model2b)
```

### Warm vs Cold Months - PM 5yrs Pre-Diagnosis Low vs High and Baseline DLCO
Looking at PM5yr_dich vs baseline DLCO

Simplest model just looking at PM in 5yrs pre-diagnosis
```{r}
DLCO_model1a <- lm(dlco_pct ~ PM5yr_dich, data=PM_warm)
summary(DLCO_model1a)
confint(DLCO_model1a)
```

Simplest model just looking at PM in 5yrs pre-diagnosis
```{r}
DLCO_model1b <- lm(dlco_pct ~ PM5yr_dich, data=PM_cold)
summary(DLCO_model1b)
confint(DLCO_model1b)
```

Complete model
```{r}
DLCO_model2a <- lm(dlco_pct ~ PM5yr_dich + sex + age_dx + smokeHx + dich_Race + avg_s + site, data=PM_warm)
summary(DLCO_model2a)
confint(DLCO_model2a)
```

Complete model
```{r}
DLCO_model2b <- lm(dlco_pct ~ PM5yr_dich + sex + age_dx + smokeHx + dich_Race + avg_s + site, data=PM_cold)
summary(DLCO_model2b)
confint(DLCO_model2b)
```


