---
title: "SS Analysis with All Cohorts - Post-Publication Time-Weighted Analysis"
author: "Gillian Goobie"
date: "2022/12/27"
output: 
  html_document: 
    toc: yes
    toc_float: yes
    number_sections: yes
    code_folding: hide
  bookdown::html_document2:
    toc: yes
    toc_float: yes
    number_sections: yes
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE, warning=FALSE, echo=F}
library(tidyverse)
library(readxl)
library(here)
library(writexl)
library(knitr)
library(lintr)
library(tidync)
library(maps)
library(ncmeta)
library(devtools)
library(RNetCDF)
library(ncdf4)
library(survival)
library(survminer)
library(lubridate)
library(cmprsk)
library(riskRegression)
library(prodlim)
library(RColorBrewer)
library(prodlim)
library(lme4)
library(performance)
library(psycho)
library(report)
library(rms)
library(splines)
library(Greg)
library(rstpm2)
library(pspline)
library(raster)
library(data.table)
library(intervalaverage)
```

# Simmons - Importing Datasets
Here I am importing the file which contains monthly SS level estimates by satellite at nearest lon/lat to Simmons patient residential addresses. These are linked to the patient ID.
```{r}
outfile1 <- here("MonthlySSData/Simm_fILD_2000_2017_SS_2021_11_05.xlsx")
SS <- read_excel(outfile1)
```

Here I am importing the file which I used for my ADI work that contains the baseline clinical and demographic data for 1425 patients who have ADI
```{r}
outfile2 <- here("MonthlySSData/Simmons_fILDPts_BaselineData_2022_03_28.xlsx")
Simm <- read_excel(outfile2)
```

# Simmons - Simplifying SS Dataframe
```{r}
SS <- SS %>% dplyr::select(!c(nrow, dist, lon, lat))
colnames(SS)
```

```{r}
SS <- SS %>% 
  pivot_longer(cols=c(2:217), names_to="SS_date", names_prefix="SS_")
```

```{r}
SSx <- SS 
SSx$SS_date <- gsub("jan", "01-01-20", SSx$SS_date)
SSx$SS_date <- gsub("feb", "01-02-20", SSx$SS_date)
SSx$SS_date <- gsub("mar", "01-03-20", SSx$SS_date)
SSx$SS_date <- gsub("apr", "01-04-20", SSx$SS_date)
SSx$SS_date <- gsub("may", "01-05-20", SSx$SS_date)
SSx$SS_date <- gsub("jun", "01-06-20", SSx$SS_date)
SSx$SS_date <- gsub("jul", "01-07-20", SSx$SS_date)
SSx$SS_date <- gsub("aug", "01-08-20", SSx$SS_date)
SSx$SS_date <- gsub("sep", "01-09-20", SSx$SS_date)
SSx$SS_date <- gsub("oct", "01-10-20", SSx$SS_date)
SSx$SS_date <- gsub("nov", "01-11-20", SSx$SS_date)
SSx$SS_date <- gsub("dec", "01-12-20", SSx$SS_date)

SSx$SS_date <- format(as.Date(SSx$SS_date, format="%d-%m-%Y"),"%Y-%m-%d")
SSx$SS_date <- as.Date(SSx$SS_date)
SS <- SSx
rm(SSx)
```


# Simmons - Simplifying Simm Dataframe
## Simmons - Death/Transplant/Censoring Date
Extracting year of diagnosis and year of death/transplant/censoring
```{r}
#Start with the year of diagnosis
Simm <- Simm %>% 
  mutate(dx_yrmo = format(as.Date(Simm$dx_date, format="%Y-%m-%d"),"%Y-%m"))
Simm <- Simm %>% 
  mutate(dx_yr = format(as.Date(Simm$dx_date, format="%Y-%m-%d"),"%Y"))
Simm$dx_yr <- as.numeric(Simm$dx_yr)

#Then the year of death or lung transplant
Simm <- Simm %>% 
  mutate(deathORtx_date = if_else(!is.na(tx_date), tx_date, death_date))
Simm <- Simm %>% 
  mutate(deathORtx_yrmo = format(as.Date(Simm$deathORtx_date, format="%Y-%m-%d"),"%Y-%m"))

#Then the year the records were last updated (i.e. year of censoring)
Simm <- Simm %>% 
  mutate(DeathTxCensor_date = if_else(!is.na(deathORtx_date), deathORtx_date, last_updated))
Simm <- Simm %>% 
  mutate(censor_yrmo = format(as.Date(Simm$DeathTxCensor_date, format="%Y-%m-%d"),"%Y-%m"))
```

## Simmons - Removing Unnecessary Columns and Correcting Dates
```{r}
Simm <- Simm %>% dplyr::select(!c(ADI_state, lat, lon, City, State, zip5, UPMC_lastvisit, Simmons_lastvisit, pkyrs, fev1_pre, fev1_pct, fvc_pre, dlco_pre, pft_timefromdx, ethnicity, dx_type, deathORtx_date, deathORtx_yrmo))
Simm <- Simm %>% 
  mutate_at(c("dob","death_date", "last_updated", "tx_date", "dx_date", "consent_date", "pft_date", "DeathTxCensor_date"), as.Date)
str(Simm)
```

## Simmons - Correcting Smoking
Need to correct smoking variables
```{r}
Simm$smokeHx <- as.character(Simm$smokeHx) 
Simm <- Simm %>% mutate(smokeHx1=if_else(is.na(smokeHx), "Unknown", smokeHx))

#now need to make new dich_smoking category
Simm$dich_smoking <- as.character(Simm$dich_smoking) 
Simm <- Simm %>% mutate(dich_smoking1=if_else(is.na(dich_smoking),"Unknown", dich_smoking))
```

Now need to remove old smoking variables and rename new ones
```{r}
Simm <- Simm %>% dplyr::select(-c(smokeHx, dich_smoking))
Simm <- Simm %>% rename(c("smokeHx"="smokeHx1", "dich_smoking"="dich_smoking1"))
Simm$smokeHx <- as.factor(Simm$smokeHx)
Simm$dich_smoking <- as.factor(Simm$dich_smoking)
Simm$smokeHx <- fct_relevel(Simm$smokeHx, c("Never","Former","Always","Unknown"))
Simm$dich_smoking <- fct_relevel(Simm$dich_smoking, c("Never","Ever","Unknown"))
```

## Simmons - Correcting Factors
Need to correct other factor variables
```{r}
Simm$sex <- fct_relevel(Simm$sex, c("M","F"))
Simm$race <- fct_relevel(Simm$race, c("W","B","A","N","U"))
Simm$dich_Race <- fct_relevel(Simm$dich_Race, c("White","Non-White"))
Simm$dx <- fct_relevel(Simm$dx, c("IPF"))
Simm$dx_group <- fct_relevel(Simm$dx_group, c("IPF"))
Simm <- Simm %>% mutate_at(c("status","died", "txed", "deadORtx"), as.factor)
```

## Simmons - Creating New Variables
### Simmons - IPF vs Other Diagnosis
```{r}
Simm <- Simm %>% mutate(dx_IPF=ifelse(dx=="IPF", "IPF", "not_IPF"))  
Simm$dx_IPF <- fct_relevel(Simm$dx_IPF, c("IPF"))
```

### Simmons - Disadvantage Distribution
Creating this empirical cumulative distribution will allow us to combine the analyses of all three cohorts even though the measurements for disadvantage are different between the three.
```{r}
plot(ecdf(Simm$ADI_nat))
Simm$disadv <- ecdf(Simm$ADI_nat)(Simm$ADI_nat)
```




# Simmons - Modifying Simmons Dataset
Take down Simmons dataset to the 1424 patients with complete data (currently at 1425)
```{r}
IDs <- as.data.table(unique(SS$ID))
IDs <- IDs %>% rename("ID"="V1")
Simm <- left_join(IDs, Simm, by="ID")
```


```{r}
Simm <- Simm %>% mutate(days_DeathTxCensor=(time_DeathTxCensor*365.25))
```
Longest time_DeathTxCensor= 20.427105yrs =7461days

30-day longest time interval would be 7470 days
```{r}
start <- seq(1, 7441, by = 30)
start
end <- seq(30, 7470, by = 30)
end
```

Repeat the list of intervals 1424 times (number of patients in Simmons)
```{r}
start <- rep(start, times=1424)
end <- rep(end, times=1424)
intervals <- as.data.frame(cbind(start, end))
```

Add ID column to intervals
```{r}
IDs <- rep(Simm$ID, each=249)
intervals <- as.data.frame(cbind(IDs, intervals))
intervals <- intervals %>% rename("ID"="IDs")
```

Join Simm and intervals
```{r}
Simm <- left_join(intervals, Simm, by="ID")
```

Determine if event occurred during interval
```{r}
Simm <- Simm %>% mutate(event=if_else((days_DeathTxCensor>=start & days_DeathTxCensor<=end), 1, 0))
```

Now will add date intervals for 5yr start and end times)
```{r}
Simm <- Simm %>% mutate(end_5yr=(dx_date + days(end)))
Simm <- Simm %>% mutate(start_5yr=(end_5yr - days(1826)))
```

Now need to remove any rows where end_5yr is > date_DeathTxCensor (Amanda indicated need to remove rows where start_5yr > date_DeathTxCensor, but I don't think this is correct because then we'd be assigning exposures to patients where they were alive/not transplanted for <5yrs for that exposure - need to double check with her)
```{r}
Simm <- Simm %>% filter(!(end_5yr>DeathTxCensor_date & event!=1))
```



# Simmons - Creating Time-Weighted Intervals of Exposures
```{r}
SS <- SS %>% mutate(start=as.IDate(SS_date))
SS <- SS %>% mutate(end=as.IDate(SS_date + months(1) - days(1)))
SS <- as.data.table(SS)
str(SS)
```

Creating a list of intervals we want to calculate exposures for:
```{r}
Simm_intervals <- Simm %>% dplyr::select(ID, start_5yr, end_5yr)
Simm_intervals <- Simm_intervals %>% mutate_at(c("start_5yr", "end_5yr"), as.IDate)
Simm_intervals <- Simm_intervals %>% rename(c("start"="start_5yr", "end"="end_5yr"))
Simm_intervals <- as.data.table(Simm_intervals)
str(Simm_intervals)
```

```{r}
SS_5yrWtedAvg <- intervalaverage(x=SS,
                y=Simm_intervals,
                interval_vars=c("start","end"),
                value_vars=c("value"),
                group_vars="ID",
                required_percentage = 0.01)
```
So this produces a row for each interval for each patient

Select only the necessary columns, but will keep the original dataframe to interrogate data missingess if needed
```{r}
SS_5yrWtedAvgx <- SS_5yrWtedAvg %>% dplyr::select(ID, value, start, end)
SS_5yrWtedAvgx <- SS_5yrWtedAvgx %>% rename("SS"="value", "start_5yr"="start", "end_5yr"="end")
SS_5yrWtedAvgx <- SS_5yrWtedAvgx %>% mutate_at(c("start_5yr", "end_5yr"), as.Date)
str(SS_5yrWtedAvgx)
```

Join to Simm
```{r}
Simm <- left_join(Simm, SS_5yrWtedAvgx, by=c("ID", "start_5yr", "end_5yr"))
```


# PFF - Importing Datasets
Here I am importing the file which contains monthly SS level estimates by satellite at nearest lon/lat to PFF patient residential addresses. These are linked to the patient ID.
```{r}
outfile1 <- here("MonthlySSData/PFF_fILD_2000_2017_SS_2021_11_05.xlsx")
SS <- read_excel(outfile1)
```

Match up SS ID to SSID from matching file
```{r}
outfile2 <- here("MonthlySSData/PFF_fILD_PM25_BaselineData_2021_10_20.xlsx")
SSx <- read_excel(outfile2)
SSx <- SSx %>% dplyr::select(ID, SSID)
```

Merge SS and SS
```{r}
SS <- left_join(SS, SSx, by="ID")
```


Here I am importing the file containing the complete baseline clinical and demographic data for 1905 PFF patients
```{r}
outfile3 <- here("PFF_fILDPts_BaselineData_ConsentDateReference_2022_08_01.xlsx")
PFF <- read_excel(outfile3)
```

# PFF - Merging Datasets
I used a inner_join here so that the complete "PFF" dataframe only includes patients with fILD that have all baseline demographics and SS data.
```{r}
str(PFF$ID)
PFF <- inner_join(PFF, SS, by=c("SSID"))
```
This results in 1905 complete records

Reorder so "ID" is the first column
```{r}
PFF <- PFF %>% dplyr::select(ID, everything(.))
```

# PFF - Making dx_yr column
```{r}
PFF <- PFF %>% 
  mutate(dx_yr = format(as.Date(PFF$dx_date, format="%Y-%m-%d"),"%Y"))
PFF$dx_yr <- as.numeric(PFF$dx_yr)
```

# PFF - Creating Site Variable
```{r}
PFF$site <- substr(PFF$SSID, 1,3)
PFF$site <- as.factor(PFF$site)
str(PFF$site)
```



# PFF - Simplifying SS Dataframe
Reorder so "ID" is the first column
```{r}
SS <- SS %>% dplyr::select(ID, everything(.))
SS <- SS %>% dplyr::select(!SSID)
```

```{r}
SS <- SS %>% dplyr::select(!c(nrow, dist, lon, lat))
colnames(SS)
```

```{r}
SS <- SS %>% 
  pivot_longer(cols=c(2:217), names_to="SS_date", names_prefix="SS_")
```

```{r}
SSx <- SS 
SSx$SS_date <- gsub("jan", "01-01-20", SSx$SS_date)
SSx$SS_date <- gsub("feb", "01-02-20", SSx$SS_date)
SSx$SS_date <- gsub("mar", "01-03-20", SSx$SS_date)
SSx$SS_date <- gsub("apr", "01-04-20", SSx$SS_date)
SSx$SS_date <- gsub("may", "01-05-20", SSx$SS_date)
SSx$SS_date <- gsub("jun", "01-06-20", SSx$SS_date)
SSx$SS_date <- gsub("jul", "01-07-20", SSx$SS_date)
SSx$SS_date <- gsub("aug", "01-08-20", SSx$SS_date)
SSx$SS_date <- gsub("sep", "01-09-20", SSx$SS_date)
SSx$SS_date <- gsub("oct", "01-10-20", SSx$SS_date)
SSx$SS_date <- gsub("nov", "01-11-20", SSx$SS_date)
SSx$SS_date <- gsub("dec", "01-12-20", SSx$SS_date)

SSx$SS_date <- format(as.Date(SSx$SS_date, format="%d-%m-%Y"),"%Y-%m-%d")
SSx$SS_date <- as.Date(SSx$SS_date)
SS <- SSx
rm(SSx)
```


# PFF - Simplifying PFF Dataframe
## PFF - Death/Transplant/Censor Date
Extracting year of diagnosis and year of death/transplant/censoring
```{r}
#Start with the year of diagnosis
PFF <- PFF %>% 
  mutate(dx_yrmo = format(as.Date(PFF$dx_date, format="%Y-%m-%d"),"%Y-%m"))
PFF <- PFF %>% 
  mutate(dx_yr = format(as.Date(PFF$dx_date, format="%Y-%m-%d"),"%Y"))
PFF$dx_yr <- as.numeric(PFF$dx_yr)

#Then the year of death or lung transplant
PFF <- PFF %>% 
  mutate(deathORtx_date = if_else(!is.na(tx_date), tx_date, death_date))
PFF <- PFF %>% 
  mutate(deathORtx_yrmo = format(as.Date(PFF$deathORtx_date, format="%Y-%m-%d"),"%Y-%m"))

#Then the year the records were last updated (i.e. year of censoring)
PFF <- PFF %>% 
  mutate(DeathTxCensor_date = if_else(!is.na(deathORtx_date), deathORtx_date, censor_date))
PFF <- PFF %>% 
  mutate(censor_yrmo = format(as.Date(PFF$DeathTxCensor_date, format="%Y-%m-%d"),"%Y-%m"))
```

## PFF - Removing Unnecessary Columns and Correcting Dates
```{r}
PFF <- PFF %>% dplyr::select(c(ID, site, age_dx, sex, smokeHx, race, dich_Race, pct_belowpoverty, dx_group, dx, dx_date, death_date, tx_date, DeathTxCensor_date, censor_date, fvc_date, dlco_date, fvc_pct, dlco_pct, status, deadORtx, time_DeathTxCensor, dx_yr))
PFF <- PFF %>% 
  mutate_at(c("death_date", "censor_date", "tx_date", "dx_date", "fvc_date", "dlco_date", "DeathTxCensor_date"), as.Date)
str(PFF)
```

## PFF - Correcting Smoking
Need to correct smoking variables
```{r}
PFF$smokeHx <- as.character(PFF$smokeHx) 
PFF <- PFF %>% mutate(smokeHx1=if_else(is.na(smokeHx), "Unknown", smokeHx))

```

Now need to remove old smoking variables and rename new ones
```{r}
PFF <- PFF %>% dplyr::select(-c(smokeHx))
PFF <- PFF %>% rename(c("smokeHx"="smokeHx1"))
PFF$smokeHx <- as.factor(PFF$smokeHx)
PFF$smokeHx <- fct_relevel(PFF$smokeHx, c("Never","Ever","Unknown"))
```

## PFF - Correcting Factors
Need to correct other factor variables
```{r}
PFF$sex <- fct_relevel(PFF$sex, c("Male","Female"))
PFF$race <- fct_relevel(PFF$race, c("W","B","A","N","U"))
PFF$dich_Race <- fct_relevel(PFF$dich_Race, c("White","Non-White"))
PFF$dx <- fct_relevel(PFF$dx, c("IPF"))
PFF$dx_group <- fct_relevel(PFF$dx_group, c("IPF"))
PFF <- PFF %>% mutate_at(c("status", "site"), as.factor)
```

## PFF - Creating New Variables
### PFF - IPF vs Other Diagnosis
```{r}
PFF <- PFF %>% mutate(dx_IPF=ifelse(dx=="IPF", "IPF", "not_IPF"))  
PFF$dx_IPF <- fct_relevel(PFF$dx_IPF, c("IPF"))
```

### PFF - Disadvantage Distribution
Creating this empirical cumulative distribution will allow us to combine the analyses of all three cohorts even though the measurements for disadvantage are different between the three.
```{r}
plot(ecdf(PFF$pct_belowpoverty))
PFF$disadv <- ecdf(PFF$pct_belowpoverty)(PFF$pct_belowpoverty)
```

# PFF - Modifying PFF Dataset
Take down PFF dataset to the 1905 patients with complete data
```{r}
IDs <- as.data.table(unique(SS$ID))
IDs <- IDs %>% rename("ID"="V1")
PFF <- left_join(IDs, PFF, by="ID")
```


```{r}
PFF <- PFF %>% mutate(days_DeathTxCensor=(time_DeathTxCensor*365.25))
```
Longest time_DeathTxCensor= 1848days

30-day longest time interval would be 1860 days
```{r}
start <- seq(1, 1831, by = 30)
start
end <- seq(30, 1860, by = 30)
end
```

Repeat the list of intervals 1905 times (number of patients in PFF)
```{r}
start <- rep(start, times=1905)
end <- rep(end, times=1905)
intervals <- as.data.frame(cbind(start, end))
```

Add ID column to intervals
```{r}
IDs <- rep(PFF$ID, each=62)
intervals <- as.data.frame(cbind(IDs, intervals))
intervals <- intervals %>% rename("ID"="IDs")
```

Join PFF and intervals
```{r}
PFF <- left_join(intervals, PFF, by="ID")
```

Determine if event occurred during interval
```{r}
PFF <- PFF %>% mutate(event=if_else((days_DeathTxCensor>=start & days_DeathTxCensor<=end), 1, 0))
```

Now will add date intervals for 5yr start and end times)
```{r}
PFF <- PFF %>% mutate(end_5yr=(dx_date + days(end)))
PFF <- PFF %>% mutate(start_5yr=(end_5yr - days(1826)))
```

Now need to remove any rows where end_5yr is > date_DeathTxCensor (Amanda indicated need to remove rows where start_5yr > date_DeathTxCensor, but I don't think this is correct because then we'd be assigning exposures to patients where they were alive/not transplanted for <5yrs for that exposure - need to double check with her)
```{r}
PFF <- PFF %>% filter(!(end_5yr>DeathTxCensor_date & event!=1))
```



# PFF - Creating Time-Weighted Intervals of Exposures
```{r}
SS <- SS %>% mutate(start=as.IDate(SS_date))
SS <- SS %>% mutate(end=as.IDate(SS_date + months(1) - days(1)))
SS <- as.data.table(SS)
str(SS)
```

Creating a list of intervals we want to calculate exposures for:
```{r}
PFF_intervals <- PFF %>% dplyr::select(ID, start_5yr, end_5yr)
PFF_intervals <- PFF_intervals %>% mutate_at(c("start_5yr", "end_5yr"), as.IDate)
PFF_intervals <- PFF_intervals %>% rename(c("start"="start_5yr", "end"="end_5yr"))
PFF_intervals <- as.data.table(PFF_intervals)
str(PFF_intervals)
```

```{r}
SS_5yrWtedAvg <- intervalaverage(x=SS,
                y=PFF_intervals,
                interval_vars=c("start","end"),
                value_vars=c("value"),
                group_vars="ID",
                required_percentage = 0.01)
```
So this produces a row for each interval for each patient

Select only the necessary columns, but will keep the original dataframe to interrogate data missingess if needed
```{r}
SS_5yrWtedAvgx <- SS_5yrWtedAvg %>% dplyr::select(ID, value, start, end)
SS_5yrWtedAvgx <- SS_5yrWtedAvgx %>% rename("SS"="value", "start_5yr"="start", "end_5yr"="end")
SS_5yrWtedAvgx <- SS_5yrWtedAvgx %>% mutate_at(c("start_5yr", "end_5yr"), as.Date)
str(SS_5yrWtedAvgx)
```

Join to PFF
```{r}
PFF <- left_join(PFF, SS_5yrWtedAvgx, by=c("ID", "start_5yr", "end_5yr"))
```




# CARE-PF - Importing CARE-PF Datasets
Here I am importing the file which contains monthly SS level estimates by satellite at nearest lon/lat to CAREPF patient residential addresses. These are linked to the patient ID.
```{r}
outfile1 <- here("MonthlySSData/CARE_fILD_2000_2017_SS_2021_11_05.xlsx")
SS <- read_excel(outfile1)
```

Here I am importing the file which I used for my CIMD work that contains the baseline clinical and demographic data for patients who have CIMD
```{r}
outfile2 <- here("MonthlySSData/CAREPF_fILDPts_BaselineData_2021_10_22.xlsx")
CARE <- read_excel(outfile2)
```

# CARE-PF - Simplifying SS Dataframe
```{r}
SS <- SS %>% dplyr::select(!c(nrow, dist, lon, lat))
colnames(SS)
```

```{r}
SS <- SS %>% 
  pivot_longer(cols=c(2:217), names_to="SS_date", names_prefix="SS_")
```

```{r}
SSx <- SS 
SSx$SS_date <- gsub("jan", "01-01-20", SSx$SS_date)
SSx$SS_date <- gsub("feb", "01-02-20", SSx$SS_date)
SSx$SS_date <- gsub("mar", "01-03-20", SSx$SS_date)
SSx$SS_date <- gsub("apr", "01-04-20", SSx$SS_date)
SSx$SS_date <- gsub("may", "01-05-20", SSx$SS_date)
SSx$SS_date <- gsub("jun", "01-06-20", SSx$SS_date)
SSx$SS_date <- gsub("jul", "01-07-20", SSx$SS_date)
SSx$SS_date <- gsub("aug", "01-08-20", SSx$SS_date)
SSx$SS_date <- gsub("sep", "01-09-20", SSx$SS_date)
SSx$SS_date <- gsub("oct", "01-10-20", SSx$SS_date)
SSx$SS_date <- gsub("nov", "01-11-20", SSx$SS_date)
SSx$SS_date <- gsub("dec", "01-12-20", SSx$SS_date)

SSx$SS_date <- format(as.Date(SSx$SS_date, format="%d-%m-%Y"),"%Y-%m-%d")
SSx$SS_date <- as.Date(SSx$SS_date)
SS <- SSx
rm(SSx)
```


# CARE-PF - Simplifying CARE Dataframe
## CARE-PF - Death/Transplant/Censor Date
Extracting year of diagnosis and year of death/transplant/censoring
```{r}
#Start with the year of diagnosis
CARE <- CARE %>% 
  mutate(dx_yrmo = format(as.Date(CARE$dx_date, format="%Y-%m-%d"),"%Y-%m"))
CARE <- CARE %>% 
  mutate(dx_yr = format(as.Date(CARE$dx_date, format="%Y-%m-%d"),"%Y"))
CARE$dx_yr <- as.numeric(CARE$dx_yr)

#Then the year of death or lung transplant
CARE <- CARE %>% 
  mutate(deathORtx_date = if_else(!is.na(tx_date), tx_date, death_date))
CARE <- CARE %>% 
  mutate(deathORtx_yrmo = format(as.Date(CARE$deathORtx_date, format="%Y-%m-%d"),"%Y-%m"))

#Then the year the records were last updated (i.e. year of censoring)
CARE <- CARE %>% 
  mutate(DeathTxCensor_date = if_else(!is.na(deathORtx_date), deathORtx_date, last_updated))
CARE <- CARE %>% 
  mutate(censor_yrmo = format(as.Date(CARE$DeathTxCensor_date, format="%Y-%m-%d"),"%Y-%m"))
```

## CARE-PF - Removing Unnecessary Columns and Correcting Dates
```{r}
CARE <- CARE %>% dplyr::select(c(ID, site, age_dx, sex, smokeHx, dich_smoking, race, dich_Race, avg_s, dx_group, dx, dob, dx_date, death_date, tx_date, DeathTxCensor_date, last_updated, pft_date, fvc_pct, dlco_pct, pft_timefromdx, status, died, txed, deadORtx, time_DeathTxCensor, dx_yr))
CARE <- CARE %>% 
  mutate_at(c("dob","death_date", "last_updated", "tx_date", "dx_date", "pft_date", "DeathTxCensor_date"), as.Date)
str(CARE)
```

## CARE-PF - Correcting Smoking
Need to correct smoking variables
```{r}
CARE$smokeHx <- as.character(CARE$smokeHx) 
CARE <- CARE %>% mutate(smokeHx1=if_else(is.na(smokeHx), "Unknown", smokeHx))

#now need to make new dich_smoking category
CARE$dich_smoking <- as.character(CARE$dich_smoking) 
CARE <- CARE %>% mutate(dich_smoking1=if_else(is.na(dich_smoking),"Unknown", dich_smoking))
```

Now need to remove old smoking variables and rename new ones
```{r}
CARE <- CARE %>% dplyr::select(-c(smokeHx, dich_smoking))
CARE <- CARE %>% rename(c("smokeHx"="smokeHx1", "dich_smoking"="dich_smoking1"))
CARE$smokeHx <- as.factor(CARE$smokeHx)
CARE$dich_smoking <- as.factor(CARE$dich_smoking)
CARE$smokeHx <- fct_relevel(CARE$smokeHx, c("Never","Former","Always","Unknown"))
CARE$dich_smoking <- fct_relevel(CARE$dich_smoking, c("Never","Ever","Unknown"))
```

## CARE-PF - Correcting Factors
Need to correct other factor variables
```{r}
CARE$sex <- fct_relevel(CARE$sex, c("M","F"))
CARE$race <- fct_relevel(CARE$race, c("W","B","A","N","U"))
CARE$dich_Race <- fct_relevel(CARE$dich_Race, c("White","Non-White"))
CARE$dx <- fct_relevel(CARE$dx, c("IPF"))
CARE$dx_group <- fct_relevel(CARE$dx_group, c("IPF"))
CARE <- CARE %>% mutate_at(c("status","died", "txed", "deadORtx", "site"), as.factor)
```

## CARE-PF - Creating New Variables
### CARE-PF - IPF vs Other Diagnosis
```{r}
CARE <- CARE %>% mutate(dx_IPF=ifelse(dx=="IPF", "IPF", "not_IPF"))  
CARE$dx_IPF <- fct_relevel(CARE$dx_IPF, c("IPF"))
```

### CARE-PF - Disadvantage Distribution
Creating this empirical cumulative distribution will allow us to combine the analyses of all three cohorts even though the measurements for disadvantage are different between the three.
```{r}
plot(ecdf(CARE$avg_s))
CARE$disadv <- ecdf(CARE$avg_s)(CARE$avg_s)
```

# CARE-PF - Modifying CARE Dataset
Take down CARE dataset to the 3389 patients with complete data
```{r}
IDs <- as.data.table(unique(SS$ID))
IDs <- IDs %>% rename("ID"="V1")
CARE <- left_join(IDs, CARE, by="ID")
```


```{r}
CARE <- CARE %>% mutate(days_DeathTxCensor=(time_DeathTxCensor*365.25))
```
Longest time_DeathTxCensor= 9546days

30-day longest time interval would be 9570 days
```{r}
start <- seq(1, 9541, by = 30)
start
end <- seq(30, 9570, by = 30)
end
```

Repeat the list of intervals 3389 times (number of patients in CARE)
```{r}
start <- rep(start, times=3389)
end <- rep(end, times=3389)
intervals <- as.data.frame(cbind(start, end))
```

Add ID column to intervals
```{r}
IDs <- rep(CARE$ID, each=319)
intervals <- as.data.frame(cbind(IDs, intervals))
intervals <- intervals %>% rename("ID"="IDs")
```

Join CARE and intervals
```{r}
CARE <- left_join(intervals, CARE, by="ID")
```

Determine if event occurred during interval
```{r}
CARE <- CARE %>% mutate(event=if_else((days_DeathTxCensor>=start & days_DeathTxCensor<=end), 1, 0))
```

Now will add date intervals for 5yr start and end times)
```{r}
CARE <- CARE %>% mutate(end_5yr=(dx_date + days(end)))
CARE <- CARE %>% mutate(start_5yr=(end_5yr - days(1826)))
```

Now need to remove any rows where end_5yr is > date_DeathTxCensor (Amanda indicated need to remove rows where start_5yr > date_DeathTxCensor, but I don't think this is correct because then we'd be assigning exposures to patients where they were alive/not transplanted for <5yrs for that exposure - need to double check with her)
```{r}
CARE <- CARE %>% filter(!(end_5yr>DeathTxCensor_date & event!=1))
```



# CARE-PF - Creating Time-Weighted Intervals of Exposures
```{r}
SS <- SS %>% mutate(start=as.IDate(SS_date))
SS <- SS %>% mutate(end=as.IDate(SS_date + months(1) - days(1)))
SS <- as.data.table(SS)
str(SS)
```

Creating a list of intervals we want to calculate exposures for:
```{r}
CARE_intervals <- CARE %>% dplyr::select(ID, start_5yr, end_5yr)
CARE_intervals <- CARE_intervals %>% mutate_at(c("start_5yr", "end_5yr"), as.IDate)
CARE_intervals <- CARE_intervals %>% rename(c("start"="start_5yr", "end"="end_5yr"))
CARE_intervals <- as.data.table(CARE_intervals)
str(CARE_intervals)
```

```{r}
SS_5yrWtedAvg <- intervalaverage(x=SS,
                y=CARE_intervals,
                interval_vars=c("start","end"),
                value_vars=c("value"),
                group_vars="ID",
                required_percentage = 0.01)
```
So this produces a row for each interval for each patient

Select only the necessary columns, but will keep the original dataframe to interrogate data missingess if needed
```{r}
SS_5yrWtedAvgx <- SS_5yrWtedAvg %>% dplyr::select(ID, value, start, end)
SS_5yrWtedAvgx <- SS_5yrWtedAvgx %>% rename("SS"="value", "start_5yr"="start", "end_5yr"="end")
SS_5yrWtedAvgx <- SS_5yrWtedAvgx %>% mutate_at(c("start_5yr", "end_5yr"), as.Date)
str(SS_5yrWtedAvgx)
```

Join to CARE
```{r}
CARE <- left_join(CARE, SS_5yrWtedAvgx, by=c("ID", "start_5yr", "end_5yr"))
```


# Remove unnecessary dataframes
```{r}
rm(CARE_intervals, intervals, PFF_intervals, SS, SS_5yrWtedAvg, SS_5yrWtedAvgx, SS, Simm_intervals, end, start, IDs)
```

# Make Dichotomized SS Cutpoints
```{r}
Simm$SS_dich <- cut(Simm$SS,
                      breaks=c(0, 0.25, 100),
                      labels=c("Low", "High"))
summary(Simm$SS_dich)
class(Simm$SS_dich)
str(Simm$SS_dich)
```

```{r}
PFF$SS_dich <- cut(PFF$SS,
                      breaks=c(0, 0.25, 100),
                      labels=c("Low", "High"))
summary(PFF$SS_dich)
class(PFF$SS_dich)
str(PFF$SS_dich)
```

```{r}
CARE$SS_dich <- cut(CARE$SS,
                      breaks=c(0, 0.25, 100),
                      labels=c("Low", "High"))
summary(CARE$SS_dich)
class(CARE$SS_dich)
str(CARE$SS_dich)
```

# Combine all into one dataframe
```{r}
colnames(Simm)
colnames(PFF)
colnames(CARE)
```

```{r}
Simm <- Simm %>% mutate(site=1)
Simm <- Simm %>% mutate(dlco_date=pft_date)
Simm <- Simm %>% rename(c("fvc_date"="pft_date"))
Simm <- Simm %>% dplyr::select(ID, start, end, site, age_dx, sex, race, dich_Race, smokeHx, dx_IPF, disadv, dx_group, dx, dx_yr, dx_date, death_date, tx_date, DeathTxCensor_date, last_updated, fvc_date, dlco_date, fvc_pct, dlco_pct, status, deadORtx, time_DeathTxCensor, days_DeathTxCensor, event, start_5yr, end_5yr, SS, SS_dich)
```

```{r}
PFF <- PFF %>% rename(c("last_updated"="censor_date"))
PFF <- PFF %>% dplyr::select(ID, start, end, site, age_dx, sex, race, dich_Race, smokeHx, dx_IPF, disadv, dx_group, dx, dx_yr, dx_date, death_date, tx_date, DeathTxCensor_date, last_updated, fvc_date, dlco_date, fvc_pct, dlco_pct, status, deadORtx, time_DeathTxCensor, days_DeathTxCensor, event, start_5yr, end_5yr, SS, SS_dich)
```

```{r}
CARE <- CARE %>% mutate(dlco_date=pft_date)
CARE <- CARE %>% rename(c("fvc_date"="pft_date"))
CARE <- CARE %>% dplyr::select(ID, start, end, site, age_dx, sex, race, dich_Race, smokeHx, dx_IPF, disadv, dx_group, dx, dx_yr, dx_date, death_date, tx_date, DeathTxCensor_date, last_updated, fvc_date, dlco_date, fvc_pct, dlco_pct, status, deadORtx, time_DeathTxCensor, days_DeathTxCensor, event, start_5yr, end_5yr, SS, SS_dich)
```



# Survival Models
## SS Continuous
### Undajusted (except dx_yr and site)
Simmons
```{r}
coxPH_model1 <- coxph(Surv(start, end, event==1) ~ SS + dx_yr + site, data=Simm, id=ID)
summary(coxPH_model1)
```

PFF
```{r}
coxPH_model1 <- coxph(Surv(start, end, event==1) ~ SS + dx_yr + site, data=PFF, id=ID)
summary(coxPH_model1)
```

CARE-PF
```{r}
coxPH_model1 <- coxph(Surv(start, end, event==1) ~ SS + dx_yr + site, data=CARE, id=ID)
summary(coxPH_model1)
```


### Fully-Adjusted
Simmons
```{r}
coxPH_model2 <- coxph(Surv(start, end, event==1) ~ SS + dx_yr + age_dx + sex + dich_Race + smokeHx + disadv + site, data=Simm, id=ID)
summary(coxPH_model2)
```

PFF
```{r}
coxPH_model2 <- coxph(Surv(start, end, event==1) ~ SS + dx_yr + age_dx + sex + dich_Race + smokeHx + disadv + site, data=PFF, id=ID)
summary(coxPH_model2)
```

CARE-PF
```{r}
coxPH_model2 <- coxph(Surv(start, end, event==1) ~ SS + dx_yr + age_dx + sex + dich_Race + smokeHx + disadv + site, data=CARE, id=ID)
summary(coxPH_model2)
```


### Variable HR for SS Continuous
Want a plot with x-axis of SS level and y-axis of HR - unsure if it makes sense to use these plots for the time-weighted exposure approach we have here.

Unadjusted model + dx_yr + site
```{r}
#First need to make dataframe that only includes patients with time_DeathTxCensor
Simmx <- Simm %>% filter(!is.na(SS) & !is.na(time_DeathTxCensor) & !is.na(dx_yr) & !is.na(site) & SS!=0)
PFFx <- PFF %>% filter(!is.na(SS) & !is.na(time_DeathTxCensor) & !is.na(dx_yr) & !is.na(site) & SS!=0)
CAREx <- CARE %>% filter(!is.na(SS) & !is.na(time_DeathTxCensor) & !is.na(dx_yr) & !is.na(site) & SS!=0)

#Then make survival function
surv1 <- Surv(Simmx$start, Simmx$end, Simmx$event==1)
fit1 <- coxph(surv1 ~ pspline(Simmx$SS, df=3) + Simmx$dx_yr + Simmx$site)
predicted1 <- predict(fit1, type="terms", se.fit=T, terms=1)

surv2 <- Surv(PFFx$start, PFFx$end, PFFx$event==1)
fit2 <- coxph(surv2 ~ pspline(PFFx$SS, df=3) + PFFx$dx_yr + PFFx$site)
predicted2 <- predict(fit2, type="terms", se.fit=T, terms=1)

surv3 <- Surv(CAREx$start, CAREx$end, CAREx$event==1)
fit3 <- coxph(surv3 ~ pspline(CAREx$SS, df=3) + CAREx$dx_yr + CAREx$site)
predicted3 <- predict(fit3, type="terms", se.fit=T, terms=1)
```

```{r, error=T}
#Then plot
plot(Simmx$SS, exp(predicted1$fit), type="n", xlim=c(0,2.5), ylim=c(0,3))
lines(sm.spline(Simmx$SS, exp(predicted1$fit)), col = "darkgoldenrod1", lty = 1 )
lines(sm.spline(Simmx$SS, exp(predicted1$fit + 1.96 * predicted1$se)), col = "lightgoldenrod1", lty = 2 )
lines(sm.spline(Simmx$SS, exp(predicted1$fit - 1.96 * predicted1$se)), col = "lightgoldenrod1", lty = 2 )

plot(PFFx$SS, exp(predicted2$fit), type="n", xlim=c(0,2.5), ylim=c(0,3))
lines(sm.spline(PFFx$SS, exp(predicted2$fit)), col = "royalblue1", lty = 1)
lines(sm.spline(PFFx$SS, exp(predicted2$fit + 1.96 * predicted2$se)), col = "steelblue1", lty = 2 )
lines(sm.spline(PFFx$SS, exp(predicted2$fit - 1.96 * predicted2$se)), col = "steelblue1", lty = 2 )

plot(CAREx$SS, exp(predicted3$fit), type="n", xlim=c(0,2.5), ylim=c(0,3))
lines(sm.spline(CAREx$SS, exp(predicted3$fit)), col = "firebrick1", lty = 1)
lines(sm.spline(CAREx$SS, exp(predicted3$fit + 1.96 * predicted3$se)), col = "coral1", lty = 2 )
lines(sm.spline(CAREx$SS, exp(predicted3$fit - 1.96 * predicted3$se)), col = "coral1", lty = 2 )
```

Complete model + dx_yr
```{r}
#First need to make dataframe that only includes patients with time_DeathTxCensor
Simmx <- Simm %>% filter(!is.na(SS) & !is.na(time_DeathTxCensor) & !is.na(dx_yr) & !is.na(time_DeathTxCensor) & !is.na(age_dx) & !is.na(sex) & !is.na(smokeHx) & !is.na(dich_Race) & !is.na(disadv) & !is.na(site) & SS!=0)
PFFx <- PFF %>% filter(!is.na(SS) & !is.na(time_DeathTxCensor) & !is.na(dx_yr) & !is.na(time_DeathTxCensor) & !is.na(age_dx) & !is.na(sex) & !is.na(smokeHx) & !is.na(dich_Race) & !is.na(disadv) & !is.na(site) & SS!=0)
CAREx <- CARE %>% filter(!is.na(SS) & !is.na(time_DeathTxCensor) & !is.na(dx_yr) & !is.na(time_DeathTxCensor) & !is.na(age_dx) & !is.na(sex) & !is.na(smokeHx) & !is.na(dich_Race) & !is.na(disadv) & !is.na(site) & SS!=0)

#Then make survival function
surv1 <- Surv(Simmx$start, Simmx$end, Simmx$event==1)
fit1 <- coxph(surv1 ~ pspline(Simmx$SS, df=3) + Simmx$dx_yr + Simmx$age_dx + Simmx$sex + Simmx$smokeHx + Simmx$dich_Race + Simmx$disadv + Simmx$site)
predicted1 <- predict(fit1, type="terms", se.fit=T, terms=1)

surv2 <- Surv(PFFx$start, PFFx$end, PFFx$event==1)
fit2 <- coxph(surv2 ~ pspline(PFFx$SS, df=3) + PFFx$dx_yr + PFFx$age_dx + PFFx$sex + PFFx$smokeHx + PFFx$dich_Race + PFFx$disadv + PFFx$site)
predicted2 <- predict(fit2, type="terms", se.fit=T, terms=1)

surv3 <- Surv(CAREx$start, CAREx$end, CAREx$event==1)
fit3 <- coxph(surv3 ~ pspline(CAREx$SS, df=3) + CAREx$dx_yr + CAREx$age_dx + CAREx$sex + CAREx$smokeHx + CAREx$dich_Race + CAREx$disadv + CAREx$site)
predicted3 <- predict(fit3, type="terms", se.fit=T, terms=1)
```

```{r, error=T}
#Then plot
plot(Simmx$SS, exp(predicted1$fit), type="n", xlim=c(0,2.5), ylim=c(0,3))
lines(sm.spline(Simmx$SS, exp(predicted1$fit)), col = "darkgoldenrod1", lty = 1 )
lines(sm.spline(Simmx$SS, exp(predicted1$fit + 1.96 * predicted1$se)), col = "lightgoldenrod1", lty = 2 )
lines(sm.spline(Simmx$SS, exp(predicted1$fit - 1.96 * predicted1$se)), col = "lightgoldenrod1", lty = 2 )

plot(PFFx$SS, exp(predicted2$fit), type="n", xlim=c(0,2.5), ylim=c(0,3))
lines(sm.spline(PFFx$SS, exp(predicted2$fit)), col = "royalblue1", lty = 1)
lines(sm.spline(PFFx$SS, exp(predicted2$fit + 1.96 * predicted2$se)), col = "steelblue1", lty = 2 )
lines(sm.spline(PFFx$SS, exp(predicted2$fit - 1.96 * predicted2$se)), col = "steelblue1", lty = 2 )

plot(CAREx$SS, exp(predicted3$fit), type="n", xlim=c(0,2.5), ylim=c(0,3))
lines(sm.spline(CAREx$SS, exp(predicted3$fit)), col = "firebrick1", lty = 1)
lines(sm.spline(CAREx$SS, exp(predicted3$fit + 1.96 * predicted3$se)), col = "coral1", lty = 2 )
lines(sm.spline(CAREx$SS, exp(predicted3$fit - 1.96 * predicted3$se)), col = "coral1", lty = 2 )
```

```{r, error=T}
#Then plot superimposed
plot(Simmx$SS, exp(predicted1$fit), type="n", xlim=c(0,2.5), ylim=c(0,3))
lines(sm.spline(Simmx$SS, exp(predicted1$fit)), col = "darkgoldenrod1", lty = 1 )
lines(sm.spline(Simmx$SS, exp(predicted1$fit + 1.96 * predicted1$se)), col = "lightgoldenrod1", lty = 2 )
lines(sm.spline(Simmx$SS, exp(predicted1$fit - 1.96 * predicted1$se)), col = "lightgoldenrod1", lty = 2 )

lines(sm.spline(PFFx$SS, exp(predicted2$fit)), col = "royalblue1", lty = 1)
lines(sm.spline(PFFx$SS, exp(predicted2$fit + 1.96 * predicted2$se)), col = "steelblue1", lty = 2 )
lines(sm.spline(PFFx$SS, exp(predicted2$fit - 1.96 * predicted2$se)), col = "steelblue1", lty = 2 )

lines(sm.spline(CAREx$SS, exp(predicted3$fit)), col = "firebrick1", lty = 1)
lines(sm.spline(CAREx$SS, exp(predicted3$fit + 1.96 * predicted3$se)), col = "coral1", lty = 2 )
lines(sm.spline(CAREx$SS, exp(predicted3$fit - 1.96 * predicted3$se)), col = "coral1", lty = 2 )
```


# Combining Cohorts
```{r}
colnames(Simm)==colnames(PFF)
colnames(PFF)==colnames(CARE)
```

## Creating Cohort Column
```{r}
Simm <- Simm %>% mutate(cohort="Simmons")
PFF <- PFF %>% mutate(cohort="PFF")
CARE <- CARE %>% mutate(cohort="CARE-PF")
```

## Fixing Columns in PFF
Fixing sex
```{r}
PFF <- PFF %>% mutate(sex1=if_else(sex=="Male", "M", "F"))
PFF <- PFF %>% dplyr::select(!sex)
PFF <- PFF %>% rename(c("sex"="sex1"))
PFF <- PFF %>% dplyr::select(ID, start, end, site, age_dx, sex, race, dich_Race, smokeHx, dx_IPF, disadv, dx_group, dx, dx_yr, dx_date, death_date, tx_date, DeathTxCensor_date, last_updated, fvc_date, dlco_date, fvc_pct, dlco_pct, status, deadORtx, time_DeathTxCensor, days_DeathTxCensor, event, start_5yr, end_5yr, SS, SS_dich, cohort)
```

Fixing ID - some of the IDs are shared between PFF and Simm
```{r}
#First want to check how many IDs are shared between Simm and PFF, Simm and CARE, and PFF and CARE
intersect(Simm$ID, PFF$ID)
intersect(Simm$ID, CARE$ID)
intersect(CARE$ID, PFF$ID)

PFF$ID <- paste0(2000, PFF$ID)
intersect(Simm$ID, PFF$ID)
```
No overlaps between Simm/CARE or PFF/CARE, but two with Simm/PFF, so add "2000" to front of PFF IDs


## Joining rows
```{r}
All <- rbind(Simm, PFF)
All <- rbind(All, CARE)
str(All)
```

## Correct Site and Cohort to Factors
```{r}
All$cohort <- as.factor(All$cohort)
All$site <- as.factor(All$site)
All$event <- as.factor(All$event)
```


# One-Stage Meta-Analysis
## Continuous Models
```{r}
coxPH_model1 <- coxph(Surv(start, end, event==1) ~ SS + dx_yr + site + cluster(cohort), data=All, id=ID)
summary(coxPH_model1)
```

```{r}
coxPH_model2 <- coxph(Surv(start, end, event==1) ~ SS + dx_yr + age_dx + sex + dich_Race + smokeHx + disadv + site + cluster(cohort), data=All, id=ID)
summary(coxPH_model2)
```

## Variable HR Spline Models
Base model
```{r}
#First need to make dataframe that only includes patients with a value for event
Allx <- All %>% filter(!is.na(SS) & !is.na(start) & !is.na(end) & !is.na(event) & !is.na(dx_yr) & !is.na(cohort) & !is.na(site) & SS!=0)

#Then make survival function
surv1 <- Surv(Allx$start, Allx$end, Allx$event==1)
fit1 <- coxph(surv1 ~ pspline(Allx$SS, df=3) + Allx$dx_yr + cluster(Allx$cohort) + Allx$site)
predicted <- predict(fit1, type="terms", se.fit=T, terms=1)
```

```{r}
#Then plot
plot(Allx$SS, exp(predicted$fit), type="n", xlim=c(0,2.5), ylim=c(0,2))
lines(sm.spline(Allx$SS, exp(predicted$fit)), col = "red" , lty = 1 )
lines(sm.spline(Allx$SS, exp(predicted$fit + 1.96 * predicted$se)), col = "orange" , lty = 2 )
lines(sm.spline(Allx$SS, exp(predicted$fit - 1.96 * predicted$se)), col = "orange" , lty = 2 )
```

Complete model
```{r}
#First need to make dataframe that only includes patients with time_DeathTxCensor
Allx <- All %>% filter(!is.na(SS) & !is.na(time_DeathTxCensor) & !is.na(dx_yr) & !is.na(time_DeathTxCensor) & !is.na(age_dx) & !is.na(sex) & !is.na(smokeHx) & !is.na(dich_Race) & !is.na(disadv) & !is.na(site) & SS!=0)

#Then make survival function
surv1 <- Surv(Allx$start, Allx$end, Allx$event==1)
fit1 <- coxph(surv1 ~ pspline(Allx$SS, df=3) + Allx$dx_yr + Allx$age_dx + Allx$sex + Allx$smokeHx + Allx$dich_Race + Allx$disadv + cluster(Allx$cohort) + Allx$site)
predicted <- predict(fit1, type="terms", se.fit=T, terms=1)
```

```{r}
#Then plot
plot(Allx$SS, exp(predicted$fit), type="n", xlim=c(0,2), ylim=c(0,2))
lines(sm.spline(Allx$SS, exp(predicted$fit)), col = "red" , lty = 1 )
lines(sm.spline(Allx$SS, exp(predicted$fit + 1.96 * predicted$se)), col = "orange" , lty = 2 )
lines(sm.spline(Allx$SS, exp(predicted$fit - 1.96 * predicted$se)), col = "orange" , lty = 2 )
```



