---
title: "Monthly Satellite SO4 Analysis with CARE-PF Data - Revision 1 Final Analysis"
author: "Gillian Goobie"
date: "07/31/2022"
output: 
  html_document: 
    toc: yes
    toc_float: yes
    number_sections: yes
    code_folding: hide
  bookdown::html_document2:
    toc: yes
    toc_float: yes
    number_sections: yes
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE, warning=FALSE, echo=F}
library(tidyverse)
library(readxl)
library(here)
library(writexl)
library(knitr)
library(lintr)
library(tidync)
library(maps)
library(ncmeta)
library(devtools)
library(RNetCDF)
library(ncdf4)
library(survival)
library(survminer)
library(lubridate)
library(cmprsk)
library(riskRegression)
library(prodlim)
library(RColorBrewer)
library(prodlim)
library(lme4)
library(performance)
library(psycho)
library(report)
library(rms)
library(splines)
library(Greg)
library(rstpm2)
library(pspline)
library(smoothHR)
library(raster)
```

# Importing Datasets
Here I am importing the file which contains monthly SO4 level estimates by satellite at nearest lon/lat to CAREPF patient residential addresses. These are linked to the patient ID.
```{r}
outfile1 <- here("MonthlySO4Data/CARE_fILD_2000_2017_SO4_2021_11_05.xlsx")
SO4 <- read_excel(outfile1)
```

Here I am importing the file which I used for my CIMD work that contains the baseline clinical and demographic data for 3389 patients who have CIMD and full records otherwise
```{r}
outfile2 <- here("MonthlySO4Data/CAREPF_fILDPts_BaselineData_2021_10_22.xlsx")
CARE <- read_excel(outfile2)
```

# Merging Datasets
I used a inner_join here so that the complete "SO4" dataframe only includes patients with fILD that have all baseline demographics and CIMD available as well as SO4 data.
```{r}
SO4 <- inner_join(CARE, SO4, by="ID")
```

# Extracting year of diagnosis and year of death/transplant/censoring
Since the SO4 Satellite data is only averages by month, I want to create a column for each patient where just the month/year of diagnosis and the month/year of death or lung transplant is extracted. Then I can use this to determine the average SO4 exposure in the 5yrs prior to diagnosis and in the years from diagnosis until death/tx.
```{r}
#Start with the year of diagnosis
SO4 <- SO4 %>% 
  mutate(dx_yrmo = format(as.Date(SO4$dx_date, format="%Y-%m-%d"),"%Y-%m"))
SO4 <- SO4 %>% 
  mutate(dx_yr = format(as.Date(SO4$dx_date, format="%Y-%m-%d"),"%Y"))
SO4$dx_yr <- as.numeric(SO4$dx_yr)

#Then the year of death or lung transplant
SO4 <- SO4 %>% 
  mutate(deathORtx_date = if_else(!is.na(tx_date), tx_date, death_date))
SO4 <- SO4 %>% 
  mutate(deathORtx_yrmo = format(as.Date(SO4$deathORtx_date, format="%Y-%m-%d"),"%Y-%m"))

#Then the year the records were last updated (i.e. year of censoring)
SO4 <- SO4 %>% 
  mutate(DeathTxCensor_date = if_else(!is.na(deathORtx_date), deathORtx_date, last_updated))
SO4 <- SO4 %>% 
  mutate(censor_yrmo = format(as.Date(SO4$DeathTxCensor_date, format="%Y-%m-%d"),"%Y-%m"))
```

# Pivoting to Long Format
First we need to convert the SO4 dataframe into the long rather than the wide format, which will allow us to use it more easily in R's tidyverse as this is "tidy" formatting.
```{r}
SO4 <- SO4 %>% 
  pivot_longer(cols=c(66:280), names_to="SO4_date", names_prefix="SO4_")
```

## Convert date SO4_date to same format as above
```{r}
SO4x <- SO4 
SO4x$SO4_date <- gsub("jan", "01-01-20", SO4x$SO4_date)
SO4x$SO4_date <- gsub("feb", "01-02-20", SO4x$SO4_date)
SO4x$SO4_date <- gsub("mar", "01-03-20", SO4x$SO4_date)
SO4x$SO4_date <- gsub("apr", "01-04-20", SO4x$SO4_date)
SO4x$SO4_date <- gsub("may", "01-05-20", SO4x$SO4_date)
SO4x$SO4_date <- gsub("jun", "01-06-20", SO4x$SO4_date)
SO4x$SO4_date <- gsub("jul", "01-07-20", SO4x$SO4_date)
SO4x$SO4_date <- gsub("aug", "01-08-20", SO4x$SO4_date)
SO4x$SO4_date <- gsub("sep", "01-09-20", SO4x$SO4_date)
SO4x$SO4_date <- gsub("oct", "01-10-20", SO4x$SO4_date)
SO4x$SO4_date <- gsub("nov", "01-11-20", SO4x$SO4_date)
SO4x$SO4_date <- gsub("dec", "01-12-20", SO4x$SO4_date)

SO4x$SO4_date <- format(as.Date(SO4x$SO4_date, format="%d-%m-%Y"),"%Y-%m-%d")
SO4x$SO4_date <- as.Date(SO4x$SO4_date)
SO4 <- SO4x
```


# Convert Date Columns to Date Format
Next I need to convert all date columns to proper format
```{r}
SO4 <- SO4 %>% 
  mutate_at(c("dob", "dx_date", "initial_visit_date", "formal_dx_date", "reg_update", "death_date", "tx_date", "pft_date", "last_updated", "deathORtx_date", "DeathTxCensor_date"), as.Date)
str(SO4)
```

# Creating Exposure Variables
Now I'm creating new variables where I am matching up SO4 averages per year to years of major events for patients (year of diagnosis, year of death/lung transplant/censoring, etc)
## 5yrs Pre-Death/Transplant/Censoring
Here I am calculating the average SO4 value in the 5yrs prior to death/lung transplant/censoring.
```{r}
SO4 <- SO4 %>% 
  group_by(ID) %>% 
  mutate(SO4_5yrPreCensor = mean(value[ymd(SO4_date)>=(ymd(DeathTxCensor_date) - years(5)) & 
                                        ymd(SO4_date)<=ymd(DeathTxCensor_date)]))
           
           
#mean(value[ymd(SO4_date) == ymd(DeathTxCensor_date) | ymd(SO4_date) == (ymd(DeathTxCensor_date) - years(1))]))
```

## 5 years Prior to Diagnosis
Here I am calculating the average SO4 value in the year of diagnosis and the 5 years leading up to diagnosis (so if all years available, it will be the average of 6 years).
```{r}
SO4 <- SO4 %>% 
  group_by(ID) %>% 
  mutate(SO4_5yrPreDx = mean(value[ymd(SO4_date) <= ymd(dx_date) & 
                                    ymd(SO4_date) >= (ymd(dx_date)-years(5))]))
```

## Remove Dataframes not in use
```{r}
rm(list=c("SO4x", "CARE"))
```

## Getting Rid of Duplicated IDs
Now that we have calculated the patient-specific SO4 exposures, I can get rid of all rows other than the first row for each patient.
```{r}
SO4 <-  SO4 %>% 
  distinct_at(vars(ID), .keep_all=T)
```
This takes us down to our  observations that we had before.

# Creating High vs Low Exposure Groups
Cutoffs are based on calculations performed where the median total SO4 mass across all three cohorts was averaged, then the median constituent component proportion of total SO4 across all three cohorts was averaged. Subsequently, the average proportion of each constituent across the three cohorts was multiplied by the average of the total SO4 mass medians across the three cohorts to get the approximate median total mass of each constituent across all three cohorts, which was used to define the low vs high exposure cut points.

## SO4 5yrs Pre-Censor High vs Low
```{r}
SO4$SO45yrCensor_dich <- cut(SO4$SO4_5yrPreCensor,
                      breaks=c(0, 1.04, 50),
                      labels=c("Low", "High"))
summary(SO4$SO45yrCensor_dich)
class(SO4$SO45yrCensor_dich)
```

## SO4 5yrs Pre-Diagnosis High vs Low
```{r}
SO4$SO45yr_dich <- cut(SO4$SO4_5yrPreDx,
                      breaks=c(0, 1.51, 50),
                      labels=c("Low", "High"))
summary(SO4$SO45yr_dich)
class(SO4$SO45yr_dich)
```


# Creating Other New Variables
## IPF vs Other Diagnosis
```{r}
SO4 <- SO4 %>% 
  mutate(dx_IPF=ifelse(dx=="IPF", "IPF", "not_IPF"))    
```

## Changing site to a factor variable
```{r}
SO4$site <- as.factor(SO4$site)
```

# Releveling Factors
Here I am releveling factors so that they are in an intuitive order for my later analyses.
```{r}
SO4$sex <- fct_relevel(SO4$sex, c("M","F"))
SO4$race <- fct_relevel(SO4$race, c("W","B","A","N","U"))
SO4$ethnicity <- fct_relevel(SO4$ethnicity, c("N","H","D","U"))
SO4$dich_Race <- fct_relevel(SO4$dich_Race, c("White","Non-White"))
SO4$smokeHx <- fct_relevel(SO4$smokeHx, c("Never","Former","Always"))

#For dx and dx_group, I just want IPF to be first and then the rest of the categories are alphabetical
SO4$dx <- fct_relevel(SO4$dx, c("IPF"))
SO4$dx_group <- fct_relevel(SO4$dx_group, c("IPF"))
SO4$dx_IPF <- fct_relevel(SO4$dx_IPF, c("IPF"))
str(SO4)
```

# Creating Datasets with only Metropolitan Patients
```{r}
SO4_metro <- SO4 %>% filter(metro=="metropolitan")
```
This results in a dataset of  patients

# Creating Datasets with only IPF Patients
```{r}
SO4_IPF <- SO4 %>% filter(dx=="IPF")
```
This results in a dataset of  pts

# Creating Datasets with only CTD Patients
```{r}
SO4_CTD <- SO4 %>% filter(dx_group=="CTD-ILD")
```
This results in a dataset of  pts

# Creating individual site subgroups
```{r}
SO4_toronto <- SO4 %>% filter(site==101)
SO4_ab <- SO4 %>% filter(site==102)
SO4_sph <- SO4 %>% filter(site==103)
SO4_vgh <- SO4 %>%  filter(site==104)
SO4_bc <- SO4 %>% filter(site==103 | site==104)
SO4_laval <- SO4 %>% filter(site==105)
SO4_hamilton <- SO4 %>% filter(site==106)
SO4_on <- SO4 %>% filter(site==101 | site==106)
SO4_mcgill <- SO4 %>% filter(site==107)
SO4_qc <- SO4 %>% filter(site==105 | site==107)
SO4_sk <- SO4 %>% filter(site==108)
```


# Export Excel File with SO4 Data Matched to Patients and Full Cohort Data
```{r}
write_xlsx(SO4, path="Final_CAREPF_fILD_SO4_BaselineData_2022_02_12.xlsx")
```

# Exploratory Data Analysis
This is a function that allows me to make tables which summarize the count and percentages of each level of factor variables
```{r}
n_prop_tbl <- function(x) {
  tbl <- table(x)
  res <- cbind(tbl, round(prop.table(tbl)*100,2))
  colnames(res) <-  c('Count', 'Percentage')
  res
}
```


## SO4 Breakdown
```{r}
summary(SO4$SO4_5yrPreCensor)
summary(SO4$SO4_5yrPreDx)
```


## Sex Breakdown
```{r}
n_prop_tbl(SO4$sex)
```

## Race and Ethnicity Breakdown 
```{r}
n_prop_tbl(SO4$race)

n_prop_tbl(SO4$ethnicity)
```

## Smoking History Breakdown
```{r}
n_prop_tbl(SO4$smokeHx)
summary(SO4$pkyrs)
```


## Diagnostic Group Breakdown
```{r}
n_prop_tbl(SO4$dx_group)
```


## Province and Site Breakdown
```{r}
n_prop_tbl(SO4$province)
n_prop_tbl(SO4$site)
```

### Metropolitan Breakdown
```{r}
n_prop_tbl(SO4$metro)
```
69% are metropolitan

## Age at Diagnosis Breakdown
```{r}
shapiro.test(SO4$age_dx)
#Shapiro tests tells us that age_dx is not normally distributed, so should report median, IQR
summary(SO4$age_dx)
```

## CIMD Breakdown
```{r}
summary(SO4$avg_s)
```


## Vital Status Breakdown
```{r}
n_prop_tbl(SO4$status)
```
Majority of the patients in CARE-PF are still alive.

## Baseline Lung Function Breakdown
```{r}
shapiro.test(SO4$fvc_pct)
summary(SO4$fvc_pct)
sd(SO4$fvc_pct, na.rm=T)
```

```{r}
shapiro.test(SO4$dlco_pct)
summary(SO4$dlco_pct)
sd(SO4$dlco_pct, na.rm=T)
```

## Time to Censoring Breakdown
```{r}
summary(SO4$time_DeathTxCensor)
sd(SO4$time_DeathTxCensor, na.rm=T)
```


## Breakdown of Characteristics by Low vs High
### SO4_5yrPreCensor Low vs High
```{r}
table(SO4$sex, SO4$SO45yrCensor_dich)
table(SO4$race, SO4$SO45yrCensor_dich)
table(SO4$dich_Race, SO4$SO45yrCensor_dich)
table(SO4$dx_group, SO4$SO45yrCensor_dich)
table(SO4$smokeHx, SO4$SO45yrCensor_dich)
table(SO4$metro, SO4$SO45yrCensor_dich)
table(SO4$province, SO4$SO45yrCensor_dich)
table(SO4$status, SO4$SO45yrCensor_dich)
group_by(SO4, SO45yrCensor_dich) %>%
  summarise(
    count = n(),
    mean = mean(age_dx, na.rm = TRUE),
    sd = sd(age_dx, na.rm = TRUE),
    median = median(age_dx, na.rm = TRUE),
    quantile = quantile(age_dx, probs=c(0.25), na.rm = TRUE),
    quantile2 = quantile(age_dx, probs=c(0.75), na.rm = TRUE)
  )
group_by(SO4, SO45yrCensor_dich) %>%
  summarise(
    count = n(),
    mean = mean(avg_s, na.rm = TRUE),
    sd = sd(avg_s, na.rm = TRUE),
    median = median(avg_s, na.rm = TRUE),
    quantile = quantile(avg_s, probs=c(0.25), na.rm = TRUE),
    quantile2 = quantile(avg_s, probs=c(0.75), na.rm = TRUE)
  )
group_by(SO4, SO45yrCensor_dich) %>%
  summarise(
    count = n(),
    mean = mean(fvc_pct, na.rm = TRUE),
    sd = sd(fvc_pct, na.rm = TRUE),
    median = median(fvc_pct, na.rm = TRUE),
    quantile = quantile(fvc_pct, probs=c(0.25), na.rm = TRUE),
    quantile2 = quantile(fvc_pct, probs=c(0.75), na.rm = TRUE)
  )
group_by(SO4, SO45yrCensor_dich) %>%
  summarise(
    count = n(),
    mean = mean(dlco_pct, na.rm = TRUE),
    sd = sd(dlco_pct, na.rm = TRUE),
    median = median(dlco_pct, na.rm = TRUE),
    quantile = quantile(dlco_pct, probs=c(0.25), na.rm = TRUE),
    quantile2 = quantile(dlco_pct, probs=c(0.75), na.rm = TRUE)
  )
group_by(SO4, SO45yrCensor_dich) %>%
  summarise(
    count = n(),
    mean = mean(time_DeathTxCensor, na.rm = TRUE),
    sd = sd(time_DeathTxCensor, na.rm = TRUE),
    median = median(time_DeathTxCensor, na.rm = TRUE),
    quantile = quantile(time_DeathTxCensor, probs=c(0.25), na.rm = TRUE),
    quantile2 = quantile(time_DeathTxCensor, probs=c(0.75), na.rm = TRUE)
  )
```

### SO4_5yrPreDx Low vs High
```{r}
table(SO4$sex, SO4$SO45yr_dich)
table(SO4$race, SO4$SO45yr_dich)
table(SO4$dich_Race, SO4$SO45yr_dich)
table(SO4$dx_group, SO4$SO45yr_dich)
table(SO4$smokeHx, SO4$SO45yr_dich)
table(SO4$metro, SO4$SO45yr_dich)
table(SO4$province, SO4$SO45yr_dich)
table(SO4$status, SO4$SO45yr_dich)
group_by(SO4, SO45yr_dich) %>%
  summarise(
    count = n(),
    mean = mean(age_dx, na.rm = TRUE),
    sd = sd(age_dx, na.rm = TRUE),
    median = median(age_dx, na.rm = TRUE),
    quantile = quantile(age_dx, probs=c(0.25), na.rm = TRUE),
    quantile2 = quantile(age_dx, probs=c(0.75), na.rm = TRUE)
  )
group_by(SO4, SO45yr_dich) %>%
  summarise(
    count = n(),
    mean = mean(avg_s, na.rm = TRUE),
    sd = sd(avg_s, na.rm = TRUE),
    median = median(avg_s, na.rm = TRUE),
    quantile = quantile(avg_s, probs=c(0.25), na.rm = TRUE),
    quantile2 = quantile(avg_s, probs=c(0.75), na.rm = TRUE)
  )
group_by(SO4, SO45yr_dich) %>%
  summarise(
    count = n(),
    mean = mean(fvc_pct, na.rm = TRUE),
    sd = sd(fvc_pct, na.rm = TRUE),
    median = median(fvc_pct, na.rm = TRUE),
    quantile = quantile(fvc_pct, probs=c(0.25), na.rm = TRUE),
    quantile2 = quantile(fvc_pct, probs=c(0.75), na.rm = TRUE)
  )
group_by(SO4, SO45yr_dich) %>%
  summarise(
    count = n(),
    mean = mean(dlco_pct, na.rm = TRUE),
    sd = sd(dlco_pct, na.rm = TRUE),
    median = median(dlco_pct, na.rm = TRUE),
    quantile = quantile(dlco_pct, probs=c(0.25), na.rm = TRUE),
    quantile2 = quantile(dlco_pct, probs=c(0.75), na.rm = TRUE)
  )
group_by(SO4, SO45yr_dich) %>%
  summarise(
    count = n(),
    mean = mean(time_DeathTxCensor, na.rm = TRUE),
    sd = sd(time_DeathTxCensor, na.rm = TRUE),
    median = median(time_DeathTxCensor, na.rm = TRUE),
    quantile = quantile(time_DeathTxCensor, probs=c(0.25), na.rm = TRUE),
    quantile2 = quantile(time_DeathTxCensor, probs=c(0.75), na.rm = TRUE)
  )
```


## Visual EDA
### Histograms
Histogram of SO4 Exposure 5yr Pre-Death/Transplant/Censoring
```{r}
(SO4 %>% ggplot(aes(x=SO4_5yrPreCensor))+
   geom_histogram(fill="blue", color="black")+
   labs(x="Average SO4 Level 5yr Prior to Death/Transplant/Censoring", y="Number of patients with fILD", title="Average SO4 5yr Prior to Death/Transplant/Censoring")+
   theme(plot.title = element_text(hjust = 0.5))+
   xlim(0,20))
```


Histogram of SO4 Exposure 5yr Pre-Dx
```{r}
(SO4 %>% ggplot(aes(x=SO4_5yrPreDx))+
   geom_histogram(fill="blue", color="black")+
   labs(x="Average SO4 Level 5yr Prior to Diagnosis", y="Number of patients with fILD", title="Average SO4 5yr Prior to Diagnosis")+
   theme(plot.title = element_text(hjust = 0.5))+
   xlim(0,20))
```


### Violin plots of SO4 vs other demographic variables
Violin plot wrapping boxplot to visualize low vs high SO45yrCensor_dich vs CIMD
```{r}
(SO4 %>% ggplot(aes(x=SO45yrCensor_dich, y=avg_s, fill=SO45yrCensor_dich))+
   geom_boxplot(width=0.2, color="black", alpha=1.0)+
   geom_violin(width=1.0, alpha=0.5)+
   labs(x="Low vs High SO4 in 5yrs Pre-Censoring", y="National CIMD", title="National CIMD by SO4 Exposure")+
   theme(plot.title = element_text(hjust = 0.5))+
   ylim(-2.5,2.5))
```

Violin plot wrapping boxplot to visualize sex vs SO4_5yrPreCensor
```{r}
(SO4 %>% ggplot(aes(x=sex, y=SO4_5yrPreCensor, fill=sex))+
   geom_boxplot(width=0.2, color="black", alpha=1.0)+
   geom_violin(width=1.0, alpha=0.5)+
   labs(x="Sex", y="Average SO4 5 Years Pre-Censor", title="SO4 Exposure Pre-Censor by Sex")+
   theme(plot.title = element_text(hjust = 0.5))+
   ylim(0,20))
```


Violin plot wrapping boxplot to visualize sex vs SO4_5yrPreDx
```{r}
(SO4 %>% ggplot(aes(x=sex, y=SO4_5yrPreDx, fill=sex))+
   geom_boxplot(width=0.2, color="black", alpha=1.0)+
   geom_violin(width=1.0, alpha=0.5)+
   labs(x="Sex", y="Average SO4 5 Years Pre-Diagnosis", title="SO4 Exposure Pre-Diagnosis by Sex")+
   theme(plot.title = element_text(hjust = 0.5))+
   ylim(0,20))
```

Violin plot wrapping boxplot to visualize dich_Race vs SO4_5yrPreCensor
```{r}
(SO4 %>% ggplot(aes(x=dich_Race, y=SO4_5yrPreCensor, fill=dich_Race))+
   geom_boxplot(width=0.2, color="black", alpha=1.0)+
   geom_violin(width=1.0, alpha=0.5)+
   labs(x="Race", y="Average SO4 5 Years Pre-Censoring", title="SO4 Exposure Pre-Censoring by Race")+
   theme(plot.title = element_text(hjust = 0.5))+
   ylim(0,20))
```


Violin plot wrapping boxplot to visualize race vs SO4_5yrPreCensor
```{r}
(SO4 %>% ggplot(aes(x=race, y=SO4_5yrPreCensor, fill=race))+
   geom_boxplot(width=0.2, color="black", alpha=1.0)+
   geom_violin(width=1.0, alpha=0.5)+
   labs(x="Race", y="Average SO4 5 Years Pre-Censoring", title="SO4 Exposure Pre-Censoring by Race")+
   theme(plot.title = element_text(hjust = 0.5))+
   ylim(0,20))
```

Violin plot wrapping boxplot to visualize dich_Race vs SO4_5yrPreDx
```{r}
(SO4 %>% ggplot(aes(x=dich_Race, y=SO4_5yrPreDx, fill=dich_Race))+
   geom_boxplot(width=0.2, color="black", alpha=1.0)+
   geom_violin(width=1.0, alpha=0.5)+
   labs(x="Race", y="Average SO4 5 Years Pre-Diagnosis", title="SO4 Exposure Pre-Diagnosis by Race")+
   theme(plot.title = element_text(hjust = 0.5))+
   ylim(0,20))
```

Violin plot wrapping boxplot to visualize race vs SO4_5yrPreDx
```{r}
(SO4 %>% ggplot(aes(x=race, y=SO4_5yrPreDx, fill=race))+
   geom_boxplot(width=0.2, color="black", alpha=1.0)+
   geom_violin(width=1.0, alpha=0.5)+
   labs(x="Race", y="Average SO4 5 Years Pre-Diagnosis", title="SO4 Exposure Pre-Diagnosis by Race")+
   theme(plot.title = element_text(hjust = 0.5))+
   ylim(0,20))
```

Violin plot wrapping boxplot to visualize metro vs SO4_5yrPreCensor
```{r}
(SO4 %>% ggplot(aes(x=metro, y=SO4_5yrPreCensor, fill=metro))+
   geom_boxplot(width=0.2, color="black", alpha=1.0)+
   geom_violin(width=1.0, alpha=0.5)+
   labs(x="Residential Location", y="Average SO4 in 5yrs Pre-Censoring", title="SO4 Exposure in 5yrs Pre-Censoring by Residential Location")+
   theme(plot.title = element_text(hjust = 0.5))+
   ylim(0,20))
```


Violin plot wrapping boxplot to visualize metro vs SO4_5yrPreDx
```{r}
(SO4 %>% ggplot(aes(x=metro, y=SO4_5yrPreDx, fill=metro))+
   geom_boxplot(width=0.2, color="black", alpha=1.0)+
   geom_violin(width=1.0, alpha=0.5)+
   labs(x="Residential Location", y="Average SO4 in 5yrs Pre-Diagnosis", title="SO4 Exposure in 5yrs Pre-Diagnosis by Residential Location")+
   theme(plot.title = element_text(hjust = 0.5))+
   ylim(0,20))
```

### Scatterplots of SO4 over time
Scatterplot showing average SO4 5yr pre-death/transplant/censoring
```{r}
(SO4 %>% ggplot(aes(x=dx_date, y=SO4_5yrPreCensor))+
   geom_point()+
   labs(x="Date of Diagnosis", y="SO4 Exposure 5yr Prior Death/Transplant/Censoring", title="SO4 5yr Prior to Death/Transplant/Censoring vs Date of Diagnosis")+
   geom_smooth(method="lm", se = FALSE) +
   theme(plot.title = element_text(hjust = 0.5))+
   ylim(0,20))
```


```{r}
(SO4 %>% ggplot(aes(x=dx_date, y=SO4_5yrPreCensor, color=site))+
   geom_point(alpha=0.5)+
   labs(x="Date of Diagnosis", y="SO4 Exposure 5yr Prior Death/Transplant/Censoring", title="SO4 5yr Prior to Death/Transplant/Censoring vs Date of Diagnosis")+
   geom_smooth(method="lm", se = FALSE) +
   theme(plot.title = element_text(hjust = 0.5))+
   ylim(0,20))
```

Scatterplot showing average SO4 during 5yrs pre-diagnosis level by year of diagnosis
```{r}
(SO4 %>% ggplot(aes(x=dx_date, y=SO4_5yrPreDx))+
   geom_point()+
   labs(x="Date of Diagnosis", y="SO4 Exposure During 5yrs Pre-Diagnosis", title="Pre-Diagnosis SO4 vs Date of Diagnosis")+
   geom_smooth(method="lm", se = FALSE) +
   theme(plot.title = element_text(hjust = 0.5))+
   ylim(0,20)+
   scale_x_date(limits = as.Date(c("2000-01-01","2020-01-01"))))
```


Scatterplot showing average SO4 during 5yrs pre-diagnosis level by year of diagnosis
```{r}
(SO4 %>% ggplot(aes(x=dx_date, y=SO4_5yrPreDx, color=site))+
   geom_point(alpha=0.5)+
   labs(x="Date of Diagnosis", y="SO4 Exposure During 5yrs Pre-Diagnosis", title="Pre-Diagnosis SO4 vs Date of Diagnosis")+
   geom_smooth(method="lm", se = FALSE) +
   theme(plot.title = element_text(hjust = 0.5))+
   ylim(0,20)+
   scale_x_date(limits = as.Date(c("2000-01-01","2020-01-01"))))
```


# Cox PH Models for Association between SO4 and Death or Lung Transplant
## SO4 5yrs Pre-Censoring Models
### Continuous SO4 5yrs pre-censoring until Death/Transplant/Censoring
Start with the simplest model just looking at SO4 in the year of and 5yrs prior to death/transplant/censoring
```{r}
coxPH_model1 <- coxph(Surv(time_DeathTxCensor, deadORtx==1) ~ SO4_5yrPreCensor, data=SO4, id=ID)
summary(coxPH_model1)
```


Complete model
```{r}
coxPH_model2 <- coxph(Surv(time_DeathTxCensor, deadORtx==1) ~ SO4_5yrPreCensor + sex + age_dx + smokeHx + dich_Race + avg_s + site, data=SO4, id=ID)
summary(coxPH_model2)
```


Complete model + dx_yr
```{r}
coxPH_model3 <- coxph(Surv(time_DeathTxCensor, deadORtx==1) ~ SO4_5yrPreCensor + sex + age_dx + smokeHx + dich_Race + avg_s + site + dx_yr, data=SO4, id=ID)
summary(coxPH_model3)
```



### Smoothed HR Function for 5yr Pre-Censor Continuous
Base model
```{r}
coxPH_model1_tvc <- cox.tvc(coxPH_model1, "SO4_5yrPreCensor")
print(coxPH_model1_tvc)
plot(coxPH_model1_tvc)
```


Complete model (all covariates except lung function)
```{r}
coxPH_model2_tvc <- cox.tvc(coxPH_model2, "SO4_5yrPreCensor")
print(coxPH_model2_tvc)
plot(coxPH_model2_tvc)
```


```{r}
coxPH_model3_tvc <- cox.tvc(coxPH_model3, "SO4_5yrPreCensor")
print(coxPH_model3_tvc)
plot(coxPH_model3_tvc)
```


### Variable HR for SO4 risk in 5yr Pre-Cenosr
Want a plot with x-axis of SO4 level and y-axis of HR

Unadjusted model
```{r}
#First need to make dataframe that only includes patients with time_DeathTxCensor
SO4x <- SO4 %>% filter(!is.na(SO4_5yrPreCensor) & !is.na(time_DeathTxCensor))

#Then make survival function
surv1 <- Surv(SO4x$time_DeathTxCensor, SO4x$deadORtx==1)
fit1 <- coxph(surv1 ~ pspline(SO4x$SO4_5yrPreCensor, df=3))
predicted <- predict(fit1, type="terms", se.fit=T, terms=1)

#Then plot
plot(SO4x$SO4_5yrPreCensor, exp(predicted$fit), type="n", xlim=c(0,6), ylim=c(0,10))
lines(sm.spline(SO4x$SO4_5yrPreCensor, exp(predicted$fit)), col = "red" , lty = 1 )
lines(sm.spline(SO4x$SO4_5yrPreCensor, exp(predicted$fit + 1.96 * predicted$se)), col = "orange" , lty = 2 )
lines(sm.spline(SO4x$SO4_5yrPreCensor, exp(predicted$fit - 1.96 * predicted$se)), col = "orange" , lty = 2 )
```

Complete model
```{r}
#First need to make dataframe that only includes patients with time_DeathTxCensor
SO4x <- SO4 %>% filter(!is.na(SO4_5yrPreCensor) & !is.na(time_DeathTxCensor) & !is.na(age_dx) & !is.na(sex) & !is.na(smokeHx) & !is.na(dich_Race) & !is.na(avg_s) & !is.na(site))

#Then survival function
surv1 <- Surv(SO4x$time_DeathTxCensor, SO4x$deadORtx==1)
fit1 <- coxph(surv1 ~ pspline(SO4x$SO4_5yrPreCensor, df=3) + SO4x$age_dx + SO4x$sex + SO4x$smokeHx + SO4x$dich_Race + SO4x$avg_s + SO4x$site)
predicted <- predict(fit1, type="terms", se.fit=T, terms=1)

#Then plot
plot(SO4x$SO4_5yrPreCensor, exp(predicted$fit), type="n", xlim=c(0,6), ylim=c(0,10))
lines(sm.spline(SO4x$SO4_5yrPreCensor, exp(predicted$fit)), col = "red" , lty = 1 )
lines(sm.spline(SO4x$SO4_5yrPreCensor, exp(predicted$fit + 1.96 * predicted$se)), col = "orange" , lty = 2 )
lines(sm.spline(SO4x$SO4_5yrPreCensor, exp(predicted$fit - 1.96 * predicted$se)), col = "orange" , lty = 2 )
```

Complete model + dx_yr
```{r}
#First need to make dataframe that only includes patients with time_DeathTxCensor
SO4x <- SO4 %>% filter(!is.na(SO4_5yrPreCensor) & !is.na(time_DeathTxCensor) & !is.na(age_dx) & !is.na(sex) & !is.na(smokeHx) & !is.na(dich_Race) & !is.na(avg_s) & !is.na(site) & !is.na(dx_yr))

#Then survival function
surv1 <- Surv(SO4x$time_DeathTxCensor, SO4x$deadORtx==1)
fit1 <- coxph(surv1 ~ pspline(SO4x$SO4_5yrPreCensor, df=4) + SO4x$age_dx + SO4x$sex + SO4x$smokeHx + SO4x$dich_Race + SO4x$avg_s + SO4x$site)
predicted <- predict(fit1, type="terms", se.fit=T, terms=1)

#Then plot
plot(SO4x$SO4_5yrPreCensor, exp(predicted$fit), type="n", xlim=c(0,6), ylim=c(0,10))
lines(sm.spline(SO4x$SO4_5yrPreCensor, exp(predicted$fit)), col = "red" , lty = 1 )
lines(sm.spline(SO4x$SO4_5yrPreCensor, exp(predicted$fit + 1.96 * predicted$se)), col = "orange" , lty = 2 )
lines(sm.spline(SO4x$SO4_5yrPreCensor, exp(predicted$fit - 1.96 * predicted$se)), col = "orange" , lty = 2 )
```


### SO4 Low vs High 5yrs Pre-Censoring
Base Model:
```{r}
coxPH_model7 <- coxph(Surv(time_DeathTxCensor, deadORtx==1) ~ SO45yrCensor_dich, data=SO4, id=ID)
summary(coxPH_model7)
```

Complete model
```{r}
coxPH_model8 <- coxph(Surv(time_DeathTxCensor, deadORtx==1) ~ SO45yrCensor_dich + sex + age_dx + smokeHx + dich_Race + avg_s + site, data=SO4, id=ID)
summary(coxPH_model8)
```

Complete model + dx_yr
```{r}
coxPH_model9 <- coxph(Surv(time_DeathTxCensor, deadORtx==1) ~ SO45yrCensor_dich + sex + age_dx + smokeHx + dich_Race + avg_s + site + dx_yr, data=SO4, id=ID)
summary(coxPH_model9)
```


Now I will make a Kaplan-Meier survival curve, discretized by SO4 during disease deciles:
```{r, fig.height=5, fig.width=7}
surv1 <- survfit(Surv(time_DeathTxCensor, deadORtx==1) ~ SO45yrCensor_dich, data=SO4)
(ggsurvplot(surv1, 
            palette=c("dodgerblue4","firebrick4"),
            censor=TRUE,
            pval=TRUE,
            xlim=c(0,10),
            legend="right",
           data=SO4,
           legend.title="SO4 Exposure",
           legend.labs=c("Low", "High"),
           risk.table=T,
           title="Time to Death or Transplant by SO4 5yrs Pre-Censoring",
           xlab="Time (years)",
           ylab="Survival Probability (Death or Lung Transplant)",
           ggtheme = theme_grey()
           ))
```


Now we want to check the proportionality assumption by plotting the scaled Schoenfield residuals against time for each covariate in our final model.
```{r}
residtest <- cox.zph(coxPH_model8)
residtest
ggcoxzph(residtest)
```


Now I will present a forest plot with the model
```{r}
ggforest(coxPH_model8)
```



# Association between SO4 and Baseline FVC
## Visual EDA 5yrs pre-Dx
Scatterplot of continuous SO4 5yrs pre-diagnosis vs FVC, with sex represented by the color of points
```{r}
(SO4 %>% ggplot(aes(x=SO4_5yrPreDx, y=fvc_pct, color=sex))+
   geom_point()+
   labs(x="SO4 5yrs Pre-Diagnosis", y="Baseline FVC % Predicted", title="Baseline FVC % Predicted by SO4 5yrs Pre-Diagnosis")+
   geom_smooth(method="lm", se = FALSE) +
   theme(plot.title = element_text(hjust = 0.5))+
   xlim(0,7))
```


Violin plot wrapping boxplot to visualize SO4 5yrs pre-dx low vs high vs FVC
```{r}
SO4b <- SO4 %>% filter(!is.na(SO45yr_dich))
(SO4b %>% ggplot(aes(x=SO45yr_dich, y=fvc_pct, fill=SO45yr_dich))+
   geom_boxplot(width=0.2, color="black", alpha=1.0)+
   geom_violin(width=1.0, alpha=0.5)+
   labs(x="SO4 5yrs Pre-Diagnosis Low vs High", y="Baseline FVC % Predicted", title="Baseline FVC % Predicted by SO4 5yrs Pre-Diagnosis Low vs High")+
   theme_light()+
   theme(legend.position = "none", plot.title = element_text(hjust = 0.5))+
   scale_fill_brewer(type="seq", palette="YlOrRd"))
```



## SO4 5yrs Pre-Diagnosis and FVC Models
Looking at continuous SO4_5yrPreDx vs baseline FVC

Simplest model just looking at SO4 in 5yrs pre-diagnosis
```{r}
FVC_model1 <- lm(fvc_pct ~ SO4_5yrPreDx, data=SO4)
summary(FVC_model1)
confint(FVC_model1)
```

Complete model
```{r}
FVC_model2 <- lm(fvc_pct ~ SO4_5yrPreDx + sex + age_dx + smokeHx + dich_Race + avg_s + site, data=SO4)
summary(FVC_model2)
confint(FVC_model2)
```

Complete model + dx_yr
```{r}
FVC_model3 <- lm(fvc_pct ~ SO4_5yrPreDx + sex + age_dx + smokeHx + dich_Race + avg_s + site + dx_yr, data=SO4)
summary(FVC_model3)
confint(FVC_model3)
```


### SO4 5yrs Pre-Diagnosis Low vs High and Baseline FVC
Looking at SO45yr_dich vs baseline FVC

Simplest model just looking at SO4 in 5yrs pre-diagnosis
```{r}
FVC_model1 <- lm(fvc_pct ~ SO45yr_dich, data=SO4)
summary(FVC_model1)
confint(FVC_model1)
```

Complete model
```{r}
FVC_model2 <- lm(fvc_pct ~ SO45yr_dich + sex + age_dx + smokeHx + dich_Race + avg_s + site, data=SO4)
summary(FVC_model2)
confint(FVC_model2)
```

Complete model + dx_yr
```{r}
FVC_model3 <- lm(fvc_pct ~ SO45yr_dich + sex + age_dx + smokeHx + dich_Race + avg_s + site + dx_yr, data=SO4)
summary(FVC_model3)
confint(FVC_model3)
```

# Association between SO4 and Baseline DLCO
## Visual EDA 5yrs pre-Dx
Scatterplot of continuous SO4 5yrs pre-diagnosis vs DLCO, with sex represented by the color of points
```{r}
(SO4 %>% ggplot(aes(x=SO4_5yrPreDx, y=dlco_pct, color=sex))+
   geom_point()+
   labs(x="SO4 5yrs Pre-Diagnosis", y="Baseline DLCO % Predicted", title="Baseline DLCO % Predicted by SO4 5yrs Pre-Diagnosis")+
   geom_smooth(method="lm", se = FALSE) +
   theme(plot.title = element_text(hjust = 0.5))+
   xlim(0,12))
```


Violin plot wrapping boxplot to visualize SO4 5yrs pre-dx low vs high vs dlco
```{r}
(SO4b %>% ggplot(aes(x=SO45yr_dich, y=dlco_pct, fill=SO45yr_dich))+
   geom_boxplot(width=0.2, color="black", alpha=1.0)+
   geom_violin(width=1.0, alpha=0.5)+
   labs(x="SO4 5yrs Pre-Diagnosis Low vs High", y="Baseline DLCO % Predicted", title="Baseline DLCO % Predicted by SO4 5yrs Pre-Diagnosis Low vs High")+
   theme_light()+
   theme(legend.position = "none", plot.title = element_text(hjust = 0.5))+
   scale_fill_brewer(type="seq", palette="YlOrRd"))
```



## SO4 5yrs Pre-Diagnosis and DLCO Models
### SO4 5yrs Pre-Diagnosis and Baseline DLCO
Looking at continuous SO4_5yrPreDx vs baseline DLCO

Simplest model just looking at SO4 in 5yrs pre-diagnosis
```{r}
DLCO_model1 <- lm(dlco_pct ~ SO4_5yrPreDx, data=SO4)
summary(DLCO_model1)
confint(DLCO_model1)
```

Complete model
```{r}
DLCO_model2 <- lm(dlco_pct ~ SO4_5yrPreDx + sex + age_dx + smokeHx + dich_Race + avg_s + site, data=SO4)
summary(DLCO_model2)
confint(DLCO_model2)
```

Complete model + dx_yr
```{r}
DLCO_model3 <- lm(dlco_pct ~ SO4_5yrPreDx + sex + age_dx + smokeHx + dich_Race + avg_s + site + dx_yr, data=SO4)
summary(DLCO_model3)
confint(DLCO_model3)
```


### SO4 5yrs Pre-Diagnosis Low vs High and Baseline DLCO
Looking at SO45yr_dich vs baseline DLCO

Simplest model just looking at SO4 in 5yrs pre-diagnosis
```{r}
DLCO_model1 <- lm(dlco_pct ~ SO45yr_dich, data=SO4)
summary(DLCO_model1)
confint(DLCO_model1)
```

Complete model
```{r}
DLCO_model2 <- lm(dlco_pct ~ SO45yr_dich + sex + age_dx + smokeHx + dich_Race + avg_s + site, data=SO4)
summary(DLCO_model2)
confint(DLCO_model2)
```


Complete model + dx_yr
```{r}
DLCO_model3 <- lm(dlco_pct ~ SO45yr_dich + sex + age_dx + smokeHx + dich_Race + avg_s + site + dx_yr, data=SO4)
summary(DLCO_model3)
confint(DLCO_model3)
```


# IPF-Only - Cox PH Models for Association between SO4 and Death or Lung Transplant
## IPF-Only - SO4 5yrs Pre-Censoring Models
### IPF-Only - Continuous SO4 5yrs pre-censoring until Death/Transplant/Censoring
Start with the simplest model just looking at SO4 in the year of and 5yrs prior to death/transplant/censoring
```{r}
coxPH_model1 <- coxph(Surv(time_DeathTxCensor, deadORtx==1) ~ SO4_5yrPreCensor, data=SO4_IPF, id=ID)
summary(coxPH_model1)
```


Complete model
```{r}
coxPH_model2 <- coxph(Surv(time_DeathTxCensor, deadORtx==1) ~ SO4_5yrPreCensor + sex + age_dx + smokeHx + dich_Race + avg_s + site, data=SO4_IPF, id=ID)
summary(coxPH_model2)
```


Complete model + dx_yr
```{r}
coxPH_model3 <- coxph(Surv(time_DeathTxCensor, deadORtx==1) ~ SO4_5yrPreCensor + sex + age_dx + smokeHx + dich_Race + avg_s + site + dx_yr, data=SO4_IPF, id=ID)
summary(coxPH_model3)
```



### IPF-Only - Smoothed HR Function for 5yr Pre-Censor Continuous
Base model
```{r}
coxPH_model1_tvc <- cox.tvc(coxPH_model1, "SO4_5yrPreCensor")
print(coxPH_model1_tvc)
plot(coxPH_model1_tvc)
```


Complete model (all covariates except lung function)
```{r}
coxPH_model2_tvc <- cox.tvc(coxPH_model2, "SO4_5yrPreCensor")
print(coxPH_model2_tvc)
plot(coxPH_model2_tvc)
```


```{r}
#coxPH_model3_tvc <- cox.tvc(coxPH_model3, "SO4_5yrPreCensor")
#print(coxPH_model3_tvc)
#plot(coxPH_model3_tvc)
```
Error with the bove code "Error in optim(par = c(SO4_5yrPreCensor = 4.98043221986785, sexF = -0.272936265634474, : 
initial value in 'vmmin' is not finite"

### IPF-Only - Variable HR for SO4 risk in 5yr Pre-Cenosr
Want a plot with x-axis of SO4 level and y-axis of HR

Unadjusted model
```{r}
#First need to make dataframe that only includes patients with time_DeathTxCensor
SO4x <- SO4_IPF %>% filter(!is.na(SO4_5yrPreCensor) & !is.na(time_DeathTxCensor))

#Then make survival function
surv1 <- Surv(SO4x$time_DeathTxCensor, SO4x$deadORtx==1)
fit1 <- coxph(surv1 ~ pspline(SO4x$SO4_5yrPreCensor, df=3))
predicted <- predict(fit1, type="terms", se.fit=T, terms=1)

#Then plot
plot(SO4x$SO4_5yrPreCensor, exp(predicted$fit), type="n", xlim=c(0,6), ylim=c(0,10))
lines(sm.spline(SO4x$SO4_5yrPreCensor, exp(predicted$fit)), col = "red" , lty = 1 )
lines(sm.spline(SO4x$SO4_5yrPreCensor, exp(predicted$fit + 1.96 * predicted$se)), col = "orange" , lty = 2 )
lines(sm.spline(SO4x$SO4_5yrPreCensor, exp(predicted$fit - 1.96 * predicted$se)), col = "orange" , lty = 2 )
```

Complete model
```{r}
#First need to make dataframe that only includes patients with time_DeathTxCensor
SO4x <- SO4_IPF %>% filter(!is.na(SO4_5yrPreCensor) & !is.na(time_DeathTxCensor) & !is.na(age_dx) & !is.na(sex) & !is.na(smokeHx) & !is.na(dich_Race) & !is.na(avg_s) & !is.na(site))

#Then survival function
surv1 <- Surv(SO4x$time_DeathTxCensor, SO4x$deadORtx==1)
fit1 <- coxph(surv1 ~ pspline(SO4x$SO4_5yrPreCensor, df=3) + SO4x$age_dx + SO4x$sex + SO4x$smokeHx + SO4x$dich_Race + SO4x$avg_s + SO4x$site)
predicted <- predict(fit1, type="terms", se.fit=T, terms=1)

#Then plot
plot(SO4x$SO4_5yrPreCensor, exp(predicted$fit), type="n", xlim=c(0,6), ylim=c(0,10))
lines(sm.spline(SO4x$SO4_5yrPreCensor, exp(predicted$fit)), col = "red" , lty = 1 )
lines(sm.spline(SO4x$SO4_5yrPreCensor, exp(predicted$fit + 1.96 * predicted$se)), col = "orange" , lty = 2 )
lines(sm.spline(SO4x$SO4_5yrPreCensor, exp(predicted$fit - 1.96 * predicted$se)), col = "orange" , lty = 2 )
```

Complete Model + dx_yr
```{r}
#First need to make dataframe that only includes patients with time_DeathTxCensor
SO4x <- SO4_IPF %>% filter(!is.na(SO4_5yrPreCensor) & !is.na(time_DeathTxCensor) & !is.na(age_dx) & !is.na(sex) & !is.na(smokeHx) & !is.na(dich_Race) & !is.na(avg_s) & !is.na(site) & !is.na(dx_yr))

#Then survival function
surv1 <- Surv(SO4x$time_DeathTxCensor, SO4x$deadORtx==1)
fit1 <- coxph(surv1 ~ pspline(SO4x$SO4_5yrPreCensor, df=4) + SO4x$age_dx + SO4x$sex + SO4x$smokeHx + SO4x$dich_Race + SO4x$avg_s + SO4x$site + SO4x$dx_yr)
predicted <- predict(fit1, type="terms", se.fit=T, terms=1)

#Then plot
plot(SO4x$SO4_5yrPreCensor, exp(predicted$fit), type="n", xlim=c(0,6), ylim=c(0,10))
lines(sm.spline(SO4x$SO4_5yrPreCensor, exp(predicted$fit)), col = "red" , lty = 1 )
lines(sm.spline(SO4x$SO4_5yrPreCensor, exp(predicted$fit + 1.96 * predicted$se)), col = "orange" , lty = 2 )
lines(sm.spline(SO4x$SO4_5yrPreCensor, exp(predicted$fit - 1.96 * predicted$se)), col = "orange" , lty = 2 )
```


### IPF-Only - SO4 Low vs High 5yrs Pre-Censoring
Base Model:
```{r}
coxPH_model7 <- coxph(Surv(time_DeathTxCensor, deadORtx==1) ~ SO45yrCensor_dich, data=SO4_IPF, id=ID)
summary(coxPH_model7)
```


Complete model
```{r}
coxPH_model8 <- coxph(Surv(time_DeathTxCensor, deadORtx==1) ~ SO45yrCensor_dich + sex + age_dx + smokeHx + dich_Race + avg_s + site, data=SO4_IPF, id=ID)
summary(coxPH_model8)
```


Complete model + dx_yr
```{r}
coxPH_model9 <- coxph(Surv(time_DeathTxCensor, deadORtx==1) ~ SO45yrCensor_dich + sex + age_dx + smokeHx + dich_Race + avg_s + site + dx_yr, data=SO4_IPF, id=ID)
summary(coxPH_model9)
```


Now I will make a Kaplan-Meier survival curve, discretized by SO4 during disease deciles:
```{r, fig.height=5, fig.width=7}
surv1 <- survfit(Surv(time_DeathTxCensor, deadORtx==1) ~ SO45yrCensor_dich, data=SO4_IPF)
(ggsurvplot(surv1, 
            palette=c("dodgerblue4","firebrick4"),
            censor=TRUE,
            pval=TRUE,
            xlim=c(0,10),
            legend="right",
           data=SO4_IPF,
           legend.title="SO4 Exposure",
           legend.labs=c("Low", "High"),
           risk.table=T,
           title="Time to Death or Transplant by SO4 5yrs Pre-Censoring",
           xlab="Time (years)",
           ylab="Survival Probability (Death or Lung Transplant)",
           ggtheme = theme_grey()
           ))
```


Now we want to check the proportionality assumption by plotting the scaled Schoenfield residuals against time for each covariate in our final model.
```{r}
residtest <- cox.zph(coxPH_model9)
residtest
ggcoxzph(residtest)
```


Now I will present a forest plot with the model
```{r}
ggforest(coxPH_model9)
```



# IPF-Only - Association between SO4 and Baseline FVC
## IPF-Only - Visual EDA 5yrs pre-Dx
Scatterplot of continuous SO4 5yrs pre-diagnosis vs FVC, with sex represented by the color of points
```{r}
(SO4 %>% ggplot(aes(x=SO4_5yrPreDx, y=fvc_pct, color=sex))+
   geom_point()+
   labs(x="SO4 5yrs Pre-Diagnosis", y="Baseline FVC % Predicted", title="Baseline FVC % Predicted by SO4 5yrs Pre-Diagnosis")+
   geom_smooth(method="lm", se = FALSE) +
   theme(plot.title = element_text(hjust = 0.5))+
   xlim(0,7))
```


Violin plot wrapping boxplot to visualize SO4 5yrs pre-dx low vs high vs FVC
```{r}
SO4b <- SO4 %>% filter(!is.na(SO45yr_dich))
(SO4b %>% ggplot(aes(x=SO45yr_dich, y=fvc_pct, fill=SO45yr_dich))+
   geom_boxplot(width=0.2, color="black", alpha=1.0)+
   geom_violin(width=1.0, alpha=0.5)+
   labs(x="SO4 5yrs Pre-Diagnosis Low vs High", y="Baseline FVC % Predicted", title="Baseline FVC % Predicted by SO4 5yrs Pre-Diagnosis Low vs High")+
   theme_light()+
   theme(legend.position = "none", plot.title = element_text(hjust = 0.5))+
   scale_fill_brewer(type="seq", palette="YlOrRd"))
```



## IPF-Only - SO4 5yrs Pre-Diagnosis and FVC Models
Looking at continuous SO4_5yrPreDx vs baseline FVC

Simplest model just looking at SO4 in 5yrs pre-diagnosis
```{r}
FVC_model1 <- lm(fvc_pct ~ SO4_5yrPreDx, data=SO4_IPF)
summary(FVC_model1)
confint(FVC_model1)
```

Complete model
```{r}
FVC_model2 <- lm(fvc_pct ~ SO4_5yrPreDx + sex + age_dx + smokeHx + dich_Race + avg_s + site, data=SO4_IPF)
summary(FVC_model2)
confint(FVC_model2)
```

Complete model + dx_yr
```{r}
FVC_model3 <- lm(fvc_pct ~ SO4_5yrPreDx + sex + age_dx + smokeHx + dich_Race + avg_s + site + dx_yr, data=SO4_IPF)
summary(FVC_model3)
confint(FVC_model3)
```


### IPF-Only - SO4 5yrs Pre-Diagnosis Low vs High and Baseline FVC
Looking at SO45yr_dich vs baseline FVC

Simplest model just looking at SO4 in 5yrs pre-diagnosis
```{r}
FVC_model1 <- lm(fvc_pct ~ SO45yr_dich, data=SO4_IPF)
summary(FVC_model1)
confint(FVC_model1)
```

Complete model
```{r}
FVC_model2 <- lm(fvc_pct ~ SO45yr_dich + sex + age_dx + smokeHx + dich_Race + avg_s + site, data=SO4_IPF)
summary(FVC_model2)
confint(FVC_model2)
```

Complete model + dx_yr
```{r}
FVC_model3 <- lm(fvc_pct ~ SO45yr_dich + sex + age_dx + smokeHx + dich_Race + avg_s + site + dx_yr, data=SO4_IPF)
summary(FVC_model3)
confint(FVC_model3)
```

# IPF-Only - Association between SO4 and Baseline DLCO
## IPF-Only - Visual EDA 5yrs pre-Dx
Scatterplot of continuous SO4 5yrs pre-diagnosis vs DLCO, with sex represented by the color of points
```{r}
(SO4 %>% ggplot(aes(x=SO4_5yrPreDx, y=dlco_pct, color=sex))+
   geom_point()+
   labs(x="SO4 5yrs Pre-Diagnosis", y="Baseline DLCO % Predicted", title="Baseline DLCO % Predicted by SO4 5yrs Pre-Diagnosis")+
   geom_smooth(method="lm", se = FALSE) +
   theme(plot.title = element_text(hjust = 0.5))+
   xlim(0,12))
```


Violin plot wrapping boxplot to visualize SO4 5yrs pre-dx low vs high vs DLCO
```{r}
(SO4b %>% ggplot(aes(x=SO45yr_dich, y=dlco_pct, fill=SO45yr_dich))+
   geom_boxplot(width=0.2, color="black", alpha=1.0)+
   geom_violin(width=1.0, alpha=0.5)+
   labs(x="SO4 5yrs Pre-Diagnosis Low vs High", y="Baseline DLCO % Predicted", title="Baseline DLCO % Predicted by SO4 5yrs Pre-Diagnosis Low vs High")+
   theme_light()+
   theme(legend.position = "none", plot.title = element_text(hjust = 0.5))+
   scale_fill_brewer(type="seq", palette="YlOrRd"))
```



## IPF-Only - SO4 5yrs Pre-Diagnosis and DLCO Models
### IPF-Only - SO4 5yrs Pre-Diagnosis and Baseline DLCO
Looking at continuous SO4_5yrPreDx vs baseline DLCO

Simplest model just looking at SO4 in 5yrs pre-diagnosis
```{r}
DLCO_model1 <- lm(dlco_pct ~ SO4_5yrPreDx, data=SO4_IPF)
summary(DLCO_model1)
confint(DLCO_model1)
```

Complete model
```{r}
DLCO_model2 <- lm(dlco_pct ~ SO4_5yrPreDx + sex + age_dx + smokeHx + dich_Race + avg_s + site, data=SO4_IPF)
summary(DLCO_model2)
confint(DLCO_model2)
```

Complete model + dx_yr
```{r}
DLCO_model3 <- lm(dlco_pct ~ SO4_5yrPreDx + sex + age_dx + smokeHx + dich_Race + avg_s + site + dx_yr, data=SO4_IPF)
summary(DLCO_model3)
confint(DLCO_model3)
```


### IPF-Only - SO4 5yrs Pre-Diagnosis Low vs High and Baseline DLCO
Looking at SO45yr_dich vs baseline DLCO

Simplest model just looking at SO4 in 5yrs pre-diagnosis
```{r}
DLCO_model1 <- lm(dlco_pct ~ SO45yr_dich, data=SO4_IPF)
summary(DLCO_model1)
confint(DLCO_model1)
```

Complete model
```{r}
DLCO_model2 <- lm(dlco_pct ~ SO45yr_dich + sex + age_dx + smokeHx + dich_Race + avg_s + site, data=SO4_IPF)
summary(DLCO_model2)
confint(DLCO_model2)
```


Complete model + dx_yr
```{r}
DLCO_model3 <- lm(dlco_pct ~ SO45yr_dich + sex + age_dx + smokeHx + dich_Race + avg_s + site + dx_yr, data=SO4_IPF)
summary(DLCO_model3)
confint(DLCO_model3)
```



# Pre-2015 vs Post-2015 Models
```{r}
pre2015 <- SO4 %>% filter(dx_yr<=2015)

post2015 <- SO4 %>% filter(dx_yr>2015)
```

# Pre/Post-2015 Only - Cox PH Models for Association between SO4 and Death or Lung Transplant
## Pre/Post-2015 Only - SO4 5yrs Pre-Censoring Models
### Pre/Post-2015 Only - Continuous SO4 5yrs pre-censoring until Death/Transplant/Censoring
Start with the simplest model just looking at SO4 in the year of and 5yrs prior to death/transplant/censoring
```{r}
coxPH_model1a <- coxph(Surv(time_DeathTxCensor, deadORtx==1) ~ SO4_5yrPreCensor, data=pre2015, id=ID)
summary(coxPH_model1a)
```

```{r}
coxPH_model1b <- coxph(Surv(time_DeathTxCensor, deadORtx==1) ~ SO4_5yrPreCensor, data=post2015, id=ID)
summary(coxPH_model1b)
```

Complete model
```{r}
coxPH_model2a <- coxph(Surv(time_DeathTxCensor, deadORtx==1) ~ SO4_5yrPreCensor + sex + age_dx + smokeHx + dich_Race + avg_s + site, data=pre2015, id=ID)
summary(coxPH_model2a)
```

Complete model
```{r}
coxPH_model2b <- coxph(Surv(time_DeathTxCensor, deadORtx==1) ~ SO4_5yrPreCensor + sex + age_dx + smokeHx + dich_Race + avg_s + site, data=post2015, id=ID)
summary(coxPH_model2b)
```

### Pre/Post-2015 Only - Variable HR for SO4 risk in 5yr Pre-Cenosr
Want a plot with x-axis of SO4 level and y-axis of HR

Unadjusted model - Pre-2015
```{r}
#First need to make dataframe that only includes patients with time_DeathTxCensor
pre2015x <- pre2015 %>% filter(!is.na(SO4_5yrPreCensor) & !is.na(time_DeathTxCensor))

#Then make survival function
surv1 <- Surv(pre2015x$time_DeathTxCensor, pre2015x$deadORtx==1)
fit1 <- coxph(surv1 ~ pspline(pre2015x$SO4_5yrPreCensor, df=3))
predicted <- predict(fit1, type="terms", se.fit=T, terms=1)

#Then plot
plot(pre2015x$SO4_5yrPreCensor, exp(predicted$fit), type="n", xlim=c(0,6), ylim=c(0,10))
lines(sm.spline(pre2015x$SO4_5yrPreCensor, exp(predicted$fit)), col = "red" , lty = 1 )
lines(sm.spline(pre2015x$SO4_5yrPreCensor, exp(predicted$fit + 1.96 * predicted$se)), col = "orange" , lty = 2 )
lines(sm.spline(pre2015x$SO4_5yrPreCensor, exp(predicted$fit - 1.96 * predicted$se)), col = "orange" , lty = 2 )
```

Unadjusted model - Post-2015
```{r}
#First need to make dataframe that only includes patients with time_DeathTxCensor
post2015x <- post2015 %>% filter(!is.na(SO4_5yrPreCensor) & !is.na(time_DeathTxCensor))

#Then make survival function
surv1 <- Surv(post2015x$time_DeathTxCensor, post2015x$deadORtx==1)
fit1 <- coxph(surv1 ~ pspline(post2015x$SO4_5yrPreCensor, df=3))
predicted <- predict(fit1, type="terms", se.fit=T, terms=1)

#Then plot
plot(post2015x$SO4_5yrPreCensor, exp(predicted$fit), type="n", xlim=c(0,6), ylim=c(0,10))
lines(sm.spline(post2015x$SO4_5yrPreCensor, exp(predicted$fit)), col = "red" , lty = 1 )
lines(sm.spline(post2015x$SO4_5yrPreCensor, exp(predicted$fit + 1.96 * predicted$se)), col = "orange" , lty = 2 )
lines(sm.spline(post2015x$SO4_5yrPreCensor, exp(predicted$fit - 1.96 * predicted$se)), col = "orange" , lty = 2 )
```

Complete model Pre-2015
```{r}
#First need to make dataframe that only includes patients with time_DeathTxCensor
pre2015x <- pre2015 %>% filter(!is.na(SO4_5yrPreCensor) & !is.na(time_DeathTxCensor) & !is.na(age_dx) & !is.na(sex) & !is.na(smokeHx) & !is.na(dich_Race) & !is.na(avg_s) & !is.na(site))

#Then survival function
surv1 <- Surv(pre2015x$time_DeathTxCensor, pre2015x$deadORtx==1)
fit1 <- coxph(surv1 ~ pspline(pre2015x$SO4_5yrPreCensor, df=3) + pre2015x$age_dx + pre2015x$sex + pre2015x$smokeHx + pre2015x$dich_Race + pre2015x$avg_s + pre2015x$site)
predicted <- predict(fit1, type="terms", se.fit=T, terms=1)

#Then plot
plot(pre2015x$SO4_5yrPreCensor, exp(predicted$fit), type="n", xlim=c(0,6), ylim=c(0,10))
lines(sm.spline(pre2015x$SO4_5yrPreCensor, exp(predicted$fit)), col = "red" , lty = 1 )
lines(sm.spline(pre2015x$SO4_5yrPreCensor, exp(predicted$fit + 1.96 * predicted$se)), col = "orange" , lty = 2 )
lines(sm.spline(pre2015x$SO4_5yrPreCensor, exp(predicted$fit - 1.96 * predicted$se)), col = "orange" , lty = 2 )
```

Complete model Post-2015
```{r}
#First need to make dataframe that only includes patients with time_DeathTxCensor
post2015x <- post2015 %>% filter(!is.na(SO4_5yrPreCensor) & !is.na(time_DeathTxCensor) & !is.na(age_dx) & !is.na(sex) & !is.na(smokeHx) & !is.na(dich_Race) & !is.na(avg_s) & !is.na(site))

#Then survival function
surv1 <- Surv(post2015x$time_DeathTxCensor, post2015x$deadORtx==1)
fit1 <- coxph(surv1 ~ pspline(post2015x$SO4_5yrPreCensor, df=3) + post2015x$age_dx + post2015x$sex + post2015x$smokeHx + post2015x$dich_Race + post2015x$avg_s + post2015x$site)
predicted <- predict(fit1, type="terms", se.fit=T, terms=1)

#Then plot
plot(post2015x$SO4_5yrPreCensor, exp(predicted$fit), type="n", xlim=c(0,6), ylim=c(0,10))
lines(sm.spline(post2015x$SO4_5yrPreCensor, exp(predicted$fit)), col = "red" , lty = 1 )
lines(sm.spline(post2015x$SO4_5yrPreCensor, exp(predicted$fit + 1.96 * predicted$se)), col = "orange" , lty = 2 )
lines(sm.spline(post2015x$SO4_5yrPreCensor, exp(predicted$fit - 1.96 * predicted$se)), col = "orange" , lty = 2 )
```

### Pre/Post-2015 Only - SO4 Low vs High 5yrs Pre-Censoring
Base Model:
```{r}
coxPH_model7a <- coxph(Surv(time_DeathTxCensor, deadORtx==1) ~ SO45yrCensor_dich, data=pre2015, id=ID)
summary(coxPH_model7a)
```

Base Model:
```{r}
coxPH_model7b <- coxph(Surv(time_DeathTxCensor, deadORtx==1) ~ SO45yrCensor_dich, data=post2015, id=ID)
summary(coxPH_model7b)
```

Complete model
```{r}
coxPH_model8a <- coxph(Surv(time_DeathTxCensor, deadORtx==1) ~ SO45yrCensor_dich + sex + age_dx + smokeHx + dich_Race + avg_s + site, data=pre2015, id=ID)
summary(coxPH_model8a)
```

Complete model
```{r}
coxPH_model8b <- coxph(Surv(time_DeathTxCensor, deadORtx==1) ~ SO45yrCensor_dich + sex + age_dx + smokeHx + dich_Race + avg_s + site, data=post2015, id=ID)
summary(coxPH_model8b)
```


Now I will make a Kaplan-Meier survival curve, discretized by SO4 during disease deciles:
```{r, fig.height=5, fig.width=7}
surv1 <- survfit(Surv(time_DeathTxCensor, deadORtx==1) ~ SO45yrCensor_dich, data=pre2015)
(ggsurvplot(surv1, 
            palette=c("dodgerblue4","firebrick4"),
            censor=TRUE,
            pval=TRUE,
            xlim=c(0,10),
            legend="right",
           data=SO4_IPF,
           legend.title="SO4 Exposure",
           legend.labs=c("Low", "High"),
           risk.table=T,
           title="Time to Death or Transplant by SO4 5yrs Pre-Censoring",
           xlab="Time (years)",
           ylab="Survival Probability (Death or Lung Transplant)",
           ggtheme = theme_grey()
           ))
```

Now I will make a Kaplan-Meier survival curve, discretized by SO4 during disease deciles:
```{r, fig.height=5, fig.width=7}
surv1 <- survfit(Surv(time_DeathTxCensor, deadORtx==1) ~ SO45yrCensor_dich, data=post2015)
(ggsurvplot(surv1, 
            palette=c("dodgerblue4","firebrick4"),
            censor=TRUE,
            pval=TRUE,
            xlim=c(0,10),
            legend="right",
           data=SO4_IPF,
           legend.title="SO4 Exposure",
           legend.labs=c("Low", "High"),
           risk.table=T,
           title="Time to Death or Transplant by SO4 5yrs Pre-Censoring",
           xlab="Time (years)",
           ylab="Survival Probability (Death or Lung Transplant)",
           ggtheme = theme_grey()
           ))
```

Now we want to check the proportionality assumption by plotting the scaled Schoenfield residuals against time for each covariate in our final model.
```{r}
residtest <- cox.zph(coxPH_model8a)
residtest
ggcoxzph(residtest)
```

```{r}
residtest <- cox.zph(coxPH_model8b)
residtest
ggcoxzph(residtest)
```


Now I will present a forest plot with the model
```{r}
ggforest(coxPH_model8a)
```

```{r}
ggforest(coxPH_model8b)
```


# Pre/Post-2015 Only - Association between SO4 and Baseline FVC
## Pre/Post-2015 Only - Visual EDA 5yrs pre-Dx
Scatterplot of continuous SO4 5yrs pre-diagnosis vs FVC, with sex represented by the color of points - Pre-2015
```{r}
(pre2015 %>% ggplot(aes(x=SO4_5yrPreDx, y=fvc_pct, color=sex))+
   geom_point()+
   labs(x="SO4 5yrs Pre-Diagnosis", y="Baseline FVC % Predicted", title="Baseline FVC % Predicted by SO4 5yrs Pre-Diagnosis")+
   geom_smooth(method="lm", se = FALSE) +
   theme(plot.title = element_text(hjust = 0.5))+
   xlim(5,20))
```

Scatterplot of continuous SO4 5yrs pre-diagnosis vs FVC, with sex represented by the color of points - Post-2015
```{r}
(post2015 %>% ggplot(aes(x=SO4_5yrPreDx, y=fvc_pct, color=sex))+
   geom_point()+
   labs(x="SO4 5yrs Pre-Diagnosis", y="Baseline FVC % Predicted", title="Baseline FVC % Predicted by SO4 5yrs Pre-Diagnosis")+
   geom_smooth(method="lm", se = FALSE) +
   theme(plot.title = element_text(hjust = 0.5))+
   xlim(5,20))
```

Violin plot wrapping boxplot to visualize SO4 5yrs pre-dx low vs high vs FVC - Pre-2015
```{r}
pre2015b <- pre2015 %>% filter(!is.na(SO45yr_dich))
(pre2015b %>% ggplot(aes(x=SO45yr_dich, y=fvc_pct, fill=SO45yr_dich))+
   geom_boxplot(width=0.2, color="black", alpha=1.0)+
   geom_violin(width=1.0, alpha=0.5)+
   labs(x="SO4 5yrs Pre-Diagnosis Low vs High", y="Baseline FVC % Predicted", title="Baseline FVC % Predicted by SO4 5yrs Pre-Diagnosis Low vs High")+
   theme_light()+
   theme(legend.position = "none", plot.title = element_text(hjust = 0.5))+
   scale_fill_brewer(type="seq", palette="YlOrRd"))
```

Violin plot wrapping boxplot to visualize SO4 5yrs pre-dx low vs high vs FVC - Post-2015
```{r}
post2015b <- post2015 %>% filter(!is.na(SO45yr_dich))
(post2015b %>% ggplot(aes(x=SO45yr_dich, y=fvc_pct, fill=SO45yr_dich))+
   geom_boxplot(width=0.2, color="black", alpha=1.0)+
   geom_violin(width=1.0, alpha=0.5)+
   labs(x="SO4 5yrs Pre-Diagnosis Low vs High", y="Baseline FVC % Predicted", title="Baseline FVC % Predicted by SO4 5yrs Pre-Diagnosis Low vs High")+
   theme_light()+
   theme(legend.position = "none", plot.title = element_text(hjust = 0.5))+
   scale_fill_brewer(type="seq", palette="YlOrRd"))
```



## Pre/Post-2015 Only - SO4 5yrs Pre-Diagnosis and FVC Models
Looking at continuous SO4_5yrPreDx vs baseline FVC

Simplest model just looking at SO4 in 5yrs pre-diagnosis
```{r}
FVC_model1a <- lm(fvc_pct ~ SO4_5yrPreDx, data=pre2015)
summary(FVC_model1a)
confint(FVC_model1a)
```

Simplest model just looking at SO4 in 5yrs pre-diagnosis
```{r}
FVC_model1b <- lm(fvc_pct ~ SO4_5yrPreDx, data=post2015)
summary(FVC_model1b)
confint(FVC_model1b)
```

Complete model
```{r}
FVC_model2a <- lm(fvc_pct ~ SO4_5yrPreDx + sex + age_dx + smokeHx + dich_Race + avg_s + site, data=pre2015)
summary(FVC_model2a)
confint(FVC_model2a)
```

Complete model
```{r}
FVC_model2b <- lm(fvc_pct ~ SO4_5yrPreDx + sex + age_dx + smokeHx + dich_Race + avg_s + site, data=post2015)
summary(FVC_model2b)
confint(FVC_model2b)
```


### Pre/Post-2015 Only - SO4 5yrs Pre-Diagnosis Low vs High and Baseline FVC
Looking at SO45yr_dich vs baseline FVC

Simplest model just looking at SO4 in 5yrs pre-diagnosis
```{r}
FVC_model1a <- lm(fvc_pct ~ SO45yr_dich, data=pre2015)
summary(FVC_model1a)
confint(FVC_model1a)
```

Simplest model just looking at SO4 in 5yrs pre-diagnosis
```{r}
FVC_model1b <- lm(fvc_pct ~ SO45yr_dich, data=post2015)
summary(FVC_model1b)
confint(FVC_model1b)
```

Complete model
```{r}
FVC_model2a <- lm(fvc_pct ~ SO45yr_dich + sex + age_dx + smokeHx + dich_Race + avg_s + site, data=pre2015)
summary(FVC_model2a)
confint(FVC_model2a)
```

Complete model
```{r}
FVC_model2b <- lm(fvc_pct ~ SO45yr_dich + sex + age_dx + smokeHx + dich_Race + avg_s + site, data=post2015)
summary(FVC_model2b)
confint(FVC_model2b)
```

# Pre/Post-2015 Only - Association between SO4 and Baseline DLCO
## Pre/Post-2015 Only - Visual EDA 5yrs pre-Dx
Scatterplot of continuous SO4 5yrs pre-diagnosis vs DLCO, with sex represented by the color of points - Pre-2015
```{r}
(pre2015 %>% ggplot(aes(x=SO4_5yrPreDx, y=dlco_pct, color=sex))+
   geom_point()+
   labs(x="SO4 5yrs Pre-Diagnosis", y="Baseline DLCO % Predicted", title="Baseline DLCO % Predicted by SO4 5yrs Pre-Diagnosis")+
   geom_smooth(method="lm", se = FALSE) +
   theme(plot.title = element_text(hjust = 0.5))+
   xlim(5,20))
```

Scatterplot of continuous SO4 5yrs pre-diagnosis vs DLCO, with sex represented by the color of points - Post-2015
```{r}
(post2015 %>% ggplot(aes(x=SO4_5yrPreDx, y=dlco_pct, color=sex))+
   geom_point()+
   labs(x="SO4 5yrs Pre-Diagnosis", y="Baseline DLCO % Predicted", title="Baseline DLCO % Predicted by SO4 5yrs Pre-Diagnosis")+
   geom_smooth(method="lm", se = FALSE) +
   theme(plot.title = element_text(hjust = 0.5))+
   xlim(5,20))
```

Violin plot wrapping boxplot to visualize SO4 5yrs pre-dx low vs high vs DLCO - Pre-2015
```{r}
(pre2015b %>% ggplot(aes(x=SO45yr_dich, y=dlco_pct, fill=SO45yr_dich))+
   geom_boxplot(width=0.2, color="black", alpha=1.0)+
   geom_violin(width=1.0, alpha=0.5)+
   labs(x="SO4 5yrs Pre-Diagnosis Low vs High", y="Baseline DLCO % Predicted", title="Baseline DLCO % Predicted by SO4 5yrs Pre-Diagnosis Low vs High")+
   theme_light()+
   theme(legend.position = "none", plot.title = element_text(hjust = 0.5))+
   scale_fill_brewer(type="seq", palette="YlOrRd"))
```

Violin plot wrapping boxplot to visualize SO4 5yrs pre-dx low vs high vs DLCO - Post-2015
```{r}
(post2015b %>% ggplot(aes(x=SO45yr_dich, y=dlco_pct, fill=SO45yr_dich))+
   geom_boxplot(width=0.2, color="black", alpha=1.0)+
   geom_violin(width=1.0, alpha=0.5)+
   labs(x="SO4 5yrs Pre-Diagnosis Low vs High", y="Baseline DLCO % Predicted", title="Baseline DLCO % Predicted by SO4 5yrs Pre-Diagnosis Low vs High")+
   theme_light()+
   theme(legend.position = "none", plot.title = element_text(hjust = 0.5))+
   scale_fill_brewer(type="seq", palette="YlOrRd"))
```

## Pre/Post-2015 Only - SO4 5yrs Pre-Diagnosis and DLCO Models
### Pre/Post-2015 Only - SO4 5yrs Pre-Diagnosis and Baseline DLCO
Looking at continuous SO4_5yrPreDx vs baseline DLCO

Simplest model just looking at SO4 in 5yrs pre-diagnosis
```{r}
DLCO_model1a <- lm(dlco_pct ~ SO4_5yrPreDx, data=pre2015)
summary(DLCO_model1a)
confint(DLCO_model1a)
```

Simplest model just looking at SO4 in 5yrs pre-diagnosis
```{r}
DLCO_model1b <- lm(dlco_pct ~ SO4_5yrPreDx, data=post2015)
summary(DLCO_model1b)
confint(DLCO_model1b)
```

Complete model
```{r}
DLCO_model2a <- lm(dlco_pct ~ SO4_5yrPreDx + sex + age_dx + smokeHx + dich_Race + avg_s + site, data=pre2015)
summary(DLCO_model2a)
confint(DLCO_model2a)
```

Complete model
```{r}
DLCO_model2b <- lm(dlco_pct ~ SO4_5yrPreDx + sex + age_dx + smokeHx + dich_Race + avg_s + site, data=post2015)
summary(DLCO_model2b)
confint(DLCO_model2b)
```

### Pre/Post-2015 Only - SO4 5yrs Pre-Diagnosis Low vs High and Baseline DLCO
Looking at SO45yr_dich vs baseline DLCO

Simplest model just looking at SO4 in 5yrs pre-diagnosis
```{r}
DLCO_model1a <- lm(dlco_pct ~ SO45yr_dich, data=pre2015)
summary(DLCO_model1a)
confint(DLCO_model1a)
```

Simplest model just looking at SO4 in 5yrs pre-diagnosis
```{r}
DLCO_model1b <- lm(dlco_pct ~ SO45yr_dich, data=post2015)
summary(DLCO_model1b)
confint(DLCO_model1b)
```

Complete model
```{r}
DLCO_model2a <- lm(dlco_pct ~ SO45yr_dich + sex + age_dx + smokeHx + dich_Race + avg_s + site, data=pre2015)
summary(DLCO_model2a)
confint(DLCO_model2a)
```

Complete model
```{r}
DLCO_model2b <- lm(dlco_pct ~ SO45yr_dich + sex + age_dx + smokeHx + dich_Race + avg_s + site, data=post2015)
summary(DLCO_model2b)
confint(DLCO_model2b)
```



