---
title: "Monthly Satellite SO4 Analysis with Complete PFF Data using Consent Date as Reference - Updated Dichotomized SO4"
author: "Gillian Goobie"
date: "02/12/2022"
output: 
  bookdown::html_document2:
    toc: yes
    toc_float: yes
    number_sections: yes
    code_folding: hide
  html_document: 
    toc: yes
    toc_float: yes
    number_sections: yes
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE, warning=FALSE, echo=F}
library(tidyverse)
library(readxl)
library(here)
library(writexl)
library(knitr)
library(lintr)
library(tidync)
library(maps)
library(ncmeta)
library(devtools)
library(RNetCDF)
library(ncdf4)
library(survival)
library(survminer)
library(lubridate)
library(cmprsk)
library(riskRegression)
library(prodlim)
library(RColorBrewer)
library(prodlim)
library(lme4)
library(performance)
library(psycho)
library(report)
library(rms)
library(splines)
library(Greg)
library(rstpm2)
library(pspline)
```

# Importing Datasets
Here I am importing the file which contains monthly SO4 level estimates by satellite at nearest lon/lat to PFF patient residential addresses. These are linked to the patient ID.
```{r}
outfile1 <- here("MonthlySO4Data/PFF_fILD_2000_2017_SO4_2021_11_05.xlsx")
SO4 <- read_excel(outfile1)
```

Match up SO4 ID to SSID from matching file
```{r}
outfile2 <- here("MonthlySO4Data/PFF_fILD_PM25_BaselineData_2021_10_20.xlsx")
PM <- read_excel(outfile2)
PM <- PM %>% dplyr::select(ID, SSID)
```

Merge SO4 and PM
```{r}
SO4 <- left_join(SO4, PM, by="ID")
```



Here I am importing the file containing the complete baseline clinical and demographic data for 1905 PFF patients
```{r}
outfile3 <- here("PFF_fILDPts_BaselineData_ConsentDateReference_2021_11_17.xlsx")
PFF <- read_excel(outfile3)
```

# Merging Datasets
I used a inner_join here so that the complete "SO4" dataframe only includes patients with fILD that have all baseline demographics and SO4 data.
```{r}
SO4 <- inner_join(PFF, SO4, by="SSID")
```
This results in 1905 complete records


# Pivoting to Long Format
First we need to convert the SO4 dataframe into the long rather than the wide format, which will allow us to use it more easily in R's tidyverse as this is "tidy" formatting.
```{r}
SO4 <- SO4 %>% 
  pivot_longer(cols=c(45:260), names_to="SO4_date", names_prefix="SO4_")
```

## Convert date SO4_date to same format as above
```{r}
SO4x <- SO4 
SO4x$SO4_date <- gsub("jan", "01-01-20", SO4x$SO4_date)
SO4x$SO4_date <- gsub("feb", "01-02-20", SO4x$SO4_date)
SO4x$SO4_date <- gsub("mar", "01-03-20", SO4x$SO4_date)
SO4x$SO4_date <- gsub("apr", "01-04-20", SO4x$SO4_date)
SO4x$SO4_date <- gsub("may", "01-05-20", SO4x$SO4_date)
SO4x$SO4_date <- gsub("jun", "01-06-20", SO4x$SO4_date)
SO4x$SO4_date <- gsub("jul", "01-07-20", SO4x$SO4_date)
SO4x$SO4_date <- gsub("aug", "01-08-20", SO4x$SO4_date)
SO4x$SO4_date <- gsub("sep", "01-09-20", SO4x$SO4_date)
SO4x$SO4_date <- gsub("oct", "01-10-20", SO4x$SO4_date)
SO4x$SO4_date <- gsub("nov", "01-11-20", SO4x$SO4_date)
SO4x$SO4_date <- gsub("dec", "01-12-20", SO4x$SO4_date)

SO4x$SO4_date <- format(as.Date(SO4x$SO4_date, format="%d-%m-%Y"),"%Y-%m-%d")
SO4x$SO4_date <- as.Date(SO4x$SO4_date)
SO4 <- SO4x
```


# Convert Date Columns to Date Format
Next I need to convert all date columns to proper format
```{r}
SO4 <- SO4 %>% 
  mutate_at(c("dx_date", "consent_date", "death_date", "tx_date", "sample_date", "fvc_date", "dlco_date", "censor_date", "deathORtx_date", "DeathTxCensor_date"), as.Date)
str(SO4)
```

# Creating Exposure Variables
Now I'm creating new variables where I am matching up SO4 averages per year to years of major events for patients (year of diagnosis, year of death/lung transplant/censoring, etc)

## 5yrs Pre-Death/Transplant/Censoring
Here I am calculating the average SO4 value in the 5yrs prior to death/lung transplant/censoring.
```{r}
SO4 <- SO4 %>% 
  group_by(ID) %>% 
  mutate(SO4_5yrPreCensor = mean(value[ymd(SO4_date)>=(ymd(DeathTxCensor_date) - years(5)) & 
                                        ymd(SO4_date)<=ymd(DeathTxCensor_date)]))
           
           
#mean(value[ymd(SO4_date) == ymd(DeathTxCensor_date) | ymd(SO4_date) == (ymd(DeathTxCensor_date) - years(1))]))
```

## 5 years Prior to Diagnosis
Here I am calculating the average SO4 value in the year of diagnosis and the 5 years leading up to diagnosis (so if all years available, it will be the average of 6 years).
```{r}
SO4 <- SO4 %>% 
  group_by(ID) %>% 
  mutate(SO4_5yrPreDx = mean(value[ymd(SO4_date) <= ymd(dx_date) & 
                                    ymd(SO4_date) >= (ymd(dx_date)-years(5))]))
```

## Remove Dataframes not in use
```{r}
rm(list=c("SO4x", "PM", "PFF"))
```

## Getting Rid of Duplicated IDs
Now that we have calculated the patient-specific SO4 exposures, I can get rid of all rows other than the first row for each patient.
```{r}
SO4 <-  SO4 %>% 
  distinct_at(vars(ID), .keep_all=T)
```
This takes us down to our 1905 observations that we had before.

## Remove UPitt patients
```{r}
SO4 <- SO4 %>% filter(!str_detect(SSID, "^08R"))
```
This now takes us down to 1870 observations after we remove UPitt patients which may be doubled up in our Simmons analyses


# Creating High vs Low Exposure Groups
Cutoffs are based on calculations performed where the median total PM2.5 mass across all three cohorts was averaged, then the median constituent component proportion of total PM2.5 across all three cohorts was averaged. Subsequently, the average proportion of each constituent across the three cohorts was multiplied by the average of the total PM2.5 mass medians across the three cohorts to get the approximate median total mass of each constituent across all three cohorts, which was used to define the low vs high exposure cut points.

## SO4 5yrs Pre-Censor High vs Low
```{r}
SO4$SO45yrCensor_dich <- cut(SO4$SO4_5yrPreCensor,
                      breaks=c(0, 1.23, 50),
                      labels=c("Low", "High"))
summary(SO4$SO45yrCensor_dich)
class(SO4$SO45yrCensor_dich)
```


## SO4 5yrs Pre-Diagnosis High vs Low
```{r}
SO4$SO45yr_dich <- cut(SO4$SO4_5yrPreDx,
                      breaks=c(0, 1.40, 50),
                      labels=c("Low", "High"))
summary(SO4$SO45yr_dich)
class(SO4$SO45yr_dich)
```



# Creating Other New Variables
## IPF vs Other Diagnosis
```{r}
SO4 <- SO4 %>% 
  mutate(dx_IPF=ifelse(dx=="IPF", "IPF", "not_IPF"))    
```

# Releveling Factors
Here I am releveling factors so that they are in an intuitive order for my later analyses.
```{r}
SO4$sex <- fct_relevel(SO4$sex, c("Male","Female"))
SO4$race <- fct_relevel(SO4$race, c("W","B","A","I","P","U"))
SO4$ethnicity <- fct_relevel(SO4$ethnicity, c("N","H","U"))
SO4$dich_Race <- fct_relevel(SO4$dich_Race, c("White","Non-White"))
SO4$smokeHx <- fct_relevel(SO4$smokeHx, c("Never","Ever"))

#For dx and dx_group, I just want IPF to be first and then the rest of the categories are alphabetical
SO4$dx <- fct_relevel(SO4$dx, c("IPF"))
SO4$dx_group <- fct_relevel(SO4$dx_group, c("IPF"))
SO4$dx_IPF <- fct_relevel(SO4$dx_IPF, c("IPF"))
str(SO4)
```


# Creating Metropolitan Category
Using https://www.ers.usda.gov/data-products/rural-urban-commuting-area-codes.aspx "2010 Rural-Urban Commuting Area Codes ZIP code file", which will allow me to match RUCA code to 5 digit zip code to determine if the patient lives in an area considered urban or rural. The interpretation of the code is below:

Primary RUCA codes, 2010
Code	Classification description
1	Metropolitan area core: primary flow within an urbanized area (UA)
2	Metropolitan area high commuting: primary flow 30% or more to a UA
3	Metropolitan area low commuting: primary flow 10% to 30% to a UA
4	Micropolitan area core: primary flow within an urban cluster of 10,000 to 49,999 (large UC)
5	Micropolitan high commuting: primary flow 30% or more to a large UC
6	Micropolitan low commuting: primary flow 10% to 30% to a large UC
7	Small town core: primary flow within an urban cluster of 2,500 to 9,999 (small UC)
8	Small town high commuting: primary flow 30% or more to a small UC
9	Small town low commuting: primary flow 10% to 30% to a small UC
10	Rural areas: primary flow to a tract outside a UA or UC
99	Not coded: Census tract has zero population and no rural-urban identifier information

I will classify RUCA codes 1-6 as urban and 7-10 as rural for now.

First I need to upload the RUCA to zip code file
```{r}
outfile3 <- here("RUCA2010zipcode_2021_10_22.xlsx")
ruca <- read_excel(outfile3, sheet="Data")
```

Then I will simplify the dataframe to include only what I need
```{r}
ruca <- ruca %>% dplyr::select(ZIP_CODE, RUCA1)
ruca <- ruca %>% rename("zip"="ZIP_CODE", "ruca"="RUCA1")
```

Then I will merge with my PFF dataframe
```{r}
SO4 <- left_join(SO4, ruca, by="zip")
```

Then I will make rural vs urban column
```{r}
SO4 <- SO4 %>% mutate(metro=ifelse(ruca>6, "rural", ifelse(ruca<4, "metropolitan", "micropolitan")))
```

# Putting ID column first
```{r}
SO4 <- SO4 %>% dplyr::select("ID", everything())
```

# Creating Datasets with only Metropolitan Patients
```{r}
SO4_metro <- SO4 %>% filter(metro=="metropolitan")
```
This results in a dataset of 1637 patients

# Creating Datasets with only IPF Patients
```{r}
SO4_IPF <- SO4 %>% filter(dx=="IPF")
```
This results in a dataset of 1217 pts

# Creating Datasets with only CTD Patients
```{r}
SO4_CTD <- SO4 %>% filter(dx_group=="CTD-ILD")
```
This results in a dataset of 321 pts

# Export Excel File with SO4 Data Matched to Patients and Full Cohort Data
```{r}
write_xlsx(SO4, path="Final_PFF_fILD_SO4_BaselineData_2022_02_12.xlsx")
```

# Exploratory Data Analysis
This is a function that allows me to make tables which summarize the count and percentages of each level of factor variables
```{r}
n_prop_tbl <- function(x) {
  tbl <- table(x)
  res <- cbind(tbl, round(prop.table(tbl)*100,2))
  colnames(res) <-  c('Count', 'Percentage')
  res
}
```


## SO4 Breakdown
```{r}
summary(SO4$SO4_5yrPreCensor)
summary(SO4$SO4_5yrPreDx)
```

## Sex Breakdown
```{r}
n_prop_tbl(SO4$sex)
```

## Race and Ethnicity Breakdown 
```{r}
n_prop_tbl(SO4$race)

n_prop_tbl(SO4$ethnicity)
```

## Smoking History Breakdown
```{r}
n_prop_tbl(SO4$smokeHx)
```


## Diagnostic Group Breakdown
```{r}
n_prop_tbl(SO4$dx_group)
```


## State Breakdown
```{r}
n_prop_tbl(SO4$state)
```

## Metropolitan Breakdown
```{r}
n_prop_tbl(SO4$metro)
```
Overwhelming majority of the patients are considered "metropolitan".

## Age at Diagnosis Breakdown
```{r}
shapiro.test(SO4$age_dx)
#Shapiro tests tells us that age_dx is not normally distributed, so should report median, IQR
summary(SO4$age_dx)
```
Not normally distributed.

## Vital Status Breakdown
```{r}
n_prop_tbl(SO4$status)
```
22.8% died, 13.7% transplanted.

## Baseline Lung Function Breakdown
```{r}
shapiro.test(SO4$fvc_pct)
summary(SO4$fvc_pct)
sd(SO4$fvc_pct, na.rm=T)
```
FVC not normally distributed.

```{r}
shapiro.test(SO4$dlco_pct)
summary(SO4$dlco_pct)
sd(SO4$dlco_pct, na.rm=T)
```
DLCO not normally distributed.

## Time to Censoring Breakdown
```{r}
summary(SO4$time_DeathTxCensor)
sd(SO4$time_DeathTxCensor, na.rm=T)
```
Average time to censoring is 2.5yrs, which is lower than CARE-PF or Simmons. This is shortened by the fact that we are using the consent_date rather than the dx_date as the reference date. 


## Breakdown of Characteristics by Low vs High
### SO4_5yrPreCensor Low vs High
```{r}
table(SO4$sex, SO4$SO45yrCensor_dich)
table(SO4$race, SO4$SO45yrCensor_dich)
table(SO4$dich_Race, SO4$SO45yrCensor_dich)
table(SO4$dx_group, SO4$SO45yrCensor_dich)
table(SO4$smokeHx, SO4$SO45yrCensor_dich)
table(SO4$metro, SO4$SO45yrCensor_dich)
table(SO4$state, SO4$SO45yrCensor_dich)
table(SO4$status, SO4$SO45yrCensor_dich)
group_by(SO4, SO45yrCensor_dich) %>%
  summarise(
    count = n(),
    mean = mean(age_dx, na.rm = TRUE),
    sd = sd(age_dx, na.rm = TRUE),
    median = median(age_dx, na.rm = TRUE),
    quantile = quantile(age_dx, probs=c(0.25), na.rm = TRUE),
    quantile2 = quantile(age_dx, probs=c(0.75), na.rm = TRUE)
  )
group_by(SO4, SO45yrCensor_dich) %>%
  summarise(
    count = n(),
    mean = mean(fvc_pct, na.rm = TRUE),
    sd = sd(fvc_pct, na.rm = TRUE),
    median = median(fvc_pct, na.rm = TRUE),
    quantile = quantile(fvc_pct, probs=c(0.25), na.rm = TRUE),
    quantile2 = quantile(fvc_pct, probs=c(0.75), na.rm = TRUE)
  )
group_by(SO4, SO45yrCensor_dich) %>%
  summarise(
    count = n(),
    mean = mean(dlco_pct, na.rm = TRUE),
    sd = sd(dlco_pct, na.rm = TRUE),
    median = median(dlco_pct, na.rm = TRUE),
    quantile = quantile(dlco_pct, probs=c(0.25), na.rm = TRUE),
    quantile2 = quantile(dlco_pct, probs=c(0.75), na.rm = TRUE)
  )
group_by(SO4, SO45yrCensor_dich) %>%
  summarise(
    count = n(),
    mean = mean(time_DeathTxCensor, na.rm = TRUE),
    sd = sd(time_DeathTxCensor, na.rm = TRUE),
    median = median(time_DeathTxCensor, na.rm = TRUE),
    quantile = quantile(time_DeathTxCensor, probs=c(0.25), na.rm = TRUE),
    quantile2 = quantile(time_DeathTxCensor, probs=c(0.75), na.rm = TRUE)
  )
```

### SO4_5yrDx Low vs High
```{r}
table(SO4$sex, SO4$SO45yr_dich)
table(SO4$race, SO4$SO45yr_dich)
table(SO4$dich_Race, SO4$SO45yr_dich)
table(SO4$dx_group, SO4$SO45yr_dich)
table(SO4$smokeHx, SO4$SO45yr_dich)
table(SO4$metro, SO4$SO45yr_dich)
table(SO4$state, SO4$SO45yr_dich)
table(SO4$status, SO4$SO45yr_dich)
group_by(SO4, SO45yr_dich) %>%
  summarise(
    count = n(),
    mean = mean(age_dx, na.rm = TRUE),
    sd = sd(age_dx, na.rm = TRUE),
    median = median(age_dx, na.rm = TRUE),
    quantile = quantile(age_dx, probs=c(0.25), na.rm = TRUE),
    quantile2 = quantile(age_dx, probs=c(0.75), na.rm = TRUE)
  )
group_by(SO4, SO45yr_dich) %>%
  summarise(
    count = n(),
    mean = mean(fvc_pct, na.rm = TRUE),
    sd = sd(fvc_pct, na.rm = TRUE),
    median = median(fvc_pct, na.rm = TRUE),
    quantile = quantile(fvc_pct, probs=c(0.25), na.rm = TRUE),
    quantile2 = quantile(fvc_pct, probs=c(0.75), na.rm = TRUE)
  )
group_by(SO4, SO45yr_dich) %>%
  summarise(
    count = n(),
    mean = mean(dlco_pct, na.rm = TRUE),
    sd = sd(dlco_pct, na.rm = TRUE),
    median = median(dlco_pct, na.rm = TRUE),
    quantile = quantile(dlco_pct, probs=c(0.25), na.rm = TRUE),
    quantile2 = quantile(dlco_pct, probs=c(0.75), na.rm = TRUE)
  )
group_by(SO4, SO45yr_dich) %>%
  summarise(
    count = n(),
    mean = mean(time_DeathTxCensor, na.rm = TRUE),
    sd = sd(time_DeathTxCensor, na.rm = TRUE),
    median = median(time_DeathTxCensor, na.rm = TRUE),
    quantile = quantile(time_DeathTxCensor, probs=c(0.25), na.rm = TRUE),
    quantile2 = quantile(time_DeathTxCensor, probs=c(0.75), na.rm = TRUE)
  )
```


## Visual EDA
### Histograms
Histogram of SO4 Exposure 5yr Pre-Death/Transplant/Censoring
```{r}
(SO4 %>% ggplot(aes(x=SO4_5yrPreCensor))+
   geom_histogram(fill="blue", color="black")+
   labs(x="Average SO4 Level 5yr Prior to Death/Transplant/Censoring", y="Number of patients with fILD", title="Average SO4 5yr Prior to Death/Transplant/Censoring")+
   theme(plot.title = element_text(hjust = 0.5))+
   xlim(0,10))
```
Much wider distribution than CARE-PF or PFF, and much higher median exposures.

Histogram of SO4 Exposure 5yr Pre-Dx
```{r}
(SO4 %>% ggplot(aes(x=SO4_5yrPreDx))+
   geom_histogram(fill="blue", color="black")+
   labs(x="Average SO4 Level 5yr Prior to Diagnosis", y="Number of patients with fILD", title="Average SO4 5yr Prior to Diagnosis")+
   theme(plot.title = element_text(hjust = 0.5))+
   xlim(0,10))
```


### Violin plots of SO4 vs other demographic variables
Violin plot wrapping boxplot to visualize sex vs SO4_5yrPreCensor
```{r}
(SO4 %>% ggplot(aes(x=sex, y=SO4_5yrPreCensor, fill=sex))+
   geom_boxplot(width=0.2, color="black", alpha=1.0)+
   geom_violin(width=1.0, alpha=0.5)+
   labs(x="Sex", y="Average SO4 5 Years Pre-Censor", title="SO4 Exposure Pre-Censor by Sex")+
   theme(plot.title = element_text(hjust = 0.5))+
   ylim(0,3))
```
SO4 appears lower in females.

Violin plot wrapping boxplot to visualize sex vs SO4_5yrPreDx
```{r}
(SO4 %>% ggplot(aes(x=sex, y=SO4_5yrPreDx, fill=sex))+
   geom_boxplot(width=0.2, color="black", alpha=1.0)+
   geom_violin(width=1.0, alpha=0.5)+
   labs(x="Sex", y="Average SO4 5 Years Pre-Diagnosis", title="SO4 Exposure Pre-Diagnosis by Sex")+
   theme(plot.title = element_text(hjust = 0.5))+
   ylim(0,6))
```

Violin plot wrapping boxplot to visualize dich_Race vs SO4_5yrPreCensor
```{r}
(SO4 %>% ggplot(aes(x=dich_Race, y=SO4_5yrPreCensor, fill=dich_Race))+
   geom_boxplot(width=0.2, color="black", alpha=1.0)+
   geom_violin(width=1.0, alpha=0.5)+
   labs(x="Race", y="Average SO4 5 Years Pre-Censoring", title="SO4 Exposure Pre-Censoring by Race")+
   theme(plot.title = element_text(hjust = 0.5))+
   ylim(0,3))
```
No clear difference by race

Violin plot wrapping boxplot to visualize race vs SO4_5yrPreCensor
```{r}
(SO4 %>% ggplot(aes(x=race, y=SO4_5yrPreCensor, fill=race))+
   geom_boxplot(width=0.2, color="black", alpha=1.0)+
   geom_violin(width=1.0, alpha=0.5)+
   labs(x="Race", y="Average SO4 5 Years Pre-Censoring", title="SO4 Exposure Pre-Censoring by Race")+
   theme(plot.title = element_text(hjust = 0.5))+
   ylim(0,2.5))
```

Violin plot wrapping boxplot to visualize dich_Race vs SO4_5yrPreDx
```{r}
(SO4 %>% ggplot(aes(x=dich_Race, y=SO4_5yrPreDx, fill=dich_Race))+
   geom_boxplot(width=0.2, color="black", alpha=1.0)+
   geom_violin(width=1.0, alpha=0.5)+
   labs(x="Race", y="Average SO4 5 Years Pre-Diagnosis", title="SO4 Exposure Pre-Diagnosis by Race")+
   theme(plot.title = element_text(hjust = 0.5))+
   ylim(0,6))
```

Violin plot wrapping boxplot to visualize race vs SO4_5yrPreDx
```{r}
(SO4 %>% ggplot(aes(x=race, y=SO4_5yrPreDx, fill=race))+
   geom_boxplot(width=0.2, color="black", alpha=1.0)+
   geom_violin(width=1.0, alpha=0.5)+
   labs(x="Race", y="Average SO4 5 Years Pre-Diagnosis", title="SO4 Exposure Pre-Diagnosis by Race")+
   theme(plot.title = element_text(hjust = 0.5))+
   ylim(0,6))
```

Violin plot wrapping boxplot to visualize metro vs SO4_5yrPreCensor
```{r}
(SO4 %>% ggplot(aes(x=metro, y=SO4_5yrPreCensor, fill=metro))+
   geom_boxplot(width=0.2, color="black", alpha=1.0)+
   geom_violin(width=1.0, alpha=0.5)+
   labs(x="Residential Location", y="Average SO4 in 5yrs Pre-Censoring", title="SO4 Exposure in 5yrs Pre-Censoring by Residential Location")+
   theme(plot.title = element_text(hjust = 0.5))+
   ylim(0,3))
```
Pretty clear difference in SO4 exposures by residential location with patients in metro having the highest SO4 exposure, then micro, then rural.

Violin plot wrapping boxplot to visualize metro vs SO4_5yrPreDx
```{r}
(SO4 %>% ggplot(aes(x=metro, y=SO4_5yrPreDx, fill=metro))+
   geom_boxplot(width=0.2, color="black", alpha=1.0)+
   geom_violin(width=1.0, alpha=0.5)+
   labs(x="Residential Location", y="Average SO4 in 5yrs Pre-Diagnosis", title="SO4 Exposure in 5yrs Pre-Diagnosis by Residential Location")+
   theme(plot.title = element_text(hjust = 0.5))+
   ylim(0,6))
```


### Scatterplots of SO4 over time
Scatterplot showing average SO4 5yr pre-death/transplant/censoring
```{r}
(SO4 %>% ggplot(aes(x=dx_date, y=SO4_5yrPreCensor))+
   geom_point()+
   labs(x="Date of Diagnosis", y="SO4 Exposure 5yr Prior Death/Transplant/Censoring", title="SO4 5yr Prior to Death/Transplant/Censoring vs Date of Diagnosis")+
   geom_smooth(method="lm", se = FALSE) +
   theme(plot.title = element_text(hjust = 0.5))+
   ylim(0,2.5))
```
Clear decrease in SO4 exposures during 5yrs pre-censoring over time, with patients that were diagnosed later having lower exposures over this period.

Scatterplot showing average SO4 5yr pre-diagnosis
```{r}
(SO4 %>% ggplot(aes(x=dx_date, y=SO4_5yrPreDx))+
   geom_point()+
   labs(x="Date of Diagnosis", y="SO4 Exposure 5yr Prior Diagnosis", title="SO4 5yr Prior to Diagnosis vs Date of Diagnosis")+
   geom_smooth(method="lm", se = FALSE) +
   theme(plot.title = element_text(hjust = 0.5))+
   ylim(0,5))
```


# Cox PH Models for Association between SO4 and Death or Lung Transplant
## SO4 5yrs Pre-Censoring Models
### Continuous SO4 5yrs pre-censoring until Death/Transplant/Censoring
Start with the simplest model just looking at SO4 in the year of and 5yrs prior to death/transplant/censoring
```{r}
coxPH_model1 <- coxph(Surv(time_DeathTxCensor, deadORtx==1) ~ SO4_5yrPreCensor, data=SO4, id=ID)
summary(coxPH_model1)
```
**Significant assoc between increasing SO4 in 5yrs pre-censoring and increased mort in this model.**

Partial model
```{r}
coxPH_model2 <- coxph(Surv(time_DeathTxCensor, deadORtx==1) ~ SO4_5yrPreCensor + sex + age_dx + smokeHx + dich_Race + dx_IPF + metro, data=SO4, id=ID)
summary(coxPH_model2)
```
**Significant assoc between increasing SO4 in 5yrs pre-censoring and increased mort in this model.**

Complete model
```{r}
coxPH_model3 <- coxph(Surv(time_DeathTxCensor, deadORtx==1) ~ SO4_5yrPreCensor + sex + age_dx + smokeHx + dich_Race + dx_IPF + metro + fvc_pct + dlco_pct, data=SO4, id=ID)
summary(coxPH_model3)
```
**Significant assoc between increasing SO4 in 5yrs pre-censoring and increased mort in this model with HR=1.57.**


### Smoothed HR Function for 5yr Pre-Censor Continuous
Base model
```{r}
#coxPH_model1_tvc <- cox.tvc(coxPH_model1, "SO4_5yrPreCensor")
#print(coxPH_model1_tvc)
#plot(coxPH_model1_tvc)
```
Issues running this code in PFF cohort.

Partial model (all covariates except lung function)
```{r}
#coxPH_model2_tvc <- cox.tvc(coxPH_model2, "SO4_5yrPreCensor")
#print(coxPH_model2_tvc)
#plot(coxPH_model2_tvc)
```
Issues running this code in PFF cohort.

```{r}
#coxPH_model3_tvc <- cox.tvc(coxPH_model3, "SO4_5yrPreCensor")
#print(coxPH_model3_tvc)
#plot(coxPH_model3_tvc)
```
Issues running this code in PFF cohort.

### Variable HR for SO4 risk in 5yr Pre-Cenosr
Want a plot with x-axis of SO4 level and y-axis of HR

Unadjusted model
```{r}
#First need to make dataframe that only includes patients with time_DeathTxCensor
SO4x <- SO4 %>% filter(!is.na(SO4_5yrPreCensor) & !is.na(time_DeathTxCensor))

#Then make survival function
surv1 <- Surv(SO4x$time_DeathTxCensor, SO4x$deadORtx==1)
fit1 <- coxph(surv1 ~ pspline(SO4x$SO4_5yrPreCensor, df=4))
predicted <- predict(fit1, type="terms", se.fit=T, terms=1)

#Then plot
plot(SO4x$SO4_5yrPreCensor, exp(predicted$fit), type="n", xlim=c(0,6), ylim=c(0,10))
lines(sm.spline(SO4x$SO4_5yrPreCensor, exp(predicted$fit)), col = "red" , lty = 1 )
lines(sm.spline(SO4x$SO4_5yrPreCensor, exp(predicted$fit + 1.96 * predicted$se)), col = "orange" , lty = 2 )
lines(sm.spline(SO4x$SO4_5yrPreCensor, exp(predicted$fit - 1.96 * predicted$se)), col = "orange" , lty = 2 )
```

Partially-adjusted model
```{r}
#First need to make dataframe that only includes patients with time_DeathTxCensor
SO4x <- SO4 %>% filter(!is.na(SO4_5yrPreCensor) & !is.na(time_DeathTxCensor) & !is.na(age_dx) & !is.na(sex) & !is.na(smokeHx) & !is.na(dich_Race) & !is.na(dx_IPF) & !is.na(metro))

#Then survival function
surv1 <- Surv(SO4x$time_DeathTxCensor, SO4x$deadORtx==1)
fit1 <- coxph(surv1 ~ pspline(SO4x$SO4_5yrPreCensor, df=4) + SO4x$age_dx + SO4x$sex + SO4x$smokeHx + SO4x$dich_Race + SO4x$dx_IPF + SO4x$metro)
predicted <- predict(fit1, type="terms", se.fit=T, terms=1)

#Then plot
plot(SO4x$SO4_5yrPreCensor, exp(predicted$fit), type="n", xlim=c(0,6), ylim=c(0,10))
lines(sm.spline(SO4x$SO4_5yrPreCensor, exp(predicted$fit)), col = "red" , lty = 1 )
lines(sm.spline(SO4x$SO4_5yrPreCensor, exp(predicted$fit + 1.96 * predicted$se)), col = "orange" , lty = 2 )
lines(sm.spline(SO4x$SO4_5yrPreCensor, exp(predicted$fit - 1.96 * predicted$se)), col = "orange" , lty = 2 )
```

Fully-adjusted model
```{r}
#First need to make dataframe that only includes patients with time_DeathTxCensor
SO4x <- SO4 %>% filter(!is.na(SO4_5yrPreCensor) & !is.na(time_DeathTxCensor) & !is.na(age_dx) & !is.na(sex) & !is.na(smokeHx) & !is.na(dich_Race) & !is.na(dx_IPF) & !is.na(metro) & !is.na(fvc_pct) & !is.na(dlco_pct))

#Then survival function
surv1 <- Surv(SO4x$time_DeathTxCensor, SO4x$deadORtx==1)
fit1 <- coxph(surv1 ~ pspline(SO4x$SO4_5yrPreCensor, df=4) + SO4x$age_dx + SO4x$sex + SO4x$smokeHx + SO4x$dich_Race + SO4x$dx_IPF + SO4x$metro + SO4x$fvc_pct + SO4x$dlco_pct)
predicted <- predict(fit1, type="terms", se.fit=T, terms=1)

#Then plot
plot(SO4x$SO4_5yrPreCensor, exp(predicted$fit), type="n", xlim=c(0,6), ylim=c(0,10))
lines(sm.spline(SO4x$SO4_5yrPreCensor, exp(predicted$fit)), col = "red" , lty = 1 )
lines(sm.spline(SO4x$SO4_5yrPreCensor, exp(predicted$fit + 1.96 * predicted$se)), col = "orange" , lty = 2 )
lines(sm.spline(SO4x$SO4_5yrPreCensor, exp(predicted$fit - 1.96 * predicted$se)), col = "orange" , lty = 2 )
```


### SO4 Low vs High 5yrs Pre-Censoring
Base Model:
```{r}
coxPH_model7 <- coxph(Surv(time_DeathTxCensor, deadORtx==1) ~ SO45yrCensor_dich, data=SO4, id=ID)
summary(coxPH_model7)
```
**High SO4 exposure in the 5yrs pre-censoring assoc with increased risk of mort in this model.**

Partial model
```{r}
coxPH_model8 <- coxph(Surv(time_DeathTxCensor, deadORtx==1) ~ SO45yrCensor_dich + sex + age_dx + smokeHx + dich_Race + dx_IPF + metro, data=SO4, id=ID)
summary(coxPH_model8)
```
**High SO4 exposure in the 5yrs pre-censoring assoc with increased risk of mort in this model.**

Complete model
```{r}
coxPH_model9 <- coxph(Surv(time_DeathTxCensor, deadORtx==1) ~ SO45yrCensor_dich + sex + age_dx + smokeHx + dich_Race + dx_IPF + metro + fvc_pct + dlco_pct, data=SO4, id=ID)
summary(coxPH_model9)
```
**High SO4 exposure in the 5yrs pre-censoring assoc with increased risk of mort in this model with HR=5.79.**

Now I will make a Kaplan-Meier survival curve, discretized by SO4 during disease deciles:
```{r, fig.height=5, fig.width=7}
surv1 <- survfit(Surv(time_DeathTxCensor, deadORtx==1) ~ SO45yrCensor_dich, data=SO4)
(ggsurvplot(surv1, 
            palette=c("dodgerblue4","firebrick4"),
            censor=TRUE,
            pval=TRUE,
            xlim=c(0,5),
            legend="right",
           data=SO4,
           legend.title="SO4 Exposure",
           legend.labs=c("Low", "High"),
           risk.table=T,
           title="Time to Death or Transplant by SO4 5yrs Pre-Censoring",
           xlab="Time (years)",
           ylab="Survival Probability (Death or Lung Transplant)",
           ggtheme = theme_grey()
           ))
```
High exposure curve clearly lower in this time period.

Now we want to check the proportionality assumption by plotting the scaled Schoenfield residuals against time for each covariate in our final model.
```{r}
residtest <- cox.zph(coxPH_model9)
residtest
ggcoxzph(residtest)
```
Proportionality is met for SO45yrCensor_dich

Now I will present a forest plot with the model
```{r}
ggforest(coxPH_model9)
```
Clear difference between high vs low exposure.


# Association between SO4 and Baseline FVC
## Visual EDA 5yrs pre-Dx
Scatterplot of continuous SO4 5yrs pre-diagnosis vs FVC, with sex represented by the color of points
```{r}
(SO4 %>% ggplot(aes(x=SO4_5yrPreDx, y=fvc_pct, color=sex))+
   geom_point()+
   labs(x="SO4 5yrs Pre-Diagnosis", y="Baseline FVC % Predicted", title="Baseline FVC % Predicted by SO4 5yrs Pre-Diagnosis")+
   geom_smooth(method="lm", se = FALSE) +
   theme(plot.title = element_text(hjust = 0.5))+
   xlim(0,7))
```
Looks like in both women and men there might be a trend towards lower baseline FVC with higher SO4 in 5yrs pre-diagnosis, which is consistent with US cohort findings.

Violin plot wrapping boxplot to visualize SO4 5yrs pre-dx low vs high vs FVC
```{r}
SO4b <- SO4 %>% filter(!is.na(SO45yr_dich))
(SO4b %>% ggplot(aes(x=SO45yr_dich, y=fvc_pct, fill=SO45yr_dich))+
   geom_boxplot(width=0.2, color="black", alpha=1.0)+
   geom_violin(width=1.0, alpha=0.5)+
   labs(x="SO4 5yrs Pre-Diagnosis Low vs High", y="Baseline FVC % Predicted", title="Baseline FVC % Predicted by SO4 5yrs Pre-Diagnosis Low vs High")+
   theme_light()+
   theme(legend.position = "none", plot.title = element_text(hjust = 0.5))+
   scale_fill_brewer(type="seq", palette="YlOrRd"))
```
Slightly decreasing trend with increasing SO4 quartile


## SO4 5yrs Pre-Diagnosis and FVC Models
Looking at continuous SO4_5yrPreDx vs baseline FVC

Simplest model just looking at SO4 in 5yrs pre-diagnosis
```{r}
FVC_model1 <- lm(fvc_pct ~ SO4_5yrPreDx, data=SO4)
summary(FVC_model1)
confint(FVC_model1)
```
**In base model, higher SO4 in the 5yrs pre-diagnosis assoc with lower baseline FVC**

Partial model
```{r}
FVC_model2 <- lm(fvc_pct ~ SO4_5yrPreDx + sex + age_dx + smokeHx, data=SO4)
summary(FVC_model2)
confint(FVC_model2)
```
**In this model, higher SO4 in the 5yrs pre-diagnosis assoc with lower baseline FVC**

Complete model
```{r}
FVC_model3 <- lm(fvc_pct ~ SO4_5yrPreDx + sex + age_dx + smokeHx + dich_Race + dx_IPF + metro, data=SO4)
summary(FVC_model3)
confint(FVC_model3)
```
**Sig remains in complete model with metro included, indicating that for each 1ug/m3 increase in pre-diagnosis SO4, the baseline FVC is ~1.44% lower**

### SO4 5yrs Pre-Diagnosis Low vs High and Baseline FVC
Looking at SO45yr_dich vs baseline FVC

Simplest model just looking at SO4 in 5yrs pre-diagnosis
```{r}
FVC_model1 <- lm(fvc_pct ~ SO45yr_dich, data=SO4)
summary(FVC_model1)
confint(FVC_model1)
```
No assoc.

Partial model
```{r}
FVC_model2 <- lm(fvc_pct ~ SO45yr_dich + sex + age_dx + smokeHx, data=SO4)
summary(FVC_model2)
confint(FVC_model2)
```
No assoc.

Complete model
```{r}
FVC_model3 <- lm(fvc_pct ~ SO45yr_dich + sex + age_dx + smokeHx + dich_Race + dx_IPF + metro, data=SO4)
summary(FVC_model3)
confint(FVC_model3)
```
No assoc.


# Association between SO4 and Baseline DLCO
## Visual EDA 5yrs pre-Dx
Scatterplot of continuous SO4 5yrs pre-diagnosis vs DLCO, with sex represented by the color of points
```{r}
(SO4 %>% ggplot(aes(x=SO4_5yrPreDx, y=dlco_pct, color=sex))+
   geom_point()+
   labs(x="SO4 5yrs Pre-Diagnosis", y="Baseline dlco % Predicted", title="Baseline dlco % Predicted by SO4 5yrs Pre-Diagnosis")+
   geom_smooth(method="lm", se = FALSE) +
   theme(plot.title = element_text(hjust = 0.5))+
   xlim(0,12))
```
No clear trend here, although in women looks like possibly higher baseline DLCO with higher SO4

Violin plot wrapping boxplot to visualize SO4 5yrs pre-dx low vs high vs dlco
```{r}
(SO4b %>% ggplot(aes(x=SO45yr_dich, y=dlco_pct, fill=SO45yr_dich))+
   geom_boxplot(width=0.2, color="black", alpha=1.0)+
   geom_violin(width=1.0, alpha=0.5)+
   labs(x="SO4 5yrs Pre-Diagnosis Low vs High", y="Baseline DLCO % Predicted", title="Baseline DLCO % Predicted by SO4 5yrs Pre-Diagnosis Low vs High")+
   theme_light()+
   theme(legend.position = "none", plot.title = element_text(hjust = 0.5))+
   scale_fill_brewer(type="seq", palette="YlOrRd"))
```
No clear trend with increasing SO4 quartile


## SO4 5yrs Pre-Diagnosis and DLCO Models
### SO4 5yrs Pre-Diagnosis and Baseline DLCO
Looking at continuous SO4_5yrPreDx vs baseline DLCO

Simplest model just looking at SO4 in 5yrs pre-diagnosis
```{r}
DLCO_model1 <- lm(dlco_pct ~ SO4_5yrPreDx, data=SO4)
summary(DLCO_model1)
confint(DLCO_model1)
```
No sig assoc.

Partial model
```{r}
DLCO_model2 <- lm(dlco_pct ~ SO4_5yrPreDx + sex + age_dx + smokeHx, data=SO4)
summary(DLCO_model2)
confint(DLCO_model2)
```
No sig assoc.

Complete model
```{r}
DLCO_model3 <- lm(dlco_pct ~ SO4_5yrPreDx + sex + age_dx + smokeHx + dich_Race + dx_IPF + metro, data=SO4)
summary(DLCO_model3)
confint(DLCO_model3)
```
No sig assoc.


### SO4 5yrs Pre-Diagnosis Low vs High and Baseline DLCO
Looking at SO45yr_dich vs baseline DLCO

Simplest model just looking at SO4 in 5yrs pre-diagnosis
```{r}
DLCO_model1 <- lm(dlco_pct ~ SO45yr_dich, data=SO4)
summary(DLCO_model1)
confint(DLCO_model1)
```
No assoc.

Partial model
```{r}
DLCO_model2 <- lm(dlco_pct ~ SO45yr_dich + sex + age_dx + smokeHx, data=SO4)
summary(DLCO_model2)
confint(DLCO_model2)
```
No assoc.

Complete model
```{r}
DLCO_model3 <- lm(dlco_pct ~ SO45yr_dich + sex + age_dx + smokeHx + dich_Race + dx_IPF + metro, data=SO4)
summary(DLCO_model3)
confint(DLCO_model3)
```
