---
title: "Monthly Satellite Soil Analysis with CARE-PF Data - Updated Dichotomized Soil"
author: "Gillian Goobie"
date: "02/12/2022"
output: 
  html_document: 
    toc: yes
    toc_float: yes
    number_sections: yes
    code_folding: hide
  bookdown::html_document2:
    toc: yes
    toc_float: yes
    number_sections: yes
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE, warning=FALSE, echo=F}
library(tidyverse)
library(readxl)
library(here)
library(writexl)
library(knitr)
library(lintr)
library(tidync)
library(maps)
library(ncmeta)
library(devtools)
library(RNetCDF)
library(ncdf4)
library(survival)
library(survminer)
library(lubridate)
library(cmprsk)
library(riskRegression)
library(prodlim)
library(RColorBrewer)
library(prodlim)
library(lme4)
library(performance)
library(psycho)
library(report)
library(rms)
library(splines)
library(Greg)
library(rstpm2)
library(pspline)
library(smoothHR)
library(raster)
```

# Importing Datasets
Here I am importing the file which contains monthly Soil level estimates by satellite at nearest lon/lat to CAREPF patient residential addresses. These are linked to the patient ID.
```{r}
outfile1 <- here("MonthlySoilData/CARE_fILD_2000_2017_Soil_2021_11_05.xlsx")
Soil <- read_excel(outfile1)
```

Here I am importing the file which I used for my CIMD work that contains the baseline clinical and demographic data for 3389 patients who have CIMD and full records otherwise
```{r}
outfile2 <- here("MonthlySoilData/CAREPF_fILDPts_BaselineData_2021_10_22.xlsx")
CARE <- read_excel(outfile2)
```

# Merging Datasets
I used a inner_join here so that the complete "Soil" dataframe only includes patients with fILD that have all baseline demographics and CIMD available as well as Soil data.
```{r}
Soil <- inner_join(CARE, Soil, by="ID")
```

# Extracting year of diagnosis and year of death/transplant/censoring
Since the Soil Satellite data is only averages by month, I want to create a column for each patient where just the month/year of diagnosis and the month/year of death or lung transplant is extracted. Then I can use this to determine the average Soil exposure in the 5yrs prior to diagnosis and in the years from diagnosis until death/tx.
```{r}
#Start with the year of diagnosis
Soil <- Soil %>% 
  mutate(dx_yrmo = format(as.Date(Soil$dx_date, format="%Y-%m-%d"),"%Y-%m"))
Soil <- Soil %>% 
  mutate(dx_yr = format(as.Date(Soil$dx_date, format="%Y-%m-%d"),"%Y"))
Soil$dx_yr <- as.numeric(Soil$dx_yr)

#Then the year of death or lung transplant
Soil <- Soil %>% 
  mutate(deathORtx_date = if_else(!is.na(tx_date), tx_date, death_date))
Soil <- Soil %>% 
  mutate(deathORtx_yrmo = format(as.Date(Soil$deathORtx_date, format="%Y-%m-%d"),"%Y-%m"))

#Then the year the records were last updated (i.e. year of censoring)
Soil <- Soil %>% 
  mutate(DeathTxCensor_date = if_else(!is.na(deathORtx_date), deathORtx_date, last_updated))
Soil <- Soil %>% 
  mutate(censor_yrmo = format(as.Date(Soil$DeathTxCensor_date, format="%Y-%m-%d"),"%Y-%m"))
```

# Pivoting to Long Format
First we need to convert the Soil dataframe into the long rather than the wide format, which will allow us to use it more easily in R's tidyverse as this is "tidy" formatting.
```{r}
Soil <- Soil %>% 
  pivot_longer(cols=c(66:281), names_to="Soil_date", names_prefix="soil_")
```

## Convert date Soil_date to same format as above
```{r}
Soilx <- Soil 
Soilx$Soil_date <- gsub("jan", "01-01-20", Soilx$Soil_date)
Soilx$Soil_date <- gsub("feb", "01-02-20", Soilx$Soil_date)
Soilx$Soil_date <- gsub("mar", "01-03-20", Soilx$Soil_date)
Soilx$Soil_date <- gsub("apr", "01-04-20", Soilx$Soil_date)
Soilx$Soil_date <- gsub("may", "01-05-20", Soilx$Soil_date)
Soilx$Soil_date <- gsub("jun", "01-06-20", Soilx$Soil_date)
Soilx$Soil_date <- gsub("jul", "01-07-20", Soilx$Soil_date)
Soilx$Soil_date <- gsub("aug", "01-08-20", Soilx$Soil_date)
Soilx$Soil_date <- gsub("sep", "01-09-20", Soilx$Soil_date)
Soilx$Soil_date <- gsub("oct", "01-10-20", Soilx$Soil_date)
Soilx$Soil_date <- gsub("nov", "01-11-20", Soilx$Soil_date)
Soilx$Soil_date <- gsub("dec", "01-12-20", Soilx$Soil_date)

Soilx$Soil_date <- format(as.Date(Soilx$Soil_date, format="%d-%m-%Y"),"%Y-%m-%d")
Soilx$Soil_date <- as.Date(Soilx$Soil_date)
Soil <- Soilx
```


# Convert Date Columns to Date Format
Next I need to convert all date columns to proper format
```{r}
Soil <- Soil %>% 
  mutate_at(c("dob", "dx_date", "initial_visit_date", "formal_dx_date", "reg_update", "death_date", "tx_date", "pft_date", "last_updated", "deathORtx_date", "DeathTxCensor_date"), as.Date)
str(Soil)
```

# Creating Exposure Variables
Now I'm creating new variables where I am matching up Soil averages per year to years of major events for patients (year of diagnosis, year of death/lung transplant/censoring, etc)
## 5yrs Pre-Death/Transplant/Censoring
Here I am calculating the average Soil value in the 5yrs prior to death/lung transplant/censoring.
```{r}
Soil <- Soil %>% 
  group_by(ID) %>% 
  mutate(Soil_5yrPreCensor = mean(value[ymd(Soil_date)>=(ymd(DeathTxCensor_date) - years(5)) & 
                                        ymd(Soil_date)<=ymd(DeathTxCensor_date)]))
           
           
#mean(value[ymd(Soil_date) == ymd(DeathTxCensor_date) | ymd(Soil_date) == (ymd(DeathTxCensor_date) - years(1))]))
```

## 5 years Prior to Diagnosis
Here I am calculating the average Soil value in the year of diagnosis and the 5 years leading up to diagnosis (so if all years available, it will be the average of 6 years).
```{r}
Soil <- Soil %>% 
  group_by(ID) %>% 
  mutate(Soil_5yrPreDx = mean(value[ymd(Soil_date) <= ymd(dx_date) & 
                                    ymd(Soil_date) >= (ymd(dx_date)-years(5))]))
```

## Remove Dataframes not in use
```{r}
rm(list=c("Soilx", "CARE"))
```

## Getting Rid of Duplicated IDs
Now that we have calculated the patient-specific Soil exposures, I can get rid of all rows other than the first row for each patient.
```{r}
Soil <-  Soil %>% 
  distinct_at(vars(ID), .keep_all=T)
```
This takes us down to our  observations that we had before.

# Creating High vs Low Exposure Groups
Cutoffs are based on calculations performed where the median total PM2.5 mass across all three cohorts was averaged, then the median constituent component proportion of total PM2.5 across all three cohorts was averaged. Subsequently, the average proportion of each constituent across the three cohorts was multiplied by the average of the total PM2.5 mass medians across the three cohorts to get the approximate median total mass of each constituent across all three cohorts, which was used to define the low vs high exposure cut points.

## Soil 5yrs Pre-Censor High vs Low
```{r}
Soil$Soil5yrCensor_dich <- cut(Soil$Soil_5yrPreCensor,
                      breaks=c(0, 0.44, 50),
                      labels=c("Low", "High"))
summary(Soil$Soil5yrCensor_dich)
class(Soil$Soil5yrCensor_dich)
```

## Soil 5yrs Pre-Diagnosis High vs Low
```{r}
Soil$Soil5yr_dich <- cut(Soil$Soil_5yrPreDx,
                      breaks=c(0, 0.50, 50),
                      labels=c("Low", "High"))
summary(Soil$Soil5yr_dich)
class(Soil$Soil5yr_dich)
```


# Creating Other New Variables
## IPF vs Other Diagnosis
```{r}
Soil <- Soil %>% 
  mutate(dx_IPF=ifelse(dx=="IPF", "IPF", "not_IPF"))    
```

## Changing site to a factor variable
```{r}
Soil$site <- as.factor(Soil$site)
```

# Releveling Factors
Here I am releveling factors so that they are in an intuitive order for my later analyses.
```{r}
Soil$sex <- fct_relevel(Soil$sex, c("M","F"))
Soil$race <- fct_relevel(Soil$race, c("W","B","A","N","U"))
Soil$ethnicity <- fct_relevel(Soil$ethnicity, c("N","H","D","U"))
Soil$dich_Race <- fct_relevel(Soil$dich_Race, c("White","Non-White"))
Soil$smokeHx <- fct_relevel(Soil$smokeHx, c("Never","Former","Always"))

#For dx and dx_group, I just want IPF to be first and then the rest of the categories are alphabetical
Soil$dx <- fct_relevel(Soil$dx, c("IPF"))
Soil$dx_group <- fct_relevel(Soil$dx_group, c("IPF"))
Soil$dx_IPF <- fct_relevel(Soil$dx_IPF, c("IPF"))
str(Soil)
```

# Creating Datasets with only Metropolitan Patients
```{r}
Soil_metro <- Soil %>% filter(metro=="metropolitan")
```
This results in a dataset of  patients

# Creating Datasets with only IPF Patients
```{r}
Soil_IPF <- Soil %>% filter(dx=="IPF")
```
This results in a dataset of  pts

# Creating Datasets with only CTD Patients
```{r}
Soil_CTD <- Soil %>% filter(dx_group=="CTD-ILD")
```
This results in a dataset of  pts

# Creating individual site subgroups
```{r}
Soil_toronto <- Soil %>% filter(site==101)
Soil_ab <- Soil %>% filter(site==102)
Soil_sph <- Soil %>% filter(site==103)
Soil_vgh <- Soil %>%  filter(site==104)
Soil_bc <- Soil %>% filter(site==103 | site==104)
Soil_laval <- Soil %>% filter(site==105)
Soil_hamilton <- Soil %>% filter(site==106)
Soil_on <- Soil %>% filter(site==101 | site==106)
Soil_mcgill <- Soil %>% filter(site==107)
Soil_qc <- Soil %>% filter(site==105 | site==107)
Soil_sk <- Soil %>% filter(site==108)
```


# Export Excel File with Soil Data Matched to Patients and Full Cohort Data
```{r}
write_xlsx(Soil, path="Final_CAREPF_fILD_Soil_BaselineData_2022_02_12.xlsx")
```

# Exploratory Data Analysis
This is a function that allows me to make tables which summarize the count and percentages of each level of factor variables
```{r}
n_prop_tbl <- function(x) {
  tbl <- table(x)
  res <- cbind(tbl, round(prop.table(tbl)*100,2))
  colnames(res) <-  c('Count', 'Percentage')
  res
}
```


## Soil Breakdown
```{r}
summary(Soil$Soil_5yrPreCensor)
summary(Soil$Soil_5yrPreDx)
```


## Sex Breakdown
```{r}
n_prop_tbl(Soil$sex)
```

## Race and Ethnicity Breakdown 
```{r}
n_prop_tbl(Soil$race)

n_prop_tbl(Soil$ethnicity)
```

## Smoking History Breakdown
```{r}
n_prop_tbl(Soil$smokeHx)
summary(Soil$pkyrs)
```


## Diagnostic Group Breakdown
```{r}
n_prop_tbl(Soil$dx_group)
```


## State Breakdown
```{r}
n_prop_tbl(Soil$province)
```

### Metropolitan Breakdown
```{r}
n_prop_tbl(Soil$metro)
```
69% are metropolitan

## Age at Diagnosis Breakdown
```{r}
shapiro.test(Soil$age_dx)
#Shapiro tests tells us that age_dx is not normally distributed, so should report median, IQR
summary(Soil$age_dx)
```

## CIMD Breakdown
```{r}
summary(Soil$avg_s)
```


## Vital Status Breakdown
```{r}
n_prop_tbl(Soil$status)
```
Majority of the patients in CARE-PF are still alive.

## Baseline Lung Function Breakdown
```{r}
shapiro.test(Soil$fvc_pct)
summary(Soil$fvc_pct)
sd(Soil$fvc_pct, na.rm=T)
```

```{r}
shapiro.test(Soil$dlco_pct)
summary(Soil$dlco_pct)
sd(Soil$dlco_pct, na.rm=T)
```

## Time to Censoring Breakdown
```{r}
summary(Soil$time_DeathTxCensor)
sd(Soil$time_DeathTxCensor, na.rm=T)
```


## Breakdown of Characteristics by Low vs High
### Soil_5yrPreCensor Low vs High
```{r}
table(Soil$sex, Soil$Soil5yrCensor_dich)
table(Soil$race, Soil$Soil5yrCensor_dich)
table(Soil$dich_Race, Soil$Soil5yrCensor_dich)
table(Soil$dx_group, Soil$Soil5yrCensor_dich)
table(Soil$smokeHx, Soil$Soil5yrCensor_dich)
table(Soil$metro, Soil$Soil5yrCensor_dich)
table(Soil$province, Soil$Soil5yrCensor_dich)
table(Soil$status, Soil$Soil5yrCensor_dich)
group_by(Soil, Soil5yrCensor_dich) %>%
  summarise(
    count = n(),
    mean = mean(age_dx, na.rm = TRUE),
    sd = sd(age_dx, na.rm = TRUE),
    median = median(age_dx, na.rm = TRUE),
    quantile = quantile(age_dx, probs=c(0.25), na.rm = TRUE),
    quantile2 = quantile(age_dx, probs=c(0.75), na.rm = TRUE)
  )
group_by(Soil, Soil5yrCensor_dich) %>%
  summarise(
    count = n(),
    mean = mean(avg_s, na.rm = TRUE),
    sd = sd(avg_s, na.rm = TRUE),
    median = median(avg_s, na.rm = TRUE),
    quantile = quantile(avg_s, probs=c(0.25), na.rm = TRUE),
    quantile2 = quantile(avg_s, probs=c(0.75), na.rm = TRUE)
  )
group_by(Soil, Soil5yrCensor_dich) %>%
  summarise(
    count = n(),
    mean = mean(fvc_pct, na.rm = TRUE),
    sd = sd(fvc_pct, na.rm = TRUE),
    median = median(fvc_pct, na.rm = TRUE),
    quantile = quantile(fvc_pct, probs=c(0.25), na.rm = TRUE),
    quantile2 = quantile(fvc_pct, probs=c(0.75), na.rm = TRUE)
  )
group_by(Soil, Soil5yrCensor_dich) %>%
  summarise(
    count = n(),
    mean = mean(dlco_pct, na.rm = TRUE),
    sd = sd(dlco_pct, na.rm = TRUE),
    median = median(dlco_pct, na.rm = TRUE),
    quantile = quantile(dlco_pct, probs=c(0.25), na.rm = TRUE),
    quantile2 = quantile(dlco_pct, probs=c(0.75), na.rm = TRUE)
  )
group_by(Soil, Soil5yrCensor_dich) %>%
  summarise(
    count = n(),
    mean = mean(time_DeathTxCensor, na.rm = TRUE),
    sd = sd(time_DeathTxCensor, na.rm = TRUE),
    median = median(time_DeathTxCensor, na.rm = TRUE),
    quantile = quantile(time_DeathTxCensor, probs=c(0.25), na.rm = TRUE),
    quantile2 = quantile(time_DeathTxCensor, probs=c(0.75), na.rm = TRUE)
  )
```

### Soil_5yrPreDx Low vs High
```{r}
table(Soil$sex, Soil$Soil5yr_dich)
table(Soil$race, Soil$Soil5yr_dich)
table(Soil$dich_Race, Soil$Soil5yr_dich)
table(Soil$dx_group, Soil$Soil5yr_dich)
table(Soil$smokeHx, Soil$Soil5yr_dich)
table(Soil$metro, Soil$Soil5yr_dich)
table(Soil$province, Soil$Soil5yr_dich)
table(Soil$status, Soil$Soil5yr_dich)
group_by(Soil, Soil5yr_dich) %>%
  summarise(
    count = n(),
    mean = mean(age_dx, na.rm = TRUE),
    sd = sd(age_dx, na.rm = TRUE),
    median = median(age_dx, na.rm = TRUE),
    quantile = quantile(age_dx, probs=c(0.25), na.rm = TRUE),
    quantile2 = quantile(age_dx, probs=c(0.75), na.rm = TRUE)
  )
group_by(Soil, Soil5yr_dich) %>%
  summarise(
    count = n(),
    mean = mean(avg_s, na.rm = TRUE),
    sd = sd(avg_s, na.rm = TRUE),
    median = median(avg_s, na.rm = TRUE),
    quantile = quantile(avg_s, probs=c(0.25), na.rm = TRUE),
    quantile2 = quantile(avg_s, probs=c(0.75), na.rm = TRUE)
  )
group_by(Soil, Soil5yr_dich) %>%
  summarise(
    count = n(),
    mean = mean(fvc_pct, na.rm = TRUE),
    sd = sd(fvc_pct, na.rm = TRUE),
    median = median(fvc_pct, na.rm = TRUE),
    quantile = quantile(fvc_pct, probs=c(0.25), na.rm = TRUE),
    quantile2 = quantile(fvc_pct, probs=c(0.75), na.rm = TRUE)
  )
group_by(Soil, Soil5yr_dich) %>%
  summarise(
    count = n(),
    mean = mean(dlco_pct, na.rm = TRUE),
    sd = sd(dlco_pct, na.rm = TRUE),
    median = median(dlco_pct, na.rm = TRUE),
    quantile = quantile(dlco_pct, probs=c(0.25), na.rm = TRUE),
    quantile2 = quantile(dlco_pct, probs=c(0.75), na.rm = TRUE)
  )
group_by(Soil, Soil5yr_dich) %>%
  summarise(
    count = n(),
    mean = mean(time_DeathTxCensor, na.rm = TRUE),
    sd = sd(time_DeathTxCensor, na.rm = TRUE),
    median = median(time_DeathTxCensor, na.rm = TRUE),
    quantile = quantile(time_DeathTxCensor, probs=c(0.25), na.rm = TRUE),
    quantile2 = quantile(time_DeathTxCensor, probs=c(0.75), na.rm = TRUE)
  )
```


## Visual EDA
### Histograms
Histogram of Soil Exposure 5yr Pre-Death/Transplant/Censoring
```{r}
(Soil %>% ggplot(aes(x=Soil_5yrPreCensor))+
   geom_histogram(fill="blue", color="black")+
   labs(x="Average Soil Level 5yr Prior to Death/Transplant/Censoring", y="Number of patients with fILD", title="Average Soil 5yr Prior to Death/Transplant/Censoring")+
   theme(plot.title = element_text(hjust = 0.5))+
   xlim(0,10))
```


Histogram of Soil Exposure 5yr Pre-Dx
```{r}
(Soil %>% ggplot(aes(x=Soil_5yrPreDx))+
   geom_histogram(fill="blue", color="black")+
   labs(x="Average Soil Level 5yr Prior to Diagnosis", y="Number of patients with fILD", title="Average Soil 5yr Prior to Diagnosis")+
   theme(plot.title = element_text(hjust = 0.5))+
   xlim(0,10))
```


### Violin plots of Soil vs other demographic variables
Violin plot wrapping boxplot to visualize low vs high PM5yrCensor_dich vs CIMD
```{r}
(Soil %>% ggplot(aes(x=Soil5yrCensor_dich, y=avg_s, fill=Soil5yrCensor_dich))+
   geom_boxplot(width=0.2, color="black", alpha=1.0)+
   geom_violin(width=1.0, alpha=0.5)+
   labs(x="Low vs High Soil in 5yrs Pre-Censoring", y="National CIMD", title="National CIMD by Soil Exposure")+
   theme(plot.title = element_text(hjust = 0.5))+
   ylim(-2.5,2.5))
```

Violin plot wrapping boxplot to visualize sex vs Soil_5yrPreCensor
```{r}
(Soil %>% ggplot(aes(x=sex, y=Soil_5yrPreCensor, fill=sex))+
   geom_boxplot(width=0.2, color="black", alpha=1.0)+
   geom_violin(width=1.0, alpha=0.5)+
   labs(x="Sex", y="Average Soil 5 Years Pre-Censor", title="Soil Exposure Pre-Censor by Sex")+
   theme(plot.title = element_text(hjust = 0.5))+
   ylim(0,2))
```


Violin plot wrapping boxplot to visualize sex vs Soil_5yrPreDx
```{r}
(Soil %>% ggplot(aes(x=sex, y=Soil_5yrPreDx, fill=sex))+
   geom_boxplot(width=0.2, color="black", alpha=1.0)+
   geom_violin(width=1.0, alpha=0.5)+
   labs(x="Sex", y="Average Soil 5 Years Pre-Diagnosis", title="Soil Exposure Pre-Diagnosis by Sex")+
   theme(plot.title = element_text(hjust = 0.5))+
   ylim(0,2))
```

Violin plot wrapping boxplot to visualize dich_Race vs Soil_5yrPreCensor
```{r}
(Soil %>% ggplot(aes(x=dich_Race, y=Soil_5yrPreCensor, fill=dich_Race))+
   geom_boxplot(width=0.2, color="black", alpha=1.0)+
   geom_violin(width=1.0, alpha=0.5)+
   labs(x="Race", y="Average Soil 5 Years Pre-Censoring", title="Soil Exposure Pre-Censoring by Race")+
   theme(plot.title = element_text(hjust = 0.5))+
   ylim(0,2))
```


Violin plot wrapping boxplot to visualize race vs Soil_5yrPreCensor
```{r}
(Soil %>% ggplot(aes(x=race, y=Soil_5yrPreCensor, fill=race))+
   geom_boxplot(width=0.2, color="black", alpha=1.0)+
   geom_violin(width=1.0, alpha=0.5)+
   labs(x="Race", y="Average Soil 5 Years Pre-Censoring", title="Soil Exposure Pre-Censoring by Race")+
   theme(plot.title = element_text(hjust = 0.5))+
   ylim(0,2))
```

Violin plot wrapping boxplot to visualize dich_Race vs Soil_5yrPreDx
```{r}
(Soil %>% ggplot(aes(x=dich_Race, y=Soil_5yrPreDx, fill=dich_Race))+
   geom_boxplot(width=0.2, color="black", alpha=1.0)+
   geom_violin(width=1.0, alpha=0.5)+
   labs(x="Race", y="Average Soil 5 Years Pre-Diagnosis", title="Soil Exposure Pre-Diagnosis by Race")+
   theme(plot.title = element_text(hjust = 0.5))+
   ylim(0,2))
```

Violin plot wrapping boxplot to visualize race vs Soil_5yrPreDx
```{r}
(Soil %>% ggplot(aes(x=race, y=Soil_5yrPreDx, fill=race))+
   geom_boxplot(width=0.2, color="black", alpha=1.0)+
   geom_violin(width=1.0, alpha=0.5)+
   labs(x="Race", y="Average Soil 5 Years Pre-Diagnosis", title="Soil Exposure Pre-Diagnosis by Race")+
   theme(plot.title = element_text(hjust = 0.5))+
   ylim(0,2))
```

Violin plot wrapping boxplot to visualize metro vs Soil_5yrPreCensor
```{r}
(Soil %>% ggplot(aes(x=metro, y=Soil_5yrPreCensor, fill=metro))+
   geom_boxplot(width=0.2, color="black", alpha=1.0)+
   geom_violin(width=1.0, alpha=0.5)+
   labs(x="Residential Location", y="Average Soil in 5yrs Pre-Censoring", title="Soil Exposure in 5yrs Pre-Censoring by Residential Location")+
   theme(plot.title = element_text(hjust = 0.5))+
   ylim(0,2))
```


Violin plot wrapping boxplot to visualize metro vs Soil_5yrPreDx
```{r}
(Soil %>% ggplot(aes(x=metro, y=Soil_5yrPreDx, fill=metro))+
   geom_boxplot(width=0.2, color="black", alpha=1.0)+
   geom_violin(width=1.0, alpha=0.5)+
   labs(x="Residential Location", y="Average Soil in 5yrs Pre-Diagnosis", title="Soil Exposure in 5yrs Pre-Diagnosis by Residential Location")+
   theme(plot.title = element_text(hjust = 0.5))+
   ylim(0,2))
```

### Scatterplots of Soil over time
Scatterplot showing average Soil 5yr pre-death/transplant/censoring
```{r}
(Soil %>% ggplot(aes(x=dx_date, y=Soil_5yrPreCensor))+
   geom_point()+
   labs(x="Date of Diagnosis", y="Soil Exposure 5yr Prior Death/Transplant/Censoring", title="Soil 5yr Prior to Death/Transplant/Censoring vs Date of Diagnosis")+
   geom_smooth(method="lm", se = FALSE) +
   theme(plot.title = element_text(hjust = 0.5))+
   ylim(0,2))
```


```{r}
(Soil %>% ggplot(aes(x=dx_date, y=Soil_5yrPreCensor, color=site))+
   geom_point(alpha=0.5)+
   labs(x="Date of Diagnosis", y="Soil Exposure 5yr Prior Death/Transplant/Censoring", title="Soil 5yr Prior to Death/Transplant/Censoring vs Date of Diagnosis")+
   geom_smooth(method="lm", se = FALSE) +
   theme(plot.title = element_text(hjust = 0.5))+
   ylim(0,2))
```

Scatterplot showing average Soil during 5yrs pre-diagnosis level by year of diagnosis
```{r}
(Soil %>% ggplot(aes(x=dx_date, y=Soil_5yrPreDx))+
   geom_point()+
   labs(x="Date of Diagnosis", y="Soil Exposure During 5yrs Pre-Diagnosis", title="Pre-Diagnosis Soil vs Date of Diagnosis")+
   geom_smooth(method="lm", se = FALSE) +
   theme(plot.title = element_text(hjust = 0.5))+
   ylim(0,2)+
   scale_x_date(limits = as.Date(c("2000-01-01","2020-01-01"))))
```


Scatterplot showing average Soil during 5yrs pre-diagnosis level by year of diagnosis
```{r}
(Soil %>% ggplot(aes(x=dx_date, y=Soil_5yrPreDx, color=site))+
   geom_point(alpha=0.5)+
   labs(x="Date of Diagnosis", y="Soil Exposure During 5yrs Pre-Diagnosis", title="Pre-Diagnosis Soil vs Date of Diagnosis")+
   geom_smooth(method="lm", se = FALSE) +
   theme(plot.title = element_text(hjust = 0.5))+
   ylim(0,2)+
   scale_x_date(limits = as.Date(c("2000-01-01","2020-01-01"))))
```


# Cox PH Models for Association between Soil and Death or Lung Transplant
## Soil 5yrs Pre-Censoring Models
### Continuous Soil 5yrs pre-censoring until Death/Transplant/Censoring
Start with the simplest model just looking at Soil in the year of and 5yrs prior to death/transplant/censoring
```{r}
coxPH_model1 <- coxph(Surv(time_DeathTxCensor, deadORtx==1) ~ Soil_5yrPreCensor, data=Soil, id=ID)
summary(coxPH_model1)
```


Partial model
```{r}
coxPH_model2 <- coxph(Surv(time_DeathTxCensor, deadORtx==1) ~ Soil_5yrPreCensor + sex + age_dx + smokeHx + dich_Race + dx_IPF + avg_s + metro, data=Soil, id=ID)
summary(coxPH_model2)
```


Complete model
```{r}
coxPH_model3 <- coxph(Surv(time_DeathTxCensor, deadORtx==1) ~ Soil_5yrPreCensor + sex + age_dx + smokeHx + dich_Race + dx_IPF + avg_s + metro + fvc_pct + dlco_pct, data=Soil, id=ID)
summary(coxPH_model3)
```



### Smoothed HR Function for 5yr Pre-Censor Continuous
Base model
```{r}
coxPH_model1_tvc <- cox.tvc(coxPH_model1, "Soil_5yrPreCensor")
print(coxPH_model1_tvc)
plot(coxPH_model1_tvc)
```


Partial model (all covariates except lung function)
```{r}
coxPH_model2_tvc <- cox.tvc(coxPH_model2, "Soil_5yrPreCensor")
print(coxPH_model2_tvc)
plot(coxPH_model2_tvc)
```


```{r}
coxPH_model3_tvc <- cox.tvc(coxPH_model3, "Soil_5yrPreCensor")
print(coxPH_model3_tvc)
plot(coxPH_model3_tvc)
```


### Variable HR for Soil risk in 5yr Pre-Cenosr
Want a plot with x-axis of Soil level and y-axis of HR

Unadjusted model
```{r}
#First need to make dataframe that only includes patients with time_DeathTxCensor
Soilx <- Soil %>% filter(!is.na(Soil_5yrPreCensor) & !is.na(time_DeathTxCensor))

#Then make survival function
surv1 <- Surv(Soilx$time_DeathTxCensor, Soilx$deadORtx==1)
fit1 <- coxph(surv1 ~ pspline(Soilx$Soil_5yrPreCensor, df=4))
predicted <- predict(fit1, type="terms", se.fit=T, terms=1)

#Then plot
plot(Soilx$Soil_5yrPreCensor, exp(predicted$fit), type="n", xlim=c(0,3), ylim=c(0,2))
lines(sm.spline(Soilx$Soil_5yrPreCensor, exp(predicted$fit)), col = "red" , lty = 1 )
lines(sm.spline(Soilx$Soil_5yrPreCensor, exp(predicted$fit + 1.96 * predicted$se)), col = "orange" , lty = 2 )
lines(sm.spline(Soilx$Soil_5yrPreCensor, exp(predicted$fit - 1.96 * predicted$se)), col = "orange" , lty = 2 )
```

Partially-adjusted model
```{r}
#First need to make dataframe that only includes patients with time_DeathTxCensor
Soilx <- Soil %>% filter(!is.na(Soil_5yrPreCensor) & !is.na(time_DeathTxCensor) & !is.na(age_dx) & !is.na(sex) & !is.na(smokeHx) & !is.na(dich_Race) & !is.na(dx_IPF) & !is.na(avg_s) & !is.na(metro))

#Then survival function
surv1 <- Surv(Soilx$time_DeathTxCensor, Soilx$deadORtx==1)
fit1 <- coxph(surv1 ~ pspline(Soilx$Soil_5yrPreCensor, df=4) + Soilx$age_dx + Soilx$sex + Soilx$smokeHx + Soilx$dich_Race + Soilx$dx_IPF + Soilx$avg_s + Soilx$metro)
predicted <- predict(fit1, type="terms", se.fit=T, terms=1)

#Then plot
plot(Soilx$Soil_5yrPreCensor, exp(predicted$fit), type="n", xlim=c(0,3), ylim=c(0,2))
lines(sm.spline(Soilx$Soil_5yrPreCensor, exp(predicted$fit)), col = "red" , lty = 1 )
lines(sm.spline(Soilx$Soil_5yrPreCensor, exp(predicted$fit + 1.96 * predicted$se)), col = "orange" , lty = 2 )
lines(sm.spline(Soilx$Soil_5yrPreCensor, exp(predicted$fit - 1.96 * predicted$se)), col = "orange" , lty = 2 )
```

Fully-adjusted model
```{r}
#First need to make dataframe that only includes patients with time_DeathTxCensor
Soilx <- Soil %>% filter(!is.na(Soil_5yrPreCensor) & !is.na(time_DeathTxCensor) & !is.na(age_dx) & !is.na(sex) & !is.na(smokeHx) & !is.na(dich_Race) & !is.na(dx_IPF) & !is.na(avg_s) & !is.na(metro) & !is.na(fvc_pct) & !is.na(dlco_pct))

#Then survival function
surv1 <- Surv(Soilx$time_DeathTxCensor, Soilx$deadORtx==1)
fit1 <- coxph(surv1 ~ pspline(Soilx$Soil_5yrPreCensor, df=4) + Soilx$age_dx + Soilx$sex + Soilx$smokeHx + Soilx$dich_Race + Soilx$dx_IPF + Soilx$avg_s + Soilx$metro + Soilx$fvc_pct + Soilx$dlco_pct)
predicted <- predict(fit1, type="terms", se.fit=T, terms=1)

#Then plot
plot(Soilx$Soil_5yrPreCensor, exp(predicted$fit), type="n", xlim=c(0,3), ylim=c(0,2))
lines(sm.spline(Soilx$Soil_5yrPreCensor, exp(predicted$fit)), col = "red" , lty = 1 )
lines(sm.spline(Soilx$Soil_5yrPreCensor, exp(predicted$fit + 1.96 * predicted$se)), col = "orange" , lty = 2 )
lines(sm.spline(Soilx$Soil_5yrPreCensor, exp(predicted$fit - 1.96 * predicted$se)), col = "orange" , lty = 2 )
```


### Soil Low vs High 5yrs Pre-Censoring
Base Model:
```{r}
coxPH_model7 <- coxph(Surv(time_DeathTxCensor, deadORtx==1) ~ Soil5yrCensor_dich, data=Soil, id=ID)
summary(coxPH_model7)
```


Partial model
```{r}
coxPH_model8 <- coxph(Surv(time_DeathTxCensor, deadORtx==1) ~ Soil5yrCensor_dich + sex + age_dx + smokeHx + dich_Race + avg_s + dx_IPF + metro, data=Soil, id=ID)
summary(coxPH_model8)
```


Complete model
```{r}
coxPH_model9 <- coxph(Surv(time_DeathTxCensor, deadORtx==1) ~ Soil5yrCensor_dich + sex + age_dx + smokeHx + dich_Race + avg_s + dx_IPF + metro + fvc_pct + dlco_pct, data=Soil, id=ID)
summary(coxPH_model9)
```


Now I will make a Kaplan-Meier survival curve, discretized by Soil during disease deciles:
```{r, fig.height=5, fig.width=7}
surv1 <- survfit(Surv(time_DeathTxCensor, deadORtx==1) ~ Soil5yrCensor_dich, data=Soil)
(ggsurvplot(surv1, 
            palette=c("dodgerblue4","firebrick4"),
            censor=TRUE,
            pval=TRUE,
            xlim=c(0,10),
            legend="right",
           data=Soil,
           legend.title="Soil Exposure",
           legend.labs=c("Low", "High"),
           risk.table=T,
           title="Time to Death or Transplant by Soil 5yrs Pre-Censoring",
           xlab="Time (years)",
           ylab="Survival Probability (Death or Lung Transplant)",
           ggtheme = theme_grey()
           ))
```


Now we want to check the proportionality assumption by plotting the scaled Schoenfield residuals against time for each covariate in our final model.
```{r}
residtest <- cox.zph(coxPH_model9)
residtest
ggcoxzph(residtest)
```


Now I will present a forest plot with the model
```{r}
ggforest(coxPH_model9)
```



# Association between Soil and Baseline FVC
## Visual EDA 5yrs pre-Dx
Scatterplot of continuous Soil 5yrs pre-diagnosis vs FVC, with sex represented by the color of points
```{r}
(Soil %>% ggplot(aes(x=Soil_5yrPreDx, y=fvc_pct, color=sex))+
   geom_point()+
   labs(x="Soil 5yrs Pre-Diagnosis", y="Baseline FVC % Predicted", title="Baseline FVC % Predicted by Soil 5yrs Pre-Diagnosis")+
   geom_smooth(method="lm", se = FALSE) +
   theme(plot.title = element_text(hjust = 0.5))+
   xlim(0,1))
```


Violin plot wrapping boxplot to visualize Soil 5yrs pre-dx low vs high vs FVC
```{r}
Soilb <- Soil %>% filter(!is.na(Soil5yr_dich))
(Soilb %>% ggplot(aes(x=Soil5yr_dich, y=fvc_pct, fill=Soil5yr_dich))+
   geom_boxplot(width=0.2, color="black", alpha=1.0)+
   geom_violin(width=1.0, alpha=0.5)+
   labs(x="Soil 5yrs Pre-Diagnosis Low vs High", y="Baseline FVC % Predicted", title="Baseline FVC % Predicted by Soil 5yrs Pre-Diagnosis Low vs High")+
   theme_light()+
   theme(legend.position = "none", plot.title = element_text(hjust = 0.5))+
   scale_fill_brewer(type="seq", palette="YlOrRd"))
```



## Soil 5yrs Pre-Diagnosis and FVC Models
Looking at continuous Soil_5yrPreDx vs baseline FVC

Simplest model just looking at Soil in 5yrs pre-diagnosis
```{r}
FVC_model1 <- lm(fvc_pct ~ Soil_5yrPreDx, data=Soil)
summary(FVC_model1)
confint(FVC_model1)
```


Partial model
```{r}
FVC_model2 <- lm(fvc_pct ~ Soil_5yrPreDx + sex + age_dx + smokeHx, data=Soil)
summary(FVC_model2)
confint(FVC_model2)
```


Complete model
```{r}
FVC_model3 <- lm(fvc_pct ~ Soil_5yrPreDx + sex + age_dx + smokeHx + dich_Race + dx_IPF + avg_s + metro, data=Soil)
summary(FVC_model3)
confint(FVC_model3)
```


### Soil 5yrs Pre-Diagnosis Low vs High and Baseline FVC
Looking at Soil5yr_dich vs baseline FVC

Simplest model just looking at Soil in 5yrs pre-diagnosis
```{r}
FVC_model1 <- lm(fvc_pct ~ Soil5yr_dich, data=Soil)
summary(FVC_model1)
confint(FVC_model1)
```


Partial model
```{r}
FVC_model2 <- lm(fvc_pct ~ Soil5yr_dich + sex + age_dx + smokeHx, data=Soil)
summary(FVC_model2)
confint(FVC_model2)
```


Complete model
```{r}
FVC_model3 <- lm(fvc_pct ~ Soil5yr_dich + sex + age_dx + smokeHx + dich_Race + dx_IPF + avg_s + metro, data=Soil)
summary(FVC_model3)
confint(FVC_model3)
```



# Association between Soil and Baseline DLCO
## Visual EDA 5yrs pre-Dx
Scatterplot of continuous Soil 5yrs pre-diagnosis vs DLCO, with sex represented by the color of points
```{r}
(Soil %>% ggplot(aes(x=Soil_5yrPreDx, y=dlco_pct, color=sex))+
   geom_point()+
   labs(x="Soil 5yrs Pre-Diagnosis", y="Baseline dlco % Predicted", title="Baseline dlco % Predicted by Soil 5yrs Pre-Diagnosis")+
   geom_smooth(method="lm", se = FALSE) +
   theme(plot.title = element_text(hjust = 0.5))+
   xlim(0,1))
```


Violin plot wrapping boxplot to visualize Soil 5yrs pre-dx low vs high vs dlco
```{r}
(Soilb %>% ggplot(aes(x=Soil5yr_dich, y=dlco_pct, fill=Soil5yr_dich))+
   geom_boxplot(width=0.2, color="black", alpha=1.0)+
   geom_violin(width=1.0, alpha=0.5)+
   labs(x="Soil 5yrs Pre-Diagnosis Low vs High", y="Baseline DLCO % Predicted", title="Baseline DLCO % Predicted by Soil 5yrs Pre-Diagnosis Low vs High")+
   theme_light()+
   theme(legend.position = "none", plot.title = element_text(hjust = 0.5))+
   scale_fill_brewer(type="seq", palette="YlOrRd"))
```



## Soil 5yrs Pre-Diagnosis and DLCO Models
### Soil 5yrs Pre-Diagnosis and Baseline DLCO
Looking at continuous Soil_5yrPreDx vs baseline DLCO

Simplest model just looking at Soil in 5yrs pre-diagnosis
```{r}
DLCO_model1 <- lm(dlco_pct ~ Soil_5yrPreDx, data=Soil)
summary(DLCO_model1)
confint(DLCO_model1)
```


Partial model
```{r}
DLCO_model2 <- lm(dlco_pct ~ Soil_5yrPreDx + sex + age_dx + smokeHx, data=Soil)
summary(DLCO_model2)
confint(DLCO_model2)
```


Complete model
```{r}
DLCO_model3 <- lm(dlco_pct ~ Soil_5yrPreDx + sex + age_dx + smokeHx + dich_Race + dx_IPF + avg_s + metro, data=Soil)
summary(DLCO_model3)
confint(DLCO_model3)
```



### Soil 5yrs Pre-Diagnosis Low vs High and Baseline DLCO
Looking at Soil5yr_dich vs baseline DLCO

Simplest model just looking at Soil in 5yrs pre-diagnosis
```{r}
DLCO_model1 <- lm(dlco_pct ~ Soil5yr_dich, data=Soil)
summary(DLCO_model1)
confint(DLCO_model1)
```


Partial model
```{r}
DLCO_model2 <- lm(dlco_pct ~ Soil5yr_dich + sex + age_dx + smokeHx, data=Soil)
summary(DLCO_model2)
confint(DLCO_model2)
```


Complete model
```{r}
DLCO_model3 <- lm(dlco_pct ~ Soil5yr_dich + sex + age_dx + smokeHx + dich_Race + dx_IPF + avg_s + metro, data=Soil)
summary(DLCO_model3)
confint(DLCO_model3)
```


