---
title: "Soil and fibrotic ILD PFF Complete Cohort (Consent Date Reference) Lung Function Decline Analysis - Revision #1 Final Analysis"
author: "Gillian Goobie"
date: "08/02/2022"
output: 
  bookdown::html_document2:
    toc: yes
    toc_float: yes
    number_sections: yes
  html_document: 
    toc: yes
    toc_float: yes
    number_sections: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE, warning=FALSE, echo=F}
library(tidyverse)
library(readxl)
library(here)
library(nnet)
library(writexl) 
library(knitr)
library(lintr)
library(psych)
library(RColorBrewer)
library(reticulate)
library(survival)
library(survminer)
library(cmprsk)
library(riskRegression)
library(prodlim)
library(lme4)
library(performance)
library(psycho)
library(report)
library(rms)
library(splines)
library(Greg)
```


# Uploading PFT Dataframes
```{r}
#Need to import the correct sheet from the original PFF file
outfile1 <- here("PFF_FormatData_2021_10_06.xlsx")
FVC <- read_excel(outfile1, sheet="FVC", na="")

DLCO <- read_excel(outfile1, sheet="DLCO", na="")
```

## Correct formatting for FVC dataframe
```{r}
FVC <- FVC %>% dplyr::select(SSID, TestDays, pctFVC)
str(FVC)
```

### Create FVC ID variable
This will make it easier to identify any cells with issues and replace them if needed
```{r}
FVC <- FVC %>% mutate(fvc_id=1:9118)
```

## Correct formatting for DLCO dataframe
```{r}
DLCO <- DLCO %>% dplyr::select(SSID, TestDays, pctDLCOuncorr)
str(DLCO)
```

### Create DLCO ID variable
This will make it easier to identify any cells with issues and replace them if needed
```{r}
DLCO <- DLCO %>% mutate(dlco_id=1:7643)
```



# Uploading Soil Dataframe
Need to upload the file which only includes patients with fibrotic ILD
```{r}
outfile2 <- here("Final_PFF_fILD_Soil_BaselineData_2022_02_12.xlsx")
Soil <- read_excel(outfile2, na="")
```

Next we need to ensure that all variables in the Soil dataframe are correctly named and in the right class format.

## Correct date format
```{r}
Soil <- Soil %>% mutate_at(vars(dx_date, consent_date, censor_date, tx_date, death_date, deathORtx_date, DeathTxCensor_date, sample_date, fvc_date, dlco_date, Soil_date), as.Date, format="%m/%d/%Y")
str(Soil)
```

## Correct Factor Variables and Levels
```{r}
Soil <- Soil %>% mutate_at(vars(status, sex, race, ethnicity, deadORtx, dx, smokeHx, dich_Race, dx_group, dx_IPF, metro, Soil5yrCensor_dich, Soil5yr_dich), as.factor)
str(Soil)
```

## Releveling Factors
Here I am releveling factors so that they are in an intuitive order for my later analyses.
```{r}
Soil$sex <- fct_relevel(Soil$sex, c("Male","Female"))
Soil$race <- fct_relevel(Soil$race, c("W","B","A","I","U"))
Soil$ethnicity <- fct_relevel(Soil$ethnicity, c("N","H","U"))
Soil$dich_Race <- fct_relevel(Soil$dich_Race, c("White","Non-White"))
Soil$smokeHx <- fct_relevel(Soil$smokeHx, c("Never"))
Soil$Soil5yrCensor_dich <- fct_relevel(Soil$Soil5yrCensor_dich, c("Low", "High"))
Soil$Soil5yr_dich <- fct_relevel(Soil$Soil5yr_dich, c("Low", "High"))

#For dx and dx_group, I just want IPF to be first and then the rest of the categories are alphabetical
Soil$dx <- fct_relevel(Soil$dx, c("IPF"))
Soil$dx_group <- fct_relevel(Soil$dx_group, c("IPF"))
str(Soil)
```

## Remove UPitt patients
```{r}
Soil <- Soil %>% filter(!str_detect(SSID, "^08R"))
```
This now takes us down to  observations after we remove UPitt patients which may be doubled up in our Simmons analyses


# Merge FVC and DLCO Dataframes with Soil dataframe
```{r}
pft <- left_join(FVC, Soil, by=c("SSID"))
pft <- pft %>% rename("first_fvc_date"="fvc_date", "first_dlco_date"="dlco_date", "first_fvc_pct"="fvc_pct", "first_dlco_pct"="dlco_pct", "fvc_pct"="pctFVC", "TestDays"="TestDays")
```

```{r}
pft <- pft %>% mutate(fvc_date = (consent_date + TestDays))
```

```{r}
DLCO <- left_join(DLCO, Soil, by=c("SSID"))
DLCO <- DLCO %>% rename("first_fvc_date"="fvc_date", "first_fvc_pct"="fvc_pct", "first_dlco_pct"="dlco_pct", "first_dlco_date"="dlco_date", "dlco_pct"="pctDLCOuncorr")
```

```{r}
DLCO <- DLCO %>% mutate(dlco_date = (consent_date + TestDays))
str(DLCO)
```

```{r}
DLCO <- DLCO %>% dplyr::select(ID, SSID, TestDays, dlco_id, age_dx, sex, dich_Race, smokeHx, metro, site, pct_belowpoverty, dx_IPF, dx, dx_group, dlco_pct, first_fvc_pct, first_dlco_pct, Soil_5yrPreDx, Soil5yr_dich, Soil_5yrPreCensor, Soil5yrCensor_dich, consent_date, dlco_date, first_fvc_date, first_dlco_date, tx_date, death_date, DeathTxCensor_date, dx_yr)
str(DLCO)
```

```{r}
pft <- pft %>% dplyr::select(ID, SSID, fvc_id, TestDays, age_dx, sex, dich_Race, smokeHx, metro, site, pct_belowpoverty, dx_IPF, dx, dx_group, fvc_pct, first_fvc_pct, first_dlco_pct, Soil_5yrPreDx, Soil5yr_dich, Soil_5yrPreCensor, Soil5yrCensor_dich, consent_date, fvc_date, first_fvc_date, first_dlco_date, tx_date, death_date, DeathTxCensor_date, dx_yr)
str(pft)
```

```{r}
pft <- full_join(pft, DLCO, by=c("ID", "SSID", "TestDays", "age_dx", "sex", "dich_Race", "smokeHx", "metro", "site", "pct_belowpoverty", "dx_IPF", "dx", "dx_group", "first_fvc_pct", "first_dlco_pct", "Soil_5yrPreDx", "Soil5yr_dich", "Soil_5yrPreCensor", "Soil5yrCensor_dich", "consent_date", "first_fvc_date", "first_dlco_date", "tx_date", "death_date", "DeathTxCensor_date", "dx_yr"))
```


```{r}
pft <- pft %>% mutate(pft_id=1:9320)
pft$ID <- as.factor(pft$ID)
str(pft)
```


# Remove PFTs that occurred >6mo before consent_date
```{r}
pft <-  pft %>% mutate(fvc_timefromcensor=(fvc_date-consent_date)/365.25)
pft$fvc_timefromcensor <-  as.numeric(pft$fvc_timefromcensor)
summary(pft$fvc_timefromcensor)
pft <- pft %>% filter(fvc_timefromcensor>=-0.5)
summary(pft$fvc_timefromcensor)
```
This takes us from  pfts to  pfts

Next need to make dlco_timefromcensor variable
```{r}
pft <-  pft %>% mutate(dlco_timefromcensor=(dlco_date-consent_date)/365.25)
pft$dlco_timefromcensor <-  as.numeric(pft$dlco_timefromcensor)
summary(pft$dlco_timefromcensor)
```
And we can see that there are none that occurred >6mo pre-consent, so no more to exclude.

# Remove PFTs that occurred after transplant
Then will make a simpler dataframe that can be joined to pft
```{r}
txed <- Soil %>% filter(status==2) %>% dplyr::select(ID, tx_date)
```
We have  patients out of the  who were transplanted

Next will join txed to pft
```{r}
str(pft)
pft$ID <- as.numeric(pft$ID)
pft <- left_join(pft, txed, by=c("ID", "tx_date"))

#remove all pfts where pft_date>tx_date
pft <- pft %>% filter(is.na(tx_date) | fvc_date<tx_date)
```
pft data frame goes from  to  rows.

# Remove PFT Values that Don't Make Sense
There are some DLCO values that are too low that we need to remove. All the extremely high DLCOs have been removed that are incorrect and the remaining ones make sense.
```{r}
summary(pft$fvc_pct)
#FVCs look appropriate
```

```{r}
summary(pft$dlco_pct)
#Should remove DLCOs <5

pft <- pft %>% filter(is.na(dlco_pct) | dlco_pct>=5)
summary(pft$dlco_pct)
```
This takes us from to  records.

## Create Time from First PFT Variable
```{r}
pft <- pft %>% mutate(time_firstFVC=(fvc_date-first_fvc_date)/365.25)
pft$time_firstFVC <- as.numeric(pft$time_firstFVC)

pft <- pft %>% mutate(time_firstDLCO=(dlco_date-first_dlco_date)/365.25)
pft$time_firstDLCO <- as.numeric(pft$time_firstDLCO)
```

## Creating a dx_yr column
```{r}
Soil <- Soil %>% 
  mutate(dx_yr = format(as.Date(Soil$dx_date, format="%Y-%m-%d"),"%Y"))
Soil$dx_yr <- as.numeric(Soil$dx_yr)
```


## Create dataframe that excludes patients which only had one PFT
```{r}
# Determine how many PFTs each patient has
PFT_number <- pft %>% count(ID)
summary(PFT_number)
# minimum number of PFTs is 1, so we do have some patients that only had one PFT that are included in later analyses

# restrict dataframe to only include patients with one PFT
one_pft <- PFT_number %>% filter(n==1)
#144 patients only had one PFT, so we need to ensure the analyses below remain consistent when we remove patients that only have one PFT

# remove patients with only one PFT
pft_over1 <- pft[!pft$ID %in% one_pft$ID,]
#Now my pft_over1 dataframe drops from 7195 rows to 7051 rows, which is a loss of 144 individuals who only had 1 pft

#Recheck using same methods as above
PFT_number <- pft_over1 %>% count(ID)
summary(PFT_number)
#And we can see that the minimum is now 2
```



# Exploratory Data Analysis
## Visualization of Relationship between Soil and FVC Decline
5yrs pre-censoring low vs high
```{r}
#First I need to make a new dataframe that removes the Soildx_quart NAs so they don't show up in the plot
pft1 <- pft %>% filter(!is.na(Soil5yrCensor_dich))

#Then I can plot it   
(ggplot(pft1, aes(x=time_firstFVC, y=fvc_pct, group=ID)) +
   geom_line(aes(color=Soil5yrCensor_dich), alpha=0.6)+
   xlim(0,5)+
   stat_smooth(aes(group=1))+
   facet_grid(. ~Soil5yrCensor_dich)+
   labs(x="Time from First PFT (years)", y="Change in FVC % Predicted", title="Change in FVC % Predicted by Low vs High Soil 5yrs Pre-Censoring", color="Soil 5yrs \nPre-Censoring")+
   theme(plot.title = element_text(hjust = 0.5))+
   scale_color_brewer(type="seq", palette="YlOrRd"))
```



## Visualization of Relationship between Soil and DLCO Decline
Now the 5yrs pre-censoring low vs high
```{r}
#First I need to make a new dataframe that removes the Soildx_quart NAs so they don't show up in the plot
pft1 <- pft %>% filter(!is.na(Soil5yrCensor_dich))

#Then I can plot it   
(ggplot(pft1, aes(x=time_firstDLCO, y=dlco_pct, group=ID)) +
   geom_line(aes(color=Soil5yrCensor_dich), alpha=0.6)+
   xlim(0,5)+
    ylim(0,120)+
   stat_smooth(aes(group=1))+
   facet_grid(. ~Soil5yrCensor_dich)+
   labs(x="Time from First PFT (years)", y="Change in DLCO % Predicted", title="Change in DLCO % Predicted by Low vs High Soil 5yrs Pre-Censoring", color="Soil 5yrs \nPre-Censoring")+
   theme(plot.title = element_text(hjust = 0.5))+
   scale_color_brewer(type="seq", palette="YlOrRd"))
```


## Descriptive Stats for FVC Decline
```{r}
#Look at the mean and total variance of fvc_pct
pft %>% summarise(mean_fvc=mean(fvc_pct, na.rm=T))
tot_var_fvc <- pft %>% dplyr::select(fvc_pct) %>% summarise(var_fvc_pct=var(fvc_pct, na.rm=T))
tot_var_fvc

#Look at the between subject variance
bw_subj_var_fvc <- pft %>% group_by(ID) %>% mutate(mean_fvc_pct=mean(fvc_pct)) %>% ungroup() %>% summarise(bw_subj_var_fvc=var(mean_fvc_pct, na.rm=T))
bw_subj_var_fvc

fvc_ICC <- bw_subj_var_fvc/tot_var_fvc
fvc_ICC
```
fvc_ICC=0.89 indicates that ~89% of the variance is between subject variance. I.e. there is more difference in fvc_pct between patients than within any single patient.

## Number of PFTs
```{r}
n_pfts <- pft %>% count(ID)
n_pfts
summary(n_pfts$n)
```
This shows that the median # of PFTs is 5 and the mean is 6 (similar to Simmons)

# Association between Soil 5yrs Pre-Diagnosis and FVC Decline
Linear mixed models with random intercept and slope will be used to evaluate the impact of Soil During Disease on FVC decline.
## Continuous Soil 5yrs Pre-Diagnosis and FVC Models with Random Intercept and Random Slope
### Base Model
```{r}
FVC_lmer1 <- lmer(fvc_pct ~ 1 + time_firstFVC*Soil_5yrPreDx + (1 + time_firstFVC| ID), data=pft)
summary(FVC_lmer1)
icc(FVC_lmer1)
report(FVC_lmer1)
```

### Complete Model
```{r}
FVC_lmer2 <- lmer(fvc_pct ~ 1 + time_firstFVC*Soil_5yrPreDx + time_firstFVC*dich_Race + time_firstFVC*sex + time_firstFVC*age_dx + time_firstFVC*smokeHx + time_firstFVC*pct_belowpoverty + time_firstFVC*site +(1 + time_firstFVC | ID), data=pft)
summary(FVC_lmer2)
icc(FVC_lmer2)
report(FVC_lmer2)
```

### Complete Model + dx_yr
```{r}
FVC_lmer3 <- lmer(fvc_pct ~ 1 + time_firstFVC*Soil_5yrPreDx + time_firstFVC*dich_Race + time_firstFVC*sex + time_firstFVC*age_dx + time_firstFVC*smokeHx + time_firstFVC*pct_belowpoverty + time_firstFVC*site +time_firstFVC*dx_yr + (1 + time_firstFVC | ID), data=pft)
summary(FVC_lmer3)
icc(FVC_lmer3)
report(FVC_lmer3)
```


## Dichotomized Soil 5yrs Pre-Diagnosis and FVC Models with Random Intercept and Random Slope
### Base Model
```{r}
FVC_lmer1 <- lmer(fvc_pct ~ 1 + time_firstFVC*Soil5yr_dich + (1 + time_firstFVC| ID), data=pft)
summary(FVC_lmer1)
icc(FVC_lmer1)
report(FVC_lmer1)
```

### Complete Model
```{r}
FVC_lmer2 <- lmer(fvc_pct ~ 1 + time_firstFVC*Soil5yr_dich + time_firstFVC*dich_Race + time_firstFVC*sex + time_firstFVC*age_dx + time_firstFVC*smokeHx + time_firstFVC*pct_belowpoverty + time_firstFVC*site +(1 + time_firstFVC | ID), data=pft)
summary(FVC_lmer2)
icc(FVC_lmer2)
report(FVC_lmer2)
```

### Complete Model + dx_yr
```{r}
FVC_lmer3 <- lmer(fvc_pct ~ 1 + time_firstFVC*Soil5yr_dich + time_firstFVC*dich_Race + time_firstFVC*sex + time_firstFVC*age_dx + time_firstFVC*smokeHx + time_firstFVC*pct_belowpoverty + time_firstFVC*site +time_firstFVC*dx_yr + (1 + time_firstFVC | ID), data=pft)
summary(FVC_lmer3)
icc(FVC_lmer3)
report(FVC_lmer3)
```




# Association between Soil 5yrs Pre-Censoring and FVC Decline
Linear mixed models with random intercept and slope will be used to evaluate the impact of Soil During Disease on FVC decline.

## Continuous Soil 5yrs Pre-Censoring and FVC Models with Random Intercept and Random Slope
### Base Model
```{r}
FVC_lmer1 <- lmer(fvc_pct ~ 1 + time_firstFVC*Soil_5yrPreCensor + (1 + time_firstFVC| ID), data=pft)
summary(FVC_lmer1)
icc(FVC_lmer1)
report(FVC_lmer1)
```


### Complete Model
```{r}
FVC_lmer2 <- lmer(fvc_pct ~ 1 + time_firstFVC*Soil_5yrPreCensor + time_firstFVC*dich_Race + time_firstFVC*sex + time_firstFVC*age_dx + time_firstFVC*smokeHx + time_firstFVC*pct_belowpoverty + time_firstFVC*site +(1 + time_firstFVC| ID), data=pft)
summary(FVC_lmer2)
icc(FVC_lmer2)
report(FVC_lmer2)
```

### Complete Model + dx_yr
```{r}
FVC_lmer3 <- lmer(fvc_pct ~ 1 + time_firstFVC*Soil_5yrPreCensor + time_firstFVC*dich_Race + time_firstFVC*sex + time_firstFVC*age_dx + time_firstFVC*smokeHx + time_firstFVC*pct_belowpoverty + time_firstFVC*site +time_firstFVC*dx_yr + (1 + time_firstFVC| ID), data=pft)
summary(FVC_lmer3)
icc(FVC_lmer3)
report(FVC_lmer3)
```


## Dichotomized Soil 5yrs Pre-Censoring and FVC Models with Random Intercept and Random Slope
### Base Model
```{r}
FVC_lmer1 <- lmer(fvc_pct ~ 1 + time_firstFVC*Soil5yrCensor_dich + (1 + time_firstFVC| ID), data=pft)
summary(FVC_lmer1)
icc(FVC_lmer1)
report(FVC_lmer1)
```

### Complete Model
```{r}
FVC_lmer2 <- lmer(fvc_pct ~ 1 + time_firstFVC*Soil5yrCensor_dich + time_firstFVC*dich_Race + time_firstFVC*sex + time_firstFVC*age_dx + time_firstFVC*smokeHx + time_firstFVC*pct_belowpoverty + time_firstFVC*site +(1 + time_firstFVC| ID), data=pft)
summary(FVC_lmer2)
icc(FVC_lmer2)
report(FVC_lmer2)
```

### Complete Model + dx_yr
```{r}
FVC_lmer3 <- lmer(fvc_pct ~ 1 + time_firstFVC*Soil5yrCensor_dich + time_firstFVC*dich_Race + time_firstFVC*sex + time_firstFVC*age_dx + time_firstFVC*smokeHx + time_firstFVC*pct_belowpoverty + time_firstFVC*site +time_firstFVC*dx_yr + (1 + time_firstFVC| ID), data=pft)
summary(FVC_lmer3)
icc(FVC_lmer3)
report(FVC_lmer3)
```


# Association between Soil 5yrs Pre-Diagnosis and DLCO Decline
Linear mixed models with random intercept and slope will be used to evaluate the impact of Soil During Disease on dlco decline.
## Continuous Soil 5yrs Pre-Diagnosis and DLCO Models with Random Intercept and Random Slope
### Base Model
```{r}
DLCO_lmer1 <- lmer(dlco_pct ~ 1 + time_firstDLCO*Soil_5yrPreDx + (1 + time_firstDLCO| ID), data=pft)
summary(DLCO_lmer1)
icc(DLCO_lmer1)
report(DLCO_lmer1)
```

### Complete Model
```{r}
DLCO_lmer2 <- lmer(dlco_pct ~ 1 + time_firstDLCO*Soil_5yrPreDx + time_firstDLCO*dich_Race + time_firstDLCO*sex + time_firstDLCO*age_dx + time_firstDLCO*smokeHx + time_firstDLCO*pct_belowpoverty + time_firstDLCO*site +(1 + time_firstDLCO | ID), data=pft)
summary(DLCO_lmer2)
icc(DLCO_lmer2)
report(DLCO_lmer2)
```

### Complete Model + dx_yr
```{r}
DLCO_lmer3 <- lmer(dlco_pct ~ 1 + time_firstDLCO*Soil_5yrPreDx + time_firstDLCO*dich_Race + time_firstDLCO*sex + time_firstDLCO*age_dx + time_firstDLCO*smokeHx + time_firstDLCO*pct_belowpoverty + time_firstDLCO*site +time_firstDLCO*dx_yr + (1 + time_firstDLCO | ID), data=pft)
summary(DLCO_lmer3)
icc(DLCO_lmer3)
report(DLCO_lmer3)
```


## Dichotomized Soil 5yrs Pre-Diagnosis and DLCO Models with Random Intercept and Random Slope
### Base Model
```{r}
DLCO_lmer1 <- lmer(dlco_pct ~ 1 + time_firstDLCO*Soil5yr_dich + (1 + time_firstDLCO| ID), data=pft)
summary(DLCO_lmer1)
icc(DLCO_lmer1)
report(DLCO_lmer1)
```

### Complete Model
```{r}
DLCO_lmer2 <- lmer(dlco_pct ~ 1 + time_firstDLCO*Soil5yr_dich + time_firstDLCO*dich_Race + time_firstDLCO*sex + time_firstDLCO*age_dx + time_firstDLCO*smokeHx + time_firstDLCO*pct_belowpoverty + time_firstDLCO*site +(1 + time_firstDLCO | ID), data=pft)
summary(DLCO_lmer2)
icc(DLCO_lmer2)
report(DLCO_lmer2)
```

### Complete Model + dx_yr
```{r}
DLCO_lmer3 <- lmer(dlco_pct ~ 1 + time_firstDLCO*Soil5yr_dich + time_firstDLCO*dich_Race + time_firstDLCO*sex + time_firstDLCO*age_dx + time_firstDLCO*smokeHx + time_firstDLCO*pct_belowpoverty + time_firstDLCO*site +time_firstDLCO*dx_yr + (1 + time_firstDLCO | ID), data=pft)
summary(DLCO_lmer3)
icc(DLCO_lmer3)
report(DLCO_lmer3)
```




# Association between Soil 5yrs Pre-Censoring and DLCO Decline
Linear mixed models with random intercept and slope will be used to evaluate the impact of Soil During Disease on dlco decline.
## Continuous Soil 5yrs Pre-Censoring and DLCO Models with Random Intercept and Random Slope
### Base Model
```{r}
DLCO_lmer1 <- lmer(dlco_pct ~ 1 + time_firstDLCO*Soil_5yrPreCensor + (1 + time_firstDLCO| ID), data=pft)
summary(DLCO_lmer1)
icc(DLCO_lmer1)
report(DLCO_lmer1)
```

### Complete Model
```{r}
DLCO_lmer2 <- lmer(dlco_pct ~ 1 + time_firstDLCO*Soil_5yrPreCensor + time_firstDLCO*dich_Race + time_firstDLCO*sex + time_firstDLCO*age_dx + time_firstDLCO*smokeHx + time_firstDLCO*pct_belowpoverty + time_firstDLCO*site +(1 + time_firstDLCO | ID), data=pft)
summary(DLCO_lmer2)
icc(DLCO_lmer2)
report(DLCO_lmer2)
```

### Complete Model + dx_yr
```{r}
DLCO_lmer3 <- lmer(dlco_pct ~ 1 + time_firstDLCO*Soil_5yrPreCensor + time_firstDLCO*dich_Race + time_firstDLCO*sex + time_firstDLCO*age_dx + time_firstDLCO*smokeHx + time_firstDLCO*pct_belowpoverty + time_firstDLCO*site +time_firstDLCO*dx_yr + (1 + time_firstDLCO | ID), data=pft)
summary(DLCO_lmer3)
icc(DLCO_lmer3)
report(DLCO_lmer3)
```


## Dichotomized Soil 5yrs Pre-Censoring and DLCO Models with Random Intercept and Random Slope
### Base Model
```{r}
DLCO_lmer1 <- lmer(dlco_pct ~ 1 + time_firstDLCO*Soil5yrCensor_dich + (1 + time_firstDLCO| ID), data=pft)
summary(DLCO_lmer1)
icc(DLCO_lmer1)
report(DLCO_lmer1)
```

### Complete Model
```{r}
DLCO_lmer2 <- lmer(dlco_pct ~ 1 + time_firstDLCO*Soil5yrCensor_dich + time_firstDLCO*dich_Race + time_firstDLCO*sex + time_firstDLCO*age_dx + time_firstDLCO*smokeHx + time_firstDLCO*pct_belowpoverty + time_firstDLCO*site +(1 + time_firstDLCO | ID), data=pft)
summary(DLCO_lmer2)
icc(DLCO_lmer2)
report(DLCO_lmer2)
```

### Complete Model + dx_yr
```{r}
DLCO_lmer3 <- lmer(dlco_pct ~ 1 + time_firstDLCO*Soil5yrCensor_dich + time_firstDLCO*dich_Race + time_firstDLCO*sex + time_firstDLCO*age_dx + time_firstDLCO*smokeHx + time_firstDLCO*pct_belowpoverty + time_firstDLCO*site +time_firstDLCO*dx_yr + (1 + time_firstDLCO | ID), data=pft)
summary(DLCO_lmer3)
icc(DLCO_lmer3)
report(DLCO_lmer3)
```

# Making IPF-Only Cohort
```{r}
pft_IPF <- pft %>% filter(dx=="IPF")
```

# IPF-Only - Association between Soil 5yrs Pre-Diagnosis and FVC Decline
Linear mixed models with random intercept and slope will be used to evaluate the impact of Soil During Disease on FVC decline.

## IPF-Only - Continuous Soil 5yrs Pre-Diagnosis and FVC Models with Random Intercept and Random Slope
### IPF-Only - Base Model
```{r}
FVC_lmer1 <- lmer(fvc_pct ~ 1 + time_firstFVC*Soil_5yrPreDx + (1 + time_firstFVC| ID), data=pft_IPF)
summary(FVC_lmer1)
icc(FVC_lmer1)
report(FVC_lmer1)
```


### IPF-Only - Complete Model
```{r}
FVC_lmer2 <- lmer(fvc_pct ~ 1 + time_firstFVC*Soil_5yrPreDx + time_firstFVC*dich_Race + time_firstFVC*sex + time_firstFVC*age_dx + time_firstFVC*smokeHx + time_firstFVC*pct_belowpoverty + time_firstFVC*site +(1 + time_firstFVC| ID), data=pft_IPF)
summary(FVC_lmer2)
icc(FVC_lmer2)
report(FVC_lmer2)
```

### IPF-Only - Complete Model + dx_yr
```{r}
FVC_lmer3 <- lmer(fvc_pct ~ 1 + time_firstFVC*Soil_5yrPreDx + time_firstFVC*dich_Race + time_firstFVC*sex + time_firstFVC*age_dx + time_firstFVC*smokeHx + time_firstFVC*pct_belowpoverty + time_firstFVC*site +time_firstFVC*dx_yr + (1 + time_firstFVC| ID), data=pft_IPF)
summary(FVC_lmer3)
icc(FVC_lmer3)
report(FVC_lmer3)
```


## IPF-Only - Dichotomized Soil 5yrs Pre-Diagnosis and FVC Models with Random Intercept and Random Slope
### IPF-Only - Base Model
```{r}
FVC_lmer1 <- lmer(fvc_pct ~ 1 + time_firstFVC*Soil5yr_dich + (1 + time_firstFVC| ID), data=pft_IPF)
summary(FVC_lmer1)
icc(FVC_lmer1)
report(FVC_lmer1)
```

### IPF-Only - Complete Model
```{r}
FVC_lmer2 <- lmer(fvc_pct ~ 1 + time_firstFVC*Soil5yr_dich + time_firstFVC*dich_Race + time_firstFVC*sex + time_firstFVC*age_dx + time_firstFVC*smokeHx + time_firstFVC*pct_belowpoverty + time_firstFVC*site +(1 + time_firstFVC| ID), data=pft_IPF)
summary(FVC_lmer2)
icc(FVC_lmer2)
report(FVC_lmer2)
```

### IPF-Only - Complete Model + dx_yr
```{r}
FVC_lmer3 <- lmer(fvc_pct ~ 1 + time_firstFVC*Soil5yr_dich + time_firstFVC*dich_Race + time_firstFVC*sex + time_firstFVC*age_dx + time_firstFVC*smokeHx + time_firstFVC*pct_belowpoverty + time_firstFVC*site +time_firstFVC*dx_yr + (1 + time_firstFVC| ID), data=pft_IPF)
summary(FVC_lmer3)
icc(FVC_lmer3)
report(FVC_lmer3)
```



# IPF-Only - Association between Soil 5yrs Pre-Censoring and FVC Decline
Linear mixed models with random intercept and slope will be used to evaluate the impact of Soil During Disease on FVC decline.

## IPF-Only - Continuous Soil 5yrs Pre-Censoring and FVC Models with Random Intercept and Random Slope
### IPF-Only - Base Model
```{r}
FVC_lmer1 <- lmer(fvc_pct ~ 1 + time_firstFVC*Soil_5yrPreCensor + (1 + time_firstFVC| ID), data=pft_IPF)
summary(FVC_lmer1)
icc(FVC_lmer1)
report(FVC_lmer1)
```


### IPF-Only - Complete Model
```{r}
FVC_lmer2 <- lmer(fvc_pct ~ 1 + time_firstFVC*Soil_5yrPreCensor + time_firstFVC*dich_Race + time_firstFVC*sex + time_firstFVC*age_dx + time_firstFVC*smokeHx + time_firstFVC*pct_belowpoverty + time_firstFVC*site +(1 + time_firstFVC| ID), data=pft_IPF)
summary(FVC_lmer2)
icc(FVC_lmer2)
report(FVC_lmer2)
```

### IPF-Only - Complete Model + dx_yr
```{r}
FVC_lmer3 <- lmer(fvc_pct ~ 1 + time_firstFVC*Soil_5yrPreCensor + time_firstFVC*dich_Race + time_firstFVC*sex + time_firstFVC*age_dx + time_firstFVC*smokeHx + time_firstFVC*pct_belowpoverty + time_firstFVC*site +time_firstFVC*dx_yr + (1 + time_firstFVC| ID), data=pft_IPF)
summary(FVC_lmer3)
icc(FVC_lmer3)
report(FVC_lmer3)
```


## IPF-Only - Dichotomized Soil 5yrs Pre-Censoring and FVC Models with Random Intercept and Random Slope
### IPF-Only - Base Model
```{r}
FVC_lmer1 <- lmer(fvc_pct ~ 1 + time_firstFVC*Soil5yrCensor_dich + (1 + time_firstFVC| ID), data=pft_IPF)
summary(FVC_lmer1)
icc(FVC_lmer1)
report(FVC_lmer1)
```

### IPF-Only - Complete Model
```{r}
FVC_lmer2 <- lmer(fvc_pct ~ 1 + time_firstFVC*Soil5yrCensor_dich + time_firstFVC*dich_Race + time_firstFVC*sex + time_firstFVC*age_dx + time_firstFVC*smokeHx + time_firstFVC*pct_belowpoverty + time_firstFVC*site +(1 + time_firstFVC| ID), data=pft_IPF)
summary(FVC_lmer2)
icc(FVC_lmer2)
report(FVC_lmer2)
```

### IPF-Only - Complete Model + dx_yr
```{r}
FVC_lmer3 <- lmer(fvc_pct ~ 1 + time_firstFVC*Soil5yrCensor_dich + time_firstFVC*dich_Race + time_firstFVC*sex + time_firstFVC*age_dx + time_firstFVC*smokeHx + time_firstFVC*pct_belowpoverty + time_firstFVC*site +time_firstFVC*dx_yr + (1 + time_firstFVC| ID), data=pft_IPF)
summary(FVC_lmer3)
icc(FVC_lmer3)
report(FVC_lmer3)
```


# IPF-Only - Association between Soil 5yrs Pre-Diagnosis and DLCO Decline
Linear mixed models with random intercept and slope will be used to evaluate the impact of Soil During Disease on dlco decline.
## IPF-Only - Continuous Soil 5yrs Pre-Diagnosis and DLCO Models with Random Intercept and Random Slope
### IPF-Only - Base Model
```{r}
DLCO_lmer1 <- lmer(dlco_pct ~ 1 + time_firstDLCO*Soil_5yrPreDx + (1 + time_firstDLCO| ID), data=pft_IPF)
summary(DLCO_lmer1)
icc(DLCO_lmer1)
report(DLCO_lmer1)
```

### IPF-Only - Complete Model
```{r}
DLCO_lmer2 <- lmer(dlco_pct ~ 1 + time_firstDLCO*Soil_5yrPreDx + time_firstDLCO*dich_Race + time_firstDLCO*sex + time_firstDLCO*age_dx + time_firstDLCO*smokeHx + time_firstDLCO*pct_belowpoverty + time_firstDLCO*site +(1 + time_firstDLCO | ID), data=pft_IPF)
summary(DLCO_lmer2)
icc(DLCO_lmer2)
report(DLCO_lmer2)
```

### IPF-Only - Complete Model + dx_yr
```{r}
DLCO_lmer3 <- lmer(dlco_pct ~ 1 + time_firstDLCO*Soil_5yrPreDx + time_firstDLCO*dich_Race + time_firstDLCO*sex + time_firstDLCO*age_dx + time_firstDLCO*smokeHx + time_firstDLCO*pct_belowpoverty + time_firstDLCO*site +time_firstDLCO*dx_yr + (1 + time_firstDLCO | ID), data=pft_IPF)
summary(DLCO_lmer3)
icc(DLCO_lmer3)
report(DLCO_lmer3)
```


## IPF-Only - Dichotomized Soil 5yrs Pre-Diagnosis and DLCO Models with Random Intercept and Random Slope
### IPF-Only - Base Model
```{r}
DLCO_lmer1 <- lmer(dlco_pct ~ 1 + time_firstDLCO*Soil5yr_dich + (1 + time_firstDLCO| ID), data=pft_IPF)
summary(DLCO_lmer1)
icc(DLCO_lmer1)
report(DLCO_lmer1)
```

### IPF-Only - Complete Model
```{r}
DLCO_lmer2 <- lmer(dlco_pct ~ 1 + time_firstDLCO*Soil5yr_dich + time_firstDLCO*dich_Race + time_firstDLCO*sex + time_firstDLCO*age_dx + time_firstDLCO*smokeHx + time_firstDLCO*pct_belowpoverty + time_firstDLCO*site +(1 + time_firstDLCO | ID), data=pft_IPF)
summary(DLCO_lmer2)
icc(DLCO_lmer2)
report(DLCO_lmer2)
```

### IPF-Only - Complete Model + dx_yr
```{r}
DLCO_lmer3 <- lmer(dlco_pct ~ 1 + time_firstDLCO*Soil5yr_dich + time_firstDLCO*dich_Race + time_firstDLCO*sex + time_firstDLCO*age_dx + time_firstDLCO*smokeHx + time_firstDLCO*pct_belowpoverty + time_firstDLCO*site +time_firstDLCO*dx_yr + (1 + time_firstDLCO | ID), data=pft_IPF)
summary(DLCO_lmer3)
icc(DLCO_lmer3)
report(DLCO_lmer3)
```




# IPF-Only - Association between Soil 5yrs Pre-Censoring and DLCO Decline
Linear mixed models with random intercept and slope will be used to evaluate the impact of Soil During Disease on dlco decline.
## IPF-Only - Continuous Soil 5yrs Pre-Censoring and DLCO Models with Random Intercept and Random Slope
### IPF-Only - Base Model
```{r}
DLCO_lmer1 <- lmer(dlco_pct ~ 1 + time_firstDLCO*Soil_5yrPreCensor + (1 + time_firstDLCO| ID), data=pft_IPF)
summary(DLCO_lmer1)
icc(DLCO_lmer1)
report(DLCO_lmer1)
```

### IPF-Only - Complete Model
```{r}
DLCO_lmer2 <- lmer(dlco_pct ~ 1 + time_firstDLCO*Soil_5yrPreCensor + time_firstDLCO*dich_Race + time_firstDLCO*sex + time_firstDLCO*age_dx + time_firstDLCO*smokeHx + time_firstDLCO*pct_belowpoverty + time_firstDLCO*site +(1 + time_firstDLCO | ID), data=pft_IPF)
summary(DLCO_lmer2)
icc(DLCO_lmer2)
report(DLCO_lmer2)
```

### IPF-Only - Complete Model + dx_yr
```{r}
DLCO_lmer3 <- lmer(dlco_pct ~ 1 + time_firstDLCO*Soil_5yrPreCensor + time_firstDLCO*dich_Race + time_firstDLCO*sex + time_firstDLCO*age_dx + time_firstDLCO*smokeHx + time_firstDLCO*pct_belowpoverty + time_firstDLCO*site +time_firstDLCO*dx_yr + (1 + time_firstDLCO | ID), data=pft_IPF)
summary(DLCO_lmer3)
icc(DLCO_lmer3)
report(DLCO_lmer3)
```


## IPF-Only - Dichotomized Soil 5yrs Pre-Censoring and DLCO Models with Random Intercept and Random Slope
### IPF-Only - Base Model
```{r}
DLCO_lmer1 <- lmer(dlco_pct ~ 1 + time_firstDLCO*Soil5yrCensor_dich + (1 + time_firstDLCO| ID), data=pft_IPF)
summary(DLCO_lmer1)
icc(DLCO_lmer1)
report(DLCO_lmer1)
```

### IPF-Only - Complete Model
```{r}
DLCO_lmer2 <- lmer(dlco_pct ~ 1 + time_firstDLCO*Soil5yrCensor_dich + time_firstDLCO*dich_Race + time_firstDLCO*sex + time_firstDLCO*age_dx + time_firstDLCO*smokeHx + time_firstDLCO*pct_belowpoverty + time_firstDLCO*site +(1 + time_firstDLCO | ID), data=pft_IPF)
summary(DLCO_lmer2)
icc(DLCO_lmer2)
report(DLCO_lmer2)
```

### IPF-Only - Complete Model + dx_yr
```{r}
DLCO_lmer3 <- lmer(dlco_pct ~ 1 + time_firstDLCO*Soil5yrCensor_dich + time_firstDLCO*dich_Race + time_firstDLCO*sex + time_firstDLCO*age_dx + time_firstDLCO*smokeHx + time_firstDLCO*pct_belowpoverty + time_firstDLCO*site +time_firstDLCO*dx_yr + (1 + time_firstDLCO | ID), data=pft_IPF)
summary(DLCO_lmer3)
icc(DLCO_lmer3)
report(DLCO_lmer3)
```



# Pre-2015 vs Post-2015 Models
```{r}
pre2015 <- pft %>% filter(dx_yr<=2015)

post2015 <- pft %>% filter(dx_yr>2015)
```


# Pre/Post-2015 - Association between Soil 5yrs Pre-Diagnosis and FVC Decline
Linear mixed models with random intercept and slope will be used to evaluate the impact of Soil During Disease on FVC decline.

## Pre/Post-2015 - Continuous Soil 5yrs Pre-Diagnosis and FVC Models with Random Intercept and Random Slope
### Pre/Post-2015 - Base Model
```{r}
FVC_lmer1a <- lmer(fvc_pct ~ 1 + time_firstFVC*Soil_5yrPreDx + (1 + time_firstFVC| ID), data=pre2015)
summary(FVC_lmer1a)
icc(FVC_lmer1a)
report(FVC_lmer1a)
```

```{r}
FVC_lmer1b <- lmer(fvc_pct ~ 1 + time_firstFVC*Soil_5yrPreDx + (1 + time_firstFVC| ID), data=post2015)
summary(FVC_lmer1b)
icc(FVC_lmer1b)
report(FVC_lmer1b)
```

### Pre/Post-2015 - Complete Model
```{r}
FVC_lmer2a <- lmer(fvc_pct ~ 1 + time_firstFVC*Soil_5yrPreDx + time_firstFVC*dich_Race + time_firstFVC*sex + time_firstFVC*age_dx + time_firstFVC*smokeHx + time_firstFVC*pct_belowpoverty + time_firstFVC*site +(1 + time_firstFVC| ID), data=pre2015)
summary(FVC_lmer2a)
icc(FVC_lmer2a)
report(FVC_lmer2a)
```

```{r}
FVC_lmer2b <- lmer(fvc_pct ~ 1 + time_firstFVC*Soil_5yrPreDx + time_firstFVC*dich_Race + time_firstFVC*sex + time_firstFVC*age_dx + time_firstFVC*smokeHx + time_firstFVC*pct_belowpoverty + time_firstFVC*site +(1 + time_firstFVC| ID), data=post2015)
summary(FVC_lmer2b)
icc(FVC_lmer2b)
report(FVC_lmer2b)
```


## Pre/Post-2015 - Dichotomized Soil 5yrs Pre-Diagnosis and FVC Models with Random Intercept and Random Slope
### Pre/Post-2015 - Base Model
```{r}
FVC_lmer1a <- lmer(fvc_pct ~ 1 + time_firstFVC*Soil5yr_dich + (1 + time_firstFVC| ID), data=pre2015)
summary(FVC_lmer1a)
icc(FVC_lmer1a)
report(FVC_lmer1a)
```

```{r}
FVC_lmer1b <- lmer(fvc_pct ~ 1 + time_firstFVC*Soil5yr_dich + (1 + time_firstFVC| ID), data=post2015)
summary(FVC_lmer1b)
icc(FVC_lmer1b)
report(FVC_lmer1b)
```

### Pre/Post-2015 - Complete Model
```{r}
FVC_lmer2a <- lmer(fvc_pct ~ 1 + time_firstFVC*Soil5yr_dich + time_firstFVC*dich_Race + time_firstFVC*sex + time_firstFVC*age_dx + time_firstFVC*smokeHx + time_firstFVC*pct_belowpoverty + time_firstFVC*site +(1 + time_firstFVC| ID), data=pre2015)
summary(FVC_lmer2a)
icc(FVC_lmer2a)
report(FVC_lmer2a)
```

```{r}
FVC_lmer2b <- lmer(fvc_pct ~ 1 + time_firstFVC*Soil5yr_dich + time_firstFVC*dich_Race + time_firstFVC*sex + time_firstFVC*age_dx + time_firstFVC*smokeHx + time_firstFVC*pct_belowpoverty + time_firstFVC*site +(1 + time_firstFVC| ID), data=post2015)
summary(FVC_lmer2b)
icc(FVC_lmer2b)
report(FVC_lmer2b)
```


# Pre/Post-2015 - Association between Soil 5yrs Pre-Censoring and FVC Decline
Linear mixed models with random intercept and slope will be used to evaluate the impact of Soil During Disease on FVC decline.

## Pre/Post-2015 - Continuous Soil 5yrs Pre-Censoring and FVC Models with Random Intercept and Random Slope
### Pre/Post-2015 - Base Model
```{r}
FVC_lmer1a <- lmer(fvc_pct ~ 1 + time_firstFVC*Soil_5yrPreCensor + (1 + time_firstFVC| ID), data=pre2015)
summary(FVC_lmer1a)
icc(FVC_lmer1a)
report(FVC_lmer1a)
```

```{r}
FVC_lmer1b <- lmer(fvc_pct ~ 1 + time_firstFVC*Soil_5yrPreCensor + (1 + time_firstFVC| ID), data=post2015)
summary(FVC_lmer1b)
icc(FVC_lmer1b)
report(FVC_lmer1b)
```

### Pre/Post-2015 - Complete Model
```{r}
FVC_lmer2a <- lmer(fvc_pct ~ 1 + time_firstFVC*Soil_5yrPreCensor + time_firstFVC*dich_Race + time_firstFVC*sex + time_firstFVC*age_dx + time_firstFVC*smokeHx + time_firstFVC*pct_belowpoverty + time_firstFVC*site +(1 + time_firstFVC| ID), data=pre2015)
summary(FVC_lmer2a)
icc(FVC_lmer2a)
report(FVC_lmer2a)
```

```{r}
FVC_lmer2b <- lmer(fvc_pct ~ 1 + time_firstFVC*Soil_5yrPreCensor + time_firstFVC*dich_Race + time_firstFVC*sex + time_firstFVC*age_dx + time_firstFVC*smokeHx + time_firstFVC*pct_belowpoverty + time_firstFVC*site +(1 + time_firstFVC| ID), data=post2015)
summary(FVC_lmer2b)
icc(FVC_lmer2b)
report(FVC_lmer2b)
```

## Pre/Post-2015 - Dichotomized Soil 5yrs Pre-Censoring and FVC Models with Random Intercept and Random Slope
### Pre/Post-2015 - Base Model
```{r}
FVC_lmer1a <- lmer(fvc_pct ~ 1 + time_firstFVC*Soil5yrCensor_dich + (1 + time_firstFVC| ID), data=pre2015)
summary(FVC_lmer1a)
icc(FVC_lmer1a)
report(FVC_lmer1a)
```

```{r}
FVC_lmer1b <- lmer(fvc_pct ~ 1 + time_firstFVC*Soil5yrCensor_dich + (1 + time_firstFVC| ID), data=post2015)
summary(FVC_lmer1b)
icc(FVC_lmer1b)
report(FVC_lmer1b)
```

### Pre/Post-2015 - Complete Model
```{r}
FVC_lmer2a <- lmer(fvc_pct ~ 1 + time_firstFVC*Soil5yrCensor_dich + time_firstFVC*dich_Race + time_firstFVC*sex + time_firstFVC*age_dx + time_firstFVC*smokeHx + time_firstFVC*pct_belowpoverty + time_firstFVC*site +(1 + time_firstFVC| ID), data=pre2015)
summary(FVC_lmer2a)
icc(FVC_lmer2a)
report(FVC_lmer2a)
```

```{r}
FVC_lmer2b <- lmer(fvc_pct ~ 1 + time_firstFVC*Soil5yrCensor_dich + time_firstFVC*dich_Race + time_firstFVC*sex + time_firstFVC*age_dx + time_firstFVC*smokeHx + time_firstFVC*pct_belowpoverty + time_firstFVC*site +(1 + time_firstFVC| ID), data=post2015)
summary(FVC_lmer2b)
icc(FVC_lmer2b)
report(FVC_lmer2b)
```


# Pre/Post-2015 - Association between Soil 5yrs Pre-Diagnosis and DLCO Decline
Linear mixed models with random intercept and slope will be used to evaluate the impact of Soil During Disease on dlco decline.
## Pre/Post-2015 - Continuous Soil 5yrs Pre-Diagnosis and DLCO Models with Random Intercept and Random Slope
### Pre/Post-2015 - Base Model
```{r}
DLCO_lmer1a <- lmer(dlco_pct ~ 1 + time_firstDLCO*Soil_5yrPreDx + (1 + time_firstDLCO| ID), data=pre2015)
summary(DLCO_lmer1a)
icc(DLCO_lmer1a)
report(DLCO_lmer1a)
```

```{r}
DLCO_lmer1b <- lmer(dlco_pct ~ 1 + time_firstDLCO*Soil_5yrPreDx + (1 + time_firstDLCO| ID), data=post2015)
summary(DLCO_lmer1b)
icc(DLCO_lmer1b)
report(DLCO_lmer1b)
```


### Pre/Post-2015 - Complete Model
```{r}
DLCO_lmer2a <- lmer(dlco_pct ~ 1 + time_firstDLCO*Soil_5yrPreDx + time_firstDLCO*dich_Race + time_firstDLCO*sex + time_firstDLCO*age_dx + time_firstDLCO*smokeHx + time_firstDLCO*pct_belowpoverty + time_firstDLCO*site +(1 + time_firstDLCO | ID), data=pre2015)
summary(DLCO_lmer2a)
icc(DLCO_lmer2a)
report(DLCO_lmer2a)
```

```{r}
DLCO_lmer2b <- lmer(dlco_pct ~ 1 + time_firstDLCO*Soil_5yrPreDx + time_firstDLCO*dich_Race + time_firstDLCO*sex + time_firstDLCO*age_dx + time_firstDLCO*smokeHx + time_firstDLCO*pct_belowpoverty + time_firstDLCO*site +(1 + time_firstDLCO | ID), data=post2015)
summary(DLCO_lmer2b)
icc(DLCO_lmer2b)
report(DLCO_lmer2b)
```


## Pre/Post-2015 - Dichotomized Soil 5yrs Pre-Diagnosis and DLCO Models with Random Intercept and Random Slope
### Pre/Post-2015 - Base Model
```{r}
DLCO_lmer1a <- lmer(dlco_pct ~ 1 + time_firstDLCO*Soil5yr_dich + (1 + time_firstDLCO| ID), data=pre2015)
summary(DLCO_lmer1a)
icc(DLCO_lmer1a)
report(DLCO_lmer1a)
```

```{r}
DLCO_lmer1b <- lmer(dlco_pct ~ 1 + time_firstDLCO*Soil5yr_dich + (1 + time_firstDLCO| ID), data=post2015)
summary(DLCO_lmer1b)
icc(DLCO_lmer1b)
report(DLCO_lmer1b)
```

### Pre/Post-2015 - Complete Model
```{r}
DLCO_lmer2a <- lmer(dlco_pct ~ 1 + time_firstDLCO*Soil5yr_dich + time_firstDLCO*dich_Race + time_firstDLCO*sex + time_firstDLCO*age_dx + time_firstDLCO*smokeHx + time_firstDLCO*pct_belowpoverty + time_firstDLCO*site +(1 + time_firstDLCO | ID), data=pre2015)
summary(DLCO_lmer2a)
icc(DLCO_lmer2a)
report(DLCO_lmer2a)
```

```{r}
DLCO_lmer2b <- lmer(dlco_pct ~ 1 + time_firstDLCO*Soil5yr_dich + time_firstDLCO*dich_Race + time_firstDLCO*sex + time_firstDLCO*age_dx + time_firstDLCO*smokeHx + time_firstDLCO*pct_belowpoverty + time_firstDLCO*site +(1 + time_firstDLCO | ID), data=post2015)
summary(DLCO_lmer2b)
icc(DLCO_lmer2b)
report(DLCO_lmer2b)
```


# Pre/Post-2015 - Association between Soil 5yrs Pre-Censoring and DLCO Decline
Linear mixed models with random intercept and slope will be used to evaluate the impact of Soil During Disease on dlco decline.
## Pre/Post-2015 - Continuous Soil 5yrs Pre-Censoring and DLCO Models with Random Intercept and Random Slope
### Pre/Post-2015 - Base Model
```{r}
DLCO_lmer1a <- lmer(dlco_pct ~ 1 + time_firstDLCO*Soil_5yrPreCensor + (1 + time_firstDLCO| ID), data=pre2015)
summary(DLCO_lmer1a)
icc(DLCO_lmer1a)
report(DLCO_lmer1a)
```

```{r}
DLCO_lmer1b <- lmer(dlco_pct ~ 1 + time_firstDLCO*Soil_5yrPreCensor + (1 + time_firstDLCO| ID), data=post2015)
summary(DLCO_lmer1b)
icc(DLCO_lmer1b)
report(DLCO_lmer1b)
```

### Pre/Post-2015 - Complete Model
```{r}
DLCO_lmer2a <- lmer(dlco_pct ~ 1 + time_firstDLCO*Soil_5yrPreCensor + time_firstDLCO*dich_Race + time_firstDLCO*sex + time_firstDLCO*age_dx + time_firstDLCO*smokeHx + time_firstDLCO*pct_belowpoverty + time_firstDLCO*site +(1 + time_firstDLCO | ID), data=pre2015)
summary(DLCO_lmer2a)
icc(DLCO_lmer2a)
report(DLCO_lmer2a)
```

```{r}
DLCO_lmer2b <- lmer(dlco_pct ~ 1 + time_firstDLCO*Soil_5yrPreCensor + time_firstDLCO*dich_Race + time_firstDLCO*sex + time_firstDLCO*age_dx + time_firstDLCO*smokeHx + time_firstDLCO*pct_belowpoverty + time_firstDLCO*site +(1 + time_firstDLCO | ID), data=post2015)
summary(DLCO_lmer2b)
icc(DLCO_lmer2b)
report(DLCO_lmer2b)
```


## Pre/Post-2015 - Dichotomized Soil 5yrs Pre-Censoring and DLCO Models with Random Intercept and Random Slope
### Pre/Post-2015 - Base Model
```{r}
DLCO_lmer1a <- lmer(dlco_pct ~ 1 + time_firstDLCO*Soil5yrCensor_dich + (1 + time_firstDLCO| ID), data=pre2015)
summary(DLCO_lmer1a)
icc(DLCO_lmer1a)
report(DLCO_lmer1a)
```

```{r}
DLCO_lmer1b <- lmer(dlco_pct ~ 1 + time_firstDLCO*Soil5yrCensor_dich + (1 + time_firstDLCO| ID), data=post2015)
summary(DLCO_lmer1b)
icc(DLCO_lmer1b)
report(DLCO_lmer1b)
```

### Pre/Post-2015 - Complete Model
```{r}
DLCO_lmer2a <- lmer(dlco_pct ~ 1 + time_firstDLCO*Soil5yrCensor_dich + time_firstDLCO*dich_Race + time_firstDLCO*sex + time_firstDLCO*age_dx + time_firstDLCO*smokeHx + time_firstDLCO*pct_belowpoverty + time_firstDLCO*site +(1 + time_firstDLCO | ID), data=pre2015)
summary(DLCO_lmer2a)
icc(DLCO_lmer2a)
report(DLCO_lmer2a)
```

```{r}
DLCO_lmer2b <- lmer(dlco_pct ~ 1 + time_firstDLCO*Soil5yrCensor_dich + time_firstDLCO*dich_Race + time_firstDLCO*sex + time_firstDLCO*age_dx + time_firstDLCO*smokeHx + time_firstDLCO*pct_belowpoverty + time_firstDLCO*site +(1 + time_firstDLCO | ID), data=post2015)
summary(DLCO_lmer2b)
icc(DLCO_lmer2b)
report(DLCO_lmer2b)
```








