---
title: "SO4 and CARE-PF fILD Complete Cohort Lung Function Decline Analysis - Revision #1 Final Analysis"
author: "Gillian Goobie"
date: "07/31/2022"
output: 
  bookdown::html_document2:
    toc: yes
    toc_float: yes
    number_sections: yes
  html_document: 
    toc: yes
    toc_float: yes
    number_sections: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE, warning=FALSE, echo=F}
library(tidyverse)
library(readxl)
library(here)
library(nnet)
library(writexl)
library(knitr)
library(lintr)
library(psych)
library(RColorBrewer)
library(reticulate)
library(survival)
library(survminer)
library(cmprsk)
library(riskRegression)
library(prodlim)
library(lme4)
library(performance)
library(psycho)
library(report)
library(rms)
library(splines)
library(Greg)
```


# Uploading PFT Dataframe
```{r}
#Need to import the correct sheet from the original Simmons file
outfile1 <- here("MonthlySO4Data/CAREPF_AllPts_PFTs_2021_06_01.xlsx")
pft <- read_excel(outfile1, na="")
```

## Correct formatting for pft dataframe
```{r}
pft <- pft %>% mutate_at(vars(pft_date, tx_date, dx_date), as.Date, format="%m/%d/%Y")
str(pft)
```

## dplyr::select Necessary PFT Columns
```{r}
pft <- pft %>% dplyr::select(ID, pft_date, pft_id, fvc, fvc_pct, fev1, fev1_pct, dlco, dlco_pct)
str(pft)
```

# Uploading SO4 Dataframe
Need to upload the file which only includes patients with fibrotic ILD
```{r}
outfile2 <- here("Final_CAREPF_fILD_SO4_BaselineData_2022_02_12.xlsx")
SO4 <- read_excel(outfile2, na="")
```
Next we need to ensure that all variables in the SO4 dataframe are correctly named and in the right class format.

## Correct date format
```{r}
SO4 <- SO4 %>% mutate_at(vars(dob, dx_date, initial_visit_date, formal_dx_date, reg_update, death_date, tx_date, pft_date, last_updated, deathORtx_date, DeathTxCensor_date, SO4_date), as.Date, format="%m/%d/%Y")
str(SO4)
```

## dplyr::select columns that we need
```{r}
SO4 <- SO4 %>% dplyr::select(ID, age_dx, sex, smokeHx, race, dich_Race, site, avg_s, dx_IPF, metro, province, status, txed, dx, dx_group, fvc_pct, dlco_pct, pft_timefromdx, SO4_5yrPreDx, SO45yr_dich, SO4_5yrPreCensor, SO45yrCensor_dich, dx_date, death_date, tx_date, DeathTxCensor_date, pft_date)
```


## Correct Factor Variables and Levels
```{r}
SO4 <- SO4 %>% mutate_at(vars(status, sex, txed, dx, smokeHx, race, dich_Race, dx_group, dx_IPF, metro, province, site, SO45yrCensor_dich), as.factor)
str(SO4)
```

## Releveling Factors
Here I am releveling factors so that they are in an intuitive order for my later analyses.
```{r}
SO4$site <- as.factor(SO4$site)
SO4$sex <- fct_relevel(SO4$sex, c("M","F"))
SO4$race <- fct_relevel(SO4$race, c("W","B","A","N", "P", "U"))
SO4$dich_Race <- fct_relevel(SO4$dich_Race, c("White","Non-White"))
SO4$smokeHx <- fct_relevel(SO4$smokeHx, c("Never","Former","Always"))

#For dx and dx_group, I just want IPF to be first and then the rest of the categories are alphabetical
SO4$dx <- fct_relevel(SO4$dx, c("IPF"))
SO4$dx_group <- fct_relevel(SO4$dx_group, c("IPF"))
SO4$SO45yrCensor_dich <- fct_relevel(SO4$SO45yrCensor_dich, c("Low"))
SO4$SO45yr_dich <- fct_relevel(SO4$SO45yr_dich, c("Low"))
str(SO4)
```

## Creating a dx_yr column
```{r}
SO4 <- SO4 %>% 
  mutate(dx_yr = format(as.Date(SO4$dx_date, format="%Y-%m-%d"),"%Y"))
SO4$dx_yr <- as.numeric(SO4$dx_yr)
```

## Remove PFTs that occurred after transplant
Then will make a simpler dataframe that can be joined to pft
```{r}
txed <- SO4 %>% filter(txed==1) %>% dplyr::select(ID, tx_date)
```
We have 176 patients out of the 3389 who were transplanted

Next will join txed to pft
```{r}
pft <- left_join(pft, txed, by="ID")

#remove all pfts where pft_date>tx_date
pft <- pft %>% filter(is.na(tx_date) | pft_date<tx_date)
```
Our PFT dataframe does not change because this filter already happened in the file creation.

# Remove PFT Values that Don't Make Sense
There are some DLCO values that are too low that we need to remove. All the extremely high DLCOs have been removed that are incorrect and the remaining ones make sense.
```{r}
summary(pft$fvc_pct)
#The one fvc_pct of 4 is an error and should be 44
pft[pft$pft_id==28700,5] <- 44

summary(pft$dlco_pct)
#DLCOs of 0 are impossible so will restrict to no DLCOs <5, but there are no DLCOs <5 in this dataset

pft <- pft %>% filter(dlco_pct>=5 | is.na(dlco_pct))
summary(pft$dlco_pct)
```

## Merge PFT Dataframe with SO4 dataframe
```{r}
pft <- left_join(pft, SO4, by=c("ID", "tx_date"))

pft <- pft %>% rename("pft_date"="pft_date.x", "fvc_pct"="fvc_pct.x", "dlco_pct"="dlco_pct.x", "first_pft_date"="pft_date.y",  "first_fvc_pct"="fvc_pct.y", "first_dlco_pct"="dlco_pct.y")
str(pft)
```

## Remove SO4=NA rows from pft dataframe
Now I need to get rid of the rows where SO4 is NA. Looking at the original excel spreadsheet, there are a bunch of patients with longitudinal PFT data who I don't have the rest of the details on, and this is because they are patients who do not have a diagnosis of fILD. Most of them have diagnoses of sarcoidosis.
```{r}
pft <- pft %>% filter(!is.na(SO4_5yrPreCensor))
```
This filtering results in our pft dataframe going from  to  observations.

## Remove PFTs that occurred before first_pft_date
```{r}
pft <- pft %>% filter(pft_date >= first_pft_date)
```
This filtering results in our pft dataframe going from  to  observations.

## Create Time from First PFT Variable
```{r}
pft <- pft %>% mutate(time_firstPFT=(pft_date-first_pft_date)/365.25)
pft$time_firstPFT <- as.numeric(pft$time_firstPFT)
summary(pft$time_firstPFT)
#Now there are no negative time_firstPFTs
```

## Create Last PFT Date Variable
```{r}
lastPFT <- pft %>% 
  group_by(ID) %>% 
  filter(pft_date == max(pft_date))

head(lastPFT)

#Pare lastPFT down to required columns
lastPFT <- lastPFT %>% dplyr::select(ID, pft_date, fev1, fev1_pct, fvc, fvc_pct, dlco, dlco_pct)

#Rename columns in lastPFT
lastPFT <- lastPFT %>% rename(last_pft_date=pft_date, last_fev1=fev1, last_fev1_pct=fev1_pct, last_fvc=fvc, last_fvc_pct=fvc_pct, last_dlco=dlco, last_dlco_pct=dlco_pct)
head(lastPFT)

#add lastPFT to pft dataframe
pft <- left_join(pft, lastPFT, by="ID")
```


## Create dataframe that excludes patients which only had one PFT
```{r}
# Determine how many PFTs each patient has
PFT_number <- pft %>% count(ID)
summary(PFT_number)
# minimum number of PFTs is 1, so we do have some patients that only had one PFT that are included in later analyses

# restrict dataframe to only include patients with one PFT
one_pft <- PFT_number %>% filter(n==1)
# patients only had one PFT, so we need to ensure the analyses below remain consistent when we remove patients that only have one PFT

# remove patients with only one PFT
pft_over1 <- pft[!pft$ID %in% one_pft$ID,]
#Now my pft_over1 dataframe drops from  rows to  rows, which is a loss of  individuals who only had 1 pft

#Recheck using same methods as above
PFT_number <- pft_over1 %>% count(ID)
summary(PFT_number)
#And we can see that the minimum is now 2
```


# Exploratory Data Analysis
## Visualization of Relationship between SO4 and FVC Decline
Now the 5yrs pre-censoring low vs high
```{r}
#First I need to make a new dataframe that removes the SO4dx_quart NAs so they don't show up in the plot
pft1 <- pft %>% filter(!is.na(SO45yrCensor_dich))

#Then I can plot it   
(ggplot(pft1, aes(x=time_firstPFT, y=fvc_pct, group=ID)) +
   geom_line(aes(color=SO45yrCensor_dich), alpha=0.6)+
   xlim(0,5)+
   stat_smooth(aes(group=1))+
   facet_grid(. ~SO45yrCensor_dich)+
   labs(x="Time from First PFT (years)", y="Change in FVC % Predicted", title="Change in FVC % Predicted by Low vs High SO4 5yrs Pre-Censoring", color="SO4 5yrs \nPre-Censoring")+
   theme(plot.title = element_text(hjust = 0.5))+
   scale_color_brewer(type="seq", palette="YlOrRd"))
```


## Visualization of Relationship between SO4 and DLCO Decline
Now the 5yrs pre-censoring low vs high
```{r}
#First I need to make a new dataframe that removes the SO4dx_quart NAs so they don't show up in the plot
pft1 <- pft %>% filter(!is.na(SO45yrCensor_dich))

#Then I can plot it   
(ggplot(pft1, aes(x=time_firstPFT, y=dlco_pct, group=ID)) +
   geom_line(aes(color=SO45yrCensor_dich), alpha=0.6)+
   xlim(0,5)+
   stat_smooth(aes(group=1))+
   facet_grid(. ~SO45yrCensor_dich)+
   labs(x="Time from First PFT (years)", y="Change in DLCO % Predicted", title="Change in DLCO % Predicted by Low vs High SO4 5yrs Pre-Censoring", color="SO4 5yrs \nPre-Censoring")+
   theme(plot.title = element_text(hjust = 0.5))+
   scale_color_brewer(type="seq", palette="YlOrRd"))
```


## Descriptive Stats for FVC Decline
```{r}
#Look at the mean and total variance of fvc_pct
pft %>% summarise(mean_fvc=mean(fvc_pct, na.rm=T))
tot_var_fvc <- pft %>% dplyr::select(fvc_pct) %>% summarise(var_fvc_pct=var(fvc_pct, na.rm=T))
tot_var_fvc

#Look at the between subject variance
bw_subj_var_fvc <- pft %>% group_by(ID) %>% mutate(mean_fvc_pct=mean(fvc_pct)) %>% ungroup() %>% summarise(bw_subj_var_fvc=var(mean_fvc_pct, na.rm=T))
bw_subj_var_fvc

fvc_ICC <- bw_subj_var_fvc/tot_var_fvc
fvc_ICC
```
fvc_ICC=0.88 indicates that ~88% of the variance is between subject variance. I.e. there is more difference in fvc_pct between patients than within any single patient.

## Number of PFTs
```{r}
n_pfts <- pft %>% count(ID)
n_pfts
summary(n_pfts$n)
#This shows that the median # of PFTs is 5 and the mean is 6
```
Median 5 PFTs per person, mean 7



# Association between SO4 5yrs Pre-Diagnosis and FVC Decline
Linear mixed models with random intercept and slope will be used to evaluate the impact of SO4 During Disease on FVC decline.

## Continuous SO4 5yrs Pre-Diagnosis and FVC Models with Random Intercept and Random Slope
### Base Model
```{r}
FVC_lmer1 <- lmer(fvc_pct ~ 1 + time_firstPFT*SO4_5yrPreDx + (1 + time_firstPFT| ID), data=pft)
summary(FVC_lmer1)
icc(FVC_lmer1)
report(FVC_lmer1)
```

### Complete Model
```{r}
FVC_lmer2 <- lmer(fvc_pct ~ 1 + time_firstPFT*SO4_5yrPreDx + time_firstPFT*dich_Race + time_firstPFT*sex +  time_firstPFT*age_dx + time_firstPFT*smokeHx + time_firstPFT*avg_s + time_firstPFT*site + (1 + time_firstPFT| ID), data=pft)
summary(FVC_lmer2)
icc(FVC_lmer2)
report(FVC_lmer2)
```

### Complete Model + dx_yr
```{r}
FVC_lmer3 <- lmer(fvc_pct ~ 1 + time_firstPFT*SO4_5yrPreDx + time_firstPFT*dich_Race + time_firstPFT*sex +  time_firstPFT*age_dx + time_firstPFT*smokeHx + time_firstPFT*avg_s + time_firstPFT*site + time_firstPFT*dx_yr + (1 + time_firstPFT| ID), data=pft)
summary(FVC_lmer3)
icc(FVC_lmer3)
report(FVC_lmer3)
```



## Dichotomized SO4 5yrs Pre-Diagnosis and FVC Models with Random Intercept and Random Slope
### Base Model
```{r}
FVC_lmer1 <- lmer(fvc_pct ~ 1 + time_firstPFT*SO45yr_dich + (1 + time_firstPFT| ID), data=pft)
summary(FVC_lmer1)
icc(FVC_lmer1)
report(FVC_lmer1)
```


### Complete Model
```{r}
FVC_lmer2 <- lmer(fvc_pct ~ 1 + time_firstPFT*SO45yr_dich + time_firstPFT*dich_Race + time_firstPFT*sex +  time_firstPFT*age_dx + time_firstPFT*smokeHx + time_firstPFT*avg_s + time_firstPFT*site + (1 + time_firstPFT| ID), data=pft)
summary(FVC_lmer2)
icc(FVC_lmer2)
report(FVC_lmer2)
```

### Complete Model + dx_yr
```{r}
FVC_lmer3 <- lmer(fvc_pct ~ 1 + time_firstPFT*SO45yr_dich + time_firstPFT*dich_Race + time_firstPFT*sex +  time_firstPFT*age_dx + time_firstPFT*smokeHx + time_firstPFT*avg_s + time_firstPFT*site + time_firstPFT*dx_yr + (1 + time_firstPFT| ID), data=pft)
summary(FVC_lmer3)
icc(FVC_lmer3)
report(FVC_lmer3)
```



# Association between SO4 5yrs Pre-Censoring and FVC Decline
Linear mixed models with random intercept and slope will be used to evaluate the impact of SO4 During Disease on FVC decline.

## Continuous SO4 5yrs Pre-Censoring and FVC Models with Random Intercept and Random Slope
### Base Model
```{r}
FVC_lmer1 <- lmer(fvc_pct ~ 1 + time_firstPFT*SO4_5yrPreCensor + (1 + time_firstPFT| ID), data=pft)
summary(FVC_lmer1)
icc(FVC_lmer1)
report(FVC_lmer1)
```

### Complete Model
```{r}
FVC_lmer2 <- lmer(fvc_pct ~ 1 + time_firstPFT*SO4_5yrPreCensor + time_firstPFT*dich_Race + time_firstPFT*sex +  time_firstPFT*age_dx + time_firstPFT*smokeHx + time_firstPFT*avg_s + time_firstPFT*site + (1 + time_firstPFT| ID), data=pft)
summary(FVC_lmer2)
icc(FVC_lmer2)
report(FVC_lmer2)
```

### Complete Model + dx_yr
```{r}
FVC_lmer3 <- lmer(fvc_pct ~ 1 + time_firstPFT*SO4_5yrPreCensor + time_firstPFT*dich_Race + time_firstPFT*sex +  time_firstPFT*age_dx + time_firstPFT*smokeHx + time_firstPFT*avg_s + time_firstPFT*site + time_firstPFT*dx_yr + (1 + time_firstPFT| ID), data=pft)
summary(FVC_lmer3)
icc(FVC_lmer3)
report(FVC_lmer3)
```



## Dichotomized SO4 5yrs Pre-Censoring and FVC Models with Random Intercept and Random Slope
### Base Model
```{r}
FVC_lmer1 <- lmer(fvc_pct ~ 1 + time_firstPFT*SO45yrCensor_dich + (1 + time_firstPFT| ID), data=pft)
summary(FVC_lmer1)
icc(FVC_lmer1)
report(FVC_lmer1)
```


### Complete Model
```{r}
FVC_lmer2 <- lmer(fvc_pct ~ 1 + time_firstPFT*SO45yrCensor_dich + time_firstPFT*dich_Race + time_firstPFT*sex +  time_firstPFT*age_dx + time_firstPFT*smokeHx + time_firstPFT*avg_s + time_firstPFT*site + (1 + time_firstPFT| ID), data=pft)
summary(FVC_lmer2)
icc(FVC_lmer2)
report(FVC_lmer2)
```

### Complete Model + dx_yr
```{r}
FVC_lmer3 <- lmer(fvc_pct ~ 1 + time_firstPFT*SO45yrCensor_dich + time_firstPFT*dich_Race + time_firstPFT*sex +  time_firstPFT*age_dx + time_firstPFT*smokeHx + time_firstPFT*avg_s + time_firstPFT*site + time_firstPFT*dx_yr + (1 + time_firstPFT| ID), data=pft)
summary(FVC_lmer3)
icc(FVC_lmer3)
report(FVC_lmer3)
```

# Association between SO4 5yrs Pre-Diagnosis and DLCO Decline
Linear mixed models with random intercept and slope will be used to evaluate the impact of SO4 During Disease on DLCO decline.

## Continuous SO4 5yrs Pre-Diagnosis and DLCO Models with Random Intercept and Random Slope
### Base Model
```{r}
DLCO_lmer1 <- lmer(dlco_pct ~ 1 + time_firstPFT*SO4_5yrPreDx + (1 + time_firstPFT| ID), data=pft)
summary(DLCO_lmer1)
icc(DLCO_lmer1)
report(DLCO_lmer1)
```

### Complete Model
```{r}
DLCO_lmer2 <- lmer(dlco_pct ~ 1 + time_firstPFT*SO4_5yrPreDx + time_firstPFT*dich_Race + time_firstPFT*sex +  time_firstPFT*age_dx + time_firstPFT*smokeHx + time_firstPFT*avg_s + time_firstPFT*site + (1 + time_firstPFT| ID), data=pft)
summary(DLCO_lmer2)
icc(DLCO_lmer2)
report(DLCO_lmer2)
```

### Complete Model + dx_yr
```{r}
DLCO_lmer3 <- lmer(dlco_pct ~ 1 + time_firstPFT*SO4_5yrPreDx + time_firstPFT*dich_Race + time_firstPFT*sex +  time_firstPFT*age_dx + time_firstPFT*smokeHx + time_firstPFT*avg_s + time_firstPFT*site + time_firstPFT*dx_yr + (1 + time_firstPFT| ID), data=pft)
summary(DLCO_lmer3)
icc(DLCO_lmer3)
report(DLCO_lmer3)
```



## Dichotomized SO4 5yrs Pre-Diagnosis and DLCO Models with Random Intercept and Random Slope
### Base Model
```{r}
DLCO_lmer1 <- lmer(dlco_pct ~ 1 + time_firstPFT*SO45yr_dich + (1 + time_firstPFT| ID), data=pft)
summary(DLCO_lmer1)
icc(DLCO_lmer1)
report(DLCO_lmer1)
```


### Complete Model
```{r}
DLCO_lmer2 <- lmer(dlco_pct ~ 1 + time_firstPFT*SO45yr_dich + time_firstPFT*dich_Race + time_firstPFT*sex +  time_firstPFT*age_dx + time_firstPFT*smokeHx + time_firstPFT*avg_s + time_firstPFT*site + (1 + time_firstPFT| ID), data=pft)
summary(DLCO_lmer2)
icc(DLCO_lmer2)
report(DLCO_lmer2)
```

### Complete Model + dx_yr
```{r}
DLCO_lmer3 <- lmer(dlco_pct ~ 1 + time_firstPFT*SO45yr_dich + time_firstPFT*dich_Race + time_firstPFT*sex +  time_firstPFT*age_dx + time_firstPFT*smokeHx + time_firstPFT*avg_s + time_firstPFT*site + time_firstPFT*dx_yr + (1 + time_firstPFT| ID), data=pft)
summary(DLCO_lmer3)
icc(DLCO_lmer3)
report(DLCO_lmer3)
```



# Association between SO4 5yrs Pre-Censoring and DLCO Decline
Linear mixed models with random intercept and slope will be used to evaluate the impact of SO4 During Disease on DLCO decline.

## Continuous SO4 5yrs Pre-Censoring and DLCO Models with Random Intercept and Random Slope
### Base Model
```{r}
DLCO_lmer1 <- lmer(dlco_pct ~ 1 + time_firstPFT*SO4_5yrPreCensor + (1 + time_firstPFT| ID), data=pft)
summary(DLCO_lmer1)
icc(DLCO_lmer1)
report(DLCO_lmer1)
```

### Complete Model
```{r}
DLCO_lmer2 <- lmer(dlco_pct ~ 1 + time_firstPFT*SO4_5yrPreCensor + time_firstPFT*dich_Race + time_firstPFT*sex +  time_firstPFT*age_dx + time_firstPFT*smokeHx + time_firstPFT*avg_s + time_firstPFT*site + (1 + time_firstPFT| ID), data=pft)
summary(DLCO_lmer2)
icc(DLCO_lmer2)
report(DLCO_lmer2)
```

### Complete Model + dx_yr
```{r}
DLCO_lmer3 <- lmer(dlco_pct ~ 1 + time_firstPFT*SO4_5yrPreCensor + time_firstPFT*dich_Race + time_firstPFT*sex +  time_firstPFT*age_dx + time_firstPFT*smokeHx + time_firstPFT*avg_s + time_firstPFT*site + time_firstPFT*dx_yr + (1 + time_firstPFT| ID), data=pft)
summary(DLCO_lmer3)
icc(DLCO_lmer3)
report(DLCO_lmer3)
```



## Dichotomized SO4 5yrs Pre-Censoring and DLCO Models with Random Intercept and Random Slope
### Base Model
```{r}
DLCO_lmer1 <- lmer(dlco_pct ~ 1 + time_firstPFT*SO45yrCensor_dich + (1 + time_firstPFT| ID), data=pft)
summary(DLCO_lmer1)
icc(DLCO_lmer1)
report(DLCO_lmer1)
```


### Complete Model
```{r}
DLCO_lmer2 <- lmer(dlco_pct ~ 1 + time_firstPFT*SO45yrCensor_dich + time_firstPFT*dich_Race + time_firstPFT*sex +  time_firstPFT*age_dx + time_firstPFT*smokeHx + time_firstPFT*avg_s + time_firstPFT*site + (1 + time_firstPFT| ID), data=pft)
summary(DLCO_lmer2)
icc(DLCO_lmer2)
report(DLCO_lmer2)
```

### Complete Model + dx_yr
```{r}
DLCO_lmer3 <- lmer(dlco_pct ~ 1 + time_firstPFT*SO45yrCensor_dich + time_firstPFT*dich_Race + time_firstPFT*sex +  time_firstPFT*age_dx + time_firstPFT*smokeHx + time_firstPFT*avg_s + time_firstPFT*site + time_firstPFT*dx_yr + (1 + time_firstPFT| ID), data=pft)
summary(DLCO_lmer3)
icc(DLCO_lmer3)
report(DLCO_lmer3)
```



# Make IPF-Specific Dataframe
```{r}
pft_IPF <- pft %>% filter(dx_group=="IPF")
```

# IPF-Only - Association between SO4 5yrs Pre-Diagnosis and FVC Decline
Linear mixed models with random intercept and slope will be used to evaluate the impact of SO4 During Disease on FVC decline.

## IPF-Only - Continuous SO4 5yrs Pre-Diagnosis and FVC Models with Random Intercept and Random Slope
### IPF-Only - Base Model
```{r}
FVC_lmer1 <- lmer(fvc_pct ~ 1 + time_firstPFT*SO4_5yrPreDx + (1 + time_firstPFT| ID), data=pft_IPF)
summary(FVC_lmer1)
icc(FVC_lmer1)
report(FVC_lmer1)
```

### IPF-Only - Complete Model
```{r}
FVC_lmer2 <- lmer(fvc_pct ~ 1 + time_firstPFT*SO4_5yrPreDx + time_firstPFT*dich_Race + time_firstPFT*sex +  time_firstPFT*age_dx + time_firstPFT*smokeHx + time_firstPFT*avg_s + time_firstPFT*site + (1 + time_firstPFT| ID), data=pft_IPF)
summary(FVC_lmer2)
icc(FVC_lmer2)
report(FVC_lmer2)
```

### IPF-Only - Complete Model + dx_yr
```{r}
FVC_lmer3 <- lmer(fvc_pct ~ 1 + time_firstPFT*SO4_5yrPreDx + time_firstPFT*dich_Race + time_firstPFT*sex +  time_firstPFT*age_dx + time_firstPFT*smokeHx + time_firstPFT*avg_s + time_firstPFT*site + time_firstPFT*dx_yr + (1 + time_firstPFT| ID), data=pft_IPF)
summary(FVC_lmer3)
icc(FVC_lmer3)
report(FVC_lmer3)
```



## IPF-Only - Dichotomized SO4 5yrs Pre-Diagnosis and FVC Models with Random Intercept and Random Slope
### IPF-Only - Base Model
```{r}
FVC_lmer1 <- lmer(fvc_pct ~ 1 + time_firstPFT*SO45yr_dich + (1 + time_firstPFT| ID), data=pft_IPF)
summary(FVC_lmer1)
icc(FVC_lmer1)
report(FVC_lmer1)
```


### IPF-Only - Complete Model
```{r}
FVC_lmer2 <- lmer(fvc_pct ~ 1 + time_firstPFT*SO45yr_dich + time_firstPFT*dich_Race + time_firstPFT*sex +  time_firstPFT*age_dx + time_firstPFT*smokeHx + time_firstPFT*avg_s + time_firstPFT*site + (1 + time_firstPFT| ID), data=pft_IPF)
summary(FVC_lmer2)
icc(FVC_lmer2)
report(FVC_lmer2)
```

### IPF-Only - Complete Model + dx_yr
```{r}
FVC_lmer3 <- lmer(fvc_pct ~ 1 + time_firstPFT*SO45yr_dich + time_firstPFT*dich_Race + time_firstPFT*sex +  time_firstPFT*age_dx + time_firstPFT*smokeHx + time_firstPFT*avg_s + time_firstPFT*site + time_firstPFT*dx_yr + (1 + time_firstPFT| ID), data=pft_IPF)
summary(FVC_lmer3)
icc(FVC_lmer3)
report(FVC_lmer3)
```



# IPF-Only - Association between SO4 5yrs Pre-Censoring and FVC Decline
Linear mixed models with random intercept and slope will be used to evaluate the impact of SO4 During Disease on FVC decline.

## IPF-Only - Continuous SO4 5yrs Pre-Censoring and FVC Models with Random Intercept and Random Slope
### IPF-Only - Base Model
```{r}
FVC_lmer1 <- lmer(fvc_pct ~ 1 + time_firstPFT*SO4_5yrPreCensor + (1 + time_firstPFT| ID), data=pft_IPF)
summary(FVC_lmer1)
icc(FVC_lmer1)
report(FVC_lmer1)
```

### IPF-Only - Complete Model
```{r}
FVC_lmer2 <- lmer(fvc_pct ~ 1 + time_firstPFT*SO4_5yrPreCensor + time_firstPFT*dich_Race + time_firstPFT*sex +  time_firstPFT*age_dx + time_firstPFT*smokeHx + time_firstPFT*avg_s + time_firstPFT*site + (1 + time_firstPFT| ID), data=pft_IPF)
summary(FVC_lmer2)
icc(FVC_lmer2)
report(FVC_lmer2)
```

### IPF-Only - Complete Model + dx_yr
```{r}
FVC_lmer3 <- lmer(fvc_pct ~ 1 + time_firstPFT*SO4_5yrPreCensor + time_firstPFT*dich_Race + time_firstPFT*sex +  time_firstPFT*age_dx + time_firstPFT*smokeHx + time_firstPFT*avg_s + time_firstPFT*site + time_firstPFT*dx_yr + (1 + time_firstPFT| ID), data=pft_IPF)
summary(FVC_lmer3)
icc(FVC_lmer3)
report(FVC_lmer3)
```



## IPF-Only - Dichotomized SO4 5yrs Pre-Censoring and FVC Models with Random Intercept and Random Slope
### IPF-Only - Base Model
```{r}
FVC_lmer1 <- lmer(fvc_pct ~ 1 + time_firstPFT*SO45yrCensor_dich + (1 + time_firstPFT| ID), data=pft_IPF)
summary(FVC_lmer1)
icc(FVC_lmer1)
report(FVC_lmer1)
```


### IPF-Only - Complete Model
```{r}
FVC_lmer2 <- lmer(fvc_pct ~ 1 + time_firstPFT*SO45yrCensor_dich + time_firstPFT*dich_Race + time_firstPFT*sex +  time_firstPFT*age_dx + time_firstPFT*smokeHx + time_firstPFT*avg_s + time_firstPFT*site + (1 + time_firstPFT| ID), data=pft_IPF)
summary(FVC_lmer2)
icc(FVC_lmer2)
report(FVC_lmer2)
```

### IPF-Only - Complete Model + dx_yr
```{r}
FVC_lmer3 <- lmer(fvc_pct ~ 1 + time_firstPFT*SO45yrCensor_dich + time_firstPFT*dich_Race + time_firstPFT*sex +  time_firstPFT*age_dx + time_firstPFT*smokeHx + time_firstPFT*avg_s + time_firstPFT*site + time_firstPFT*dx_yr + (1 + time_firstPFT| ID), data=pft_IPF)
summary(FVC_lmer3)
icc(FVC_lmer3)
report(FVC_lmer3)
```

# IPF-Only - Association between SO4 5yrs Pre-Diagnosis and DLCO Decline
Linear mixed models with random intercept and slope will be used to evaluate the impact of SO4 During Disease on DLCO decline.

## IPF-Only - Continuous SO4 5yrs Pre-Diagnosis and DLCO Models with Random Intercept and Random Slope
### IPF-Only - Base Model
```{r}
DLCO_lmer1 <- lmer(dlco_pct ~ 1 + time_firstPFT*SO4_5yrPreDx + (1 + time_firstPFT| ID), data=pft_IPF)
summary(DLCO_lmer1)
icc(DLCO_lmer1)
report(DLCO_lmer1)
```

### IPF-Only - Complete Model
```{r}
DLCO_lmer2 <- lmer(dlco_pct ~ 1 + time_firstPFT*SO4_5yrPreDx + time_firstPFT*dich_Race + time_firstPFT*sex +  time_firstPFT*age_dx + time_firstPFT*smokeHx + time_firstPFT*avg_s + time_firstPFT*site + (1 + time_firstPFT| ID), data=pft_IPF)
summary(DLCO_lmer2)
icc(DLCO_lmer2)
report(DLCO_lmer2)
```

### IPF-Only - Complete Model + dx_yr
```{r}
DLCO_lmer3 <- lmer(dlco_pct ~ 1 + time_firstPFT*SO4_5yrPreDx + time_firstPFT*dich_Race + time_firstPFT*sex +  time_firstPFT*age_dx + time_firstPFT*smokeHx + time_firstPFT*avg_s + time_firstPFT*site + time_firstPFT*dx_yr + (1 + time_firstPFT| ID), data=pft_IPF)
summary(DLCO_lmer3)
icc(DLCO_lmer3)
report(DLCO_lmer3)
```



## IPF-Only - Dichotomized SO4 5yrs Pre-Diagnosis and DLCO Models with Random Intercept and Random Slope
### IPF-Only - Base Model
```{r}
DLCO_lmer1 <- lmer(dlco_pct ~ 1 + time_firstPFT*SO45yr_dich + (1 + time_firstPFT| ID), data=pft_IPF)
summary(DLCO_lmer1)
icc(DLCO_lmer1)
report(DLCO_lmer1)
```


### IPF-Only - Complete Model
```{r}
DLCO_lmer2 <- lmer(dlco_pct ~ 1 + time_firstPFT*SO45yr_dich + time_firstPFT*dich_Race + time_firstPFT*sex +  time_firstPFT*age_dx + time_firstPFT*smokeHx + time_firstPFT*avg_s + time_firstPFT*site + (1 + time_firstPFT| ID), data=pft_IPF)
summary(DLCO_lmer2)
icc(DLCO_lmer2)
report(DLCO_lmer2)
```

### IPF-Only - Complete Model + dx_yr
```{r}
DLCO_lmer3 <- lmer(dlco_pct ~ 1 + time_firstPFT*SO45yr_dich + time_firstPFT*dich_Race + time_firstPFT*sex +  time_firstPFT*age_dx + time_firstPFT*smokeHx + time_firstPFT*avg_s + time_firstPFT*site + time_firstPFT*dx_yr + (1 + time_firstPFT| ID), data=pft_IPF)
summary(DLCO_lmer3)
icc(DLCO_lmer3)
report(DLCO_lmer3)
```



# IPF-Only - Association between SO4 5yrs Pre-Censoring and DLCO Decline
Linear mixed models with random intercept and slope will be used to evaluate the impact of SO4 During Disease on DLCO decline.

## IPF-Only - Continuous SO4 5yrs Pre-Censoring and DLCO Models with Random Intercept and Random Slope
### IPF-Only - Base Model
```{r}
DLCO_lmer1 <- lmer(dlco_pct ~ 1 + time_firstPFT*SO4_5yrPreCensor + (1 + time_firstPFT| ID), data=pft_IPF)
summary(DLCO_lmer1)
icc(DLCO_lmer1)
report(DLCO_lmer1)
```

### IPF-Only - Complete Model
```{r}
DLCO_lmer2 <- lmer(dlco_pct ~ 1 + time_firstPFT*SO4_5yrPreCensor + time_firstPFT*dich_Race + time_firstPFT*sex +  time_firstPFT*age_dx + time_firstPFT*smokeHx + time_firstPFT*avg_s + time_firstPFT*site + (1 + time_firstPFT| ID), data=pft_IPF)
summary(DLCO_lmer2)
icc(DLCO_lmer2)
report(DLCO_lmer2)
```

### IPF-Only - Complete Model + dx_yr
```{r}
DLCO_lmer3 <- lmer(dlco_pct ~ 1 + time_firstPFT*SO4_5yrPreCensor + time_firstPFT*dich_Race + time_firstPFT*sex +  time_firstPFT*age_dx + time_firstPFT*smokeHx + time_firstPFT*avg_s + time_firstPFT*site + time_firstPFT*dx_yr + (1 + time_firstPFT| ID), data=pft_IPF)
summary(DLCO_lmer3)
icc(DLCO_lmer3)
report(DLCO_lmer3)
```



## IPF-Only - Dichotomized SO4 5yrs Pre-Censoring and DLCO Models with Random Intercept and Random Slope
### IPF-Only - Base Model
```{r}
DLCO_lmer1 <- lmer(dlco_pct ~ 1 + time_firstPFT*SO45yrCensor_dich + (1 + time_firstPFT| ID), data=pft_IPF)
summary(DLCO_lmer1)
icc(DLCO_lmer1)
report(DLCO_lmer1)
```


### IPF-Only - Complete Model
```{r}
DLCO_lmer2 <- lmer(dlco_pct ~ 1 + time_firstPFT*SO45yrCensor_dich + time_firstPFT*dich_Race + time_firstPFT*sex +  time_firstPFT*age_dx + time_firstPFT*smokeHx + time_firstPFT*avg_s + time_firstPFT*site + (1 + time_firstPFT| ID), data=pft_IPF)
summary(DLCO_lmer2)
icc(DLCO_lmer2)
report(DLCO_lmer2)
```

### IPF-Only - Complete Model + dx_yr
```{r}
DLCO_lmer3 <- lmer(dlco_pct ~ 1 + time_firstPFT*SO45yrCensor_dich + time_firstPFT*dich_Race + time_firstPFT*sex +  time_firstPFT*age_dx + time_firstPFT*smokeHx + time_firstPFT*avg_s + time_firstPFT*site + time_firstPFT*dx_yr + (1 + time_firstPFT| ID), data=pft_IPF)
summary(DLCO_lmer3)
icc(DLCO_lmer3)
report(DLCO_lmer3)
```


# Pre-2015 vs Post-2015 Models
```{r}
pre2015 <- pft %>% filter(dx_yr<=2015)

post2015 <- pft %>% filter(dx_yr>2015)
```


# Pre/Post-2015 - Association between SO4 5yrs Pre-Diagnosis and FVC Decline
Linear mixed models with random intercept and slope will be used to evaluate the impact of SO4 During Disease on FVC decline.

## Pre/Post-2015 - Continuous SO4 5yrs Pre-Diagnosis and FVC Models with Random Intercept and Random Slope
### Pre/Post-2015 - Base Model
```{r}
FVC_lmer1a <- lmer(fvc_pct ~ 1 + time_firstPFT*SO4_5yrPreDx + (1 + time_firstPFT| ID), data=pre2015)
summary(FVC_lmer1a)
icc(FVC_lmer1a)
report(FVC_lmer1a)
```

```{r}
FVC_lmer1b <- lmer(fvc_pct ~ 1 + time_firstPFT*SO4_5yrPreDx + (1 + time_firstPFT| ID), data=post2015)
summary(FVC_lmer1b)
icc(FVC_lmer1b)
report(FVC_lmer1b)
```

### Pre/Post-2015 - Complete Model
```{r}
FVC_lmer2a <- lmer(fvc_pct ~ 1 + time_firstPFT*SO4_5yrPreDx + time_firstPFT*dich_Race + time_firstPFT*sex + time_firstPFT*age_dx + time_firstPFT*smokeHx + time_firstPFT*avg_s + time_firstPFT*site + (1 + time_firstPFT| ID), data=pre2015)
summary(FVC_lmer2a)
icc(FVC_lmer2a)
report(FVC_lmer2a)
```

```{r}
FVC_lmer2b <- lmer(fvc_pct ~ 1 + time_firstPFT*SO4_5yrPreDx + time_firstPFT*dich_Race + time_firstPFT*sex + time_firstPFT*age_dx + time_firstPFT*smokeHx + time_firstPFT*avg_s + time_firstPFT*site + (1 + time_firstPFT| ID), data=post2015)
summary(FVC_lmer2b)
icc(FVC_lmer2b)
report(FVC_lmer2b)
```


## Pre/Post-2015 - Dichotomized SO4 5yrs Pre-Diagnosis and FVC Models with Random Intercept and Random Slope
### Pre/Post-2015 - Base Model
```{r}
FVC_lmer1a <- lmer(fvc_pct ~ 1 + time_firstPFT*SO45yr_dich + (1 + time_firstPFT| ID), data=pre2015)
summary(FVC_lmer1a)
icc(FVC_lmer1a)
report(FVC_lmer1a)
```

```{r}
FVC_lmer1b <- lmer(fvc_pct ~ 1 + time_firstPFT*SO45yr_dich + (1 + time_firstPFT| ID), data=post2015)
summary(FVC_lmer1b)
icc(FVC_lmer1b)
report(FVC_lmer1b)
```

### Pre/Post-2015 - Complete Model
```{r}
FVC_lmer2a <- lmer(fvc_pct ~ 1 + time_firstPFT*SO45yr_dich + time_firstPFT*dich_Race + time_firstPFT*sex + time_firstPFT*age_dx + time_firstPFT*smokeHx + time_firstPFT*avg_s + time_firstPFT*site + (1 + time_firstPFT| ID), data=pre2015)
summary(FVC_lmer2a)
icc(FVC_lmer2a)
report(FVC_lmer2a)
```

```{r}
FVC_lmer2b <- lmer(fvc_pct ~ 1 + time_firstPFT*SO45yr_dich + time_firstPFT*dich_Race + time_firstPFT*sex + time_firstPFT*age_dx + time_firstPFT*smokeHx + time_firstPFT*avg_s + time_firstPFT*site + (1 + time_firstPFT| ID), data=post2015)
summary(FVC_lmer2b)
icc(FVC_lmer2b)
report(FVC_lmer2b)
```


# Pre/Post-2015 - Association between SO4 5yrs Pre-Censoring and FVC Decline
Linear mixed models with random intercept and slope will be used to evaluate the impact of SO4 During Disease on FVC decline.

## Pre/Post-2015 - Continuous SO4 5yrs Pre-Censoring and FVC Models with Random Intercept and Random Slope
### Pre/Post-2015 - Base Model
```{r}
FVC_lmer1a <- lmer(fvc_pct ~ 1 + time_firstPFT*SO4_5yrPreCensor + (1 + time_firstPFT| ID), data=pre2015)
summary(FVC_lmer1a)
icc(FVC_lmer1a)
report(FVC_lmer1a)
```

```{r}
FVC_lmer1b <- lmer(fvc_pct ~ 1 + time_firstPFT*SO4_5yrPreCensor + (1 + time_firstPFT| ID), data=post2015)
summary(FVC_lmer1b)
icc(FVC_lmer1b)
report(FVC_lmer1b)
```

### Pre/Post-2015 - Complete Model
```{r}
FVC_lmer2a <- lmer(fvc_pct ~ 1 + time_firstPFT*SO4_5yrPreCensor + time_firstPFT*dich_Race + time_firstPFT*sex + time_firstPFT*age_dx + time_firstPFT*smokeHx + time_firstPFT*avg_s + time_firstPFT*site + (1 + time_firstPFT| ID), data=pre2015)
summary(FVC_lmer2a)
icc(FVC_lmer2a)
report(FVC_lmer2a)
```

```{r}
FVC_lmer2b <- lmer(fvc_pct ~ 1 + time_firstPFT*SO4_5yrPreCensor + time_firstPFT*dich_Race + time_firstPFT*sex + time_firstPFT*age_dx + time_firstPFT*smokeHx + time_firstPFT*avg_s + time_firstPFT*site + (1 + time_firstPFT| ID), data=post2015)
summary(FVC_lmer2b)
icc(FVC_lmer2b)
report(FVC_lmer2b)
```

## Pre/Post-2015 - Dichotomized SO4 5yrs Pre-Censoring and FVC Models with Random Intercept and Random Slope
### Pre/Post-2015 - Base Model
```{r}
FVC_lmer1a <- lmer(fvc_pct ~ 1 + time_firstPFT*SO45yrCensor_dich + (1 + time_firstPFT| ID), data=pre2015)
summary(FVC_lmer1a)
icc(FVC_lmer1a)
report(FVC_lmer1a)
```

```{r}
FVC_lmer1b <- lmer(fvc_pct ~ 1 + time_firstPFT*SO45yrCensor_dich + (1 + time_firstPFT| ID), data=post2015)
summary(FVC_lmer1b)
icc(FVC_lmer1b)
report(FVC_lmer1b)
```

### Pre/Post-2015 - Complete Model
```{r}
FVC_lmer2a <- lmer(fvc_pct ~ 1 + time_firstPFT*SO45yrCensor_dich + time_firstPFT*dich_Race + time_firstPFT*sex + time_firstPFT*age_dx + time_firstPFT*smokeHx + time_firstPFT*avg_s + time_firstPFT*site + (1 + time_firstPFT| ID), data=pre2015)
summary(FVC_lmer2a)
icc(FVC_lmer2a)
report(FVC_lmer2a)
```

```{r}
FVC_lmer2b <- lmer(fvc_pct ~ 1 + time_firstPFT*SO45yrCensor_dich + time_firstPFT*dich_Race + time_firstPFT*sex + time_firstPFT*age_dx + time_firstPFT*smokeHx + time_firstPFT*avg_s + time_firstPFT*site + (1 + time_firstPFT| ID), data=post2015)
summary(FVC_lmer2b)
icc(FVC_lmer2b)
report(FVC_lmer2b)
```


# Pre/Post-2015 - Association between SO4 5yrs Pre-Diagnosis and DLCO Decline
Linear mixed models with random intercept and slope will be used to evaluate the impact of SO4 During Disease on dlco decline.
## Pre/Post-2015 - Continuous SO4 5yrs Pre-Diagnosis and DLCO Models with Random Intercept and Random Slope
### Pre/Post-2015 - Base Model
```{r}
DLCO_lmer1a <- lmer(dlco_pct ~ 1 + time_firstPFT*SO4_5yrPreDx + (1 + time_firstPFT| ID), data=pre2015)
summary(DLCO_lmer1a)
icc(DLCO_lmer1a)
report(DLCO_lmer1a)
```

```{r}
DLCO_lmer1b <- lmer(dlco_pct ~ 1 + time_firstPFT*SO4_5yrPreDx + (1 + time_firstPFT| ID), data=post2015)
summary(DLCO_lmer1b)
icc(DLCO_lmer1b)
report(DLCO_lmer1b)
```


### Pre/Post-2015 - Complete Model
```{r}
DLCO_lmer2a <- lmer(dlco_pct ~ 1 + time_firstPFT*SO4_5yrPreDx + time_firstPFT*dich_Race + time_firstPFT*sex + time_firstPFT*age_dx + time_firstPFT*smokeHx + time_firstPFT*avg_s + time_firstPFT*site + (1 + time_firstPFT | ID), data=pre2015)
summary(DLCO_lmer2a)
icc(DLCO_lmer2a)
report(DLCO_lmer2a)
```

```{r}
DLCO_lmer2b <- lmer(dlco_pct ~ 1 + time_firstPFT*SO4_5yrPreDx + time_firstPFT*dich_Race + time_firstPFT*sex + time_firstPFT*age_dx + time_firstPFT*smokeHx + time_firstPFT*avg_s + time_firstPFT*site + (1 + time_firstPFT | ID), data=post2015)
summary(DLCO_lmer2b)
icc(DLCO_lmer2b)
report(DLCO_lmer2b)
```


## Pre/Post-2015 - Dichotomized SO4 5yrs Pre-Diagnosis and DLCO Models with Random Intercept and Random Slope
### Pre/Post-2015 - Base Model
```{r}
DLCO_lmer1a <- lmer(dlco_pct ~ 1 + time_firstPFT*SO45yr_dich + (1 + time_firstPFT| ID), data=pre2015)
summary(DLCO_lmer1a)
icc(DLCO_lmer1a)
report(DLCO_lmer1a)
```

```{r}
DLCO_lmer1b <- lmer(dlco_pct ~ 1 + time_firstPFT*SO45yr_dich + (1 + time_firstPFT| ID), data=post2015)
summary(DLCO_lmer1b)
icc(DLCO_lmer1b)
report(DLCO_lmer1b)
```

### Pre/Post-2015 - Complete Model
```{r}
DLCO_lmer2a <- lmer(dlco_pct ~ 1 + time_firstPFT*SO45yr_dich + time_firstPFT*dich_Race + time_firstPFT*sex + time_firstPFT*age_dx + time_firstPFT*smokeHx + time_firstPFT*avg_s + time_firstPFT*site + (1 + time_firstPFT | ID), data=pre2015)
summary(DLCO_lmer2a)
icc(DLCO_lmer2a)
report(DLCO_lmer2a)
```

```{r}
DLCO_lmer2b <- lmer(dlco_pct ~ 1 + time_firstPFT*SO45yr_dich + time_firstPFT*dich_Race + time_firstPFT*sex + time_firstPFT*age_dx + time_firstPFT*smokeHx + time_firstPFT*avg_s + time_firstPFT*site + (1 + time_firstPFT | ID), data=post2015)
summary(DLCO_lmer2b)
icc(DLCO_lmer2b)
report(DLCO_lmer2b)
```


# Pre/Post-2015 - Association between SO4 5yrs Pre-Censoring and DLCO Decline
Linear mixed models with random intercept and slope will be used to evaluate the impact of SO4 During Disease on dlco decline.
## Pre/Post-2015 - Continuous SO4 5yrs Pre-Censoring and DLCO Models with Random Intercept and Random Slope
### Pre/Post-2015 - Base Model
```{r}
DLCO_lmer1a <- lmer(dlco_pct ~ 1 + time_firstPFT*SO4_5yrPreCensor + (1 + time_firstPFT| ID), data=pre2015)
summary(DLCO_lmer1a)
icc(DLCO_lmer1a)
report(DLCO_lmer1a)
```

```{r}
DLCO_lmer1b <- lmer(dlco_pct ~ 1 + time_firstPFT*SO4_5yrPreCensor + (1 + time_firstPFT| ID), data=post2015)
summary(DLCO_lmer1b)
icc(DLCO_lmer1b)
report(DLCO_lmer1b)
```

### Pre/Post-2015 - Complete Model
```{r}
DLCO_lmer2a <- lmer(dlco_pct ~ 1 + time_firstPFT*SO4_5yrPreCensor + time_firstPFT*dich_Race + time_firstPFT*sex + time_firstPFT*age_dx + time_firstPFT*smokeHx + time_firstPFT*avg_s + time_firstPFT*site + (1 + time_firstPFT | ID), data=pre2015)
summary(DLCO_lmer2a)
icc(DLCO_lmer2a)
report(DLCO_lmer2a)
```

```{r}
DLCO_lmer2b <- lmer(dlco_pct ~ 1 + time_firstPFT*SO4_5yrPreCensor + time_firstPFT*dich_Race + time_firstPFT*sex + time_firstPFT*age_dx + time_firstPFT*smokeHx + time_firstPFT*avg_s + time_firstPFT*site + (1 + time_firstPFT | ID), data=post2015)
summary(DLCO_lmer2b)
icc(DLCO_lmer2b)
report(DLCO_lmer2b)
```


## Pre/Post-2015 - Dichotomized SO4 5yrs Pre-Censoring and DLCO Models with Random Intercept and Random Slope
### Pre/Post-2015 - Base Model
```{r}
DLCO_lmer1a <- lmer(dlco_pct ~ 1 + time_firstPFT*SO45yrCensor_dich + (1 + time_firstPFT| ID), data=pre2015)
summary(DLCO_lmer1a)
icc(DLCO_lmer1a)
report(DLCO_lmer1a)
```

```{r}
DLCO_lmer1b <- lmer(dlco_pct ~ 1 + time_firstPFT*SO45yrCensor_dich + (1 + time_firstPFT| ID), data=post2015)
summary(DLCO_lmer1b)
icc(DLCO_lmer1b)
report(DLCO_lmer1b)
```

### Pre/Post-2015 - Complete Model
```{r}
DLCO_lmer2a <- lmer(dlco_pct ~ 1 + time_firstPFT*SO45yrCensor_dich + time_firstPFT*dich_Race + time_firstPFT*sex + time_firstPFT*age_dx + time_firstPFT*smokeHx + time_firstPFT*avg_s + time_firstPFT*site + (1 + time_firstPFT | ID), data=pre2015)
summary(DLCO_lmer2a)
icc(DLCO_lmer2a)
report(DLCO_lmer2a)
```

```{r}
DLCO_lmer2b <- lmer(dlco_pct ~ 1 + time_firstPFT*SO45yrCensor_dich + time_firstPFT*dich_Race + time_firstPFT*sex + time_firstPFT*age_dx + time_firstPFT*smokeHx + time_firstPFT*avg_s + time_firstPFT*site + (1 + time_firstPFT | ID), data=post2015)
summary(DLCO_lmer2b)
icc(DLCO_lmer2b)
report(DLCO_lmer2b)
```













